### 2013-07-17 idp/documentation/paper.tex
aded initial IDP documentation files

\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{parskip}
\usepackage{times}
\usepackage{multirow}

\clubpenalty = 10000
\widowpenalty = 10000

\title{Optimal Fitting of Curves for Given Specifications}
\author{Julian Asamer, Julian Brunner}
\date{\today}

\begin{document}

	\maketitle

	\begin{abstract}

		\noindent TODO

	\end{abstract}

	\section{Introduction}

	\section{Goals}

	\section{Proposed Solution}

	\section{Implementation}

	\section{Evaluation}

	\section{Conclusion}

		\begin{table}[htbp]
			\centering
			\begin{tabular}{c|c|r|r|}
				\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Actual Choice} \\
				\cline{2-4}
				& Payout & \multicolumn{1}{|c|}{A and B} & \multicolumn{1}{|c|}{B only} \\
				\cline{2-4}
				\multirow{2}{*}{Predicted Choice} & A and B & 1 000 \$ & 0 \$ \\
				\cline{2-4}
				& B only & 1 001 000 \$ & 1 000 000 \$ \\
				\cline{2-4}
			\end{tabular}
			\caption{Possible payouts for Newcomb's paradox}
			\label{table:NewcombParadoxPayouts}
		\end{table}

	\begin{thebibliography}{}

		\bibitem{CranePatterson2000}
			\emph{History of the mind-body problem}\\
			Tim Crane, Sarah Patterson\\
			Routledge, 2000

	\end{thebibliography}

\end{document}

### 2013-07-17 idp/documentation/paper.tex
updated title to match official application, started working on introduction, goals and proposed solution sections

@@
 
-\title{Optimal Fitting of Curves for Given Specifications}
+\title{Optimal Fitting of Planar Curves to Prescribed Constraints}
 \author{Julian Asamer, Julian Brunner}
@@
 
+		These days, vector graphics are widely used in both artistic and industrial design.
+Artists create illustrations, character designs and even fully animated cartoons using vector graphics software.
+Vector graphics are also used in industrial design, allowing for the smooth, aerodynamic shapes observed in modern cars and airplanes.
+
+		Raster graphics images (for instance, digital photos and most digitally painted artworks) consist of a finite number of pixels (picture elements) which discretely specify the colors at each point of the image.
+In contrast, vector graphics designs consist of primitives like curves and surfaces, which describe the shapes present in the design using continuous mathematical functions.
+This has various advantages over the raster approach, since designs do not suffer from discretization artifacts and can, for instance, be arbitrarily scaled without any loss of quality.
+
+		The scope of this project is restricted to two-dimensional vector graphics, but many ideas can be applied to three-dimensional vector graphics as well.
+All elements in two-dimensional vector graphics are described in terms of curves, so the design process of curves is the single most important part of two-dimensional vector graphics.
+Unfortunately, designing curves using the tools available through current vector graphics software can be very frustrating.
+The artist is often unable to make the curve look the way they want it to.
+To make matters worse it is often unclear why this is the case, as the software seems to put all the necessary tools into the hands of the designer, yet at times, they somehow don't allow the artist to communicate their intention to the software.
+Thus, the objective of this project is to identify the exact nature and cause of these inadequacies and then specify and implement curve design tools that do not exhibit them.
+
 	\section{Goals}
 
+		\subsection{Shortcomings of Existing Curve Design Software}
+
+			% shortcomings of existing solutions (bézier, spiro (link to spiro promotion), NURBS)
+			%   TODO: maybe correspond these shortcomings and their illustrations to the stuff in the evalution section
+			%   TODO: research shortcomings that spiro splines do not solve, maybe we can have illustrations for that, too
+			%   illustration: it's easy to create parabolas instead of circle arcs
+			%   illustration: spirals need many control points and it's easy to include small bumps
+			%   illustration: curvature continuity is not guaranteed
+
+		\subsection{Properties of Good Curve Design Software}
+
+			% https://code.google.com/p/kurve/wiki/Foundation
+			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles
+
 	\section{Proposed Solution}
+
+		\subsection{Specification-Based Curves}
+
+			% moving from implementation detail to differential geometric specification
+			% implementation section of https://code.google.com/p/kurve/wiki/Foundation
+
+		\subsection{Nonlinear Optimization on Polynomial Curves}
+
+			% segmentation, separation of segmentation and specification, global positions on curve (https://code.google.com/p/kurve/wiki/SegmentationParametrization, maybe diagram)
+			% disambiguation
+			% specifications
+			%   position, direction and curvature
+			% continuity connections
+			% constant velocity
+			% fairness
+			%   derivative of projected acceleration is equal to projected jerk
 

### 2013-07-18 idp/documentation/paper.tex
started working on usability of curve design software section

@@
 
-	\section{Goals}
+	\section{Usability of Curve Design Software}
+
+		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its shortcomings so as to find a way to improve it without making the same or similar mistakes.
+In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves.
+Once such a model has been obtained, it can be used to understand the issues with current tools as well as to propose criteria for good curve design tools.
 
@@
 			%   illustration: curvature continuity is not guaranteed
+
+		\subsection{Curve Design Process}
 

### 2013-07-18 idp/documentation/paper.tex
continued working on usability section

@@
 
-	\section{Usability of Curve Design Software}
+		% TODO: make sure introduction does not overlap with start of next section regarding both content and wording
+
+	\section{Usability of Curve Design Tools}
 
@@
 In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves.
-Once such a model has been obtained, it can be used to understand the issues with current tools as well as to propose criteria for good curve design tools.
+Once such a model has been obtained, it can be used to understand the issues with current software as well as to propose criteria for good curve design tools.
 
-		\subsection{Shortcomings of Existing Curve Design Software}
+		\subsection{Curve Design Process}
+
+			By observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible.
+
+			\begin{enumerate}
+				\item obtain the source curve as either a real or a mental image of the curve
+				\item extract a set of properties that uniquely identify the source curve from it
+				\item provide these properties to the software to obtain the result curve
+			\end{enumerate}
+
+			Steps 2 and 3 are then repeated on portions of the curve in order to make small adjustments until the result curve is sufficiently similar to the source curve.
+
+			The crucial point here is that the set of properties that is used to communicate the perceived source curve in step 2 to the software in step 3 has to be formulated in the `language' the software uses to describe curves.
+Different languages may differ in their well-suitedness to be `read' from the source curve by humans, to be `spoken' by humans to the software and in their ability to `describe' curves succinctly and uniquely.
+
+		\subsection{Shortcomings of Existing Curve Design Tools}
 
@@
 
-		\subsection{Curve Design Process}
+		\subsection{Criteria for Good Cruve Design Tools}
 
-		\subsection{Properties of Good Curve Design Software}
+%			From this model, one can derive the requirements that the software tools need to fulfill in order to be useful to the artist.
+%
+%			\begin{itemize}
+%				\item The conversion "curve - defining parameters" should be easy to comprehend (follows from 2 and 5)
+%				\item The parameters should be easy to input to the software (follows from 3)
+%				\item The effects of changing parameters should be local (follows from 5)
+%			\end{itemize}
+%
+%			One may note that the choice of parameters that are used to define the result curve has the most significant influence on the usability of the resulting software tool.
+It follows that it would be advantageous to make most of the project independent of this choice.
+That way, there is room for experimenting with the parameters once the framework is established, in order to find the most suitable choice.
 

### 2013-07-18 idp/documentation/paper.tex
continued working on usability section

--- 2013-07-18 idp/documentation/paper.tex
@@
 In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves.
-Once such a model has been obtained, it can be used to understand the issues with current tools as well as to propose criteria for good curve design tools.
+Once such a model has been obtained, it can be used to understand the issues with current software as well as to propose criteria for good curve design tools.
 
-		\subsection{Shortcomings of Existing Curve Design Software}
+		\subsection{Curve Design Process}
+
+			By observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible.
+
+			\begin{enumerate}
+				\item obtain the source curve as either a real or a mental image of the curve
+				\item extract a set of properties from the source curve that uniquely identify it
+				\item provide these properties to the software
+				\item have the software turn them into the result curve
+			\end{enumerate}
+
+			Steps 2 to 3 are then repeated on portions of the curve in order to make small adjustments until the result curve is sufficiently similar to the source curve.
+
+			Step 1 is of course independent of any software that may be used by the designer.
+Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as descriptions for curves (the `language' the software uses).
+Finally, the success in step 4 depends solely on the ability of the software to turn descriptions of curves that are given to it into mathematical curves.
+
+			The influence which the design tool has on the success or failure of this whole process thus depends on the following key points.
+
+			\begin{itemize}
+				\item Can the language describe curves succinctly and uniquely?
+				\item Is it easy for humans to `read' the language from curves?
+				\item Is it easy for humans to `speak' the language to the software?
+				\item Is it possible for software to build curves from descriptions in the language?
+			\end{itemize}
+
+			At this point it seems obvious that the choice of language is the single most influential one when it comes to the usability of curve design software.
+Thus, usability analysis in the subsequent sections will focus almost exclusively on this concept of language and the key points made in the list above.
+
+		\subsection{Shortcomings of Existing Curve Design Tools}
+
+			Using the model of the curve design process established in the previous section, various tools available in current vector graphics software can be analyzed
+
+			\subsubsection{Free-Hand Curves}
+
+			\subsubsection{Bézier Splines}
+
+			\subsubsection{Spiro Splines}
 
@@
 
-		\subsection{Curve Design Process}
-
-		\subsection{Properties of Good Curve Design Software}
+		\subsection{Criteria for Good Cruve Design Tools}
 
@@
 
+			% with language being this important, it seems irresponsible to simply choose to use the mathematical parameters of the model as the language
 			% moving from implementation detail to differential geometric specification

### 2013-07-18 idp/documentation/paper.tex
continued working on usability section

@@
 \usepackage{times}
+\usepackage{mathtools}
+\usepackage{amssymb}
 \usepackage{multirow}
@@
 
-		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its shortcomings so as to find a way to improve it without making the same or similar mistakes.
+		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings so as to find a way to improve it without making the same or similar mistakes.
 In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves.
-Once such a model has been obtained, it can be used to understand the issues with current software as well as to propose criteria for good curve design tools.
+Once such a model has been obtained, it can be used to propose criteria for good curve design software as well as to analyze potential issues with existing tools.
 
@@
 
-		\subsection{Shortcomings of Existing Curve Design Tools}
+		\subsection{Existing Curve Design Tools}
 
-			Using the model of the curve design process established in the previous section, various tools available in current vector graphics software can be analyzed
+			Using the model of the curve design process as well as the criteria for good curve design software that were established in the previous section, the usability of various tools available in current vector graphics software can be analyzed.
 
-			\subsubsection{Free-Hand Curves}
+			\subsubsection{Free-Hand Curve Design}
 
-			\subsubsection{Bézier Splines}
+			\subsubsection{Common Design Tools for Bézier Splines}
 
-			\subsubsection{Spiro Splines}
+				Bézier splines are probably the most widely used type of curve in two-dimensional vector graphics.
+Bézier splines are defined piecewise using Bézier curves, which are planar parametric curves (functions \(\left[0,1\right] \rightarrow \mathbb{R}^2\)) realized using polynomials (usually cubic) in Bernstein form.
+A Bézier curve \(B\) of degree \(n\) consists of \(n + 1\) coefficients \(P_i \in \mathbb{R}^2\).
+The curve will always go through the first and the last coefficient point (\(B\left(0\right) = P_0, B\left(1\right) = P_n\)) and stay within the convex hull of the coefficient points at all other times.
+Bézier curves in a Bézier spline are usually pieced together in such a way as to make the tangent direction along the curve a continuous function.
+
+				In existing vector graphics software, Bézier splines are edited by more or less directly modifying the coefficient points of the underlying Bézier curves.
+Most vector graphics tools have functions that allow preserving the continuity of the tangent direction by automatically modifying adjacent coefficient points accordingly.
+Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more Bézier curves to the spline.
+
+			\subsubsection{Common Design Tools for Spiro Splines}
 
@@
 
-		\subsection{Criteria for Good Cruve Design Tools}
-
-			% https://code.google.com/p/kurve/wiki/Foundation
-			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles
-
-	\section{Proposed Solution}
-
 		\subsection{Specification-Based Curves}
@@
 			% implementation section of https://code.google.com/p/kurve/wiki/Foundation
+			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles
 
-		\subsection{Nonlinear Optimization on Polynomial Curves}
+	\section{Proposed Solution: Nonlinear Optimization on Polynomial Curves}
 

### 2013-07-18 idp/documentation/paper.tex
continued working on usability section

@@
 		\subsection{Curve Design Process}
+		\label{section:CurveDesignProcess}
 
@@
 
-			The influence which the design tool has on the success or failure of this whole process thus depends on the following key points.
+			The influence which the design tool has on the success or failure of this whole process thus depends on the following criteria.
 
-			\begin{itemize}
+			\begin{enumerate}
 				\item Can the language describe curves succinctly and uniquely?
-				\item Is it easy for humans to `read' the language from curves?
-				\item Is it easy for humans to `speak' the language to the software?
+				\item Is it easy for humans to `read' descriptions in the language from curves?
+				\item Is it easy for humans to `speak' descriptions in the language to the software?
 				\item Is it possible for software to build curves from descriptions in the language?
-			\end{itemize}
+			\end{enumerate}
 
 			At this point it seems obvious that the choice of language is the single most influential one when it comes to the usability of curve design software.
+Ideally, the language should be very expressive, facilitating points 1 to 3, but it should still be possible to fulfill point 4 and derive mathematical curves from a description, a process which becomes increasingly difficult with more expressive description languages.
 Thus, usability analysis in the subsequent sections will focus almost exclusively on this concept of language and the key points made in the list above.
@@
 			\subsubsection{Free-Hand Curve Design}
+
+				While many people would not consider free-hand curve design a form of vector graphics curve design, it is well-suited to demonstrate how the model of the curve design process and the concept of a communication language that the designer uses to convey the curve to the software can be used to analyze the usability of curve design tools.
+
+				Free-hand curves are created by the designer using some kind of pointing device (for instance, a mouse or a graphics tablet), to convey a large number of points that are on the curve to the software.
+For the simplicity of the argument, we will assume that the graphics software will simply connect these points using straight lines.
+Some vector graphics software will use Bézier curves instead of straight lines, but due to the large number of points involved as well as due to the low precision of free-hand input using a pointing device, this does not make much of a difference.
+
+				Applying the criteria from section \ref{section:CurveDesignProcess} to free-hand curve design, it is immediately obvious that hundreds of points along the curve are not a succinct way of describing a curve, so criterion 1 is definitely not met very well.
+Criterion 2 on the other hand seems fine, since all the points that lie on the curve are immediately apparent when looking at it.
+Point 3 is looking promising aswell, since it's easy enough to input points into some vector graphics software, although this is severely hindered by the fact that there are so many points that need to be specified.
+Criterion 4 is very well met again, since the software has no problem constructing a polygon from a few hundred points.
+
+				Note that this only applies if the software actually uses the points in the described fashion, that is, by building a curve that goes through all the points.
+If, on the other hand, the software is tasked with striking a balance between a curve that goes through as many points as possible while also remaining as smooth as possible, this has to be regarded as an entirely different communication language.
+In this case, the points the user creates using his pointing device are merely an intermediate representation, what is actually communicated to the software is the smooth motion the user makes while using the pointing device.
+The success of this approach depends largely on the specific algorithm the software uses to fit a smooth curve through the given points as well as the curve primitive that is eventually used to represent the final curve.
 
@@
 
+				% TODO - illustrations
+				% curves that are best described in terms of curvature are difficult to put down
+				%   an excessive number of control points are needed to pin down curvature
+				%     compromises smoothness
+				%     makes the curve hard to change
+				%     somewhat okay when tracing an existing image, since the shapes are already final and the source is likely smooth
+				%   examples
+				%     circle arcs (may have too much curvature in the middle, no constant curvature)
+				%     spirals (fluttershy's mane)
+				%     lugia's back line
+				% changind direction at some node usually adversely affects neighboring spline segments
+				%   changing the direction on the left side of a node with a long segment to the right screws up the right side
+				%     bézier splines do not have much degrees of freedom left after direction continuity is ensured, even though curvature continuity may still be violated
+				%     requires additional segments, tradeoff between smoothness and locality, not much tool support for retaining as much smoothness as possible while adding more freedom
+				% curvature continuity is not guaranteed
+				%   curves with lots of changes in curvature sign are affected the most
+				%   examples
+				%     lugia's belly
+
 			\subsubsection{Common Design Tools for Spiro Splines}
 
-			% shortcomings of existing solutions (bézier, spiro (link to spiro promotion), NURBS)
-			%   TODO: maybe correspond these shortcomings and their illustrations to the stuff in the evalution section
-			%   TODO: research shortcomings that spiro splines do not solve, maybe we can have illustrations for that, too
-			%   illustration: it's easy to create parabolas instead of circle arcs
-			%   illustration: spirals need many control points and it's easy to include small bumps
-			%   illustration: curvature continuity is not guaranteed
+				% TODO - illustrations
+				% would in principle be well-suited for circle arcs and spirals and other shapes with linearly changing curvature, but curvature can only be controlled indirectly using points
+				% tangent direction and curvature can only be controlled indirectly
+				%   sometimes, many nodes need to be placed to convey the desired direction/curvature information
+				%   specifying direction/curvature always comes with the unwanted side-effect of specifying additional points, possibly introducing unwanted bumpiness
+				% sometimes, nice, smooth curves can be achieved using only few points
+				%   it's not obvious where the points have to be specified such that one needs to specify as few as possible
+				%   changing those 'minimal' curves is somewhat hampered by the fact that the purpose of each point may be non-obvious
+				%     has bad locality since the points on the curve don't really specify points, but have an ulterior purpose
+				% suffers from instability issues
+				% does guarantee curvature continuity
 
@@
 
-			% with language being this important, it seems irresponsible to simply choose to use the mathematical parameters of the model as the language
+			% with language being this important, it seems irresponsible to have it determined by the mathematical parameters of the model as the language
 			% moving from implementation detail to differential geometric specification
 			% implementation section of https://code.google.com/p/kurve/wiki/Foundation
+            % TODO - diagram: process of going from user input and fairness to curve
 			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles
@@
 	\section{Evaluation}
+
+			% demonstrate euler spirals and bézier curves being embedded
+			% revisit examples made when analyzing Bézier splines and Spiro splines
 

### 2013-07-19 idp/documentation/paper.tex
continued working on usability section

@@
 				\item obtain the source curve as either a real or a mental image of the curve
-				\item extract a set of properties from the source curve that uniquely identify it
+				\item extract some properties from the source curve
 				\item provide these properties to the software
-				\item have the software turn them into the result curve
+				\item have the software derive the most likely result curve from these properties
 			\end{enumerate}
@@
 Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as descriptions for curves (the `language' the software uses).
-Finally, the success in step 4 depends solely on the ability of the software to turn descriptions of curves that are given to it into mathematical curves.
+Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve (an unlikely situation in curve design), in step 4, the design tool will have to choose the curve that the user was most likely describing from an infinite set of curves that exhibit the supplied properties.
+This notion of 'most likely' is usually modelled in terms of a fairness measure, fairer curves being regarded as more likely.
+Fairness ususally encodes intuitive notions such as smoothness using mathematical concepts such as continuity and curvature.
+Here, the software has an opportunity to influence the design process both positively and negatively depending on how the fairness measure is chosen as well as by how good the software is at reliably finding a curve that has a high fairness.
 
-			The influence which the design tool has on the success or failure of this whole process thus depends on the following criteria.
+			At this point it seems like the two ways in which curve design software can affect the design process are through the choice of language and fairness measure.
+Ideally, the language should be very expressive, enabling the user to communicate their intent to the software efficiently, while it and the fairness measure should still retain the mathematical properties that allow curves to be derived from them, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures.
+
+			From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.
 
 			\begin{enumerate}
-				\item Can the language describe curves succinctly and uniquely?
-				\item Is it easy for humans to `read' descriptions in the language from curves?
-				\item Is it easy for humans to `speak' descriptions in the language to the software?
-				\item Is it possible for software to build curves from descriptions in the language?
+				\item How well can the language describe curves?
+				\item How well can humans to `read' descriptions in the language from curves?
+				\item How well can humans to `speak' descriptions in the language to the software?
+				\item How well does the fairness measure capture the intuitive notion of smoothness?
+				\item How well can the software derive curves from descriptions and fairness?
 			\end{enumerate}
-
-			At this point it seems obvious that the choice of language is the single most influential one when it comes to the usability of curve design software.
-Ideally, the language should be very expressive, facilitating points 1 to 3, but it should still be possible to fulfill point 4 and derive mathematical curves from a description, a process which becomes increasingly difficult with more expressive description languages.
-Thus, usability analysis in the subsequent sections will focus almost exclusively on this concept of language and the key points made in the list above.
 
@@
 				%     lugia's back line
-				% changind direction at some node usually adversely affects neighboring spline segments
-				%   changing the direction on the left side of a node with a long segment to the right screws up the right side
-				%     bézier splines do not have much degrees of freedom left after direction continuity is ensured, even though curvature continuity may still be violated
-				%     requires additional segments, tradeoff between smoothness and locality, not much tool support for retaining as much smoothness as possible while adding more freedom
 				% curvature continuity is not guaranteed

### 2013-07-19 source/Kurve/Kurve.Curves/Optimization/OptimizationProblem.cs

@@ -112,16 +112,16 @@ namespace Kurve.Curves.Optimization
 					(
 						segment0CurveVelocity.Apply(Terms.Constant(1)),
 						segment1CurveVelocity.Apply(Terms.Constant(0))
-					),
-
-					from segmentIndex in Enumerable.Range(0, optimizationSegments.Segments.Count() - 1)
-					let segment0CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 0).LocalCurve.Acceleration
-					let segment1CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 1).LocalCurve.Acceleration
-					select Constraints.CreateEquality
-					(
-						segment0CurveAcceleration.Apply(Terms.Constant(1)),
-						segment1CurveAcceleration.Apply(Terms.Constant(0))
 					)
+
+//					from segmentIndex in Enumerable.Range(0, optimizationSegments.Segments.Count() - 1)
+//					let segment0CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 0).LocalCurve.Acceleration
+//					let segment1CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 1).LocalCurve.Acceleration
+//					select Constraints.CreateEquality
+//					(
+//						segment0CurveAcceleration.Apply(Terms.Constant(1)),
+//						segment1CurveAcceleration.Apply(Terms.Constant(0))
+//					)
 				)
 			)
 			.ToArray();

### 2013-07-19 source/Kurve/Kurve/Components/Controls/SegmentComponent.cs

@@ -49,6 +49,13 @@ namespace Kurve.Component
 
 				Krach.Graphics.Color color = StretchedColor(stretchFactor);
 				Drawing.DrawLine(context, Curve.GetPoint(positions.Item1), Curve.GetPoint(positions.Item2), 2, color);
+
+				Vector2Double point = 0.5 * (Curve.GetPoint(positions.Item1) + Curve.GetPoint(positions.Item2));
+				Vector2Double direction = Curve.GetDirection((positions.Item1 + positions.Item2) / 2);
+				Vector2Double angularDirection = new Vector2Double(direction.Y, -direction.X);
+				double curvature = Curve.GetCurvature((positions.Item1 + positions.Item2) / 2);
+				Vector2Double curvatureVector = 10000 * curvature * angularDirection;
+				Drawing.DrawLine(context, point, point + curvatureVector, 2, Colors.Green);
 			}
 
 			base.Draw(context);

### 2013-07-20 source/Kurve/Kurve.Curves/Optimization/OptimizationProblem.cs
reverted accidently commited files for testing curvature continuity of Bézier splines

@@ -112,16 +112,16 @@ namespace Kurve.Curves.Optimization
 					(
 						segment0CurveVelocity.Apply(Terms.Constant(1)),
 						segment1CurveVelocity.Apply(Terms.Constant(0))
-					)
+					),
 
-//					from segmentIndex in Enumerable.Range(0, optimizationSegments.Segments.Count() - 1)
-//					let segment0CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 0).LocalCurve.Acceleration
-//					let segment1CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 1).LocalCurve.Acceleration
-//					select Constraints.CreateEquality
-//					(
-//						segment0CurveAcceleration.Apply(Terms.Constant(1)),
-//						segment1CurveAcceleration.Apply(Terms.Constant(0))
-//					)
+					from segmentIndex in Enumerable.Range(0, optimizationSegments.Segments.Count() - 1)
+					let segment0CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 0).LocalCurve.Acceleration
+					let segment1CurveAcceleration = optimizationSegments.Segments.ElementAt(segmentIndex + 1).LocalCurve.Acceleration
+					select Constraints.CreateEquality
+					(
+						segment0CurveAcceleration.Apply(Terms.Constant(1)),
+						segment1CurveAcceleration.Apply(Terms.Constant(0))
+					)
 				)
 			)
 			.ToArray();

### 2013-07-20 source/Kurve/Kurve/Components/Controls/SegmentComponent.cs

@@ -50,12 +50,12 @@ namespace Kurve.Component
 				Krach.Graphics.Color color = StretchedColor(stretchFactor);
 				Drawing.DrawLine(context, Curve.GetPoint(positions.Item1), Curve.GetPoint(positions.Item2), 2, color);
 
-				Vector2Double point = 0.5 * (Curve.GetPoint(positions.Item1) + Curve.GetPoint(positions.Item2));
-				Vector2Double direction = Curve.GetDirection((positions.Item1 + positions.Item2) / 2);
-				Vector2Double angularDirection = new Vector2Double(direction.Y, -direction.X);
-				double curvature = Curve.GetCurvature((positions.Item1 + positions.Item2) / 2);
-				Vector2Double curvatureVector = 10000 * curvature * angularDirection;
-				Drawing.DrawLine(context, point, point + curvatureVector, 2, Colors.Green);
+//				Vector2Double point = 0.5 * (Curve.GetPoint(positions.Item1) + Curve.GetPoint(positions.Item2));
+//				Vector2Double direction = Curve.GetDirection((positions.Item1 + positions.Item2) / 2);
+//				Vector2Double angularDirection = new Vector2Double(direction.Y, -direction.X);
+//				double curvature = Curve.GetCurvature((positions.Item1 + positions.Item2) / 2);
+//				Vector2Double curvatureVector = 10000 * curvature * angularDirection;
+//				Drawing.DrawLine(context, point, point + curvatureVector, 2, Colors.Green);
 			}
 
 			base.Draw(context);

### 2013-07-22 idp/documentation/paper.tex
continued working on usability section

--- 2013-07-18 idp/documentation/paper.tex
@@
 \usepackage[T1]{fontenc}
+\usepackage{a4wide}
 \usepackage{parskip}
 \usepackage{times}
+\usepackage{textcomp}
 \usepackage{mathtools}
@@
 \date{\today}
+
+% TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
 
@@
 		\subsection{Curve Design Process}
+		\label{section:curve_design_process}
 
@@
 				\item obtain the source curve as either a real or a mental image of the curve
-				\item extract a set of properties from the source curve that uniquely identify it
+				\item extract some properties from the source curve
 				\item provide these properties to the software
-				\item have the software turn them into the result curve
+				\item have the software derive the most likely result curve from these properties
 			\end{enumerate}
 
-			Steps 2 to 3 are then repeated on portions of the curve in order to make small adjustments until the result curve is sufficiently similar to the source curve.
+			Steps 2 to 4 are then repeated on portions of the curve in order to make small adjustments until the result curve is sufficiently similar to the source curve.
 
@@
 Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as descriptions for curves (the `language' the software uses).
-Finally, the success in step 4 depends solely on the ability of the software to turn descriptions of curves that are given to it into mathematical curves.
+Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve, in step 4, the design tool will have to choose the curve that the user was most likely describing from a possibly infinite set of curves that exhibit the supplied properties.
+This notion of `most likely' is usually modelled in terms of a fairness measure, fairer curves being regarded as more likely.
+Fairness ususally encodes intuitive notions such as smoothness using mathematical concepts such as continuity and change in curvature.
+Here, the software has another opportunity to influence the design process both positively and negatively depending on how the fairness measure is chosen as well as by how good the software is at reliably finding a curve that has a high fairness.
 
-			The influence which the design tool has on the success or failure of this whole process thus depends on the following key points.
+			Note that there may be multiple ways to describe a given curve design tool in terms of this process.
+Instead of modelling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify.
+For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve.
+Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + (p_1 - p_0) \cdot t\), thus uniquely identifying the desired curve.
+Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
+In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the language (the language does not allow the user to request curvature continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
 
-			\begin{itemize}
-				\item Can the language describe curves succinctly and uniquely?
-				\item Is it easy for humans to `read' the language from curves?
-				\item Is it easy for humans to `speak' the language to the software?
-				\item Is it possible for software to build curves from descriptions in the language?
-			\end{itemize}
+		\subsection{Usability of Curve Design Tools}
+		\label{section:usability_curve_design_tools}
 
-			At this point it seems obvious that the choice of language is the single most influential one when it comes to the usability of curve design software.
-Thus, usability analysis in the subsequent sections will focus almost exclusively on this concept of language and the key points made in the list above.
+			At this point, it seems like the two ways in which curve design software can affect the design process are through the choice of language and fairness measure.
+Ideally, the language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
+The fairness should capture the intuitive notion of curve smoothness sufficiently well.
+These goals are competing with ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures.
+From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.
+
+			\begin{enumerate}
+				\item How efficient is the language at describing various curves?
+				\item How well can humans `read' descriptions in the language from curves?
+				\item How well can humans `speak' descriptions in the language to the software?
+				\item How well does the fairness measure capture the intuitive notion of smoothness?
+				\item How well can the software derive curves from descriptions and a fairness measure?
+			\end{enumerate}
 
@@
 
-			Using the model of the curve design process as well as the criteria for good curve design software that were established in the previous section, the usability of various tools available in current vector graphics software can be analyzed.
+			Using the model of the curve design process as well as the criteria for usability of curve design software that were established in the previous section, various tools available in current vector graphics software can be analyzed.
 
-			\subsubsection{Free-Hand Curve Design}
-
+			% overview of bézier spline design tools
+			% applying curve design model to bézier spline design tools
+			% analysis of bézier spline design tools in terms of criteria
+			%   works well: read, speak, derive
+			%   has issues
+			%     fairness
+			%       curvature continuity is not ensured
+			%       ensuring curvature continuity removes some degree of freedom, curves need more nodes
+			%     expressiveness
+			%       curves that are best described in terms of curvature
+			%         need many nodes
+			%           compromises smoothness
+			%           makes the curve hard to change
+			%           no curvature continuity guaranteed at nodes
+			%         are difficult to specify, mistakes are easy
 			\subsubsection{Common Design Tools for Bézier Splines}
@@
 				In existing vector graphics software, Bézier splines are edited by more or less directly modifying the coefficient points of the underlying Bézier curves.
+In the case of cubic Bézier curves, the first and the last coefficient points represent the curve's start and end points, while the differences between the second and the first as well as the fourth and the third coefficient points are proportional to the velocity of the curve at the start and end points, respectively.
+In that sense, the user can specify start end end points of the curve, as well as start and end velocities of the curve using the coefficient points.
 Most vector graphics tools have functions that allow preserving the continuity of the tangent direction by automatically modifying adjacent coefficient points accordingly.
+They usually do not ensure curvature continuity though.
 Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more Bézier curves to the spline.
+
+				When applying the model introduced in section \ref{section:curve_design_process} to Bézier spline design tools, there are two possibilities for modelling the fairness measure.
+One way to view things is to assume that the user, in supplying the four coefficient points of each curve segment, uniquely specifies each segment and thus the entire curve, leaving no room for choosing a curve according to some fairness measure.
+Since users usually don't think about Bézier splines that way, it is more practical to assume that the user only supplies both point and velocity at the start and end of each curve segment, with a fairness measure designed to always choose the cubic polynomial that is uniquely specified by the given properties as the fairest curve.
+
+				Applying the criteria established in section \ref{section:usability_curve_design_tools}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their language (criteria 2 and 3), a skill that can be acquired with some practice.
+Bézier splines are furthermore very simple to derive from their description, fulfilling the 5th criterion perfectly.
+Design tools for Bézier splines have some issues with both the fairness measure (criterion 4) and the expressiveness of the language (criterion 1), both of which shall be analyzed in the following paragraphs.
+
+				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guarantee curvature continuity.
+It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it very difficult to create smooth curves if the design software does not support the user in this regard.
+The problem is aggravated by the fact that curvature incontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
+In case the software allows the user to request curvature-continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
+Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with language expressiveness.
+
+				The language used to specify Bézier splines in common design tools is lacking in expressiveness.
+Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature).
+Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment.
+There is also no way to directly specify linearly changing curvature.
+This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result.
+Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created, respectively moved in order to change the overall curve.
+It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
+Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous.
+In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points, making it easy to make small mistakes.
+
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/usability_bezier.pdf}
+					\caption{Usability Analysis of Bézier Splines}
+					\label{figure:usability_bézier}
+				\end{figure}
+
+				% TODO: improve example index numbers
+				Figure \ref{figure:usability_bézier} illustrates some of these issues by showing some examples of Bézier splines as they appear in common design tools.
+Bézier nodes are displayed as filled dots, while Bézier handles are displayed as circles.
+
+				Example 1 shows two Bézier splines.
+The left curve is a circle arc, exhibiting constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
+Since the difference can be hard to see, it's easy to make mistakes when trying to create circle arcs.
+
+				Examples 2 and 3 show curves that could be described more efficiently in terms of curvature.
+Since there is no way to directly specify curvature using Bézier splines, a considerable number of nodes is needed to get close to the desired curve.
+It may be observed that the spiral in example 2 is not perfect.
+Creating perfect spirals using Bézier splines is very tedious and would require even more nodes (at least one for every 90\textdegree of winding angle covered).
+It could be speculated on whether the artist would have opted to use a perfect spiral in this example if they had had the tools efficiently do so.
+
+				% TODO: consider removing the handle markers to make curvature incontinuity more visible
+				Example 4 shows a spline with a curvature incontinuity which went unnoticed by the artist at the time of creation.
+In fact, this incontinuity would be even less visible if the curve were not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.
 
@@
 
-			% shortcomings of existing solutions (bézier, spiro (link to spiro promotion), NURBS)
-			%   TODO: maybe correspond these shortcomings and their illustrations to the stuff in the evalution section
-			%   TODO: research shortcomings that spiro splines do not solve, maybe we can have illustrations for that, too
-			%   illustration: it's easy to create parabolas instead of circle arcs
-			%   illustration: spirals need many control points and it's easy to include small bumps
-			%   illustration: curvature continuity is not guaranteed
+				% would in principle be well-suited for circle arcs and spirals and other shapes with linearly changing curvature, but curvature can only be controlled indirectly using points
+				% tangent direction and curvature can only be controlled indirectly
+				%   sometimes, many nodes need to be placed to convey the desired direction/curvature information
+				%   specifying direction/curvature always comes with the unwanted side-effect of specifying additional points, possibly introducing unwanted bumpiness
+				% sometimes, nice, smooth curves can be achieved using only few points
+				%   it's not obvious where the points have to be specified such that one needs to specify as few as possible
+				%   changing those 'minimal' curves is somewhat hampered by the fact that the purpose of each point may be non-obvious
+				%     has bad locality since the points on the curve don't really specify points, but have an ulterior purpose
+				% suffers from instability issues
+				% does guarantee curvature continuity
 
@@
 
-			% with language being this important, it seems irresponsible to simply choose to use the mathematical parameters of the model as the language
+			% with language being this important, it seems irresponsible to have it determined by the mathematical parameters of the model
 			% moving from implementation detail to differential geometric specification
 			% implementation section of https://code.google.com/p/kurve/wiki/Foundation
+            % TODO - diagram: process of going from user input and fairness to curve
 			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles
@@
 		% continuity connections
-			% constant velocity
+		% fixed length, constant velocity
+		%   fixed length is necessary if specification positions are a fraction of the arc-length of the whole curve, otherwise, any change in the specifications may change the curve length and thus move all specifications along the curve
 		% fairness
@@
 
+		% demonstrate euler spirals and bézier curves being embedded
+		% revisit examples made when analyzing Bézier splines and Spiro splines
+		% a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification
+
 	\section{Conclusion}
+
+		% possibility to have curves without fixed length? could be useful for curves which only have specifications at the start and/or end of the curve
 

### 2013-08-05 idp/documentation/paper.tex
removed comments
fixed some grammar mistakes

@@
 The fairness should capture the intuitive notion of curve smoothness sufficiently well.
-These goals are competing with ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures.
+These goals are competing with the software's ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures.
 From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.
@@
 
-			% overview of bézier spline design tools
-			% applying curve design model to bézier spline design tools
-			% analysis of bézier spline design tools in terms of criteria
-			%   works well: read, speak, derive
-			%   has issues
-			%     fairness
-			%       curvature continuity is not ensured
-			%       ensuring curvature continuity removes some degree of freedom, curves need more nodes
-			%     expressiveness
-			%       curves that are best described in terms of curvature
-			%         need many nodes
-			%           compromises smoothness
-			%           makes the curve hard to change
-			%           no curvature continuity guaranteed at nodes
-			%         are difficult to specify, mistakes are easy
 			\subsubsection{Common Design Tools for Bézier Splines}
@@
 
-				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guarantee curvature continuity.
-It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it very difficult to create smooth curves if the design software does not support the user in this regard.
+				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing curvature continuity.
+It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
 The problem is aggravated by the fact that curvature incontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
@@
 This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result.
-Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created, respectively moved in order to change the overall curve.
+Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created, or moved in order to change the overall curve.
 It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
@@
 Creating perfect spirals using Bézier splines is very tedious and would require even more nodes (at least one for every 90\textdegree of winding angle covered).
-It could be speculated on whether the artist would have opted to use a perfect spiral in this example if they had had the tools efficiently do so.
+It could be speculated on whether the artist would have opted to use a perfect spiral in this example if they had had the tools to efficiently do so.
 

### 2013-08-08 idp/documentation/paper.tex
replaced template bibliography entry
added some comments for the included packages, added hyperref package to enable pdf index, continued working on section about spiro splines

@@
 
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{a4wide}
-\usepackage{parskip}
-\usepackage{times}
-\usepackage{textcomp}
-\usepackage{mathtools}
-\usepackage{amssymb}
-\usepackage{multirow}
+\usepackage[utf8]{inputenc}    % enables use of utf8 characters in the input file
+\usepackage[T1]{fontenc}       % enables proper output of unicode characters such that text can be copy-pasted from the pdf file
+\usepackage{hyperref}          % enables pdf index, links for web addresses, intra-document links
+\usepackage{a4wide}            % reduces page margins
+\usepackage{parskip}           % replaces hanging paragraphs with spacing between paragraphs
+\usepackage{times}             % replaces Computer Modern font with something resembling Times
+\usepackage{textcomp}          % adds symbols like \textdegree
+\usepackage{mathtools}         % modern version of ams math package
+\usepackage{amssymb}           % ams math symbols
+\usepackage{multirow}          % enables multirow tables
 
+% penalizes lonely lines at the beginning or end of a page
 \clubpenalty = 10000
@@
 % TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
+% TODO: replace curvature continuity with G^2 continuity
 
@@
 		% TODO: make sure introduction does not overlap with start of next section regarding both content and wording
+		% TODO: comment on that although there has been some research into better curves (quote MVC curves), it was less focused on usability and hasn't had much impact on actual vector graphics software which continues to almost exclusively use bézier splinse to this day
 
@@
 
-			At this point, it seems like the two ways in which curve design software can affect the design process are through the choice of language and fairness measure.
+			We conclude that the two ways in which curve design software can affect the design process are through the choice of language and fairness measure.
 Ideally, the language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
@@
 It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
-The problem is aggravated by the fact that curvature incontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
+The problem is aggravated by the fact that curvature discontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
 In case the software allows the user to request curvature-continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
@@
 
-				% TODO: consider removing the handle markers to make curvature incontinuity more visible
-				Example 4 shows a spline with a curvature incontinuity which went unnoticed by the artist at the time of creation.
-In fact, this incontinuity would be even less visible if the curve were not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.
+				% TODO: consider removing the handle markers to make curvature discontinuity more visible
+				Example 4 shows a spline with a curvature discontinuity which went unnoticed by the artist at the time of creation.
+In fact, this discontinuity would be even less visible if the curve were not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.
 
 			\subsubsection{Common Design Tools for Spiro Splines}
+
+				Spiro splines were added to Inkscape in 2009.
+Even though they were integrated using path effects and thus have a somewhat cumbersome user interface, they constitute a commendable effort of incorporating results from curve research into vector graphics software, something that has hardly happened this far.
+More information on Spiro splines can be found in the dissertation of Raphael Levien \cite{Levien2009}.
+
+				Spiro splines consist of parts of the Euler spiral, which is a curve whose curvature changes linearly with the curve's arc length.
+The user can specify control points through which the software puts an interpolating spline.
+There is no way to directly specify tangent direction and/or curvature.
+The splines are guaranteed to have \(G^2\) continuity.
+
+				Again, the model from section \ref{section:curve_design_process} is applied.
+With Spiro splines, the modelling is fairly straightforward.
+The user supplies a set of points for which the software computes an interpolating spline.
+The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is \(G^2\) continuity and no excessive winding.
+
+				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_curve_design_tools}.
+The language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
+The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4).
+The software is also very good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.
+
+				The only real issue lies in the expressiveness of the language.
 
@@
 		% demonstrate euler spirals and bézier curves being embedded
+		%   minimization of normal jerk squares prefers linear change in curvature
 		% revisit examples made when analyzing Bézier splines and Spiro splines
@@
 
-		\bibitem{CranePatterson2000}
-			\emph{History of the mind-body problem}\\
-			Tim Crane, Sarah Patterson\\
-			Routledge, 2000
+		\bibitem{Levien2009}
+			\emph{From Spiral to Spline: Optimal Techniques in Interactive Curve Design}\\
+			Raphael Linus Levien\\
+			University of California, Berkeley, Fall 2009
 

### 2013-08-08 idp/documentation/paper.tex
continued working on section about spiro splines

@@
 			\subsubsection{Common Design Tools for Bézier Splines}
+			\label{section:common_design_tools_bézier_splines}
 
@@
 This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result.
-Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created, or moved in order to change the overall curve.
+Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created or moved in order to change the overall curve.
 It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
 Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous.
-In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points, making it easy to make small mistakes.
+In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.
 
@@
 				Example 1 shows two Bézier splines.
-The left curve is a circle arc, exhibiting constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
-Since the difference can be hard to see, it's easy to make mistakes when trying to create circle arcs.
+The left curve is a Bézier spline approximation of a circle arc, exhibiting nearly constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
+Since the difference can be hard to see, it's easy to make mistakes when trying to approximate circle arcs using Bézier splines.
 
@@
 The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4).
-The software is also very good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.
+The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.
 
-				The only real issue lies in the expressiveness of the language.
-
-				% would in principle be well-suited for circle arcs and spirals and other shapes with linearly changing curvature, but curvature can only be controlled indirectly using points
-				% tangent direction and curvature can only be controlled indirectly
-				%   sometimes, many nodes need to be placed to convey the desired direction/curvature information
-				%   specifying direction/curvature always comes with the unwanted side-effect of specifying additional points, possibly introducing unwanted bumpiness
-				% sometimes, nice, smooth curves can be achieved using only few points
-				%   it's not obvious where the points have to be specified such that one needs to specify as few as possible
-				%   changing those 'minimal' curves is somewhat hampered by the fact that the purpose of each point may be non-obvious
-				%     has bad locality since the points on the curve don't really specify points, but have an ulterior purpose
-				% suffers from instability issues
-				% does guarantee curvature continuity
+				The only real issue lies in the insufficient expressiveness of the language.
+Section \ref{section:common_design_tools_bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability.
+Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature.
+These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
+This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circle arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
+Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties.
+Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified.
+These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
+That way, when indirectly specifying direction and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
 

### 2013-08-11 idp/documentation/paper.tex
added illustration for spiro spline design tools

@@
 
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/usability_spiro.pdf}
+					\caption{Usability Analysis of Spiro Splines}
+					\label{figure:usability_spiro}
+				\end{figure}
+
+				% TODO: improve example index numbers
+				Figure \ref{figure:usability_spiro} illustrates some of these issues by showing some examples of Spiro splines as they appear in common design tools. Control points are displayed as filled dots.
+
+				Example 1 shows a shape that is in principle well-suited for being represented by parts of the Euler spiral.
+Unfortunately, describing the curve using Spiro splines still requires many control points.
+In contrast to the corresponding example in section \ref{section:common_design_tools_bézier_splines}, the curve in this example is close to a perfect spiral an thus does not match the source image perfectly.
+Additional control points would have to be added in order to approximate the imperfect spiral in the source image better.
+
+				Example 2 shows how some shapes can be described using Spiro splines by placing only very few control points.
+In this case, the shape consists of a circle arc with low curvature (segment a), a segment of linearly increasing curvature (segment b) and finally a circle arc with high curvature (segment c).
+This shape can be described conveniently using a Spiro spline with just 4 control points.
+
+				Unfortunately, as can be seen in example 3, this scenario does not handle modifications very well.
+For instance, increasing the curvature of segment c has to be done indirectly by moving a control point.
+Unfortunately, in this case, this also changes the direction of the curve at that point, such that additional modifications would be needed to correct this.
+The change also adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
+
 		\subsection{Specification-Based Curves}

### 2013-08-13 idp/documentation/paper.tex
continued working in idp paper

@@
 
-% penalizes lonely lines at the beginning or end of a page
+% penalizes orphan lines at the beginning or end of a page
 \clubpenalty = 10000
@@
 % TODO: replace curvature continuity with G^2 continuity
+% TODO: replace language with specification language
 
@@
 	\section{Introduction}
+	\label{section:introduction}
 
@@
 	\section{Usability of Curve Design Tools}
+	\label{section:usability_curve_design_tools}
 
@@
 
-		\subsection{Usability of Curve Design Tools}
-		\label{section:usability_curve_design_tools}
+		\subsection{Usability Analysis}
+		\label{section:usability_analysis}
 
@@
 		\subsection{Existing Curve Design Tools}
+		\label{section:existing_curve_design_tools}
 
@@
 
-				Applying the criteria established in section \ref{section:usability_curve_design_tools}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their language (criteria 2 and 3), a skill that can be acquired with some practice.
+				Applying the criteria established in section \ref{section:usability_analysis}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their language (criteria 2 and 3), a skill that can be acquired with some practice.
 Bézier splines are furthermore very simple to derive from their description, fulfilling the 5th criterion perfectly.
@@
 
-				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_curve_design_tools}.
+				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_analysis}.
 The language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
@@
 		\subsection{Specification-Based Curves}
+		\label{section:specification-based_curves}
 
-			% with language being this important, it seems irresponsible to have it determined by the mathematical parameters of the model
-			% moving from implementation detail to differential geometric specification
-			% implementation section of https://code.google.com/p/kurve/wiki/Foundation
-            % TODO - diagram: process of going from user input and fairness to curve
-			% TODO - illustration: visually introduce point, direction, curvature with tangents and osculating circles
+			Seeing how the specification language and fairness measure almost completely determine all usability aspects of the resulting curve design tool, it seems unwise to choose them based on low-level mathematical aspects of the curve as is the case with Bézier splines.
+Spiro splines are much better in that a lot of effort was put into the fairness measure.
+Unfortunately, the project has set out to solve the problem of constructing interpolating splines, thereby committing to a very restricted specification language right from the start.
+
+			We think that in order to get the best usability, using a good specification language has to be the top priority.
+Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables flexible construction of curves based on some specification and some fairness measure, as illustrated in figure \ref{figure:specification-based_curves}.
+
+			\begin{figure}[htb]
+				\centering
+				\includegraphics[width=\textwidth]{../resources/specification-based_curves.pdf}
+				\caption{Specification-Based Curves}
+				\label{figure:specification-based_curves}
+			\end{figure}
 
@@
 		% specifications
+		%   differential geometric properties
 		%   position, direction and curvature
@@
 		%   minimization of normal jerk squares prefers linear change in curvature
+		% take examples from demo
 		% revisit examples made when analyzing Bézier splines and Spiro splines

### 2013-08-15 idp/documentation/paper.tex
fixed replaced some mentions of curvature continuity with G^2 continuity

@@
 % TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
-% TODO: replace curvature continuity with G^2 continuity
 % TODO: replace language with specification language
@@
 Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
-In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the language (the language does not allow the user to request curvature continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the language (the language does not allow the user to request \(G^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
 
@@
 The curve will always go through the first and the last coefficient point (\(B\left(0\right) = P_0, B\left(1\right) = P_n\)) and stay within the convex hull of the coefficient points at all other times.
-Bézier curves in a Bézier spline are usually pieced together in such a way as to make the tangent direction along the curve a continuous function.
+Bézier curves in a Bézier spline are usually pieced together such that the resulting spline is \(G^1\) continuous.
 
@@
 In that sense, the user can specify start end end points of the curve, as well as start and end velocities of the curve using the coefficient points.
-Most vector graphics tools have functions that allow preserving the continuity of the tangent direction by automatically modifying adjacent coefficient points accordingly.
-They usually do not ensure curvature continuity though.
+Most vector graphics tools have functions that allow preserving the \(G^1\) continuity of the spline by automatically modifying adjacent coefficient points accordingly.
+They usually do not ensure \(G^2\) continuity though.
 Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more Bézier curves to the spline.
@@
 
-				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing curvature continuity.
-It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
+				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing \(G^2\) continuity.
+It is difficult for humans to choose the coefficient points in such a way that \(G^2\) continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
 The problem is aggravated by the fact that curvature discontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
-In case the software allows the user to request curvature-continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
+In case the software allows the user to request \(G^2\) continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
 Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with language expressiveness.
@@
 It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
-Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous.
+Another issue is that without software support for \(G^2\) continuity, using more nodes usually also introduces more points where curvature is not continuous.
 In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.
@@
 		%   fixed length is necessary if specification positions are a fraction of the arc-length of the whole curve, otherwise, any change in the specifications may change the curve length and thus move all specifications along the curve
-		% fairness
+		% fairness (quote paper or thesis on MVC)
 		%   derivative of projected acceleration is equal to projected jerk

### 2013-08-15 idp/documentation/paper.tex
fixed some more wording issues

@@
 
-% penalizes orphan lines at the beginning or end of a page
+% tells LaTeX to avoid orphan lines at the beginning or end of a page
 \clubpenalty = 10000
@@
 % TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
-% TODO: replace language with specification language
 
@@
 			Step 1 is of course independent of any software that may be used by the designer.
-Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as descriptions for curves (the `language' the software uses).
+Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as specifications for curves (the specification language the software uses).
 Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve, in step 4, the design tool will have to choose the curve that the user was most likely describing from a possibly infinite set of curves that exhibit the supplied properties.
@@
 Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
-In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the language (the language does not allow the user to request \(G^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(G^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
 
@@
 
-			We conclude that the two ways in which curve design software can affect the design process are through the choice of language and fairness measure.
-Ideally, the language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
+			We conclude that the two ways in which curve design software can affect the design process are through the choice of specification language and fairness measure.
+Ideally, the specification language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
 The fairness should capture the intuitive notion of curve smoothness sufficiently well.
-These goals are competing with the software's ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures.
+These goals are competing with the software's ability to derive curves from specifications and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive specification languages and more complex fairness measures.
 From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.
@@
 			\begin{enumerate}
-				\item How efficient is the language at describing various curves?
-				\item How well can humans `read' descriptions in the language from curves?
-				\item How well can humans `speak' descriptions in the language to the software?
+				\item How efficient is the specification language at describing various curves?
+				\item How well can humans `read' the specification language from curves?
+				\item How well can humans `speak' the specification language to the software?
 				\item How well does the fairness measure capture the intuitive notion of smoothness?
-				\item How well can the software derive curves from descriptions and a fairness measure?
+				\item How well can the software derive curves from specifications and a fairness measure?
 			\end{enumerate}
@@
 
-				Applying the criteria established in section \ref{section:usability_analysis}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their language (criteria 2 and 3), a skill that can be acquired with some practice.
-Bézier splines are furthermore very simple to derive from their description, fulfilling the 5th criterion perfectly.
-Design tools for Bézier splines have some issues with both the fairness measure (criterion 4) and the expressiveness of the language (criterion 1), both of which shall be analyzed in the following paragraphs.
+				Applying the criteria established in section \ref{section:usability_analysis}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice.
+Bézier splines are furthermore very simple to derive from their specification, fulfilling the 5th criterion perfectly.
+Design tools for Bézier splines have some issues with both the fairness measure (criterion 4) and the expressiveness of the specification language (criterion 1), both of which shall be analyzed in the following paragraphs.
 
@@
 In case the software allows the user to request \(G^2\) continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
-Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with language expressiveness.
+Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with specification language expressiveness.
 
-				The language used to specify Bézier splines in common design tools is lacking in expressiveness.
+				The specification language used to specify Bézier splines in common design tools is lacking in expressiveness.
 Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature).
@@
 				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_analysis}.
-The language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
+The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
 The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4).
@@
 
-				The only real issue lies in the insufficient expressiveness of the language.
+				The only real issue lies in the insufficient expressiveness of the specification language.
 Section \ref{section:common_design_tools_bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability.

### 2013-08-15 idp/documentation/paper.tex
added references for illustrations
inlined notes into LaTeX source file

@@
 % TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
+% TODO: maybe extract topical project progress from log
 
@@
 		% TODO: make sure introduction does not overlap with start of next section regarding both content and wording
+		% TODO: quote raphael levien on usability of Bézier splines
 		% TODO: comment on that although there has been some research into better curves (quote MVC curves), it was less focused on usability and hasn't had much impact on actual vector graphics software which continues to almost exclusively use bézier splinse to this day
@@
 					\includegraphics[width=\textwidth]{../resources/usability_bezier.pdf}
-					\caption{Usability Analysis of Bézier Splines}
+					\caption{Usability Analysis of Bézier Splines (left: \cite{Fluttershy}, right: \cite{Lugia})}
 					\label{figure:usability_bézier}
@@
 					\includegraphics[width=\textwidth]{../resources/usability_spiro.pdf}
-					\caption{Usability Analysis of Spiro Splines}
+					\caption{Usability Analysis of Spiro Splines (left: \cite{Fluttershy}, right: \cite{Lugia})}
 					\label{figure:usability_spiro}
@@
 
-			We think that in order to get the best usability, using a good specification language has to be the top priority.
-Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables flexible construction of curves based on some specification and some fairness measure, as illustrated in figure \ref{figure:specification-based_curves}.
+			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority.
+Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables flexible creation of curves based on some specification and some fairness measure, as illustrated in figure \ref{figure:specification-based_curves}.
 
@@
 
+		% instance of specification-based curves
 		% segmentation, separation of segmentation and specification, global positions on curve (https://code.google.com/p/kurve/wiki/SegmentationParametrization, maybe diagram)
@@
 
+		% choice of platform
+		%   C#: platform independent, high-level, possibility to interface with native code, familarity, good UI frameworks (GTK#)
+		%   ipopt
+		%   casadi
+		%   link to googlecode page
+		% numeric optimization
+		%   formalization of model using symbolic functions in casadi
+		%   optimization using ipopt
+		%   caching
+		%   keeping track of disambiguation information
+		%   input is a set of specifications, disambiguation, etc.
+		%   result is a curve whose segmentation is no longer visible
+		% user interface (include screenshots)
+		%   goal: proof of concept
+		%   interacts with numeric optimizer
+		%   specify any combination of point, direction and curvature at positions on the curve
+		%   change position of a specification
+		%     in case of pointspecifications optionally moving the point along the curve accordingly
+		%   insert/remove curve length at or between specifications
+		%   displays deviation from set velocity, such that the user knows when a segment is over-stretched, or where the curve degree/segment count is too low
+		%   loading/saving of documents
+		%     specification can be serialized from/to XML
+
 	\section{Evaluation}
+
+		% were the goals met? (TODO: systematic approach for describing expressiveness and new features)
+		%   controlling length is actually a very nice thing to have
+		%     a good way to control circle arcs in terms of radius (fluttershy's ear, mouth)
+		%     a good way to pull curves in place when they don't have direction specifications
+		%   fluttershy's mane
+		%     spirals done this way are guaranteed to be smooth and not have any unintentional bumps
+		%     comparison with bézier
+		%       source image is in flash, already made of bézier splines, easy to trace
+		%       still many more nodes needed
+		%       hard to get actual, perfect spirals right
+		%       curvature incontinuities, small bumps
+		%     somewhat hampered by poor performance and optimization stability issues, concept seems very powerful though
+		%   some curves are more easily described in terms of curvature than points and/or direction
+		% problems
+		%   local optima
+		%   bad stability, infeasible for numeric optimization
+		%   performance
+		%     possible improvements
+		%       make sure substitutions only happen where and when necessary
+		%       cache partially instantiated IpoptProblem? such that fewer substitutions have to happen when changing a single specification
+		%       use better linear solver with ipopt
+		%       properly configure ipopt options
+		%       adaptive segment density
+		%       adaptive velocity/fairness trapezoid sum
 
@@
 
+		% distinguish between further research work and turning the project into practical software)	
+		% further possibilities to instantiate overall structure (different and/or combined spline primitives, constructive instead of optimization approach)
+		% adaptive line segments, adaptive bézier approximation
+		% steps that are necessary to turn this project into a useful application
+		%   make absolutely sure the concept works by testing it with actual users
+		%   performance improvements
+		%   inkscape plugin
 		% possibility to have curves without fixed length? could be useful for curves which only have specifications at the start and/or end of the curve
+		% investigate into the calculus of variations and whether there may be a possibility to construct MVC curves based on our specification format, once the format itsself has proven to be useful
 
@@
 
+		\bibitem{Fluttershy}
+			\emph{Fluttershy}\\
+			Hasbro, Inc.\\
+			My Little Pony: Friendship is Magic, 2010
+
+		\bibitem{Lugia}
+			\emph{Lugia}\\
+			Nintendo Co., Ltd.\\
+			Pokémon, 1999
+
 	\end{thebibliography}

### 2013-08-15 idp/documentation/paper.tex
some reorganization, started working on solution section

@@
 
-	\section{Usability of Curve Design Tools}
-	\label{section:usability_curve_design_tools}
+	\section{Problem Analysis}
+	\label{section:problem_analysis}
 
@@
 In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves.
-Once such a model has been obtained, it can be used to propose criteria for good curve design software as well as to analyze potential issues with existing tools.
+Once such a model has been obtained, it can be used to propose criteria for good curve design software as well as to analyze potential usability issues with existing tools.
 
@@
 
-			By observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible.
+			Observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible.
 
@@
 
-		\subsection{Usability Analysis}
-		\label{section:usability_analysis}
+		\subsection{Usability Criteria for Curve Design Tools}
+		\label{section:usability_criteria_curve_design_tools}
 
@@
 
-			\subsubsection{Common Design Tools for Bézier Splines}
-			\label{section:common_design_tools_bézier_splines}
+			\subsubsection{Bézier Splines}
+			\label{section:bézier_splines}
 
@@
 
-				Applying the criteria established in section \ref{section:usability_analysis}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice.
+				Applying the criteria established in section \ref{section:usability_criteria_curve_design_tools}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice.
 Bézier splines are furthermore very simple to derive from their specification, fulfilling the 5th criterion perfectly.
@@
 
-			\subsubsection{Common Design Tools for Spiro Splines}
+			\subsubsection{Spiro Splines}
+			\label{section:spiro_splines}
 
@@
 
-				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_analysis}.
+				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_criteria_curve_design_tools}.
 The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
@@
 				The only real issue lies in the insufficient expressiveness of the specification language.
-Section \ref{section:common_design_tools_bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability.
+Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability.
 Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature.
@@
 Unfortunately, describing the curve using Spiro splines still requires many control points.
-In contrast to the corresponding example in section \ref{section:common_design_tools_bézier_splines}, the curve in this example is close to a perfect spiral an thus does not match the source image perfectly.
+In contrast to the corresponding example in section \ref{section:bézier_splines}, the curve in this example is close to a perfect spiral an thus does not match the source image perfectly.
 Additional control points would have to be added in order to approximate the imperfect spiral in the source image better.
@@
 
+	\section{Proposed Solution}
+	\label{section:proposed_solution}
+
+		TODO
+
 		\subsection{Specification-Based Curves}
@@
 			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority.
-Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables flexible creation of curves based on some specification and some fairness measure, as illustrated in figure \ref{figure:specification-based_curves}.
+Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on some specification and some fairness measure.
 
@@
 
-	\section{Proposed Solution: Nonlinear Optimization on Polynomial Curves}
+			Figure \ref{figure:specification-based_curves} shows such a process.
+Note that many specifics are still left undetermined in this approach since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like.
+It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single parametric curve, or if the curve should be determined by numerical optimization or by explicit construction.
+These choices will be dealt with in the following section.
 
-		% instance of specification-based curves
+		\subsection{Nonlinear Optimization on Polynomial Curves}
+		\label{section:nonlinear_optimization_polynomial_curves}
+
+			After having established the general approach to curve design tools that the usability analysis seems to suggest in section \ref{section:specification-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
+
+			The goal of this project is to figure out which specification language and fairness measure yield the best usability in the resulting curve design tool.
+In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves which are connected with \(G^2\) continuity and whose parameters are determined using nonlinear optimization on some objective function determined by the specification and the fairness measure.
+This approach allows us to conveniently experiment with different specification languages and fairness measures in order to assess their usability.
+
+			\begin{table}[htb]
+				\centering
+				\begin{tabular}{c|c|r|r|}
+					\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Segmentation Visibility} \\
+					\cline{2-4}
+					& Scenario & \multicolumn{1}{|c|}{Exposed} & \multicolumn{1}{|c|}{Opaque} \\
+					\cline{2-4}
+					\multirow{2}{*}{Position Specification} & Default & 1 & 2 \\
+					\cline{2-4}
+					& Arc-Length & 3 & 4 \\
+					\cline{2-4}
+				\end{tabular}
+				\caption{Choices for Position Specification and Segmentation Visibility}
+				\label{table:choices_position_specification_segmentation_visibility}
+			\end{table}
+
 		% segmentation, separation of segmentation and specification, global positions on curve (https://code.google.com/p/kurve/wiki/SegmentationParametrization, maybe diagram)
@@
 
-		\begin{table}[htbp]
-			\centering
-			\begin{tabular}{c|c|r|r|}
-				\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Actual Choice} \\
-				\cline{2-4}
-				& Payout & \multicolumn{1}{|c|}{A and B} & \multicolumn{1}{|c|}{B only} \\
-				\cline{2-4}
-				\multirow{2}{*}{Predicted Choice} & A and B & 1 000 \$ & 0 \$ \\
-				\cline{2-4}
-				& B only & 1 001 000 \$ & 1 000 000 \$ \\
-				\cline{2-4}
-			\end{tabular}
-			\caption{Possible payouts for Newcomb's paradox}
-			\label{table:NewcombParadoxPayouts}
-		\end{table}
-
 	\begin{thebibliography}{}

### 2013-08-16 idp/documentation/paper.tex
more work on solution section

@@
 
-		TODO
+		Having analyzed the usability aspect of curve design tools, it's time to think about how to properly address their current issues.
+In the following sections, we propose both an abstract approach for designing a usability-oriented curve design tool, as well as our specific ideas for developing a piece of software using this approach.
 
@@
 
+			\subsubsection{Specification Language}
+
+				% specification language
+				%   differential geometric properties
+				%   point, direction and curvature
+				%   some kind of position specification (ideally, arc-length) is needed for non-points (even points aren't optimal without as they can slide)
+				% maybe just leave out the whole justification with segmentation, etc. and just assume arc-length parametrization
+
+			\subsubsection{Fairness Measure}
+
+				% fairness (quote paper or thesis on MVC)
+				%   derivative of projected acceleration is equal to projected jerk (maybe this belongs to implementation)
+
+			\subsubsection{Segmentation and Parametrization}
+
+				% continuity connections
+				% segmentation, separation of segmentation and specification, global positions on curve
+				% fixed length, constant velocity
+				%   fixed length is necessary if specification positions are a fraction of the arc-length of the whole curve, otherwise, any change in the specifications may change the curve length and thus move all specifications along the curve
+
 				\begin{table}[htb]
@@
 
-		% segmentation, separation of segmentation and specification, global positions on curve (https://code.google.com/p/kurve/wiki/SegmentationParametrization, maybe diagram)
-		% disambiguation
-		% specifications
-		%   differential geometric properties
-		%   position, direction and curvature
-		% continuity connections
-		% fixed length, constant velocity
-		%   fixed length is necessary if specification positions are a fraction of the arc-length of the whole curve, otherwise, any change in the specifications may change the curve length and thus move all specifications along the curve
-		% fairness (quote paper or thesis on MVC)
-		%   derivative of projected acceleration is equal to projected jerk
+			\subsubsection{Disambiguation Information}
 

### 2013-08-18 idp/documentation/paper.tex
continued working on specification-based curves section

@@
 \usepackage{amssymb}           % ams math symbols
-\usepackage{multirow}          % enables multirow tables
 
@@
 % TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
+% TODO: expand n't contractions
+% TODO: check if all quotations are done `like this'
 % TODO: maybe extract topical project progress from log
@@
 			\subsubsection{Specification Language}
+			\label{section:specification_language}
 
-				% specification language
-				%   differential geometric properties
-				%   point, direction and curvature
-				%   some kind of position specification (ideally, arc-length) is needed for non-points (even points aren't optimal without as they can slide)
-				% maybe just leave out the whole justification with segmentation, etc. and just assume arc-length parametrization
+				The specification language we propose uses differential geometric properties of curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use.
+The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will use points on the curve, tangent angles of the curve as well as the curvature value of the curve.
+In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties.
+
+				With point specifications, it's sometimes enough to require the curve to go through all of them in order, although sometimes, this will lead to undesirable results.
+Even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points.
+When specifying tangent angles and/or curvature values without an associated point, it becomes a necessity to associate this specification with some kind of positional information in order to anchor it to some point on the curve.
+If this was not done, the specification could move freely along the curve and would thus behave very unpredictably.
+We thus chose to associate each combination of point, tangent angle and curvature value with a position on the curve.
+
+				An obvious choice for specifying a position on a curve would be by giving a parameter value for the underlying parametric curve.
+However, this would violate our design idea of using differential geometric properties of curves.
+A way of specifying a position on a curve that is independent from its parametrization is the arc length of the curve.
+Instead of directly using the arc length to give the position of a specification, we chose to specify positions as fractions of the total arc length of the curve.
+This way, the position of a specification relative to the start and end points of the curve is independent of the total arc length of the curve.
+
+				With this setup, it seems advisable to also include the curve's total arc length in the specification.
+This way, one can specify curves that don't have any specifications associated to their start and end points.
+Also, with unspecified curve length, changing any specification may inadvertently change the curve's total arc length, thus possibly adversely affecting other parts of the curve whose specification positions are given as a fraction of the curve's total arc length.
+
+				% TODO: give formal description of specification language
 
 			\subsubsection{Fairness Measure}
+			\label{section:fairness_measure}
 
@@
 
-			\subsubsection{Segmentation and Parametrization}
+			\subsubsection{Optimization Problem}
+			\label{section:optimization_problem}
 
@@
 				% segmentation, separation of segmentation and specification, global positions on curve
-				% fixed length, constant velocity
-				%   fixed length is necessary if specification positions are a fraction of the arc-length of the whole curve, otherwise, any change in the specifications may change the curve length and thus move all specifications along the curve
-
-				\begin{table}[htb]
-					\centering
-					\begin{tabular}{c|c|r|r|}
-						\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Segmentation Visibility} \\
-						\cline{2-4}
-						& Scenario & \multicolumn{1}{|c|}{Exposed} & \multicolumn{1}{|c|}{Opaque} \\
-						\cline{2-4}
-						\multirow{2}{*}{Position Specification} & Default & 1 & 2 \\
-						\cline{2-4}
-						& Arc-Length & 3 & 4 \\
-						\cline{2-4}
-					\end{tabular}
-					\caption{Choices for Position Specification and Segmentation Visibility}
-					\label{table:choices_position_specification_segmentation_visibility}
-				\end{table}
-
-			\subsubsection{Disambiguation Information}
+				% fixed length as constant speed
 

### 2013-08-18 idp/documentation/paper.tex
refined section about specification language

@@
 
+				With that in mind, it is now possible to abandon this notion of a combination of point, tangent angle and curvature value and instead talk about specification items, a specification item being either a positioned point, a positioned tangent angle or a positioned curvature value.
+A combination of point and curvature value at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
+
 				An obvious choice for specifying a position on a curve would be by giving a parameter value for the underlying parametric curve.
@@
 A way of specifying a position on a curve that is independent from its parametrization is the arc length of the curve.
-Instead of directly using the arc length to give the position of a specification, we chose to specify positions as fractions of the total arc length of the curve.
-This way, the position of a specification relative to the start and end points of the curve is independent of the total arc length of the curve.
+Instead of directly using the arc length to give the positions of specification items, we chose to specify positions as fractions of the total arc length of the curve.
+This way, the position of a specification item relative to the start and end points of the curve is independent of the total arc length of the curve.
 
 				With this setup, it seems advisable to also include the curve's total arc length in the specification.
-This way, one can specify curves that don't have any specifications associated to their start and end points.
-Also, with unspecified curve length, changing any specification may inadvertently change the curve's total arc length, thus possibly adversely affecting other parts of the curve whose specification positions are given as a fraction of the curve's total arc length.
+This way, one can specify curves that don't have any specification items associated to their start and end points.
+Also, with unspecified curve length, changing any specification item may inadvertently change the curve's total arc length, thus possibly adversely affecting other parts of the curve whose specification item positions are given as a fraction of the curve's total arc length.
 
-				% TODO: give formal description of specification language
+				Thus, the final specification language is as follows.
+
+				\begin{align*}
+					\text{Positions} & = \left[0, 1\right]\\
+					\text{Points} & = \mathbb{R}^2\\
+					\text{Directions} & = \mathbb{R}\\
+					\text{Curvatures} & = \mathbb{R}\\
+					\text{SpecificationItems} & = \text{Positions} \times \left(\text{Points} \cup \text{Directions} \cup \text{Curvatures}\right)\\
+					\text{CurveLengths} & = \mathbb{R}\\
+					\text{Specifications} & = \mathcal{P}\left(\text{SpecificationItems}\right) \times \text{CurveLengths}
+				\end{align*}
 

### 2013-08-19 idp/documentation/paper.tex
started to work on fairness measure section, formatting, small fixes

@@
 
-			Observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible.
-
+			Observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible:
 			\begin{enumerate}
@@
 These goals are competing with the software's ability to derive curves from specifications and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive specification languages and more complex fairness measures.
-From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.
 
+			From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools:
 			\begin{enumerate}
@@
 Another issue is that without software support for \(G^2\) continuity, using more nodes usually also introduces more points where curvature is not continuous.
-In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.
+In cases where the curvature is an important property of the curve (for example the constant curvature of circular arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.
 
@@
 				Example 1 shows two Bézier splines.
-The left curve is a Bézier spline approximation of a circle arc, exhibiting nearly constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
-Since the difference can be hard to see, it's easy to make mistakes when trying to approximate circle arcs using Bézier splines.
+The left curve is a Bézier spline approximation of a circular arc, exhibiting nearly constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
+Since the difference can be hard to see, it's easy to make mistakes when trying to approximate circular arcs using Bézier splines.
 
@@
 These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
-This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circle arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
+This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
 Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties.
@@
 				Example 2 shows how some shapes can be described using Spiro splines by placing only very few control points.
-In this case, the shape consists of a circle arc with low curvature (segment a), a segment of linearly increasing curvature (segment b) and finally a circle arc with high curvature (segment c).
+In this case, the shape consists of a circular arc with low curvature (segment a), a segment of linearly increasing curvature (segment b) and finally a circular arc with high curvature (segment c).
 This shape can be described conveniently using a Spiro spline with just 4 control points.
@@
 				The specification language we propose uses differential geometric properties of curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use.
-The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will use points on the curve, tangent angles of the curve as well as the curvature value of the curve.
+The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will use points on the curve (denoted \(p\)), tangent angles (denoted \(\theta\)) of the curve as well as the curvature value (denoted \(\kappa\)) of the curve.
 In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties.
@@
 
-				Thus, the final specification language is as follows.
-
+				Thus, the final specification language is as follows:
 				\begin{align*}
@@
 
-				% fairness (quote paper or thesis on MVC)
-				%   derivative of projected acceleration is equal to projected jerk (maybe this belongs to implementation)
+				% TODO: add citations
+
+				Many different fairness measures for plane curves have been proposed throughout history.
+Most of them set out to emulate the behavior of mechanical splines made from wood or steel, like those which were used for industrial design before the advent of computer aided design.
+The most direct approach of this type uses mechanics to derive a term for the energy in a bent steel band, arriving at the following functional describing the minimum energy curve (MEC):
+				\begin{displaymath}
+					\int_{0}^{1}{\kappa\!\left(t\right)}^2\;\mathrm{d}t
+				\end{displaymath}
+
+				While this approach is a good idealization of the mechanical spline, it has some disadvantages when used for curve design.
+For instance, it will not yield a circular arc when constrained by co-circular points.
+Thus, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
+One such approach is that of the minimum variation curve (MVC), whose functional is given below:
+				\begin{displaymath}
+					\int_{0}^{1}{\kappa'\!\left(t\right)}^2\;\mathrm{d}t
+				\end{displaymath}
+
+				Minimum variation curves capture the intuitive notion of smoothness very well.
+The functional will put straight lines through points that are co-linear and circular arcs through points that are co-circular when otherwise unconstrained.
+When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvture, since that will make the derivation of the curvature a constant and thereby minimize the integral over the square fo the derivative of the curvature.
 
@@
 		%   controlling length is actually a very nice thing to have
-		%     a good way to control circle arcs in terms of radius (fluttershy's ear, mouth)
+		%     a good way to control circular arcs in terms of radius (fluttershy's ear, mouth)
 		%     a good way to pull curves in place when they don't have direction specifications

### 2013-08-19 idp/documentation/paper.tex
started working on optimization section

@@
 		Having analyzed the usability aspect of curve design tools, it's time to think about how to properly address their current issues.
-In the following sections, we propose both an abstract approach for designing a usability-oriented curve design tool, as well as our specific ideas for developing a piece of software using this approach.
+In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as our specific ideas for developing a piece of software using this approach.
 
@@
 It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single parametric curve, or if the curve should be determined by numerical optimization or by explicit construction.
-These choices will be dealt with in the following section.
 
@@
 When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvture, since that will make the derivation of the curvature a constant and thereby minimize the integral over the square fo the derivative of the curvature.
+Because of these appealing properties, the MVC fairness measure was chosen.
 
@@
 
+				Having chosen a specification language and a fairness measure, it is time to formulate the optimization problem.
+Since polynomials of very high degree usually don't behave very well, an approach where multiple polynomial parametric curves are connected to a spline seems advisable.
+This way, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline.
+In the following, we will denote the \(i\)th coefficient of the \(k\)th polynomial by \(c_{k,i}\), and the vector of all coefficients by \(c\).
+
+				definitions 
+				\begin{align*}
+					p & : \left[0,1\right] \rightarrow \mathbb{R}^2 & & && \text{(point)}\\
+					v & : \left[0,1\right] \rightarrow \mathbb{R}^2, & v & = p' && \text{(velocity)}\\
+					a & : \left[0,1\right] \rightarrow \mathbb{R}^2, & a & = v' && \text{(acceleration)}
+				\end{align*}
+
+				The first problem we face is that polynomials will in general not be in natural parametrization (that is, a parametrization where the parameter \(t\) is equal to the covered arc length of the curve up to the point \(p\left(t\right)\)).
+Since we chose to give the position of a specification item as a fraction of the curve's total arc length, it would be very advantageous to have a parametrization where the parameter is at least proportional to the covered arc length of the curve.
+This property is equivalent to the first derivative of the curve (velocity) having constant euclidean norm (\(\|p'\left(t\right)\|\) is constant) and has additional useful consequences, such as the second derivative of the curve (acceleration) being orthogonal to the velocity.
+Thus, one of the objectives of the optimization is to make sure that the speed of the curve is constant.
+
+				% constant velocity
+				% fairness
+				% specification
+				% continuity connections
+
 				% continuity connections
@@
 				% fixed length as constant speed
+				% give objective and constraints function
 

### 2013-08-24 source/Kurve/Kurve.Curves/Curve.cs
updated implementation to fit new angle-based model improved performance

@@ -11,7 +11,7 @@ namespace Kurve.Curves
 	{
 		public abstract Vector2Double GetPoint(double position);
 		public abstract double GetSpeed(double position);
-		public abstract Vector2Double GetDirection(double position);
+		public abstract double GetDirection(double position);
 		public abstract double GetCurvature(double position);
 	}
 }

### 2013-08-24 source/Kurve/Kurve.Curves/FunctionTermCurves/FunctionTermCurve.cs

@@ -14,6 +14,7 @@ namespace Kurve.Curves
 		readonly FunctionTerm acceleration;
 		readonly FunctionTerm jerk;
 		readonly FunctionTerm speed;
+		readonly FunctionTerm normalizedVelocity;
 		readonly FunctionTerm direction;
 		readonly FunctionTerm curvature;
 
@@ -22,34 +23,38 @@ namespace Kurve.Curves
 		public FunctionTerm Acceleration { get { return acceleration; } }
 		public FunctionTerm Jerk { get { return jerk; } }
 		public FunctionTerm Speed { get { return speed; } }
+		public FunctionTerm NormalizedVelocity { get { return normalizedVelocity; } }
 		public FunctionTerm Direction { get { return direction; } }
 		public FunctionTerm Curvature { get { return curvature; } }
 
 		public FunctionTermCurve(FunctionTerm function)
 		{
 			if (function == null) throw new ArgumentNullException("function");
-			if (function.DomainDimension != 1) throw new ArgumentException("parameter 'function' has wrong dimension.");
+			if (function.DomainDimension != 1) throw new ArgumentException("parameter 'function' has wrong domain dimension.");
+			if (function.CodomainDimension != 2) throw new ArgumentException("parameter 'function' has wrong codomain dimension.");
 
 			ValueTerm position = Terms.Variable("t");
-
+			
 			ValueTerm point = function.Apply(position);
 			ValueTerm velocity = function.GetDerivatives().Single().Apply(position);
 			ValueTerm acceleration = function.GetDerivatives().Single().GetDerivatives().Single().Apply(position);
 			ValueTerm jerk = function.GetDerivatives().Single().GetDerivatives().Single().GetDerivatives().Single().Apply(position);
 
-			ValueTerm speed = Terms.Norm(velocity);
-			ValueTerm direction = Terms.Normalize(velocity);
-			// TODO: maybe we can get rid of some normalizations here (seems like angularDirection uses the normalized direction just to divide it by the square of the speed later)
-			ValueTerm angularDirection = Terms.Vector(direction.Select(1), Terms.Negate(direction.Select(0)));
-			ValueTerm angularAcceleration = Terms.DotProduct(acceleration, angularDirection);
-			ValueTerm curvature = Terms.Scaling(Terms.Invert(Terms.Square(speed)), angularAcceleration);
-
 			this.point = point.Abstract(position);
 			this.velocity = velocity.Abstract(position);
[…]

### 2013-08-24 source/Kurve/Kurve.Curves/FunctionTermCurves/PolynomialFunctionTermCurveTemplate.cs

@@ -10,11 +10,11 @@ namespace Kurve.Curves
 {
 	public class PolynomialFunctionTermCurveTemplate : FunctionTermCurveTemplate, IEquatable<PolynomialFunctionTermCurveTemplate>
 	{
-		readonly int degree;
+		readonly int length;
 
 		public static string XElementName { get { return "polynomial_curve_template"; } }
 
-		public int Degree { get { return degree; } }
+		public int Degree { get { return length; } }
 		public override XElement XElement
 		{
 			get
@@ -22,16 +22,16 @@ namespace Kurve.Curves
 				return new XElement
 				(
 					XElementName,
-					new XElement("degree", degree)
+					new XElement("length", length)
 				);
 			}
 		}
 
-		public PolynomialFunctionTermCurveTemplate(int degree) : base(Create(degree))
+		public PolynomialFunctionTermCurveTemplate(int length) : base(Create(length))
 		{
-			if (degree < 0) throw new ArgumentOutOfRangeException("degree");
+			if (length < 0) throw new ArgumentOutOfRangeException("length");
 
-			this.degree = degree;
+			this.length = length;
 		}
 		public PolynomialFunctionTermCurveTemplate(XElement source) : this(Create(source)) { }
 
@@ -41,7 +41,7 @@ namespace Kurve.Curves
 		}
 		public override int GetHashCode()
 		{
-			return GetType().Name.GetHashCode() ^ degree.GetHashCode();
+			return GetType().Name.GetHashCode() ^ length.GetHashCode();
 		}
 		public bool Equals(PolynomialFunctionTermCurveTemplate other)
 		{
@@ -62,13 +62,13 @@ namespace Kurve.Curves
 			if (object.ReferenceEquals(curveTemplate1, curveTemplate2)) return true;
 			if (object.ReferenceEquals(curveTemplate1, null) || object.ReferenceEquals(curveTemplate2, null)) return false;
 			
-			return curveTemplate1.degree == curveTemplate2.degree;
+			return curveTemplate1.length == curveTemplate2.length;
 		}
-		static FunctionTerm Create(int degree)
+		static FunctionTerm Create(int length)
 		{
-			if (degree < 0) throw new ArgumentOutOfRangeException("degree");
+			if (length < 0) throw new ArgumentOutOfRangeException("length");
 
-			return Terms.Polynomial(Terms.StandardPolynomialBasis(degree), 2);
[…]

### 2013-08-24 source/Kurve/Kurve.Curves/Optimization/OptimizationProblem.cs

@@ -36,33 +36,39 @@ namespace Kurve.Curves.Optimization
 			this.curveLength = Terms.Variable("curveLength");
 
 			IEnumerable<ValueTerm> variables = optimizationSegments.Parameters; 
+			
+			ValueTerm targetSpeed = Terms.Quotient(curveLength, Terms.Constant(optimizationSegments.Segments.Count()));
 
-			ValueTerm velocityError = Terms.Sum
+			ValueTerm speedError = Terms.Sum
 			(
 				from segment in optimizationSegments.Segments
 				let position = Terms.Variable("t")
 				let curve = segment.LocalCurve
 				let speed = curve.Speed.Apply(position)
-				let length = Terms.Quotient(curveLength, Terms.Constant(optimizationSegments.Segments.Count()))
-				let error = Terms.Square(Terms.Difference(speed, length))
-				select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
-			);
-			ValueTerm fairnessError = Terms.Sum
-			(
-				from segment in optimizationSegments.Segments
-				let position = Terms.Variable("t")
-				let curve = segment.LocalCurve
-				let jerk = curve.Jerk.Apply(position)
-				let acceleration = curve.Acceleration.Apply(position)
-				let error = Terms.Square(Terms.DotProduct(jerk, Terms.Normalize(acceleration)))
-				select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
+				let error = Terms.Square(Terms.Difference(speed, targetSpeed))
+				select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 32)
 			);
 
-			// TODO: figure out a more mathematically sound way to do the weighting, maybe dependent on curve length
+			ValueTerm fairnessError = Terms.Scaling
+			(
+				Terms.Exponentiation(targetSpeed, Terms.Constant(-6)),
+				Terms.Sum
+				(
+					from segment in optimizationSegments.Segments
+					let position = Terms.Variable("t")
+					let curve = segment.LocalCurve
+					let jerk = curve.Jerk.Apply(position)
+					let normal = Terms.Normal(curve.Velocity.Apply(position))
+					let error = Terms.Square(Terms.DotProduct(jerk, normal))
[…]

### 2013-08-24 source/Kurve/Kurve.Curves/Optimization/OptimizationSubstitutions.cs

@@ -90,7 +90,7 @@ namespace Kurve.Curves.Optimization
 					DirectionCurveSpecification directionCurveSpecification = directionCurveSpecifications.ElementAt(directionSpecificationTemplateIndex);
 
 					ValueTerm position = Terms.Constant(directionCurveSpecification.Position);
-					ValueTerm value = Terms.Constant(directionCurveSpecification.Direction.NormalizedVector.X, directionCurveSpecification.Direction.NormalizedVector.Y);
+					ValueTerm value = Terms.Direction(Terms.Constant(directionCurveSpecification.Direction));
 
 					foreach (Substitution substitution in directionSpecificationTemplate.GetSubstitutions(optimizationSegments.Segments, directionCurveSpecification.Position, position, value))
 						yield return substitution;

### 2013-08-24 source/Kurve/Kurve.Curves/SegmentedCurve.cs

@@ -29,7 +29,7 @@ namespace Kurve.Curves
 		{
 			return segments.ElementAt(getSegmentIndex(position)).GlobalCurve.GetSpeed(position);
 		}
-		public override Vector2Double GetDirection(double position)
+		public override double GetDirection(double position)
 		{
 			return segments.ElementAt(getSegmentIndex(position)).GlobalCurve.GetDirection(position);
 		}

### 2013-08-24 source/Kurve/Kurve.Curves/Specification/DirectionCurveSpecification.cs

@@ -3,18 +3,19 @@ using Krach.Basics;
 using Wrappers.Casadi;
 using System.Xml.Linq;
 using System.Linq;
+using Krach.Extensions;
 
 namespace Kurve.Curves
 {
 	public class DirectionCurveSpecification : CurveSpecification, IEquatable<DirectionCurveSpecification>
 	{
 		readonly double position;
-		readonly Vector2Double direction;
+		readonly double direction;
 		
 		public static string XElementName { get { return "direction_curve_specification"; } }
 
 		public override double Position { get { return position; } }
-		public Vector2Double Direction { get { return direction; } }
+		public double Direction { get { return direction; } }
 		public override XElement XElement
 		{
 			get
@@ -23,12 +24,12 @@ namespace Kurve.Curves
 				(
 					XElementName,
 					new XElement("position", position),
-					new XElement("direction", direction.XElement)
+					new XElement("direction", direction)
 				);	
 			}
 		}
 
-		public DirectionCurveSpecification(double position, Vector2Double direction)
+		public DirectionCurveSpecification(double position, double direction)
 		{
 			if (position < 0 || position > 1) throw new ArgumentOutOfRangeException("position");
 			
@@ -40,7 +41,7 @@ namespace Kurve.Curves
 			if (source == null) throw new ArgumentNullException("source");
 
 			this.position = (double)source.Element("position");
-			this.direction = new Vector2Double(source.Element("direction").Elements().Single());
+			this.direction = (double)source.Element("direction");
 		}
 
 		public override bool Equals(object obj)

### 2013-08-24 source/Kurve/Kurve.Curves/Specification/SpecificationTemplate.cs

@@ -95,7 +95,7 @@ namespace Kurve.Curves
 					(
 						segments,
 						segmentWeights,
-						(segment, segmentWeight) => Terms.Scaling(segmentWeight, Terms.Difference(segment.GlobalCurve.Direction.Apply(position), direction))
+						(segment, segmentWeight) => Terms.Scaling(segmentWeight, Terms.Difference(segment.GlobalCurve.NormalizedVelocity.Apply(position), direction))
 					)
 				)
 			);

### 2013-08-24 source/Kurve/Kurve/Components/Controls/SpecificationComponent.cs

@@ -20,7 +20,7 @@ namespace Kurve.Component
 
 		double position = 0;
 		Vector2Double point;
-		Vector2Double direction;
+		double direction;
 		double curvature;
 	
 		public event Action SpecificationChanged;
@@ -47,7 +47,7 @@ namespace Kurve.Component
 				if (specifiesPoint) point = value;
 			}
 		}
-		public Vector2Double Direction 
+		public double Direction 
 		{ 
 			get 
 			{ 
@@ -181,9 +181,7 @@ namespace Kurve.Component
 				}
 				else
 				{
-					double angle = Scalars.ArcTangent(Direction.Y, Direction.X) + 0.1 * SlowDownFactor * ((scrollDirection == ScrollDirection.Up) ? 1 : -1);
-
-					Direction = new Vector2Double(Scalars.Cosine(angle), Scalars.Sine(angle));
+					Direction += 0.1 * SlowDownFactor * ((scrollDirection == ScrollDirection.Up) ? 1 : -1);
 				}
 
 				OnSpecificationChanged();

### 2013-08-24 source/Kurve/Kurve/DiscreteCurve.cs

@@ -41,7 +41,7 @@ namespace Kurve
 		{
 			return GetItem(position).Speed;
 		}
-		public override Vector2Double GetDirection(double position)
+		public override double GetDirection(double position)
 		{
 			return GetItem(position).Direction;
 		}

### 2013-08-24 source/Kurve/Kurve/DiscreteCurveItem.cs

@@ -13,15 +13,15 @@ namespace Kurve
 	{
 		readonly Vector2Double point;
 		readonly double speed;
-		readonly Vector2Double direction;
+		readonly double direction;
 		readonly double curvature;
 
 		public Vector2Double Point { get { return point; } }
 		public double Speed { get { return speed; } }
-		public Vector2Double Direction { get { return direction; } }
+		public double Direction { get { return direction; } }
 		public double Curvature { get { return curvature; } }
 
-		public DiscreteCurveItem(Vector2Double point, double speed, Vector2Double direction, double curvature)
+		public DiscreteCurveItem(Vector2Double point, double speed, double direction, double curvature)
 		{
 			this.point = point;
 			this.speed = speed;

### 2013-08-26 idp/documentation/paper.tex
fixed some formatting, added definitions

@@
 \widowpenalty = 10000
+
+% math commands
+\DeclareMathOperator{\atan2}{atan2}
+\newcommand{\apply}[2]{#1\!\left(#2\right)}
+\renewcommand{\vector}[2]{\begin{pmatrix}#1\\\linebreak{}#2\end{pmatrix}}
 
@@
 For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve.
-Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + (p_1 - p_0) \cdot t\), thus uniquely identifying the desired curve.
+Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \left(p_1 - p_0\right) \cdot t\), thus uniquely identifying the desired curve.
 Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
@@
 A Bézier curve \(B\) of degree \(n\) consists of \(n + 1\) coefficients \(P_i \in \mathbb{R}^2\).
-The curve will always go through the first and the last coefficient point (\(B\left(0\right) = P_0, B\left(1\right) = P_n\)) and stay within the convex hull of the coefficient points at all other times.
+The curve will always go through the first and the last coefficient point (\(\apply{B}{0} = P_0, \apply{B}{1} = P_n\)) and stay within the convex hull of the coefficient points at all other times.
 Bézier curves in a Bézier spline are usually pieced together such that the resulting spline is \(G^1\) continuous.
@@
 The user can specify control points through which the software puts an interpolating spline.
-There is no way to directly specify tangent direction and/or curvature.
+There is no way to directly specify tangent angle and/or curvature.
 The splines are guaranteed to have \(G^2\) continuity.
@@
 Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability.
-Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature.
-These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
+Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify tangent angle nor curvature.
+These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of tangent angle and/or curvature.
 This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
@@
 These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
-That way, when indirectly specifying direction and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
+That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
 
@@
 For instance, increasing the curvature of segment c has to be done indirectly by moving a control point.
-Unfortunately, in this case, this also changes the direction of the curve at that point, such that additional modifications would be needed to correct this.
+Unfortunately, in this case, this also changes the tangent angle of the curve at that point, such that additional modifications would be needed to correct this.
 The change also adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
@@
 
+		\subsection{Parametric Plane Curves}
+		\label{section:parametric_plane_curves}
+
+			In order to be able to reason about curves, a few basic definitions are needed.
+
+			Definitions about vectors in the euclidean plane:
+			\begin{align*}
+				\overline{\cdot} & : \mathbb{R}^2 \rightarrow \mathbb{R}^2     & \overline{v}      & = \vector{-v_2}{v_1}                      && \text{orthogonal vector}\\
+				\alpha           & : \mathbb{R}^2 \rightarrow \mathbb{R}       & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1}                && \text{polar coordinate angle}
+			\end{align*}
+
+			Definitions about parametric curves:
+			\begin{align*}
+				p                & : \left[0,1\right] \rightarrow \mathbb{R}^2 &                   &                                           && \text{point}\\
+				s                & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{s}{t}      & = \left\|\apply{p'}{t}\right\|            && \text{speed}\\
+				\theta           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\theta}{t} & = \apply{\alpha}{\apply{p'}{t}}           && \text{direction}\\
+				\kappa           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\kappa}{t} & = \frac{\apply{\theta'}{t}}{\apply{s}{t}} && \text{curvature}
+			\end{align*}
+
+			Using these definitions, the direction \(\apply{\theta}{t}\) is equal to the tangent angle at the point \(\apply{p}{t}\) and the curvature \(\apply{\kappa}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{p}{t}\) or \(0\) if the curve is locally a straight line at the point \(\apply{p}{t}\).
+
 		\subsection{Nonlinear Optimization on Polynomial Curves}
@@
 				The specification language we propose uses differential geometric properties of curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use.
-The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will use points on the curve (denoted \(p\)), tangent angles (denoted \(\theta\)) of the curve as well as the curvature value (denoted \(\kappa\)) of the curve.
+The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will focus on the properties point, direction and curvature as defined in \ref{section:parametric_plane_curves}.
 In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties.
@@
 Even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points.
-When specifying tangent angles and/or curvature values without an associated point, it becomes a necessity to associate this specification with some kind of positional information in order to anchor it to some point on the curve.
+When specifying direction and/or curvature without an associated point, it becomes a necessity to associate this specification with some kind of positional information in order to anchor it to some point on the curve.
 If this was not done, the specification could move freely along the curve and would thus behave very unpredictably.
-We thus chose to associate each combination of point, tangent angle and curvature value with a position on the curve.
+We thus chose to associate each combination of point, direction and curvature with a position on the curve.
 
-				With that in mind, it is now possible to abandon this notion of a combination of point, tangent angle and curvature value and instead talk about specification items, a specification item being either a positioned point, a positioned tangent angle or a positioned curvature value.
-A combination of point and curvature value at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
+				With that in mind, it is now possible to abandon this notion of a combination of point, direction and curvature and instead talk about specification items, a specification item being either a positioned point, a positioned direction or a positioned curvature.
+A combination of point and curvature at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
 
@@
 				\begin{displaymath}
-					\int_{0}^{1}{\kappa\!\left(t\right)}^2\;\mathrm{d}t
+					\int_{0}^{1}\apply{\kappa}{t}^2\;\mathrm{d}t
 				\end{displaymath}
@@
 				\begin{displaymath}
-					\int_{0}^{1}{\kappa'\!\left(t\right)}^2\;\mathrm{d}t
+					\int_{0}^{1}\apply{\kappa'}{t}^2\;\mathrm{d}t
 				\end{displaymath}
@@
 The functional will put straight lines through points that are co-linear and circular arcs through points that are co-circular when otherwise unconstrained.
-When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvture, since that will make the derivation of the curvature a constant and thereby minimize the integral over the square fo the derivative of the curvature.
+When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvture, since that will make the derivation of the curvature a constant and thereby minimize the integral over the square of the derivative of the curvature.
 Because of these appealing properties, the MVC fairness measure was chosen.
@@
 
-				definitions 
-				\begin{align*}
-					p & : \left[0,1\right] \rightarrow \mathbb{R}^2 & & && \text{(point)}\\
-					v & : \left[0,1\right] \rightarrow \mathbb{R}^2, & v & = p' && \text{(velocity)}\\
-					a & : \left[0,1\right] \rightarrow \mathbb{R}^2, & a & = v' && \text{(acceleration)}
-				\end{align*}
-
-				The first problem we face is that polynomials will in general not be in natural parametrization (that is, a parametrization where the parameter \(t\) is equal to the covered arc length of the curve up to the point \(p\left(t\right)\)).
+				The first problem we face is that polynomials will in general not be in natural parametrization (that is, a parametrization where the parameter \(t\) is equal to the covered arc length of the curve up to the point \(\apply{p}{t}\)).
 Since we chose to give the position of a specification item as a fraction of the curve's total arc length, it would be very advantageous to have a parametrization where the parameter is at least proportional to the covered arc length of the curve.
-This property is equivalent to the first derivative of the curve (velocity) having constant euclidean norm (\(\|p'\left(t\right)\|\) is constant) and has additional useful consequences, such as the second derivative of the curve (acceleration) being orthogonal to the velocity.
+This property is equivalent to the speed of the curve being constant and has additional useful consequences, such as the second derivative of the curve being orthogonal to the first derivative.
 Thus, one of the objectives of the optimization is to make sure that the speed of the curve is constant.

### 2013-08-27 idp/documentation/paper.tex
continued working on optimization problem section

@@
 				s                & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{s}{t}      & = \left\|\apply{p'}{t}\right\|            && \text{speed}\\
+				l                & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{l}{t}      & = \int_{0}^{t}\apply{s}{u}\;\mathrm{d}u   && \text{covered arc length}\\
 				\theta           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\theta}{t} & = \apply{\alpha}{\apply{p'}{t}}           && \text{direction}\\
-				\kappa           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\kappa}{t} & = \frac{\apply{\theta'}{t}}{\apply{s}{t}} && \text{curvature}
+				\kappa           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\kappa}{t} & = \frac{\apply{\theta'}{t}}{\apply{s}{t}} = \frac{\apply{p''}{t}^\mathrm{T}\overline{\apply{p'}{t}}}{\apply{s}{t}^3} && \text{curvature}
 			\end{align*}
@@
 
-				The first problem we face is that polynomials will in general not be in natural parametrization (that is, a parametrization where the parameter \(t\) is equal to the covered arc length of the curve up to the point \(\apply{p}{t}\)).
-Since we chose to give the position of a specification item as a fraction of the curve's total arc length, it would be very advantageous to have a parametrization where the parameter is at least proportional to the covered arc length of the curve.
-This property is equivalent to the speed of the curve being constant and has additional useful consequences, such as the second derivative of the curve being orthogonal to the first derivative.
-Thus, one of the objectives of the optimization is to make sure that the speed of the curve is constant.
+				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need a function mapping covered arc length to parameter values (the inverse to the function \(l\) defined in section \ref{section:parametric_plane_curves}).
+Unfortunately, this function isn't very well suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
+Thus, an additional optimization goal is to find parametric curves where this function is approximately linear.
+This property is equivaent to the speed of the curve being approximately constant and has additional useful consequences, such as the second derivative of the curve being orthogonal to the first derivative.
+Since the chosen specification language already specifies the total arc length of the curve, and the parameter runs through the unit range, the target value for the speed is simply the total arc length of the curve (\(l^*\)), leading to the following error term:
+				\begin{displaymath}
+					\int_{0}^{1}\left(\apply{s}{t} - l^*\right)^2\;\mathrm{d}t
+				\end{displaymath}
 
@@
 
-				% continuity connections
 				% segmentation, separation of segmentation and specification, global positions on curve
-				% fixed length as constant speed
 				% give objective and constraints function

### 2013-08-27 idp/documentation/paper.tex
continued working on optimization problem section

@@
 \newcommand{\apply}[2]{#1\!\left(#2\right)}
-\renewcommand{\vector}[2]{\begin{pmatrix}#1\\\linebreak{}#2\end{pmatrix}}
+\newcommand{\vectorA}[1]{\begin{pmatrix}#1\end{pmatrix}}
+\newcommand{\vectorB}[2]{\begin{pmatrix}#1\\\linebreak{}#2\end{pmatrix}}
+\newcommand{\vectorC}[3]{\begin{pmatrix}#1\\\linebreak{}#2\\\linebreak{}#3\end{pmatrix}}
 
@@
 
+	% TODO: maybe we don't want to mention G^n and C^n until formally introduced later, doesn't add much value to the usability discussion anyways
 	\section{Problem Analysis}
@@
 Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
-In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(G^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(\mathrm{G}^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
 
@@
 Most vector graphics tools have functions that allow preserving the \(G^1\) continuity of the spline by automatically modifying adjacent coefficient points accordingly.
-They usually do not ensure \(G^2\) continuity though.
+They usually do not ensure \(\mathrm{G}^2\) continuity though.
 Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more Bézier curves to the spline.
@@
 
-				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing \(G^2\) continuity.
-It is difficult for humans to choose the coefficient points in such a way that \(G^2\) continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
+				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing \(\mathrm{G}^2\) continuity.
+It is difficult for humans to choose the coefficient points in such a way that \(\mathrm{G}^2\) continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
 The problem is aggravated by the fact that curvature discontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
-In case the software allows the user to request \(G^2\) continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
+In case the software allows the user to request \(\mathrm{G}^2\) continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
 Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with specification language expressiveness.
@@
 It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
-Another issue is that without software support for \(G^2\) continuity, using more nodes usually also introduces more points where curvature is not continuous.
+Another issue is that without software support for \(\mathrm{G}^2\) continuity, using more nodes usually also introduces more points where curvature is not continuous.
 In cases where the curvature is an important property of the curve (for example the constant curvature of circular arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.
@@
 There is no way to directly specify tangent angle and/or curvature.
-The splines are guaranteed to have \(G^2\) continuity.
+The splines are guaranteed to have \(\mathrm{G}^2\) continuity.
 
@@
 The user supplies a set of points for which the software computes an interpolating spline.
-The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is \(G^2\) continuity and no excessive winding.
+The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is \(\mathrm{G}^2\) continuity and no excessive winding.
 
@@
 Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified.
-These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
+These issues are somewhat alleviated by the guaranteed \(\mathrm{G}^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
 That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
@@
 			\begin{align*}
-				\overline{\cdot} & : \mathbb{R}^2 \rightarrow \mathbb{R}^2     & \overline{v}      & = \vector{-v_2}{v_1}                      && \text{orthogonal vector}\\
+				\overline{\cdot} & : \mathbb{R}^2 \rightarrow \mathbb{R}^2     & \overline{v}      & = \vectorB{-v_2}{v_1}                      && \text{orthogonal vector}\\
 				\alpha           & : \mathbb{R}^2 \rightarrow \mathbb{R}       & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1}                && \text{polar coordinate angle}
@@
 				\theta           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\theta}{t} & = \apply{\alpha}{\apply{p'}{t}}           && \text{direction}\\
-				\kappa           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\kappa}{t} & = \frac{\apply{\theta'}{t}}{\apply{s}{t}} = \frac{\apply{p''}{t}^\mathrm{T}\overline{\apply{p'}{t}}}{\apply{s}{t}^3} && \text{curvature}
+				\kappa           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\kappa}{t} & = \frac{\apply{\theta'}{t}}{\apply{s}{t}} && \text{curvature}
 			\end{align*}
@@
 			Using these definitions, the direction \(\apply{\theta}{t}\) is equal to the tangent angle at the point \(\apply{p}{t}\) and the curvature \(\apply{\kappa}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{p}{t}\) or \(0\) if the curve is locally a straight line at the point \(\apply{p}{t}\).
+
+		\subsection{Continuity}
+		\label{section:continuity}
+
+			% TODO: introduce parametric and geometric continuity
 
@@
 			The goal of this project is to figure out which specification language and fairness measure yield the best usability in the resulting curve design tool.
-In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves which are connected with \(G^2\) continuity and whose parameters are determined using nonlinear optimization on some objective function determined by the specification and the fairness measure.
-This approach allows us to conveniently experiment with different specification languages and fairness measures in order to assess their usability.
+In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves whose parameters are determined using nonlinear optimization on some objective function determined by the specification and the fairness measure.
+This approach allows us to conveniently experiment with specification languages and fairness measures in order to assess their usability.
 
@@
 The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will focus on the properties point, direction and curvature as defined in \ref{section:parametric_plane_curves}.
-In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties.
+In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties at any number of positions on the curve.
 
@@
 
-				Having chosen a specification language and a fairness measure, it is time to formulate the optimization problem.
+				Having chosen a specification language and a fairness measure, it is time to state the optimization problem.
 Since polynomials of very high degree usually don't behave very well, an approach where multiple polynomial parametric curves are connected to a spline seems advisable.
@@
 
-				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need a function mapping covered arc length to parameter values (the inverse to the function \(l\) defined in section \ref{section:parametric_plane_curves}).
+				% TODO: introduce constraints and objective
+
+				Using definitions from section \ref{section:parametric_plane_curves}...
+
+				Given specification items ..., total arc length \(l^*\)...
+
+				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need the function mapping the covered arc length to the corresponding parameter values in order to state the error terms for the specification items.
 Unfortunately, this function isn't very well suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
-Thus, an additional optimization goal is to find parametric curves where this function is approximately linear.
-This property is equivaent to the speed of the curve being approximately constant and has additional useful consequences, such as the second derivative of the curve being orthogonal to the first derivative.
-Since the chosen specification language already specifies the total arc length of the curve, and the parameter runs through the unit range, the target value for the speed is simply the total arc length of the curve (\(l^*\)), leading to the following error term:
+Additionally, we need to incorporate the specification of the total arc length of the curve into the optimization goals.
+In order to address these problems, the following error term is used as part of the objective function:
 				\begin{displaymath}
-					\int_{0}^{1}\left(\apply{s}{t} - l^*\right)^2\;\mathrm{d}t
+					\apply{se}{c} = \int_{0}^{1}\left(\apply{s_c}{t} - l^*\right)^2\;\mathrm{d}t
 				\end{displaymath}
+				For polynomial splines with sufficient degree and segment count, this error term will be brought close to zero, making the speed approximately constant with value \(l^*\).
+This, in turn, makes the covered arc length function approximately linear:
+				\begin{displaymath}
+					\apply{l_c}{t} = \int_{0}^{t}\apply{s_c}{u}\;\mathrm{d}u \approx \int_{0}^{t}l^*\;\mathrm{d}u = l^*t
+				\end{displaymath}
+				This causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the error terms for the specification items.
+It also causes the total arc length of the curve to be equal to \(l^*\).
+
+				As mentioned in section \ref{section:specification_language}, we are not concerned with properties beyond the second derivative of the curve.
+Thus, we will only require \(\mathrm{G}^2\) continuity for the whole spline.
+While the polynomial curve segments are smooth (\(\mathrm{C}^\infty\)), the resulting spline in general is not.
+The segments have to be pieced together in such a way as to ensure a certain degree of smoothness.
+Thus, to ensure \(\mathrm{G}^2\) continuity for the whole curve, it is sufficient to ensure \(\mathrm{G}^2\) continuity at the segment connection points.
+Since the constant speed requirement introduced in the previous paragraph collapses \(\mathrm{C}^2\) continuity and \(\mathrm{G}^2\) continuity, ensuring \(\mathrm{C}^2\) continuity is equivalent, yet better suited for nonlinear optimization.
+Thus, for each connection point (\(0 \leq k \leq n - 1\)), three constraints are added:
+				\begin{align*}
+					\apply{p_{c_k}}{1}   & = \apply{p_{c_{k+1}}}{0}\\
+					\apply{p'_{c_k}}{1}  & = \apply{p'_{c_{k+1}}}{0}\\
+					\apply{p''_{c_k}}{1} & = \apply{p''_{c_{k+1}}}{0}
+				\end{align*}
 
@@
 				% give objective and constraints function
+				% check code to see if anything was left out
 
@@
 		% numeric optimization
+		%   trapezoid sum to approximate integrals
 		%   formalization of model using symbolic functions in casadi

### 2013-08-27 idp/documentation/paper.tex
math formatting, continued working on optimization section

@@
 \usepackage{textcomp}          % adds symbols like \textdegree
-\usepackage{mathtools}         % modern version of ams math package
+\usepackage{graphicx}          % enables including graphics
+\usepackage{amsmath}           % ams math module
 \usepackage{amssymb}           % ams math symbols
+\usepackage{mathtools}         % fixes a few quirks with the ams packages
 
@@
 				Thus, the final specification language is as follows:
-				\begin{align*}
-					\text{Positions} & = \left[0, 1\right]\\
-					\text{Points} & = \mathbb{R}^2\\
-					\text{Directions} & = \mathbb{R}\\
-					\text{Curvatures} & = \mathbb{R}\\
-					\text{SpecificationItems} & = \text{Positions} \times \left(\text{Points} \cup \text{Directions} \cup \text{Curvatures}\right)\\
-					\text{CurveLengths} & = \mathbb{R}\\
-					\text{Specifications} & = \mathcal{P}\left(\text{SpecificationItems}\right) \times \text{CurveLengths}
-				\end{align*}
+				\begin{equation*}
+					\begin{alignedat}{2}
+						& \mathrm{Positions}          && = \left[0, 1\right]\\
+						& \mathrm{Points}             && = \mathbb{R}^2\\
+						& \mathrm{Directions}         && = \mathbb{R}\\
+						& \mathrm{Curvatures}         && = \mathbb{R}\\
+						& \mathrm{SpecificationItems} && = \mathrm{Positions} \times \left(\mathrm{Points} \cup \mathrm{Directions} \cup \mathrm{Curvatures}\right)\\
+						& \mathrm{CurveLengths}       && = \mathbb{R}\\
+						& \mathrm{Specifications}     && = \mathcal{P}\left(\mathrm{SpecificationItems}\right) \times \mathrm{CurveLengths}
+					\end{alignedat}
+				\end{equation*}
 
@@
 				Many different fairness measures for plane curves have been proposed throughout history.
-Most of them set out to emulate the behavior of mechanical splines made from wood or steel, like those which were used for industrial design before the advent of computer aided design.
-The most direct approach of this type uses mechanics to derive a term for the energy in a bent steel band, arriving at the following functional describing the minimum energy curve (MEC):
-				\begin{displaymath}
+Most of them set out to emulate the behavior of mechanical splines made from wood or steel, which were used for industrial design before the advent of computer aided design.
+The most direct approach of this type uses mechanics to derive a term for the energy in an idealized bent steel band, arriving at the following functional describing the minimum energy curve (MEC):
+				\begin{equation*}
 					\int_{0}^{1}\apply{\kappa}{t}^2\;\mathrm{d}t
-				\end{displaymath}
-
+				\end{equation*}
 				While this approach is a good idealization of the mechanical spline, it has some disadvantages when used for curve design.
 For instance, it will not yield a circular arc when constrained by co-circular points.
+
 				Thus, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
 One such approach is that of the minimum variation curve (MVC), whose functional is given below:
-				\begin{displaymath}
+				\begin{equation*}
 					\int_{0}^{1}\apply{\kappa'}{t}^2\;\mathrm{d}t
-				\end{displaymath}
-
+				\end{equation*}
 				Minimum variation curves capture the intuitive notion of smoothness very well.
@@
 
-				% TODO: introduce constraints and objective
+				% TODO: figure out proper names
 
@@
 Unfortunately, this function isn't very well suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
-Additionally, we need to incorporate the specification of the total arc length of the curve into the optimization goals.
+Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which polynomial segment each specification item will end up, something that is difficult to state in a smooth optimization problem.
+Another issue is the need to incorporate the specification of the total arc length of the curve into the optimization goals, which is also related to the covered arc length function.
 In order to address these problems, the following error term is used as part of the objective function:
-				\begin{displaymath}
-					\apply{se}{c} = \int_{0}^{1}\left(\apply{s_c}{t} - l^*\right)^2\;\mathrm{d}t
-				\end{displaymath}
+				\begin{equation*}
+					\apply{e_s}{c} = \int_{0}^{1}\left(\apply{s_c}{t} - l^*\right)^2\;\mathrm{d}t
+				\end{equation*}
 				For polynomial splines with sufficient degree and segment count, this error term will be brought close to zero, making the speed approximately constant with value \(l^*\).
-This, in turn, makes the covered arc length function approximately linear:
-				\begin{displaymath}
-					\apply{l_c}{t} = \int_{0}^{t}\apply{s_c}{u}\;\mathrm{d}u \approx \int_{0}^{t}l^*\;\mathrm{d}u = l^*t
-				\end{displaymath}
-				This causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the error terms for the specification items.
+This, in turn, makes the covered arc length function approximately linear (\(\apply{l_c}{t} \approx l^*t\)) which causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the error terms for the specification items.
 It also causes the total arc length of the curve to be equal to \(l^*\).
@@
 Thus, to ensure \(\mathrm{G}^2\) continuity for the whole curve, it is sufficient to ensure \(\mathrm{G}^2\) continuity at the segment connection points.
-Since the constant speed requirement introduced in the previous paragraph collapses \(\mathrm{C}^2\) continuity and \(\mathrm{G}^2\) continuity, ensuring \(\mathrm{C}^2\) continuity is equivalent, yet better suited for nonlinear optimization.
-Thus, for each connection point (\(0 \leq k \leq n - 1\)), three constraints are added:
-				\begin{align*}
-					\apply{p_{c_k}}{1}   & = \apply{p_{c_{k+1}}}{0}\\
-					\apply{p'_{c_k}}{1}  & = \apply{p'_{c_{k+1}}}{0}\\
-					\apply{p''_{c_k}}{1} & = \apply{p''_{c_{k+1}}}{0}
-				\end{align*}
+Since the constant speed requirement introduced in the previous paragraph collapses \(\mathrm{C}^2\) continuity and \(\mathrm{G}^2\) continuity, ensuring \(\mathrm{C}^2\) continuity is equivalent, yet better suited for optimization.
+Thus, for each connection point (\(0 \leq k \leq n - 1\)), three error terms are added:
+				\begin{equation*}
+					\begin{aligned}
+						\apply{e_a}{c} & = \apply{p_{c_k}}{1} - \apply{p_{c_{k+1}}}{0}\\
+						\apply{e_b}{c} & = \apply{p'_{c_k}}{1} - \apply{p'_{c_{k+1}}}{0}\\
+						\apply{e_c}{c} & = \apply{p''_{c_k}}{1} - \apply{p''_{c_{k+1}}}{0}
+					\end{aligned}
+				\end{equation*}
 
-				% constant velocity
-				% fairness
-				% specification
-				% continuity connections
+				Having taken care of the somewhat technical details arising from using polynomial splines, it's time to state the error terms for the specification items:
+				\begin{equation*}
+					\begin{aligned}
+						\apply{e_d}{c} & = \apply{p_c}{t_s} - p_s\\
+						\apply{e_e}{c} & = \apply{\theta_c}{t_s} - \theta_s\\
+						\apply{e_f}{c} & = \apply{\kappa_c}{t_s} - \kappa_s
+					\end{aligned}
+				\end{equation*}
+
+				Finally, the error term for the MVC fairness measure as defined in section \ref{section:fairness_measure}:
+				\begin{equation*}
+					\apply{e_f}{c} = \int_{0}^{1}\apply{\kappa'_c}{t}^2\;\mathrm{d}t
+				\end{equation*}
+
+				% TODO: introduce constraints and objective model of ipopt
 

### 2013-08-28 source/Kurve/Kurve.Curves/FunctionTermCurves/PolynomialFunctionTermCurveTemplate.cs
fixed a small bug, increased trapezoid density

@@ -74,7 +74,7 @@ namespace Kurve.Curves
 		{
 			if (source == null) throw new ArgumentNullException("source");
 
-			return (int)source.Element("degree");
+			return (int)source.Element("length");
 		}
 	}
 }

### 2013-08-28 source/Kurve/Kurve.Curves/Optimization/OptimizationProblem.cs

@@ -46,7 +46,7 @@ namespace Kurve.Curves.Optimization
 				let curve = segment.LocalCurve
 				let speed = curve.Speed.Apply(position)
 				let error = Terms.Square(Terms.Difference(speed, targetSpeed))
-				select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 32)
+				select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
 			);
 
 			ValueTerm fairnessError = Terms.Scaling
@@ -60,7 +60,7 @@ namespace Kurve.Curves.Optimization
 					let jerk = curve.Jerk.Apply(position)
 					let normal = Terms.Normal(curve.Velocity.Apply(position))
 					let error = Terms.Square(Terms.DotProduct(jerk, normal))
-					select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 32)
+					select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
 				)
 			);
 

### 2013-08-29 idp/documentation/paper.tex
reorganized proposed solution section

@@
 			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority.
-Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on some specification and some fairness measure.
+Together with a good fairness measure, it will then lead to usable curve design tools.
+Since it is not immediately clear which specification language or which fairness measure is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on explicit specifications and a fairness measure.
 
@@
 Note that many specifics are still left undetermined in this approach since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like.
-It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single parametric curve, or if the curve should be determined by numerical optimization or by explicit construction.
+It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single curve, or if the curve should be determined by numerical optimization or by explicit construction.
 
-		\subsection{Parametric Plane Curves}
+		\subsection{Definitions}
+		\label{section:definitions}
+
+			TODO
+
+			\subsubsection{Parametric Plane Curves}
 			\label{section:parametric_plane_curves}
@@
 
-		\subsection{Continuity}
+			\subsubsection{Continuity}
 			\label{section:continuity}
@@
 
-		\subsection{Nonlinear Optimization on Polynomial Curves}
-		\label{section:nonlinear_optimization_polynomial_curves}
+		\subsection{Specification Language and Fairness Measure}
+		\label{section:specification_language_fairness_measure}
 
-			After having established the general approach to curve design tools that the usability analysis seems to suggest in section \ref{section:specification-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
-
-			The goal of this project is to figure out which specification language and fairness measure yield the best usability in the resulting curve design tool.
-In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves whose parameters are determined using nonlinear optimization on some objective function determined by the specification and the fairness measure.
-This approach allows us to conveniently experiment with specification languages and fairness measures in order to assess their usability.
+			After using the results of the usability analysis to establish the general approach to curve design tools in section \ref{section:specification-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
+The first steps are to choose both a specification language and a fairness measure.
 
@@
 				% TODO: add citations
+				% TODO: make clear that these terms are to be minimized
 
@@
 
+		\subsection{Nonlinear Optimization on Polynomial Curves}
+		\label{section:nonlinear_optimization_polynomial_curves}
+
+			Having decided on a specification language and a fairness measure in section \ref{section:specification_language_fairness_measure}, the only thing missing from the process described in section \ref{section:specification-based_curves} is a way of turning specifications into actual curves.
+In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves whose parameters are determined using nonlinear optimization based on the specification and the fairness measure.
+
 			\subsubsection{Optimization Problem}
@@
 
-				Having chosen a specification language and a fairness measure, it is time to state the optimization problem.
-Since polynomials of very high degree usually don't behave very well, an approach where multiple polynomial parametric curves are connected to a spline seems advisable.
-This way, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline.
-In the following, we will denote the \(i\)th coefficient of the \(k\)th polynomial by \(c_{k,i}\), and the vector of all coefficients by \(c\).
+				Nonlinear optimization problems are usually expressed using an objective function \(f : \mathbb{R}^n \rightarrow \mathbb{R}\), a constraint function \(g : \mathbb{R}^n \rightarrow \mathbb{R}^m\) and constraint bounds \(g_l, g_u \in \mathbb{R}^m\).
+The optimization problem is then given by the task of finding an \(x^* \in \mathbb{R}^n\) such that:
+				\begin{equation*}
+					\begin{aligned}
+						x^* = \min_{x \in \mathbb{R}^n} \apply{f}{x}\\
+						g_l \leq \apply{g}{x^*} \leq g_u
+					\end{aligned}
+				\end{equation*}
+				This allows expression of a variety of optimization goals.
+The ones that are important for us are zero-error constraints (using the constraint function and bounds to specify that an error term is zero), upper bounds on errors (using the constraint function and bounds to specify an upper bound for an error term) and minimization goals (using the objective function to express that an error term should be minimized).
 
-				% TODO: figure out proper names
+				In the following sections, we will introduce the optimization goals that enable us to use nonlinear optimization for building curves based on specifications using the specification language and fairness measure established in section \ref{section:specification_language_fairness_measure}.
+We will also state how they are expressed using constraint function and bounds as well as objective function.
 
-				Using definitions from section \ref{section:parametric_plane_curves}...
+			\subsubsection{Constant Speed}
+			\label{section:constant_speed}
 
-				Given specification items ..., total arc length \(l^*\)...
+				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need the function mapping the covered arc length to the corresponding parameter values in order to state the optimization goals for the specification items.
+Unfortunately, this function is not very well-suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
+Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which segment each specification item will end up, something that is difficult to state in a smooth optimization problem.
+Another issue is the need to incorporate the specification of the total arc length of the curve into the optimization goals, which is also related to the covered arc length function.
+In order to address these problems, the following error term is used:
+				\begin{equation*}
+					\apply{e_s}{c} = \max \left\{\left|\apply{s_c}{t} - l^*\right| \; \middle| \; t \in \left[0, 1\right]\right\}
+				\end{equation*}
+				For polynomial splines with sufficient degree and segment count, this term will be brought close to zero in the optimization, making the speed approximately constant with value \(l^*\).
+This, in turn, makes the covered arc length function approximately linear (\(\apply{l_c}{t} \approx l^*t\)) which causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the optimization goals for the specification items as well as fixing the segment for each specification item.
+It also causes the total arc length of the curve to be approximately equal to \(l^*\), as requested by the specification.
 
-				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need the function mapping the covered arc length to the corresponding parameter values in order to state the error terms for the specification items.
-Unfortunately, this function isn't very well suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
-Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which polynomial segment each specification item will end up, something that is difficult to state in a smooth optimization problem.
-Another issue is the need to incorporate the specification of the total arc length of the curve into the optimization goals, which is also related to the covered arc length function.
-In order to address these problems, the following error term is used as part of the objective function:
+				Since this goal can only be fulfilled approximately by polynomial curves, it is impossible to state it as a zero-error constraint.
+It could however be stated as both an upper bound on errors and a minimization goal.
+Both approaches were implemented and tested, with the latter proving to yield a more stable optimization process.
+
+				Since the error term given above is not well-suited for smooth optimization, the following approximation was finally used as part of the objective function:
 				\begin{equation*}
@@
 				\end{equation*}
-				For polynomial splines with sufficient degree and segment count, this error term will be brought close to zero, making the speed approximately constant with value \(l^*\).
-This, in turn, makes the covered arc length function approximately linear (\(\apply{l_c}{t} \approx l^*t\)) which causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the error terms for the specification items.
-It also causes the total arc length of the curve to be equal to \(l^*\).
+
+			\subsubsection{Continuity Connections}
+			\label{section:continuity_connections}
 
@@
 Since the constant speed requirement introduced in the previous paragraph collapses \(\mathrm{C}^2\) continuity and \(\mathrm{G}^2\) continuity, ensuring \(\mathrm{C}^2\) continuity is equivalent, yet better suited for optimization.
-Thus, for each connection point (\(0 \leq k \leq n - 1\)), three error terms are added:
+Thus, for each connection point (\(0 \leq k \leq n - 1\)), the following error terms are added:
 				\begin{equation*}
@@
 				\end{equation*}
+
+			\subsubsection{Specification Items}
+			\label{section:specification_items}
 
@@
 
+			\subsubsection{Fairness Error}
+			\label{section:fairness_error}
+
 				Finally, the error term for the MVC fairness measure as defined in section \ref{section:fairness_measure}:
@@
 
-				% TODO: introduce constraints and objective model of ipopt
+			\subsubsection{Complete Formalization}
+			\label{section:complete_formalization}
 
-				% segmentation, separation of segmentation and specification, global positions on curve
+				Since polynomials of very high degree usually don't behave very well, an approach where multiple polynomial parametric curves are connected to a spline seems advisable.
+This way, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline.
+In the following, we will denote the \(i\)th coefficient of the \(k\)th polynomial by \(c_{k,i}\), and the vector of all coefficients by \(c\).
+
+				% TODO: proper explanation of how segments are put together to a spline (wrt. parameter t)
+				% TODO: figure out proper names, use asterisks consistently
+				% TODO: where to use squares in errors terms, absolute values, etc.
+
+				The continuity connections need exhibit zero error, since it is not acceptable if the curve segments don't touch, have sharp corners or curvature discontinuities.
+The specification item error terms should also have zero error, since the user usually wants to specify exact curve behavior rather than indirectly influence its properties.
+It's clear that in many cases, the fairness error term will not be zero, and in fact, may be arbitrarily large, so the fairness error term is best stated as an optimization goal rather than a hard bound on the error.
+
+				Using definitions from section \ref{section:parametric_plane_curves}...
+
+				Given specification items ..., total arc length \(l^*\)...
+
 				% give objective and constraints function

### 2013-08-31 idp/documentation/paper.tex
introduced description and description language

@@
 
-			We conclude that the two ways in which curve design software can affect the design process are through the choice of specification language and fairness measure.
+			% TODO: properly instroduce description language, how the fairness measure determines semantics but not syntax of that language, etc., figure out when to use description and when to use specification, since they are syntactically identical
+			We conclude that the two ways in which curve design software can affect the design process are through the choice of specification language and fairness measure, which we will call the description language.
+The usability is of course also affected by the software's ability to derive curves from descriptions.
+
 			Ideally, the specification language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
 The fairness should capture the intuitive notion of curve smoothness sufficiently well.
-These goals are competing with the software's ability to derive curves from specifications and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive specification languages and more complex fairness measures.
+These goals are competing with the software's ability to derive curves from descriptions reliably and efficiently, a process which becomes increasingly difficult with more expressive specification languages and more complex fairness measures.
 
@@
 
+				% TODO: explain velocity
 				In existing vector graphics software, Bézier splines are edited by more or less directly modifying the coefficient points of the underlying Bézier curves.
@@
 
+				% TODO: consider moving away from the 'source image is traced' samples and use the actual vector graphics leading to those curves
+				% TODO: make sure the size of the node visualization is consistent, use visible colors for node visualization, maybe get rid of circles for handles
 				\begin{figure}[htb]
@@
 
-				% TODO: improve example index numbers
+				% TODO: improve example index numbers (also use font consistent with text)
 				Figure \ref{figure:usability_bézier} illustrates some of these issues by showing some examples of Bézier splines as they appear in common design tools.
@@
 
-		\subsection{Specification-Based Curves}
-		\label{section:specification-based_curves}
+		\subsection{Description-Based Curves}
+		\label{section:description-based_curves}
 
-			Seeing how the specification language and fairness measure almost completely determine all usability aspects of the resulting curve design tool, it seems unwise to choose them based on low-level mathematical aspects of the curve as is the case with Bézier splines.
+			Seeing how the description language almost completely determine all usability aspects of the resulting curve design tool, it seems unwise to choose them based on low-level mathematical aspects of the curve as is the case with Bézier splines.
 Spiro splines are much better in that a lot of effort was put into the fairness measure.
@@
 			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority.
-Together with a good fairness measure, it will then lead to usable curve design tools.
-Since it is not immediately clear which specification language or which fairness measure is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on explicit specifications and a fairness measure.
+Together with a good fairness measure, it forms a description language that, if it allows for reliable and efficient derivation of curves, will then lead to usable curve design tools.
+Since it is not immediately clear which description language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on descriptions.
 
@@
 				\centering
-				\includegraphics[width=\textwidth]{../resources/specification-based_curves.pdf}
-				\caption{Specification-Based Curves}
-				\label{figure:specification-based_curves}
+				\includegraphics[width=\textwidth]{../resources/description-based_curves.pdf}
+				\caption{Description-Based Curves}
+				\label{figure:description-based_curves}
 			\end{figure}
 
-			Figure \ref{figure:specification-based_curves} shows such a process.
+			Figure \ref{figure:description-based_curves} shows such a process.
 Note that many specifics are still left undetermined in this approach since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like.
-It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single curve, or if the curve should be determined by numerical optimization or by explicit construction.
+It has yet to be determined what the description language should be, as well as how to derive curves from them.
 
@@
 
-		\subsection{Specification Language and Fairness Measure}
-		\label{section:specification_language_fairness_measure}
+		\subsection{Description Language}
+		\label{section:description_language}
 
-			After using the results of the usability analysis to establish the general approach to curve design tools in section \ref{section:specification-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
+			After using the results of the usability analysis to establish the general approach to curve design tools in section \ref{section:description-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
 The first steps are to choose both a specification language and a fairness measure.
@@
 
-		\subsection{Nonlinear Optimization on Polynomial Curves}
-		\label{section:nonlinear_optimization_polynomial_curves}
+		\subsection{Curve Derivation}
+		\label{section:curve_derivation}
 
-			Having decided on a specification language and a fairness measure in section \ref{section:specification_language_fairness_measure}, the only thing missing from the process described in section \ref{section:specification-based_curves} is a way of turning specifications into actual curves.
+			Having decided on a specification language and a fairness measure in section \ref{section:description_language}, the only thing missing from the process described in section \ref{section:description-based_curves} is a way of turning specifications into actual curves.
 In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves whose parameters are determined using nonlinear optimization based on the specification and the fairness measure.
@@
 
-				In the following sections, we will introduce the optimization goals that enable us to use nonlinear optimization for building curves based on specifications using the specification language and fairness measure established in section \ref{section:specification_language_fairness_measure}.
+				In the following sections, we will introduce the optimization goals that enable us to use nonlinear optimization for building curves based on specifications using the specification language and fairness measure established in section \ref{section:description_language}.
 We will also state how they are expressed using constraint function and bounds as well as objective function.

### 2013-08-31 idp/documentation/paper.tex
moved definitions section
improved math notation

@@
 
+\input{math.tex}               % some math helper macros
+
 % tells LaTeX to avoid orphan lines at the beginning or end of a page
@@
 \widowpenalty = 10000
-
-% math commands
-\DeclareMathOperator{\atan2}{atan2}
-\newcommand{\apply}[2]{#1\!\left(#2\right)}
-\newcommand{\vectorA}[1]{\begin{pmatrix}#1\end{pmatrix}}
-\newcommand{\vectorB}[2]{\begin{pmatrix}#1\\\linebreak{}#2\end{pmatrix}}
-\newcommand{\vectorC}[3]{\begin{pmatrix}#1\\\linebreak{}#2\\\linebreak{}#3\end{pmatrix}}
 
@@
 % TODO: maybe extract topical project progress from log
+% TODO: consistent use of polynomial, planar, plane, parametric, curve
 
@@
 Fairness ususally encodes intuitive notions such as smoothness using mathematical concepts such as continuity and change in curvature.
-Here, the software has another opportunity to influence the design process both positively and negatively depending on how the fairness measure is chosen as well as by how good the software is at reliably finding a curve that has a high fairness.
+Here, the software has another opporunity to influence the design process both positively and negatively depending on how the fairness measure is chosen as well as by how good the software is at reliably finding a curve that has a high fairness.
 
@@
 For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve.
-Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \left(p_1 - p_0\right) \cdot t\), thus uniquely identifying the desired curve.
+Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \xp{p_1 - p_0} t\), thus uniquely identifying the desired curve.
 Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
-In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(\mathrm{G}^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(\contg{2}\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
 
@@
 				Bézier splines are probably the most widely used type of curve in two-dimensional vector graphics.
-Bézier splines are defined piecewise using Bézier curves, which are planar parametric curves (functions \(\left[0,1\right] \rightarrow \mathbb{R}^2\)) realized using polynomials (usually cubic) in Bernstein form.
-A Bézier curve \(B\) of degree \(n\) consists of \(n + 1\) coefficients \(P_i \in \mathbb{R}^2\).
-The curve will always go through the first and the last coefficient point (\(\apply{B}{0} = P_0, \apply{B}{1} = P_n\)) and stay within the convex hull of the coefficient points at all other times.
-Bézier curves in a Bézier spline are usually pieced together such that the resulting spline is \(G^1\) continuous.
+Bézier splines are defined piecewise using Bézier curves, which are planar parametric curves (functions \(\function{\unit}{\R{2}}\)) realized using polynomials (usually cubic) in Bernstein form.
+A Bézier curve \(\phi\) of degree \(n\) consists of \(n + 1\) coefficients \(c_i \in \R{2}\).
+The curve will always go through the first and the last coefficient point (\(\apply{\phi}{0} = c_0, \apply{\phi}{1} = c_n\)) and stay within the convex hull of the coefficient points at all other times.
+Bézier curves in a Bézier spline are usually pieced together such that the resulting spline is \(\contg{1}\) continuous.
 
@@
 In that sense, the user can specify start end end points of the curve, as well as start and end velocities of the curve using the coefficient points.
-Most vector graphics tools have functions that allow preserving the \(G^1\) continuity of the spline by automatically modifying adjacent coefficient points accordingly.
-They usually do not ensure \(\mathrm{G}^2\) continuity though.
+Most vector graphics tools have functions that allow preserving the \(\contg{1}\) continuity of the spline by automatically modifying adjacent coefficient points accordingly.
+They usually do not ensure \(\contg{2}\) continuity though.
 Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more Bézier curves to the spline.
@@
 
-				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing \(\mathrm{G}^2\) continuity.
-It is difficult for humans to choose the coefficient points in such a way that \(\mathrm{G}^2\) continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
+				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing \(\contg{2}\) continuity.
+It is difficult for humans to choose the coefficient points in such a way that \(\contg{2}\) continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
 The problem is aggravated by the fact that curvature discontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
-In case the software allows the user to request \(\mathrm{G}^2\) continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
+In case the software allows the user to request \(\contg{2}\) continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
 Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with specification language expressiveness.
@@
 It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
-Another issue is that without software support for \(\mathrm{G}^2\) continuity, using more nodes usually also introduces more points where curvature is not continuous.
+Another issue is that without software support for \(\contg{2}\) continuity, using more nodes usually also introduces more points where curvature is not continuous.
 In cases where the curvature is an important property of the curve (for example the constant curvature of circular arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.
@@
 There is no way to directly specify tangent angle and/or curvature.
-The splines are guaranteed to have \(\mathrm{G}^2\) continuity.
+The splines are guaranteed to have \(\contg{2}\) continuity.
 
@@
 The user supplies a set of points for which the software computes an interpolating spline.
-The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is \(\mathrm{G}^2\) continuity and no excessive winding.
+The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is \(\contg{2}\) continuity and no excessive winding.
 
@@
 Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified.
-These issues are somewhat alleviated by the guaranteed \(\mathrm{G}^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
+These issues are somewhat alleviated by the guaranteed \(\contg{2}\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
 That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
@@
 
+	\section{Definitions}
+	\label{section:definitions}
+
+		TODO
+
+		\subsection{Parametric Plane Curves}
+		\label{section:parametric_plane_curves}
+
+			In order to be able to reason about curves, a few basic definitions are needed.
+
+			Definitions about vectors in the euclidean plane:
+			\begin{align*}
+				\overline{\cdot} & : \function{\R{2}}{\R{2}} & \overline{v}      & = \vectorB{-v_2}{v_1}      && \text{orthogonal vector}\\
+				\alpha           & : \function{\R{2}}{\RR} & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1} && \text{polar coordinate angle}
+			\end{align*}
+
+			Definitions about parametric curves:
+			\begin{align*}
+				\phi    & : \function{\unit}{\R{2}} &                    &                                                && \text{point}\\
+				\sigma  & : \function{\unit}{\RR}   & \apply{\sigma}{t}  & = \xn{\apply{\phi'}{t}}                        && \text{speed}\\
+				\lambda & : \function{\unit}{\RR}   & \apply{\lambda}{t} & = \integral{0}{t}{\apply{\sigma}{u}}{u}        && \text{covered arc length}\\
+				\delta  & : \function{\unit}{\RR}   & \apply{\delta}{t}  & = \apply{\alpha}{\apply{\phi'}{t}}             && \text{direction}\\
+				\chi    & : \function{\unit}{\RR}   & \apply{\chi}{t}    & = \frac{\apply{\delta'}{t}}{\apply{\sigma}{t}} && \text{curvature}
+			\end{align*}
+
+			Using these definitions, the direction \(\apply{\delta}{t}\) is equal to the tangent angle at the point \(\apply{\phi}{t}\) and the curvature \(\apply{\chi}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{\phi}{t}\) and it is \(0\) if the curve is locally a straight line at the point \(\apply{\phi}{t}\).
+
+		\subsection{Continuity}
+		\label{section:continuity}
+
+			% TODO: introduce parametric and geometric continuity
+
 	\section{Proposed Solution}
@@
 
-		Having analyzed the usability aspect of curve design tools, it's time to think about how to properly address their current issues.
+		Having analyzed the usability aspect of curve design tools in \ref{section:problem_analysis}, it's time to think about how to properly address their current issues.
 In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as our specific ideas for developing a piece of software using this approach.
@@
 			Figure \ref{figure:description-based_curves} shows such a process.
-Note that many specifics are still left undetermined in this approach since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like.
-It has yet to be determined what the description language should be, as well as how to derive curves from them.
-
-		\subsection{Definitions}
-		\label{section:definitions}
-
-			TODO
-
-			\subsubsection{Parametric Plane Curves}
-			\label{section:parametric_plane_curves}
-
-				In order to be able to reason about curves, a few basic definitions are needed.
-
-				Definitions about vectors in the euclidean plane:
-				\begin{align*}
-					\overline{\cdot} & : \mathbb{R}^2 \rightarrow \mathbb{R}^2     & \overline{v}      & = \vectorB{-v_2}{v_1}                     && \text{orthogonal vector}\\
-					\alpha           & : \mathbb{R}^2 \rightarrow \mathbb{R}       & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1}                && \text{polar coordinate angle}
-				\end{align*}
-
-				Definitions about parametric curves:
-				\begin{align*}
-					p                & : \left[0,1\right] \rightarrow \mathbb{R}^2 &                   &                                           && \text{point}\\
-					s                & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{s}{t}      & = \left\|\apply{p'}{t}\right\|            && \text{speed}\\
-					l                & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{l}{t}      & = \int_{0}^{t}\apply{s}{u}\;\mathrm{d}u   && \text{covered arc length}\\
-					\theta           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\theta}{t} & = \apply{\alpha}{\apply{p'}{t}}           && \text{direction}\\
-					\kappa           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\kappa}{t} & = \frac{\apply{\theta'}{t}}{\apply{s}{t}} && \text{curvature}
-				\end{align*}
-
-				Using these definitions, the direction \(\apply{\theta}{t}\) is equal to the tangent angle at the point \(\apply{p}{t}\) and the curvature \(\apply{\kappa}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{p}{t}\) or \(0\) if the curve is locally a straight line at the point \(\apply{p}{t}\).
-
-			\subsubsection{Continuity}
-			\label{section:continuity}
-
-				% TODO: introduce parametric and geometric continuity
+Note that this approach doesn't give any specifics yet, since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like.
+It has yet to be determined what the description language should be, as well as how to derive curves from these descriptions.
 
@@
 
-			After using the results of the usability analysis to establish the general approach to curve design tools in section \ref{section:description-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
-The first steps are to choose both a specification language and a fairness measure.
+			After establishing the general approach to curve design tools in section \ref{section:description-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
+The first step is to choose a description language, consisting of both a specification language and a fairness measure.
+
+			% TODO: talk about insensitivity of the human visual system to properties beyond the second derivative in both sections, build everything on that premise
 
@@
 					\begin{alignedat}{2}
-						& \mathrm{Positions}          && = \left[0, 1\right]\\
-						& \mathrm{Points}             && = \mathbb{R}^2\\
-						& \mathrm{Directions}         && = \mathbb{R}\\
-						& \mathrm{Curvatures}         && = \mathbb{R}\\
+						& \mathrm{Positions}          && = \unit\\
+						& \mathrm{Points}             && = \R{2}\\
+						& \mathrm{Directions}         && = \RR\\
+						& \mathrm{Curvatures}         && = \RR\\
 						& \mathrm{SpecificationItems} && = \mathrm{Positions} \times \left(\mathrm{Points} \cup \mathrm{Directions} \cup \mathrm{Curvatures}\right)\\
-						& \mathrm{CurveLengths}       && = \mathbb{R}\\
-						& \mathrm{Specifications}     && = \mathcal{P}\left(\mathrm{SpecificationItems}\right) \times \mathrm{CurveLengths}
+						& \mathrm{CurveLengths}       && = \RR\\
+						& \mathrm{Specifications}     && = \powerset{\mathrm{SpecificationItems}} \times \mathrm{CurveLengths}
 					\end{alignedat}
@@
 				% TODO: make clear that these terms are to be minimized
+				% TODO: explain how the MVC fairness reduces unnecessary inflection points, etc.
 
@@
 				\begin{equation*}
-					\int_{0}^{1}\apply{\kappa}{t}^2\;\mathrm{d}t
+					\integral{0}{1}{\apply{\chi}{t}^2}{t}
 				\end{equation*}
@@
 				\begin{equation*}
-					\int_{0}^{1}\apply{\kappa'}{t}^2\;\mathrm{d}t
+					\integral{0}{1}{\apply{\chi'}{t}^2}{t}
 				\end{equation*}
@@
 
-			Having decided on a specification language and a fairness measure in section \ref{section:description_language}, the only thing missing from the process described in section \ref{section:description-based_curves} is a way of turning specifications into actual curves.
-In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves whose parameters are determined using nonlinear optimization based on the specification and the fairness measure.
+			Implementing the approach presented in section \ref{section:description-based_curves} requires a way of turning descriptions into actual curves.
+We chose to use nonlinear optimization on polynomial splines in order to achieve this goal.
+
+			Nonlinear optimization allowed us to stay relatively flexible with respect to the description language and enabled us to experiment with different specification languages and fairness measures, assessing their usability before settling on the choices presented in section \ref{section:description_language}.
+
+			Polynomial parametric curves were chosen since they are mathematically simple, very flexible through the choice of their degree, and are well-suited to approximate a variety of smooth curves.
+We opted to use polynomial splines since they allow approximating more complex curves without involving polynomials of excessively high degree, which are known to have poor stability.
 
@@
 
-				Nonlinear optimization problems are usually expressed using an objective function \(f : \mathbb{R}^n \rightarrow \mathbb{R}\), a constraint function \(g : \mathbb{R}^n \rightarrow \mathbb{R}^m\) and constraint bounds \(g_l, g_u \in \mathbb{R}^m\).
-The optimization problem is then given by the task of finding an \(x^* \in \mathbb{R}^n\) such that:
+				Nonlinear optimization problems are usually expressed using an objective function \(f : \function{\R{n}}{\RR}\), a constraint function \(g : \function{\R{n}}{\R{m}}\) and constraint bounds \(g_l, g_u \in \R{m}\).
+The optimization problem is then given by the task of finding an \(x^* \in \R{n}\) such that:
 				\begin{equation*}
-					\begin{aligned}
-						x^* = \min_{x \in \mathbb{R}^n} \apply{f}{x}\\
+					\begin{gathered}
+						x^* = \argmin_{x \in \R{n}} \apply{f}{x}\\
 						g_l \leq \apply{g}{x^*} \leq g_u
-					\end{aligned}
+					\end{gathered}
 				\end{equation*}
@@
 
-				In the following sections, we will introduce the optimization goals that enable us to use nonlinear optimization for building curves based on specifications using the specification language and fairness measure established in section \ref{section:description_language}.
-We will also state how they are expressed using constraint function and bounds as well as objective function.
+				In the following, we will describe how nonlinear optimization can be used for building polynomial splines based on descriptions in the description language established in section \ref{section:description_language}.
+
+				% TODO: proper explanation of how segments are put together to a spline (wrt. parameter t)
+				When using polynomial splines, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline.
+Let \(m\) be the number of segments.
+In the following, we will denote the vector of coefficients of the \(i\)th polynomial by \(c_i\), and the vector of all coefficients by \(c\).
+The functions defined in section \ref{section:parametric_plane_curves} are used with subscripts (\(\apply{\phi_c}{t}\)) to indicate their dependence on the coefficient vector \(c\).
+They may also be used with an indexed coefficient vector (\(\apply{\phi_{c_i}}{t}\)) when referring to the \(i\)th segment of the spline.
+
+				Let \(n\) be the number of specification items, with the \(i\)th item being at position \(\hat{t}_i\).
+If the \(i\)th item is a point specification, the point is denoted \(\hat{\phi}_i\), if it is a direction specification, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification, the curvature is doneted \(\hat{\chi}_i\).
+The specified total arc length for the curve is denoted \(\hat{\lambda}\).
 
@@
 Unfortunately, this function is not very well-suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
-Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which segment each specification item will end up, something that is difficult to state in a smooth optimization problem.
-Another issue is the need to incorporate the specification of the total arc length of the curve into the optimization goals, which is also related to the covered arc length function.
+Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
+We also need to state the specification of the total arc length of the curve as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
 In order to address these problems, the following error term is used:
 				\begin{equation*}
-					\apply{e_s}{c} = \max \left\{\left|\apply{s_c}{t} - l^*\right| \; \middle| \; t \in \left[0, 1\right]\right\}
+					\apply{e_a}{c} = \max_{t \in \unit} \xa{\apply{\sigma_c}{t} - \hat{\lambda}}
 				\end{equation*}
-				For polynomial splines with sufficient degree and segment count, this term will be brought close to zero in the optimization, making the speed approximately constant with value \(l^*\).
-This, in turn, makes the covered arc length function approximately linear (\(\apply{l_c}{t} \approx l^*t\)) which causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the optimization goals for the specification items as well as fixing the segment for each specification item.
-It also causes the total arc length of the curve to be approximately equal to \(l^*\), as requested by the specification.
+				For polynomial splines with sufficient degree and segment count, this term will be brought close to zero in the optimization, making the speed approximately constant with value \(\hat{\lambda}\).
+This, in turn, makes the covered arc length function approximately linear (\(\apply{\lambda_c}{t} \approx \hat{\lambda}t\)) which causes the inverse of the covered arc length function \(\lambda_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the optimization goals for the specification items as well as fixing the segment for each specification item.
+It also causes the total arc length of the curve to be approximately equal to \(\hat{\lambda}\), as requested by the specification.
 
-				Since this goal can only be fulfilled approximately by polynomial curves, it is impossible to state it as a zero-error constraint.
-It could however be stated as both an upper bound on errors and a minimization goal.
-Both approaches were implemented and tested, with the latter proving to yield a more stable optimization process.
+				Since this goal can only be fulfilled approximately by polynomial curves, it is impossible to use the error term given above in a zero-error constraint.
+It could however be used in both an upper bound on errors and in a minimization goal.
+Both approaches were implemented and tested, with the latter yielding a more stable optimization process.
 
-				Since the error term given above is not well-suited for smooth optimization, the following approximation was finally used as part of the objective function:
+				The error term given above is not well-suited for smooth optimization because it incorporates finding a maximum.
+Since the optimization problem will allow the original error term to be brought close to zero, we can approximate it using the following error term:
 				\begin{equation*}
-					\apply{e_s}{c} = \int_{0}^{1}\left(\apply{s_c}{t} - l^*\right)^2\;\mathrm{d}t
+					\apply{e_b}{c} = \integral{0}{1}{\xp{\apply{\sigma_c}{t} - \hat{\lambda}}^2}{t}
 				\end{equation*}
+				This term is then finally used as part of the objective function, thus making it a minimization goal.
 
@@
 
-				As mentioned in section \ref{section:specification_language}, we are not concerned with properties beyond the second derivative of the curve.
-Thus, we will only require \(\mathrm{G}^2\) continuity for the whole spline.
-While the polynomial curve segments are smooth (\(\mathrm{C}^\infty\)), the resulting spline in general is not.
+				As mentioned in section \ref{section:description_language}, we are not concerned with properties beyond the second derivative of the curve.
+Thus, we will only require \(\contg{2}\) continuity for the whole spline.
+While the polynomial curve segments are smooth (\(\contp{\infty}\) continuity), the resulting spline in general is not.
 The segments have to be pieced together in such a way as to ensure a certain degree of smoothness.
-Thus, to ensure \(\mathrm{G}^2\) continuity for the whole curve, it is sufficient to ensure \(\mathrm{G}^2\) continuity at the segment connection points.
-Since the constant speed requirement introduced in the previous paragraph collapses \(\mathrm{C}^2\) continuity and \(\mathrm{G}^2\) continuity, ensuring \(\mathrm{C}^2\) continuity is equivalent, yet better suited for optimization.
-Thus, for each connection point (\(0 \leq k \leq n - 1\)), the following error terms are added:
+Thus, to ensure \(\contg{2}\) continuity for the whole curve, it is sufficient to ensure \(\contg{2}\) continuity at the segment connection points.
+To do this, the following error terms are added for each connection point (\(0 \leq i \leq m - 1\)):
 				\begin{equation*}
 					\begin{aligned}
-						\apply{e_a}{c} & = \apply{p_{c_k}}{1} - \apply{p_{c_{k+1}}}{0}\\
-						\apply{e_b}{c} & = \apply{p'_{c_k}}{1} - \apply{p'_{c_{k+1}}}{0}\\
-						\apply{e_c}{c} & = \apply{p''_{c_k}}{1} - \apply{p''_{c_{k+1}}}{0}
+						\apply{e_c}{c} & = \apply{\phi_{c_i}}{1} - \apply{\phi_{c_{i+1}}}{0}\\
+						\apply{e_d}{c} & = \apply{\phi'_{c_i}}{1} - \apply{\phi'_{c_{i+1}}}{0}\\
+						\apply{e_e}{c} & = \apply{\phi''_{c_i}}{1} - \apply{\phi''_{c_{i+1}}}{0}
 					\end{aligned}
 				\end{equation*}
+				These error terms being zero implies \(\contp{2}\) continuity.
+However, since the constant speed requirement introduced in the previous paragraph collapses \(\contp{2}\) continuity and \(\contg{2}\) continuity, this is equivalent to the original requirement of ensuring only \(\contg{2}\) continuity, yet better suited for optimization.
+
+				Since it is not acceptable for the curve segments to not touch, have sharp corners or curvature discontinuities, the error terms above are added as zero-error constraints.
 
@@
 
-				Having taken care of the somewhat technical details arising from using polynomial splines, it's time to state the error terms for the specification items:
+				Having taken care of the somewhat technical details arising from using polynomial splines, it's time to state the error terms for the specification items.
+For each specification item (\(0 \leq i \leq n\)), we add one of the following error terms, depending on the type of the specification item:
 				\begin{equation*}
 					\begin{aligned}
-						\apply{e_d}{c} & = \apply{p_c}{t_s} - p_s\\
-						\apply{e_e}{c} & = \apply{\theta_c}{t_s} - \theta_s\\
-						\apply{e_f}{c} & = \apply{\kappa_c}{t_s} - \kappa_s
+						\apply{e_f}{c} & = \apply{\phi_c}{\hat{t}_i} - \hat{\phi}_i\\
+						\apply{e_g}{c} & = \apply{\delta_c}{\hat{t}_i} - \hat{\delta}_i\\
+						\apply{e_h}{c} & = \apply{\chi_c}{\hat{t}_i} - \hat{\chi}_i
 					\end{aligned}
 				\end{equation*}
+
+				The user usually wants to specify exact curve behavior rather than indirectly influencing its properties, so these error terms are also added as zero-error constraints.
 
@@
 
-				Finally, the error term for the MVC fairness measure as defined in section \ref{section:fairness_measure}:
+				Finally, the error term for the MVC fairness measure as defined in section \ref{section:fairness_measure} is used:
 				\begin{equation*}
-					\apply{e_f}{c} = \int_{0}^{1}\apply{\kappa'_c}{t}^2\;\mathrm{d}t
+					\apply{e_i}{c} = \integral{0}{1}{\apply{\chi'_c}{t}^2}{t}
 				\end{equation*}
+
+				It's clear that in many cases, this error term will not be zero, and in fact, may be arbitrarily large.
+Thus, it is not suitable to be constrained to zero or bounded error and has to be stated as a minimization goal.
 
@@
 			\label{section:complete_formalization}
-
-				Since polynomials of very high degree usually don't behave very well, an approach where multiple polynomial parametric curves are connected to a spline seems advisable.
-This way, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline.
-In the following, we will denote the \(i\)th coefficient of the \(k\)th polynomial by \(c_{k,i}\), and the vector of all coefficients by \(c\).
-
-				% TODO: proper explanation of how segments are put together to a spline (wrt. parameter t)
-				% TODO: figure out proper names, use asterisks consistently
-				% TODO: where to use squares in errors terms, absolute values, etc.
-
-				The continuity connections need exhibit zero error, since it is not acceptable if the curve segments don't touch, have sharp corners or curvature discontinuities.
-The specification item error terms should also have zero error, since the user usually wants to specify exact curve behavior rather than indirectly influence its properties.
-It's clear that in many cases, the fairness error term will not be zero, and in fact, may be arbitrarily large, so the fairness error term is best stated as an optimization goal rather than a hard bound on the error.
-
-				Using definitions from section \ref{section:parametric_plane_curves}...
-
-				Given specification items ..., total arc length \(l^*\)...
 

### 2013-08-31 idp/documentation/paper.tex
continued working on derivation section

@@
 				This allows expression of a variety of optimization goals.
-The ones that are important for us are zero-error constraints (using the constraint function and bounds to specify that an error term is zero), upper bounds on errors (using the constraint function and bounds to specify an upper bound for an error term) and minimization goals (using the objective function to express that an error term should be minimized).
+The ones that are important for us are zero-error constraints (using the constraint function and bounds to specify that an error term should be zero), upper bounds on errors (using the constraint function and bounds to specify an upper bound for an error term) and minimization goals (using the objective function to express that an error term should be minimized).
 
@@
 
-				% TODO: proper explanation of how segments are put together to a spline (wrt. parameter t)
-				When using polynomial splines, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline.
-Let \(m\) be the number of segments.
-In the following, we will denote the vector of coefficients of the \(i\)th polynomial by \(c_i\), and the vector of all coefficients by \(c\).
-The functions defined in section \ref{section:parametric_plane_curves} are used with subscripts (\(\apply{\phi_c}{t}\)) to indicate their dependence on the coefficient vector \(c\).
-They may also be used with an indexed coefficient vector (\(\apply{\phi_{c_i}}{t}\)) when referring to the \(i\)th segment of the spline.
+			\subsubsection{Sgementation and Notation}
+			\label{section:segmentation_notation}
 
-				Let \(n\) be the number of specification items, with the \(i\)th item being at position \(\hat{t}_i\).
+				Let \(m\) be the number of segments that make up the polynomial spline.
+The parametric curve functions are used with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline \(\apply{\phi}{t}\).
+Mapping the parameter \(t\) between the segments and the whole spline works as follows:
+				\begin{equation*}
+					\apply{\phi_i}{t} = \apply{\phi}{i + \frac{t}{m}}
+				\end{equation*}
+				That is, the segments are simply laid out consecutively in parameter space, scaled to fit the unit interval, making the parameter mapping function linear and static.
+
+				When using polynomial splines, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline, denoted \(c\).
+The parametric curve functions of both the whole spline and each segment are used with subscript \(c\) to indicate their dependence on the coefficient vector (\(\apply{\phi_c}{t}\), \(\apply{{\phi_i}_c}{t}\)) when used in error terms for the optimization.
+
+				Let \(n\) be the number of specification items.
+Let \(\hat{t}_i\) be the position of the \(i\)th specification item.
 If the \(i\)th item is a point specification, the point is denoted \(\hat{\phi}_i\), if it is a direction specification, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification, the curvature is doneted \(\hat{\chi}_i\).
@@
 				\begin{equation*}
-					\apply{e_a}{c} = \max_{t \in \unit} \xa{\apply{\sigma_c}{t} - \hat{\lambda}}
+					\max_{t \in \unit} \xa{\apply{\sigma_c}{t} - \hat{\lambda}}
 				\end{equation*}
@@
 				\begin{equation*}
-					\apply{e_b}{c} = \integral{0}{1}{\xp{\apply{\sigma_c}{t} - \hat{\lambda}}^2}{t}
+					\integral{0}{1}{\xp{\apply{\sigma_c}{t} - \hat{\lambda}}^2}{t}
 				\end{equation*}
@@
 				\begin{equation*}
-					\begin{aligned}
-						\apply{e_c}{c} & = \apply{\phi_{c_i}}{1} - \apply{\phi_{c_{i+1}}}{0}\\
-						\apply{e_d}{c} & = \apply{\phi'_{c_i}}{1} - \apply{\phi'_{c_{i+1}}}{0}\\
-						\apply{e_e}{c} & = \apply{\phi''_{c_i}}{1} - \apply{\phi''_{c_{i+1}}}{0}
-					\end{aligned}
+					\begin{gathered}
+						\apply{{\phi_i}_c}{1} - \apply{{\phi_{i + 1}}_c}{0}\\
+						\apply{{\phi'_i}_c}{1} - \apply{{\phi'_{i + 1}}_c}{0}\\
+						\apply{{\phi''_i}_c}{1} - \apply{{\phi''_{i + 1}}_c}{0}
+					\end{gathered}
 				\end{equation*}
@@
 				\begin{equation*}
-					\begin{aligned}
-						\apply{e_f}{c} & = \apply{\phi_c}{\hat{t}_i} - \hat{\phi}_i\\
-						\apply{e_g}{c} & = \apply{\delta_c}{\hat{t}_i} - \hat{\delta}_i\\
-						\apply{e_h}{c} & = \apply{\chi_c}{\hat{t}_i} - \hat{\chi}_i
-					\end{aligned}
+					\begin{gathered}
+						\apply{\phi_c}{\hat{t}_i} - \hat{\phi}_i\\
+						\apply{\delta_c}{\hat{t}_i} - \hat{\delta}_i\\
+						\apply{\chi_c}{\hat{t}_i} - \hat{\chi}_i
+					\end{gathered}
 				\end{equation*}
 
-				The user usually wants to specify exact curve behavior rather than indirectly influencing its properties, so these error terms are also added as zero-error constraints.
+				The user usually wants to specify exact curve behavior rather than indirectly influencing properties, so these error terms are also added as zero-error constraints.
 
@@
 				\begin{equation*}
-					\apply{e_i}{c} = \integral{0}{1}{\apply{\chi'_c}{t}^2}{t}
+					\integral{0}{1}{\apply{\chi'_c}{t}^2}{t}
 				\end{equation*}
@@
 
-			\subsubsection{Complete Formalization}
-			\label{section:complete_formalization}
+			\subsubsection{Final Assembly}
+			\label{section:final_assembly}
 
-				% give objective and constraints function
-				% check code to see if anything was left out
+				Having gathered all the necessary error terms, it's time to put them together.
+Terms with zero-error constraints, as well as terms with upper bounds on the error are added as component functions of the constraint function \(g\) together with appropriate bounds in \(g_l\) and \(g_u\).
+
+				Minimization goals, on the other hand, are assembled in a weighted sum which defines the the objective function \(f\).
 

### 2013-08-31 idp/documentation/paper.tex
removed some useless details

@@
 
+				Since the polynomial spline is uniquely determined by the coefficients of each polynomial segment, the optimization domain is a vector consisting of all these coefficients.
 Let \(m\) be the number of segments that make up the polynomial spline.
-The parametric curve functions are used with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline \(\apply{\phi}{t}\).
+The parametric curve functions are used with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline (\(\apply{\phi}{t}\)).
 Mapping the parameter \(t\) between the segments and the whole spline works as follows:
@@
 				That is, the segments are simply laid out consecutively in parameter space, scaled to fit the unit interval, making the parameter mapping function linear and static.
-
-				When using polynomial splines, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline, denoted \(c\).
-The parametric curve functions of both the whole spline and each segment are used with subscript \(c\) to indicate their dependence on the coefficient vector (\(\apply{\phi_c}{t}\), \(\apply{{\phi_i}_c}{t}\)) when used in error terms for the optimization.
 
@@
 				\begin{equation*}
-					\max_{t \in \unit} \xa{\apply{\sigma_c}{t} - \hat{\lambda}}
+					\max_{t \in \unit} \xa{\apply{\sigma}{t} - \hat{\lambda}}
 				\end{equation*}
 				For polynomial splines with sufficient degree and segment count, this term will be brought close to zero in the optimization, making the speed approximately constant with value \(\hat{\lambda}\).
-This, in turn, makes the covered arc length function approximately linear (\(\apply{\lambda_c}{t} \approx \hat{\lambda}t\)) which causes the inverse of the covered arc length function \(\lambda_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the optimization goals for the specification items as well as fixing the segment for each specification item.
+This, in turn, makes the covered arc length function approximately linear (\(\apply{\lambda}{t} \approx \hat{\lambda}t\)) which causes the inverse of the covered arc length function \(\lambda^{-1}\) to be approximately linear as well, facilitating the simple statement of the optimization goals for the specification items as well as fixing the segment for each specification item.
 It also causes the total arc length of the curve to be approximately equal to \(\hat{\lambda}\), as requested by the specification.
@@
 				\begin{equation*}
-					\integral{0}{1}{\xp{\apply{\sigma_c}{t} - \hat{\lambda}}^2}{t}
+					\integral{0}{1}{\xp{\apply{\sigma}{t} - \hat{\lambda}}^2}{t}
 				\end{equation*}
@@
 					\begin{gathered}
-						\apply{{\phi_i}_c}{1} - \apply{{\phi_{i + 1}}_c}{0}\\
-						\apply{{\phi'_i}_c}{1} - \apply{{\phi'_{i + 1}}_c}{0}\\
-						\apply{{\phi''_i}_c}{1} - \apply{{\phi''_{i + 1}}_c}{0}
+						\apply{\phi_i}{1} - \apply{\phi_{i + 1}}{0}\\
+						\apply{\phi'_i}{1} - \apply{\phi'_{i + 1}}{0}\\
+						\apply{\phi''_i}{1} - \apply{\phi''_{i + 1}}{0}
 					\end{gathered}
@@
 					\begin{gathered}
-						\apply{\phi_c}{\hat{t}_i} - \hat{\phi}_i\\
-						\apply{\delta_c}{\hat{t}_i} - \hat{\delta}_i\\
-						\apply{\chi_c}{\hat{t}_i} - \hat{\chi}_i
+						\apply{\phi}{\hat{t}_i} - \hat{\phi}_i\\
+						\apply{\delta}{\hat{t}_i} - \hat{\delta}_i\\
+						\apply{\chi}{\hat{t}_i} - \hat{\chi}_i
 					\end{gathered}
@@
 				\begin{equation*}
-					\integral{0}{1}{\apply{\chi'_c}{t}^2}{t}
+					\integral{0}{1}{\apply{\chi'}{t}^2}{t}
 				\end{equation*}

### 2013-09-02 idp/documentation/paper.tex
fixed a small grammar error in definitions section

@@
 The change also adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
-
-	\section{Definitions}
-	\label{section:definitions}
-
-		TODO
-
-		\subsection{Parametric Plane Curves}
-		\label{section:parametric_plane_curves}
-
-			In order to be able to reason about curves, a few basic definitions are needed.
-
-			Definitions about vectors in the euclidean plane:
-			\begin{align*}
-				\overline{\cdot} & : \function{\R{2}}{\R{2}} & \overline{v}      & = \vectorB{-v_2}{v_1}      && \text{orthogonal vector}\\
-				\alpha           & : \function{\R{2}}{\RR} & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1} && \text{polar coordinate angle}
-			\end{align*}
-
-			Definitions about parametric curves:
-			\begin{align*}
-				\phi    & : \function{\unit}{\R{2}} &                    &                                                && \text{point}\\
-				\sigma  & : \function{\unit}{\RR}   & \apply{\sigma}{t}  & = \xn{\apply{\phi'}{t}}                        && \text{speed}\\
-				\lambda & : \function{\unit}{\RR}   & \apply{\lambda}{t} & = \integral{0}{t}{\apply{\sigma}{u}}{u}        && \text{covered arc length}\\
-				\delta  & : \function{\unit}{\RR}   & \apply{\delta}{t}  & = \apply{\alpha}{\apply{\phi'}{t}}             && \text{direction}\\
-				\chi    & : \function{\unit}{\RR}   & \apply{\chi}{t}    & = \frac{\apply{\delta'}{t}}{\apply{\sigma}{t}} && \text{curvature}
-			\end{align*}
-
-			Using these definitions, the direction \(\apply{\delta}{t}\) is equal to the tangent angle at the point \(\apply{\phi}{t}\) and the curvature \(\apply{\chi}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{\phi}{t}\) and it is \(0\) if the curve is locally a straight line at the point \(\apply{\phi}{t}\).
-
-		\subsection{Continuity}
-		\label{section:continuity}
-
-			% TODO: introduce parametric and geometric continuity
 
@@
 
+	\appendix
+
+	\section{Definitions}
+	\label{section:definitions}
+
+		This section contains basic definitions used in the rest of the document.
+
+		\subsection{Plane Vectors}
+		\label{section:plane_vectors}
+
+			\begin{align*}
+				\overline{\cdot} & : \function{\R{2}}{\R{2}} & \overline{v}      & = \vectorB{-v_2}{v_1}      && \text{orthogonal vector}\\
+				\alpha           & : \function{\R{2}}{\RR} i & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1} && \text{polar coordinate angle}
+			\end{align*}
+
+		\subsection{Parametric Curves}
+		\label{section:parametric_curves}
+
+			Parametric curves are functions \(\phi : \function{\unit}{\R{n}}\), called the parametrization function.
+The represented curve is then the list of points that are traversed by this function.
+While there are many ways to parametrize a given curve, that is, represent it using a parametrization function, we are usually interested in those which yield a sufficiently smooth parametrization function, allowing for the concise description of curve properties using differential geometry.
+
+		\subsection{Parametric Plane Curves}
+		\label{section:parametric_plane_curves}
+
+			In the special case of parametric plane curves, that is, parametric curves in the euclidean plane, we can define the following functions:
+			\begin{align*}
+				\phi    & : \function{\unit}{\R{2}} &                    &                                                && \text{point}\\
+				\sigma  & : \function{\unit}{\RR}   & \apply{\sigma}{t}  & = \xn{\apply{\phi'}{t}}                        && \text{speed}\\
+				\lambda & : \function{\unit}{\RR}   & \apply{\lambda}{t} & = \integral{0}{t}{\apply{\sigma}{u}}{u}        && \text{covered arc length}\\
+				\delta  & : \function{\unit}{\RR}   & \apply{\delta}{t}  & = \apply{\alpha}{\apply{\phi'}{t}}             && \text{direction}\\
+				\chi    & : \function{\unit}{\RR}   & \apply{\chi}{t}    & = \frac{\apply{\delta'}{t}}{\apply{\sigma}{t}} && \text{curvature}
+			\end{align*}
+
+			Using these definitions, the direction \(\apply{\delta}{t}\) is equal to the tangent angle at the point \(\apply{\phi}{t}\) and the curvature \(\apply{\chi}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{\phi}{t}\) and it is \(0\) if the curve is locally a straight line at the point \(\apply{\phi}{t}\).
+
+		\subsection{Continuity}
+		\label{section:continuity}
+
+			The first way to define continuity of parametric curves is by simply transferring the notion of continuity from calculus as applied to the parametrization function.
+This leads to the notion of parametric continuity, denoted by \(\contp{n}\), stating that the \(n\)th derivative of the parametrization function is continuous.
+
+			One may notice that this notion of continuity is a little too strict.
+A curve may be arbitrarily smooth, while its parametrization does not even have \(\contp{1}\) continuity, since the derivatives of the parametrization function may have discontinuities that do not manifest themselves in the curve that they represent.
+If that is the case, one can always find a different parametrization of the same curve that does not have these discontinuities.
+This insight gives rise to the definition of geometric continuity, denoted by \(\contg{n}\), which only requires that there exists a reparametrization of the parametric curve which is \(\contp{n}\) continuous.
+
+			It follows immediately that every parametric curve that is \(\contp{n}\) continuous is also \(\contg{n}\) continuous.
+One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itsself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve.
+In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, we can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.
+
 	\begin{thebibliography}{}

### 2013-09-02 idp/documentation/paper.tex
worked on definitions section

@@
 One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itsself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve.
-In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, we can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.
+In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, we can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity \(\chi\) being continuous.
 

### 2013-09-02 idp/documentation/paper.tex
improved citations, started working on revising sections

--- 2013-09-02 idp/documentation/paper.tex
@@
 % TODO: check if all quotations are done `like this'
-% TODO: maybe extract topical project progress from log
 % TODO: consistent use of polynomial, planar, plane, parametric, curve
@@
 		Raster graphics images (for instance, digital photos and most digitally painted artworks) consist of a finite number of pixels (picture elements) which discretely specify the colors at each point of the image.
-In contrast, vector graphics designs consist of primitives like curves and surfaces, which describe the shapes present in the design using continuous mathematical functions.
-This has various advantages over the raster approach, since designs do not suffer from discretization artifacts and can, for instance, be arbitrarily scaled without any loss of quality.
+In contrast, vector graphics designs consist of primitives like curves and surfaces, which describe the shapes present in the design using mathematical functions.
+This has various advantages over the raster approach, since designs do not suffer from discretization artifacts and can, for instance, be arbitrarily scaled or repeatedly modified without any loss of quality.
 
@@
 All elements in two-dimensional vector graphics are described in terms of curves, so the design process of curves is the single most important part of two-dimensional vector graphics.
-Unfortunately, designing curves using the tools available through current vector graphics software can be very frustrating.
+Unfortunately, designing curves using the tools available in current vector graphics software can be very frustrating.
 The artist is often unable to make the curve look the way they want it to.
 To make matters worse it is often unclear why this is the case, as the software seems to put all the necessary tools into the hands of the designer, yet at times, they somehow don't allow the artist to communicate their intention to the software.
-Thus, the objective of this project is to identify the exact nature and cause of these inadequacies and then specify and implement curve design tools that do not exhibit them.
 
-		% TODO: make sure introduction does not overlap with start of next section regarding both content and wording
-		% TODO: quote raphael levien on usability of Bézier splines
-		% TODO: comment on that although there has been some research into better curves (quote MVC curves), it was less focused on usability and hasn't had much impact on actual vector graphics software which continues to almost exclusively use bézier splinse to this day
+		While there has been some promising research into improving curve design tools \cite{thesis-mvc} \cite{thesis-spiro}, it focuses mostly on the aspect of improving the fairness of curves, rather than usability as a whole.
+Unfortunately, it has not had much of an impact on actual vector graphics software used in practice, which continues to focus almost exclusively on Bézier splines to this day.
 
-	% TODO: maybe we don't want to mention G^n and C^n until formally introduced later, doesn't add much value to the usability discussion anyways
+		Thus, the objective of this project is to identify the exact nature and cause of the shortcomings of current curve design tools in order to use this knowledge to develop a curve design tool that does not exhibit them.
+
 	\section{Problem Analysis}
@@
 Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
-In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(\contg{2}\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request curvature continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
 
@@
 
+		% TODO: quote raphael levien on usability of Bézier splines
 			\subsubsection{Bézier Splines}
@@
 The curve will always go through the first and the last coefficient point (\(\apply{\phi}{0} = c_0, \apply{\phi}{1} = c_n\)) and stay within the convex hull of the coefficient points at all other times.
-Bézier curves in a Bézier spline are usually pieced together such that the resulting spline is \(\contg{1}\) continuous.
+Bézier curves in a Bézier spline are usually pieced together such that the resulting spline exhibits continuous direction.
 
@@
 In that sense, the user can specify start end end points of the curve, as well as start and end velocities of the curve using the coefficient points.
-Most vector graphics tools have functions that allow preserving the \(\contg{1}\) continuity of the spline by automatically modifying adjacent coefficient points accordingly.
-They usually do not ensure \(\contg{2}\) continuity though.
+Most vector graphics tools have functions that allow preserving the direction continuity of the spline by automatically modifying adjacent coefficient points accordingly.
+They usually do not ensure curvature continuity though.
 Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more Bézier curves to the spline.
@@
 
-				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing \(\contg{2}\) continuity.
-It is difficult for humans to choose the coefficient points in such a way that \(\contg{2}\) continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
+				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing curvature continuity.
+It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
 The problem is aggravated by the fact that curvature discontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
-In case the software allows the user to request \(\contg{2}\) continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
+In case the software allows the user to request curvature continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
 Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with specification language expressiveness.
@@
 It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
-Another issue is that without software support for \(\contg{2}\) continuity, using more nodes usually also introduces more points where curvature is not continuous.
+Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous.
 In cases where the curvature is an important property of the curve (for example the constant curvature of circular arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.
@@
 					\includegraphics[width=\textwidth]{../resources/usability_bezier.pdf}
-					\caption{Usability Analysis of Bézier Splines (left: \cite{Fluttershy}, right: \cite{Lugia})}
+					% TODO: silent citation or not?
+					\caption{Usability Analysis of Bézier Splines (left: \cite{fluttershy}, right: \cite{lugia})}
 					\label{figure:usability_bézier}
@@
 Even though they were integrated using path effects and thus have a somewhat cumbersome user interface, they constitute a commendable effort of incorporating results from curve research into vector graphics software, something that has hardly happened this far.
-More information on Spiro splines can be found in the dissertation of Raphael Levien \cite{Levien2009}.
+More information on Spiro splines can be found in the dissertation of Raphael Levien \cite{thesis-spiro}.
 
@@
 There is no way to directly specify tangent angle and/or curvature.
-The splines are guaranteed to have \(\contg{2}\) continuity.
+The splines are guaranteed to have curvature continuity.
 
@@
 The user supplies a set of points for which the software computes an interpolating spline.
-The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is \(\contg{2}\) continuity and no excessive winding.
+The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is curvature continuity and no excessive winding.
 
@@
 Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified.
-These issues are somewhat alleviated by the guaranteed \(\contg{2}\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
+These issues are somewhat alleviated by the guaranteed curvature continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
 That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
@@
 					\includegraphics[width=\textwidth]{../resources/usability_spiro.pdf}
-					\caption{Usability Analysis of Spiro Splines (left: \cite{Fluttershy}, right: \cite{Lugia})}
+					\caption{Usability Analysis of Spiro Splines (left: \cite{fluttershy}, right: \cite{lugia})}
 					\label{figure:usability_spiro}
@@
 
-			Seeing how the description language almost completely determine all usability aspects of the resulting curve design tool, it seems unwise to choose them based on low-level mathematical aspects of the curve as is the case with Bézier splines.
+			Seeing how the description language almost completely determines all usability aspects of the resulting curve design tool, it seems unwise to base it on low-level mathematical aspects of the curve as is the case with Bézier splines.
 Spiro splines are much better in that a lot of effort was put into the fairness measure.
@@
 We also need to state the specification of the total arc length of the curve as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
+Furthermore, the difficulty of expressing the fairness measure, which is naturally an integral over the arc length of the curve, is dependent on the complexity of the covered arc length function.
 In order to address these problems, the following error term is used:
@@
 
+	\clearpage
+
 	\appendix
@@
 
-		This section contains basic definitions used in the rest of the document.
+		This section contains definitions used in the rest of the document.
 
@@
 
+			A few basic definitions about vectors in the euclidean plane are needed:
 			\begin{align*}
@@
 
+	\clearpage
+
 	\begin{thebibliography}{}
 
-		\bibitem{Levien2009}
+		\bibitem{thesis-mvc}
+			\emph{Minimum Curvature Variation Curves, Networks, and Surfaces for Fair Free-Form Shape Design}\\
+			Hentry Packard Moreton\\
+			University of California, Berkeley, 1992
+
+		\bibitem{thesis-spiro}
 			\emph{From Spiral to Spline: Optimal Techniques in Interactive Curve Design}\\
@@
 
-		\bibitem{Fluttershy}
+		\bibitem{fluttershy}
 			\emph{Fluttershy}\\
@@
 
-		\bibitem{Lugia}
+		\bibitem{lugia}
 			\emph{Lugia}\\

### 2013-09-04 idp/documentation/paper.tex
updated illustrations
small fixes in illustrations and usability analysis section

@@
 In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves.
-Once such a model has been obtained, it can be used to propose criteria for good curve design software as well as to analyze potential usability issues with existing tools.
+Once such a model has been obtained, it can be used to propose criteria for good curve design tools as well as to analyze usability aspects of existing tools.
 
@@
 This notion of `most likely' is usually modelled in terms of a fairness measure, fairer curves being regarded as more likely.
-Fairness ususally encodes intuitive notions such as smoothness using mathematical concepts such as continuity and change in curvature.
-Here, the software has another opporunity to influence the design process both positively and negatively depending on how the fairness measure is chosen as well as by how good the software is at reliably finding a curve that has a high fairness.
+Fairness ususally encodes intuitive notions such as smoothness and minimality using mathematical concepts such as continuity and variation of curvature.
+Here, the software has another opportunity to influence the design process both positively and negatively depending on what fairness measure is chosen and how good the software is at reliably finding a curve that has a high fairness.
 
@@
 Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
-In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request curvature continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+
+			In a similar fashion, one may consider some shortcomings (for instance, insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request curvature continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
 
@@
 
-			% TODO: properly instroduce description language, how the fairness measure determines semantics but not syntax of that language, etc., figure out when to use description and when to use specification, since they are syntactically identical
-			We conclude that the two ways in which curve design software can affect the design process are through the choice of specification language and fairness measure, which we will call the description language.
-The usability is of course also affected by the software's ability to derive curves from descriptions.
+			We conclude that the two ways in which curve design tools can affect the design process, are through their choice of specification language and fairness measure, as well as their practical ability to derive curves from them.
+We call a combination of specification language and fairness measure a description language.
+One may note that while the specification language has both syntactical as well as semantical influence on the language, the fairness measure determines semantics only.
 
 			Ideally, the specification language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
-The fairness should capture the intuitive notion of curve smoothness sufficiently well.
+The fairness measure should capture the intuitive notion of curve smoothness sufficiently well and select curves that are minimal with respect to the specification.
 These goals are competing with the software's ability to derive curves from descriptions reliably and efficiently, a process which becomes increasingly difficult with more expressive specification languages and more complex fairness measures.
@@
 			\begin{enumerate}
-				\item How efficient is the specification language at describing various curves?
-				\item How well can humans `read' the specification language from curves?
-				\item How well can humans `speak' the specification language to the software?
-				\item How well does the fairness measure capture the intuitive notion of smoothness?
-				\item How well can the software derive curves from specifications and a fairness measure?
+				\item expressiveness of the specification language
+				\item human ability to `read' the specification language from curves
+				\item human ability to `speak' the specification language to the software
+				\item smoothness of curves selected by the fairness measure
+				\item minimality of curves selected by the fairness measure
+				\item software's ability to derive curves from descriptions
 			\end{enumerate}
@@
 
-			Using the model of the curve design process as well as the criteria for usability of curve design software that were established in the previous section, various tools available in current vector graphics software can be analyzed.
+			Using the model of the curve design process as well as the criteria for usability of curve design software that were established in sections \ref{section:curve_design_process} and \ref{section:usability_criteria_curve_design_tools}, various tools available in current vector graphics software can be analyzed.
 
-		% TODO: quote raphael levien on usability of Bézier splines
 			\subsubsection{Bézier Splines}
@@
 				Bézier splines are probably the most widely used type of curve in two-dimensional vector graphics.
-Bézier splines are defined piecewise using Bézier curves, which are planar parametric curves (functions \(\function{\unit}{\R{2}}\)) realized using polynomials (usually cubic) in Bernstein form.
-A Bézier curve \(\phi\) of degree \(n\) consists of \(n + 1\) coefficients \(c_i \in \R{2}\).
+Bézier splines are defined piecewise using Bézier curves, which are parametric plane curves (see also sections \ref{section:parametric_curves} and \ref{section:parametric_plane_curves}) realized using polynomials in Bernstein form.
+Cubic Bézier curves are the most frequently used ones.
+A Bézier curve \(\phi\) of degree \(n\) consists of \(n + 1\) coefficients \(c_i \in \R{2}\), called coefficient points.
 The curve will always go through the first and the last coefficient point (\(\apply{\phi}{0} = c_0, \apply{\phi}{1} = c_n\)) and stay within the convex hull of the coefficient points at all other times.
-Bézier curves in a Bézier spline are usually pieced together such that the resulting spline exhibits continuous direction.
+Bézier curves in a Bézier spline are usually pieced together such that the resulting spline's tangent direction along the curve is continuous.
 
-				% TODO: explain velocity
 				In existing vector graphics software, Bézier splines are edited by more or less directly modifying the coefficient points of the underlying Bézier curves.
@@
 In that sense, the user can specify start end end points of the curve, as well as start and end velocities of the curve using the coefficient points.
-Most vector graphics tools have functions that allow preserving the direction continuity of the spline by automatically modifying adjacent coefficient points accordingly.
+Most vector graphics tools have functions that allow preserving the directional continuity of the spline by automatically modifying adjacent coefficient points accordingly.
 They usually do not ensure curvature continuity though.
-Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more Bézier curves to the spline.
+Some tools can also modify existing coefficient points in such a way that the curve passes through a point that is chosen by the user without adding more segments to the spline.
 
 				When applying the model introduced in section \ref{section:curve_design_process} to Bézier spline design tools, there are two possibilities for modelling the fairness measure.
-One way to view things is to assume that the user, in supplying the four coefficient points of each curve segment, uniquely specifies each segment and thus the entire curve, leaving no room for choosing a curve according to some fairness measure.
-Since users usually don't think about Bézier splines that way, it is more practical to assume that the user only supplies both point and velocity at the start and end of each curve segment, with a fairness measure designed to always choose the cubic polynomial that is uniquely specified by the given properties as the fairest curve.
+One way to view things is to assume that the user, in supplying the four coefficient points of each curve segment, uniquely specifies each segment's Bézier curve and thus the entire spline, leaving no room for choosing a curve according to some fairness measure.
+Since users usually don't think about Bézier splines that way, it is more practical to assume that the user only supplies both point and velocity at the start and end of each curve segment, with a fairness measure designed to always choose the cubic Bézier curve that is uniquely specified by the given properties as the fairest curve.
 
 				Applying the criteria established in section \ref{section:usability_criteria_curve_design_tools}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice.
-Bézier splines are furthermore very simple to derive from their specification, fulfilling the 5th criterion perfectly.
-Design tools for Bézier splines have some issues with both the fairness measure (criterion 4) and the expressiveness of the specification language (criterion 1), both of which shall be analyzed in the following paragraphs.
+The curve chosen by the fairness measure is also reasonably minimal (criterion 5), since the spline's complexity is limited by the number of segments that are used, which is in direct correspondence to the specification.
+Bézier splines are furthermore very simple to derive from their description, fulfilling criterion 6 perfectly.
+Design tools for Bézier splines have some issues with both the fairness measure's smoothness (criterion 4) and the expressiveness of the specification language (criterion 1), both of which shall be analyzed in the following paragraphs.
 
-				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing curvature continuity.
+				The issue with the fairness measure is that it always selects the uniquely determined cubic Bézier curve for each segment, thus not guaranteeing curvature continuity across segments.
 It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard.
@@
 Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment.
-There is also no way to directly specify linearly changing curvature.
+There is also no way to directly specify constant or linearly changing curvature.
 This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result.
@@
 
-				% TODO: consider moving away from the 'source image is traced' samples and use the actual vector graphics leading to those curves
-				% TODO: make sure the size of the node visualization is consistent, use visible colors for node visualization, maybe get rid of circles for handles
 				\begin{figure}[htb]
@@
 					\includegraphics[width=\textwidth]{../resources/usability_bezier.pdf}
-					% TODO: silent citation or not?
-					\caption{Usability Analysis of Bézier Splines (left: \cite{fluttershy}, right: \cite{lugia})}
+					\caption{Usability Analysis of Bézier Splines (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
 					\label{figure:usability_bézier}
@@
 
-				% TODO: improve example index numbers (also use font consistent with text)
 				Figure \ref{figure:usability_bézier} illustrates some of these issues by showing some examples of Bézier splines as they appear in common design tools.
-Bézier nodes are displayed as filled dots, while Bézier handles are displayed as circles.
+Bézier nodes are displayed as red dots, while Bézier handles are displayed as green lines.
 
@@
 Since there is no way to directly specify curvature using Bézier splines, a considerable number of nodes is needed to get close to the desired curve.
-It may be observed that the spiral in example 2 is not perfect.
-Creating perfect spirals using Bézier splines is very tedious and would require even more nodes (at least one for every 90\textdegree of winding angle covered).
-It could be speculated on whether the artist would have opted to use a perfect spiral in this example if they had had the tools to efficiently do so.
 
-				% TODO: consider removing the handle markers to make curvature discontinuity more visible
 				Example 4 shows a spline with a curvature discontinuity which went unnoticed by the artist at the time of creation.
-In fact, this discontinuity would be even less visible if the curve were not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.
+In fact, this discontinuity would be even less visible if the curve was not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.
 
@@
 				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_criteria_curve_design_tools}.
-The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
-The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4).
-The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.
+The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them are needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
+The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criteria 4 and 5).
+The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 6 well enough.
+The only real issue lies in the insufficient expressiveness of the specification language (criterion 1), discussed further in the following paragraph.
 
-				The only real issue lies in the insufficient expressiveness of the specification language.
 				Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability.
-Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify tangent angle nor curvature.
-These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of tangent angle and/or curvature.
+Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature.
+These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
 This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
 Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties.
-Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified.
+Unfortunately, this is rarely practical, since it both requires a lot of time to set up and doesn't behave very well when modified.
 These issues are somewhat alleviated by the guaranteed curvature continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
@@
 					\includegraphics[width=\textwidth]{../resources/usability_spiro.pdf}
-					\caption{Usability Analysis of Spiro Splines (left: \cite{fluttershy}, right: \cite{lugia})}
+					\caption{Usability Analysis of Spiro Splines (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
 					\label{figure:usability_spiro}
@@
 
-				% TODO: improve example index numbers
-				Figure \ref{figure:usability_spiro} illustrates some of these issues by showing some examples of Spiro splines as they appear in common design tools. Control points are displayed as filled dots.
+				Figure \ref{figure:usability_spiro} illustrates some of these issues by showing some examples of Spiro splines as they appear in common design tools. Control points are displayed as red dots.
 
@@
 Unfortunately, describing the curve using Spiro splines still requires many control points.
-In contrast to the corresponding example in section \ref{section:bézier_splines}, the curve in this example is close to a perfect spiral an thus does not match the source image perfectly.
-Additional control points would have to be added in order to approximate the imperfect spiral in the source image better.
+Note that even though this example has the same number of specified on-curve points (7) as the corresponding example in section \ref{section:bézier_splines}, the latter has an additional 12 velocity specifications, proving that even though curvature cannot be specified explicitly, Spiro splines are a much better choice for curves that are best expressed in terms of curvature than Bézier splines.
 
@@
 In this case, the shape consists of a circular arc with low curvature (segment a), a segment of linearly increasing curvature (segment b) and finally a circular arc with high curvature (segment c).
-This shape can be described conveniently using a Spiro spline with just 4 control points.
+This shape can be described conveniently using a Spiro spline with just 4 carefully placed control points.
 
@@
 For instance, increasing the curvature of segment c has to be done indirectly by moving a control point.
-Unfortunately, in this case, this also changes the tangent angle of the curve at that point, such that additional modifications would be needed to correct this.
-The change also adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
+Unfortunately, this adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
 
@@
 
-		Having analyzed the usability aspect of curve design tools in \ref{section:problem_analysis}, it's time to think about how to properly address their current issues.
+		Having analyzed the usability aspect of curve design tools in section \ref{section:problem_analysis}, it's time to think about how to properly address their current issues.
 In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as our specific ideas for developing a piece of software using this approach.

### 2013-09-04 idp/documentation/paper.tex
small fixes in description language section

@@
 
-			% TODO: talk about insensitivity of the human visual system to properties beyond the second derivative in both sections, build everything on that premise
+			The human visual system seems to be fairly insensitive to properties beyond the third derivative of the curve, that is, beyond variation of curvature.
+Based on that premise, we chose to let the user specify properties up to the second derivative of the curve using the specification language, while designing the fairness measure to choose curves whose third derivative is minimal from the set of curves that fulfill the specification's requirements.
 
@@
 
-				The specification language we propose uses differential geometric properties of curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use.
-The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will focus on the properties point, direction and curvature as defined in \ref{section:parametric_plane_curves}.
-In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties at any number of positions on the curve.
+				The specification language we propose uses differential geometric properties of parametric curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use.
+The specification language will focus on the properties point, direction and curvature as defined in \ref{section:parametric_plane_curves}.
+In order to provide maximal flexibility, the specification language should allow to specify any combination of these three properties at any number of positions on the curve.
 
-				With point specifications, it's sometimes enough to require the curve to go through all of them in order, although sometimes, this will lead to undesirable results.
+				When only concerned with point specifications, it's sometimes enough to require the curve to go through all of them in order, although sometimes, this will lead to undesirable results.
 Even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points.
@@
 				With that in mind, it is now possible to abandon this notion of a combination of point, direction and curvature and instead talk about specification items, a specification item being either a positioned point, a positioned direction or a positioned curvature.
-A combination of point and curvature at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
+For example, a combination of point and curvature at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
 
@@
 
-				% TODO: add citations
-				% TODO: make clear that these terms are to be minimized
-				% TODO: explain how the MVC fairness reduces unnecessary inflection points, etc.
-
-				Many different fairness measures for plane curves have been proposed throughout history.
-Most of them set out to emulate the behavior of mechanical splines made from wood or steel, which were used for industrial design before the advent of computer aided design.
-The most direct approach of this type uses mechanics to derive a term for the energy in an idealized bent steel band, arriving at the following functional describing the minimum energy curve (MEC):
+				Different fairness measures for plane curves have been proposed throughout history.
+Many of them set out to emulate the behavior of mechanical splines made from wood or steel, which were used for industrial design before the advent of computer aided design.
+The most direct approach of this type uses mechanics to derive a term for the energy in an idealized bent steel band to minimize, arriving at the following functional describing the minimum energy curve (MEC) \cite{paper-mec}:
 				\begin{equation*}
@@
 				While this approach is a good idealization of the mechanical spline, it has some disadvantages when used for curve design.
-For instance, it will not yield a circular arc when constrained by co-circular points.
+For instance, it will not yield a circular arc when the curve is constrained by co-circular points.
+Generally speaking, the MEC fairness measure is not very sensitive to bumps and inflection points on the curve, as long as they're small.
+The human visual system on the other hand, very much is and would prefer one large circular arc over five small bumps on the curve anytime.
 
-				Thus, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
-One such approach is that of the minimum variation curve (MVC), whose functional is given below:
+				In order to adress these issues, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
+One such approach is that of the minimum variation curve (MVC) \cite{thesis-mvc}, whose functional is given below:
 				\begin{equation*}
@@
 The functional will put straight lines through points that are co-linear and circular arcs through points that are co-circular when otherwise unconstrained.
-When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvture, since that will make the derivation of the curvature a constant and thereby minimize the integral over the square of the derivative of the curvature.
-Because of these appealing properties, the MVC fairness measure was chosen.
+When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvature, since that will make the derivative of the curvature a constant and thereby minimize the integral over its square.
+In general, the MVC fairness measure is very good at avoiding bumps and inflection points when possible.
+Because of these appealing properties, the MVC fairness measure was chosen as part of our description language.
 
@@
 
+		\bibitem{paper-mec}
+			\emph{The Curve of Least Energy}\\
+			B. K. P. Horn\\
+			Massachusetts Institute of Technology, January 1981
+
 		\bibitem{fluttershy}

### 2013-09-04 idp/documentation/paper.tex
more minor fixes in the proposed solution section

@@
 
-			The human visual system seems to be fairly insensitive to properties beyond the third derivative of the curve, that is, beyond variation of curvature.
-Based on that premise, we chose to let the user specify properties up to the second derivative of the curve using the specification language, while designing the fairness measure to choose curves whose third derivative is minimal from the set of curves that fulfill the specification's requirements.
+			The human visual system seems to be fairly insensitive to properties beyond the second derivative of the curve, that is, beyond curvature.
+Based on that premise, we chose to let the user specify properties up to the second derivative of the curve using the specification language.
+The fairness measure will then select the curve whose third derivative is minimal from the set of curves that fulfill the specification's requirements.
 
@@
 
-			Polynomial parametric curves were chosen since they are mathematically simple, very flexible through the choice of their degree, and are well-suited to approximate a variety of smooth curves.
+			Polynomial curves were chosen since they are mathematically simple, very flexible through the choice of their degree, and well-suited to approximate a variety of smooth curves.
 We opted to use polynomial splines since they allow approximating more complex curves without involving polynomials of excessively high degree, which are known to have poor stability.
@@
 
-				In the following, we will describe how nonlinear optimization can be used for building polynomial splines based on descriptions in the description language established in section \ref{section:description_language}.
+				Our implementation of choice uses smooth optimization, or, more specifically, an interior point method.
+This type of algorithm further requires the functions \(f\) and \(g\) to be at least once, ideally twice continuously differentiable.
+
+				In the following, we will describe how nonlinear optimization can be used for building polynomial splines based on descriptions in the language established in section \ref{section:description_language}.
 
@@
 Let \(m\) be the number of segments that make up the polynomial spline.
-The parametric curve functions are used with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline (\(\apply{\phi}{t}\)).
+We will use the functions defined in section \ref{section:parametric_plane_curves} with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline (\(\apply{\phi}{t}\)).
 Mapping the parameter \(t\) between the segments and the whole spline works as follows:
 				\begin{equation*}
-					\apply{\phi_i}{t} = \apply{\phi}{i + \frac{t}{m}}
+					\apply{\phi}{\frac{i + t}{m}} = \apply{\phi_i}{t}
 				\end{equation*}
@@
 Unfortunately, this function is not very well-suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
-Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
+Also, since the function isn't fixed and varies with the coefficients, there is no way of telling on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
 We also need to state the specification of the total arc length of the curve as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
 Furthermore, the difficulty of expressing the fairness measure, which is naturally an integral over the arc length of the curve, is dependent on the complexity of the covered arc length function.
-In order to address these problems, the following error term is used:
+In order to address all these problems, the following error term is used:
 				\begin{equation*}
@@
 				For polynomial splines with sufficient degree and segment count, this term will be brought close to zero in the optimization, making the speed approximately constant with value \(\hat{\lambda}\).
-This, in turn, makes the covered arc length function approximately linear (\(\apply{\lambda}{t} \approx \hat{\lambda}t\)) which causes the inverse of the covered arc length function \(\lambda^{-1}\) to be approximately linear as well, facilitating the simple statement of the optimization goals for the specification items as well as fixing the segment for each specification item.
+This, in turn, makes the covered arc length function approximately linear (\(\apply{\lambda}{t} \approx \hat{\lambda}t\)), facilitating the simple statement of the optimization goals for the specification items and the fairness measure as well as fixing the segment for each specification item.
 It also causes the total arc length of the curve to be approximately equal to \(\hat{\lambda}\), as requested by the specification.
@@
 				\begin{equation*}
-					\integral{0}{1}{\xp{\apply{\sigma}{t} - \hat{\lambda}}^2}{t}
+					\hat{\lambda}^{-2}\integral{0}{1}{\xp{\apply{\sigma}{t} - \hat{\lambda}}^2}{t}
 				\end{equation*}
-				This term is then finally used as part of the objective function, thus making it a minimization goal.
+				Since this term is intended to be used as a minimization goal, the factor \(\hat{\lambda}^{-2}\) was added, which makes it scale invariant.
+This enables the error term to be statically weighted against other terms in the objective function without this weighting being shifted because of curve scaling.
 
@@
 Thus, we will only require \(\contg{2}\) continuity for the whole spline.
-While the polynomial curve segments are smooth (\(\contp{\infty}\) continuity), the resulting spline in general is not.
+While the polynomial curves used as segments are smooth (\(\contp{\infty}\) continuity), the resulting spline in general is not.
 The segments have to be pieced together in such a way as to ensure a certain degree of smoothness.
@@
 				Having taken care of the somewhat technical details arising from using polynomial splines, it's time to state the error terms for the specification items.
-For each specification item (\(0 \leq i \leq n\)), we add one of the following error terms, depending on the type of the specification item:
+Since the optimization goal in section \ref{section:constant_speed} makes sure that \(\apply{\lambda}{t} \approx \hat{\lambda}t\), we know that the whole spline is parametrized in such a way that the parameter \(t\) is a fraction of the curve's total arc length.
+This fits perfectly with the fact that the position of each specification item is also specified as a fraction of the curve's total arc length.
+Thus, for each specification item (\(0 \leq i \leq n\)), we add one of the following error terms, depending on the type of the specification item:
 				\begin{equation*}
@@
 
-				Finally, the error term for the MVC fairness measure as defined in section \ref{section:fairness_measure} is used:
+				Since constant speed was established in section \ref{section:constant_speed}, stating the error term for the fairness chosen in section \ref{section:fairness_measure} is fairly straightforward:
 				\begin{equation*}
-					\integral{0}{1}{\apply{\chi'}{t}^2}{t}
+					\hat{\lambda}^3\integral{0}{1}{\apply{\chi'}{t}^2}{t}
 				\end{equation*}
-
 				It's clear that in many cases, this error term will not be zero, and in fact, may be arbitrarily large.
 Thus, it is not suitable to be constrained to zero or bounded error and has to be stated as a minimization goal.
+Note that similar to the error term for constant speed, we added the factor \(\hat{\lambda}^3\), which causes the fairness measure to be invariant under scaling of the whole curve, enabling static weighting of error terms in the objective function.
 
@@
 				Having gathered all the necessary error terms, it's time to put them together.
-Terms with zero-error constraints, as well as terms with upper bounds on the error are added as component functions of the constraint function \(g\) together with appropriate bounds in \(g_l\) and \(g_u\).
+Terms with zero-error constraints, as well as terms with upper bounds on the error are added as component functions of the constraint function \(g\) together with appropriate bound components in \(g_l\) and \(g_u\).
 
 				Minimization goals, on the other hand, are assembled in a weighted sum which defines the the objective function \(f\).
+Since the error terms were designed to be scale invariant, this weighting can be done statically.
+The exact values of the weights are to be determined empirically.
 

### 2013-09-04 source/Kurve/Kurve.Curves/Optimization/OptimizationProblem.cs
added scale invariance factors

@@ -39,19 +39,23 @@ namespace Kurve.Curves.Optimization
 			
 			ValueTerm targetSpeed = Terms.Quotient(curveLength, Terms.Constant(optimizationSegments.Segments.Count()));
 
-			ValueTerm speedError = Terms.Sum
+			ValueTerm speedError = Terms.Scaling
 			(
-				from segment in optimizationSegments.Segments
-				let position = Terms.Variable("t")
-				let curve = segment.LocalCurve
-				let speed = curve.Speed.Apply(position)
-				let error = Terms.Square(Terms.Difference(speed, targetSpeed))
-				select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
+				Terms.Exponentiation(targetSpeed, Terms.Constant(-2)),
+				Terms.Sum
+				(
+					from segment in optimizationSegments.Segments
+					let position = Terms.Variable("t")
+					let curve = segment.LocalCurve
+					let speed = curve.Speed.Apply(position)
+					let error = Terms.Square(Terms.Difference(speed, targetSpeed))
+					select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
+				)
 			);
 
 			ValueTerm fairnessError = Terms.Scaling
 			(
-				Terms.Exponentiation(targetSpeed, Terms.Constant(-6)),
+				Terms.Exponentiation(targetSpeed, Terms.Constant(-4)),
 				Terms.Sum
 				(
 					from segment in optimizationSegments.Segments
@@ -64,11 +68,10 @@ namespace Kurve.Curves.Optimization
 				)
 			);
 
-			// TODO: figure out how to do weighting that is independent of curve size
 			ValueTerm objectiveValue = Terms.Sum
 			(
-				Terms.Product(Terms.Constant(1.0), speedError),
-				Terms.Product(Terms.Constant(100000.0), fairnessError)
+				Terms.Product(Terms.Constant(100000.0), speedError),
+				Terms.Product(Terms.Constant(1), fairnessError)
 			)
 			.Simplify();
 

### 2013-09-04 idp/documentation/paper.tex
fixed scale invariance factor, typo

@@
 
-			\subsubsection{Sgementation and Notation}
+			\subsubsection{Segmentation and Notation}
 			\label{section:segmentation_notation}
@@
 				\begin{equation*}
-					\hat{\lambda}^3\integral{0}{1}{\apply{\chi'}{t}^2}{t}
+					\hat{\lambda}^2\integral{0}{1}{\apply{\chi'}{t}^2}{t}
 				\end{equation*}
@@
 Thus, it is not suitable to be constrained to zero or bounded error and has to be stated as a minimization goal.
-Note that similar to the error term for constant speed, we added the factor \(\hat{\lambda}^3\), which causes the fairness measure to be invariant under scaling of the whole curve, enabling static weighting of error terms in the objective function.
+Note that similar to the error term for constant speed, we added the factor \(\hat{\lambda}^2\), which causes the fairness measure to be invariant under scaling of the whole curve, enabling static weighting of error terms in the objective function.
 
@@
 
-			Using these definitions, the direction \(\apply{\delta}{t}\) is equal to the tangent angle at the point \(\apply{\phi}{t}\) and the curvature \(\apply{\chi}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{\phi}{t}\) and it is \(0\) if the curve is locally a straight line at the point \(\apply{\phi}{t}\).
+			Using these definitions, the direction \(\apply{\delta}{t}\) is equal to the tangent angle at the point \(\apply{\phi}{t}\).
+If the curve is locally a straight line at the point \(\apply{\phi}{t}\), the curvature \(\apply{\chi}{t}\) is equal to \(0\).
+Otherwise, the curvature \(\apply{\chi}{t}\) is equal (up to sign) to the inverse the radius of the osculating circle at the point \(\apply{\phi}{t}\).
 

### 2013-09-07 source/Kurve/Kurve.Curves/FunctionTermCurves/FunctionTermCurve.cs
simplified some constraint terms

@@ -14,7 +14,6 @@ namespace Kurve.Curves
 		readonly FunctionTerm acceleration;
 		readonly FunctionTerm jerk;
 		readonly FunctionTerm speed;
-		readonly FunctionTerm normalizedVelocity;
 		readonly FunctionTerm direction;
 		readonly FunctionTerm curvature;
 
@@ -23,7 +22,6 @@ namespace Kurve.Curves
 		public FunctionTerm Acceleration { get { return acceleration; } }
 		public FunctionTerm Jerk { get { return jerk; } }
 		public FunctionTerm Speed { get { return speed; } }
-		public FunctionTerm NormalizedVelocity { get { return normalizedVelocity; } }
 		public FunctionTerm Direction { get { return direction; } }
 		public FunctionTerm Curvature { get { return curvature; } }
 
@@ -33,29 +31,21 @@ namespace Kurve.Curves
 			if (function.DomainDimension != 1) throw new ArgumentException("parameter 'function' has wrong domain dimension.");
 			if (function.CodomainDimension != 2) throw new ArgumentException("parameter 'function' has wrong codomain dimension.");
 
-			ValueTerm position = Terms.Variable("t");
+			this.point = function;
+			this.velocity = point.GetDerivatives().Single();
+			this.acceleration = velocity.GetDerivatives().Single();
+			this.jerk = acceleration.GetDerivatives().Single();
 			
-			ValueTerm point = function.Apply(position);
-			ValueTerm velocity = function.GetDerivatives().Single().Apply(position);
-			ValueTerm acceleration = function.GetDerivatives().Single().GetDerivatives().Single().Apply(position);
-			ValueTerm jerk = function.GetDerivatives().Single().GetDerivatives().Single().GetDerivatives().Single().Apply(position);
+			ValueTerm position = Terms.Variable("t");
 
-			this.point = point.Abstract(position);
-			this.velocity = velocity.Abstract(position);
-			this.acceleration = acceleration.Abstract(position);
-			this.jerk = jerk.Abstract(position);
+			ValueTerm speedValue = Terms.Norm(velocity.Apply(position));
+			this.speed = speedValue.Abstract(position);
 
-			ValueTerm speed = Terms.Norm(velocity);
-			this.speed = speed.Abstract(position);
+			ValueTerm directionValue = Terms.Angle(velocity.Apply(position));
+			this.direction = directionValue.Abstract(position);
 
[…]

### 2013-09-07 source/Kurve/Kurve.Curves/Optimization/OptimizationProblem.cs

@@ -37,25 +37,25 @@ namespace Kurve.Curves.Optimization
 
 			IEnumerable<ValueTerm> variables = optimizationSegments.Parameters; 
 			
-			ValueTerm targetSpeed = Terms.Quotient(curveLength, Terms.Constant(optimizationSegments.Segments.Count()));
+			ValueTerm segmentLength = Terms.Quotient(curveLength, Terms.Constant(optimizationSegments.Segments.Count()));
 
 			ValueTerm speedError = Terms.Scaling
 			(
-				Terms.Exponentiation(targetSpeed, Terms.Constant(-2)),
+				Terms.Exponentiation(segmentLength, Terms.Constant(-2)),
 				Terms.Sum
 				(
 					from segment in optimizationSegments.Segments
 					let position = Terms.Variable("t")
 					let curve = segment.LocalCurve
 					let speed = curve.Speed.Apply(position)
-					let error = Terms.Square(Terms.Difference(speed, targetSpeed))
+					let error = Terms.Square(Terms.Difference(speed, segmentLength))
 					select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
 				)
 			);
 
 			ValueTerm fairnessError = Terms.Scaling
 			(
-				Terms.Exponentiation(targetSpeed, Terms.Constant(-4)),
+				Terms.Exponentiation(segmentLength, Terms.Constant(-4)),
 				Terms.Sum
 				(
 					from segment in optimizationSegments.Segments
@@ -84,13 +84,13 @@ namespace Kurve.Curves.Optimization
 			this.directionSpecificationTemplates =
 			(
 				from directionSpecificationIndex in Enumerable.Range(0, curveSpecifications.Count(curveSpecification => curveSpecification is DirectionCurveSpecification))
-				select SpecificationTemplate.CreateDirectionSpecificationTemplate(optimizationSegments.Segments, directionSpecificationIndex)
+				select SpecificationTemplate.CreateDirectionSpecificationTemplate(optimizationSegments.Segments, curveLength, directionSpecificationIndex)
 			)
 			.ToArray();
 			this.curvatureSpecificationTemplates =
 			(
 				from curvatureSpecificationIndex in Enumerable.Range(0, curveSpecifications.Count(curveSpecification => curveSpecification is CurvatureCurveSpecification))
[…]

### 2013-09-07 source/Kurve/Kurve.Curves/Optimization/OptimizationSubstitutions.cs

@@ -90,7 +90,7 @@ namespace Kurve.Curves.Optimization
 					DirectionCurveSpecification directionCurveSpecification = directionCurveSpecifications.ElementAt(directionSpecificationTemplateIndex);
 
 					ValueTerm position = Terms.Constant(directionCurveSpecification.Position);
-					ValueTerm value = Terms.Direction(Terms.Constant(directionCurveSpecification.Direction));
+					ValueTerm value = Terms.Constant(directionCurveSpecification.Direction);
 
 					foreach (Substitution substitution in directionSpecificationTemplate.GetSubstitutions(optimizationSegments.Segments, directionCurveSpecification.Position, position, value))
 						yield return substitution;

### 2013-09-07 source/Kurve/Kurve.Curves/Specification/BasicSpecification.cs

@@ -47,7 +47,7 @@ namespace Kurve.Curves
 			this.segmentTemplate = segmentTemplate;
 			this.curveSpecifications = curveSpecifications;
 		}
-		public BasicSpecification() : this(1, 1, new PolynomialFunctionTermCurveTemplate(1), Enumerables.Create<CurveSpecification>()) { }
+		public BasicSpecification() : this(1, 1, new PolynomialFunctionTermCurveTemplate(2), Enumerables.Create<CurveSpecification>()) { }
 		public BasicSpecification(XElement source)
 		{
 			if (source == null) throw new ArgumentNullException("source");

### 2013-09-07 source/Kurve/Kurve.Curves/Specification/SpecificationTemplate.cs

@@ -70,17 +70,21 @@ namespace Kurve.Curves
 					(
 						segments,
 						segmentWeights,
-						(segment, segmentWeight) => Terms.Scaling(segmentWeight, Terms.Difference(segment.GlobalCurve.Point.Apply(position), point))
+						(segment, segmentWeight) => Terms.Scaling
+						(
+							segmentWeight,
+							Terms.Difference(segment.GlobalCurve.Point.Apply(position), point)
+						)
 					)
 				)
 			);
 
 			return new SpecificationTemplate(position, point, segmentWeights, constraint);
 		}
-		public static SpecificationTemplate CreateDirectionSpecificationTemplate(IEnumerable<Segment> segments, int index)
+		public static SpecificationTemplate CreateDirectionSpecificationTemplate(IEnumerable<Segment> segments, ValueTerm speed, int index)
 		{
 			ValueTerm position = Terms.Variable(string.Format("d_{0}_position", index));
-			ValueTerm direction = Terms.Variable(string.Format("d_{0}_direction", index), 2);
+			ValueTerm direction = Terms.Variable(string.Format("d_{0}_direction", index), 1);
 			IEnumerable<ValueTerm> segmentWeights =
 			(
 				from segmentIndex in Enumerable.Range(0, segments.Count())
@@ -95,14 +99,26 @@ namespace Kurve.Curves
 					(
 						segments,
 						segmentWeights,
-						(segment, segmentWeight) => Terms.Scaling(segmentWeight, Terms.Difference(segment.GlobalCurve.NormalizedVelocity.Apply(position), direction))
+						(segment, segmentWeight) => Terms.Scaling
+						(
+							segmentWeight,
+							Terms.Difference
+							(
+								Terms.Scaling
+								(
+									Terms.Exponentiation(speed, Terms.Constant(-1)),
+									segment.GlobalCurve.Velocity.Apply(position)
+								),
+								Terms.Direction(direction)
+							)
+						)
 					)
 				)
 			);
 
 			return new SpecificationTemplate(position, direction, segmentWeights, constraint);
 		}
-		public static SpecificationTemplate CreateCurvatureSpecificationTemplate(IEnumerable<Segment> segments, int index)
+		public static SpecificationTemplate CreateCurvatureSpecificationTemplate(IEnumerable<Segment> segments, ValueTerm speed, int index)
 		{
[…]

### 2014-01-21 idp/documentation/paper.tex
fixed a few mistakes in the definitions section

@@
 				\overline{\cdot} & : \function{\R{2}}{\R{2}} & \overline{v}      & = \vectorB{-v_2}{v_1}      && \text{orthogonal vector}\\
-				\alpha           & : \function{\R{2}}{\RR} i & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1} && \text{polar coordinate angle}
+				\alpha           & : \function{\R{2}}{\RR}   & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1} && \text{polar coordinate angle}
 			\end{align*}
@@
 
-			Parametric curves are functions \(\phi : \function{\unit}{\R{n}}\), called the parametrization function.
-The represented curve is then the list of points that are traversed by this function.
+			Parametric curves are described using functions \(\phi : \function{\unit}{\R{n}}\), called the parametrization function.
+The represented curve is then the image of this function.
 While there are many ways to parametrize a given curve, that is, represent it using a parametrization function, we are usually interested in those which yield a sufficiently smooth parametrization function, allowing for the concise description of curve properties using differential geometry.
@@
 If the curve is locally a straight line at the point \(\apply{\phi}{t}\), the curvature \(\apply{\chi}{t}\) is equal to \(0\).
-Otherwise, the curvature \(\apply{\chi}{t}\) is equal (up to sign) to the inverse the radius of the osculating circle at the point \(\apply{\phi}{t}\).
+Otherwise, the curvature \(\apply{\chi}{t}\) is equal (up to sign) to the inverse of the radius of the osculating circle at the point \(\apply{\phi}{t}\).
 
@@
 			One may notice that this notion of continuity is a little too strict.
-A curve may be arbitrarily smooth, while its parametrization does not even have \(\contp{1}\) continuity, since the derivatives of the parametrization function may have discontinuities that do not manifest themselves in the curve that they represent.
+A curve may look arbitrarily smooth, while its parametrization does not even have \(\contp{1}\) continuity, since the derivatives of the parametrization function may have discontinuities that do not manifest themselves in the curve that they represent.
 If that is the case, one can always find a different parametrization of the same curve that does not have these discontinuities.

### 2014-01-23 idp/documentation/paper.tex
More documentation (implementation, UI)

--- 2013-09-04 idp/documentation/paper.tex
@@
 
+		\subsection{Architecture}
+
+			\subsubsection{Programming Language}
+
+				The first step of the implementation of the proposed solution was the choice of the platform and development environment that would be used. 
+				As the programming language, C\# was the best option due to the following features:
+
+				\begin{itemize}
+				  	\item Platform independence - the project is not bound to any specific platform and can be used on all major operating systems, provided the libraries Kurve depends on are available. 
+				  	\item C\# is a high-level language - his reduced overhead in the development process, for example spending time on manual memory management.
+					\item Ability to interface with native code - native code can be interfaced using reasonable amounts of boilerplate code.
+As it was clear that native libraries would be used, this is a key requirement.
+					\item Familiarity - the developers were familiar with this language. This choice eliminated overhead that would have been caused by adopting an unfamiliar language.
+					\item UI frameworks - part of the project is a proof of concept user interface to work with curves. Several UI frameworks are available that interface directly with C\#.
+				\end{itemize}
+
+			\subsubsection{Libraries}
+
+				Kurve offers its own APIs in C\# for working with terms and optimization problems.
+However, the implementations of the functionalities of these APIs are provided by third-party libraries.
+Hence, Kurve relies on several libraries, including:
+
+				\begin{itemize}
+				  	\item Ipopt\footnote{https://projects.coin-or.org/Ipopt} is a library that does numeric non-linear optimization.
+This library finds optima of optimization problems based on an objective function and its first two derivatives.
+In addition, it allows the specification of hard constraint functions (and their first two derivatives) that the solutions have to match.
+This allows Kurve to use hard specifications on curves.
+					\item CasADi\footnote{http://casadi.org} provides functionality to perform automatic differentiation of terms.
+It interfaces with Ipopt directly, which makes it a perfect match for Kurve's symbolic differentiation needs.
+CasADi is written in C++, but unfortunately only C functions can be called easily from C\#.
+For that reason, a small CasADi wrapper written in C was required to implement Kurve's C\# term APIs.
+				\end{itemize}
+
+				 Kurve has strong requirements for both libraries: Terms used in the functions that need to be optimized in Kurve are quite large and plentiful, but differentiation and optimization still need to be performed in a short period of time to allow real-time editing of curves.
+Fortunately, CasADi and Ipopt are very efficient and allow Kurve to offer close to real time editing of curves.
+
+		% choice of platform
+		%   C#: platform independent, high-level, possibility to interface with native code, familarity, good UI frameworks (GTK#)
+		%   ipopt
+		%   casadi
+		%   link to googlecode page
+		% numeric optimization
+		%   trapezoid sum to approximate integrals
+		%   formalization of model using symbolic functions in casadi
+		%   optimization using ipopt
+		%   caching
+		%   keeping track of disambiguation information
+		%   input is a set of specifications, disambiguation, etc.
+		%   result is a curve whose segmentation is no longer visible
+		% user interface (include screenshots)
+		%   goal: proof of concept
+		%   interacts with numeric optimizer
+		%   specify any combination of point, direction and curvature at positions on the curve
+		%   change position of a specification
+		%     in case of pointspecifications optionally moving the point along the curve accordingly
+		%   insert/remove curve length at or between specifications
+		%   displays deviation from set velocity, such that the user knows when a segment is over-stretched, or where the curve degree/segment count is too low
+		%   loading/saving of documents
+		%     specification can be serialized from/to XML
+				For Kurve's user interface, GTK\#\footnote{http://www.mono-project.com/GtkSharp} was used.
+GTK\# is used to draw the curves and controls into a window, as well as handling keyboard and mouse events.
+
+			\subsubsection{Subsystem Decomposition}
+
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/subsystems.pdf}
+					\caption{Subsystem Decomposition of the Kurve Project}
+					\label{figure:subsystem_decomposition}
+				\end{figure}
+
+				Kurve's implementation is split into two primary subsystems: Kurve contains the user interface related code and has a dependency on GTK\# as well as Kurve.Curves; Kurve.Curves is responsible for generating curves from specifications using CasADi through Wrappers.Casadi.
+
+				Wrappers.Casadi and Wrappers.Casadi.Native contain a wrapper for CasADi using C\# and C++ code respectively.
+Wrappers.Casadi uses the functions defined in Wrappers.Casadi.Native.
+Wrappers.Casadi and Wrappers.Casadi.Native are now part of the general-purpose library Krach to allow easy reuse in other projects.
+
+		\section{Numerical Optimization}
+			The Curve.Curves project contains the code required to transform input specifications into curves.
+It relies on CasADi (through Krach.Wrappers.Casadi) to represent curves symbolically and on Ipopt for optimization.
+
+			\subsection{Specification}
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/specification_class_diagram.pdf}
+					\caption{Class Diagram of Specification-Related Classes}
+					\label{figure:specification_class_diagram}
+				\end{figure}
+
+				To fully specify a curve, several different pieces of information are needed:
+
+				\begin{itemize}
+					\item Any number of point, direction and and curvature specifications. A set of such CurveSpecifications is stored in a BasicSpecification.
+					\item The curve's total length is also part of a BasicSpecification.
+					\item The curves are represented using a number of segments, each of which is represented by some function, realized using the FunctionTermCurve class.
+Because of that, the number of segments as well as the function type that is used to represent each segment are part of the BasicSpecification as well.
+					The type of function used is represented using a FunctionTermCurveTemplate object; this template allows the creation of FunctionTermCurve objects.  
+					However, this information is only used during optimization; if the number of segments and function template yield a curve with enough degrees of freedom in respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed.
+It is possible to specify the segment count and function template to modify the degrees of freedom that the resulting curves offers (ie.
+more degrees of freedom mean the curve can satisfy more curve specifications while maintaining constant speed and good fairness, but the optimization process takes longer).
+					\item Curves could be generated from a BasicSpecification alone.
+However, if there are multiple curves that fit a BasicSpecification very well (ie.
+the optimization problem has multiple local optima), different actual curves could be generated from a BasicSpecification.
+This results in a number of problems: for example, when saving a curve, if only the BasicSpecification were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process.
+This is clearly undesirable and a way to disambiguate between the different optima is required.
+					For that reason, a full Specification consists of a BasicSpecification as well as some disambiguation information.
+That disambiguation information is the result of the optimization, which can be used as a starting position for the optimization process when a curve is loaded.
+This both ensures that the same curve will be generated and speeds the optimization process up as well.
+				\end{itemize}
+
+			\subsection{Optimization}
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/optimization_class_diagram.pdf}
+					\caption{Class Diagram of Optimization-Related Classes}
+					\label{figure:optimization_class_diagram}
+				\end{figure}
+
+				The Optimizer class handles the optimization process of a curve.
+By supplying a Specification, a concrete curve as well as a new Specification with updated disambiguation information can be obtained.
+The resulting curve is represented by a Curve object which allows access to the curve's points, speed, direction and curvature, but carries no information about the internal segmentation and function type used.
+
+				The optimization process is perfomed in a number of steps. Depending on the change of the specification, only certain steps have to be run again as previous results can be reused.
+
+				\subsubsection{Segments}
+					The first step, OptimizationSegments, creates curve segments based on the number of segments and the segment template specified in the BasicSpecification.
+A list of parameters of all segments is created; by applying concrete values to each of the parameters, a concrete curve can be obtained.
+The generated segments are cached and do not have to be regenerated if the number of segments and segment template specified in the basic specification do not change.
+
+				\subsubsection{Optimization Problem}
+					Next, the optimization problem is created.
+The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_measure}.
+Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the specifications of the curve are realized using a constraint function.
+The concrete values required by the curve specifications are not used; instead, placeholder variables are introduced.
+For the total length of the curve a placeholder is used as well.
+This allows the reuse of the optimization problem if curve specifications or curve length change.
+Thus, the optimization problem is cached and only rebuilt if the optimiztion segments change or if the number of each type of constraint on the curve changes.
+If the value that a constraint requires changes, but the type of constraint remains the same (eg.
+a point constraint requires the curve to pass through a different point than before), the optimization problem does not have to be rebuilt.
+
+				\subsubsection{Substitutions}
+					In this step, substitutions for each of the placeholder variables (ie.
+for each curve specification and the curve length) are created.
+These substitutions, when applied to an optimization problem, remove all placeholder variables and yield the optimization problem that needs to be solved to obtain a concrete curve.
+The substitutions have to be rebuilt whenever a specification changes, the total curve length changes, or any of the previous steps had to be recomputed.
+
+				\subsubsection{Solver}
+					In this step, a solver for the optimization problem is created.
+The solver used is provided by Ipopt.
+A new solver needs to be created whenever either the optimization problem or the substitutions change.
+
+				\subsubsection{Position}
+					The position represents a result of an optimization process.
+To obtain this, the solver created in the previous step is run; as the starting position, the previous optimization's position is used.
+Using the obainted position, a concrete curve can be obtained from the optimization segments.
+This curve is the desired end result returned by the Optimizer.
+
+		\section{User Interface}
+			\subsection{Implementation}
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/ui_components_class_diagram.pdf}
+					\caption{Class Diagram of the User Interface Components}
+					\label{figure:ui_components_class_diagram}
+				\end{figure}
+
+				Kurve's user interface constructed from Components. Components may contain a number of subcomponents, render themselves into a GTK context, and handle mouse and keyboards events. 
+				\begin{itemize} 
+					\item MainWindow - The MainWindow is a GTK window that contains the RootComponent.
+It triggers redraws of the RootComponent when appropriate (eg.
+when the window is resized) and forwards mouse and keyboard events to the RootComponent.
+					\item Component - Component is the abstract base class of all components that make up Kurve's user interface.
+It provides default implementations for rendering components, handling mouse and keyboard events, and provides a mechanism to trigger redraws of components.
+					\item RootComponent - The RootComponent manages a list of CurveComponents and implements functionality to add and remove curve components.
+Additionaly, the RootComponent implements save and load functionality.
+					\item BackgroundComponent - This component can display a background image. This can be used to trace existing images with the Kurve tools.
+					\item CurveComponent - This component represents a curve in the user interface.
+It consists of any number of PositionedControlComponents (any number of SpecificationComponents and a FixedPositionComponent at the start and the end of the curve each).
+This yields segments of the curve - one segment between each pair of PositionedControlComponents.
+These segments are represented by SegmentComponents.
+A CurveComponent uses a CurveOptimizer to create a curve from the specifications given by the SpecificationComponents.
+It implmenets functionality to add and remove SpecificationComponents and to change the curve template used (eg.
+change the degree of the polynomial template).
+					\item LengthControlComponent - A LengthControlComponent is a Component that can modify the length of the Curve in a certain area.
+					\item PositionedControlComponent - A PositionedControlComponent is a component that has a specific position on the curve. 
+					\item SpecificationComponent - A SpecificationComponent allows to specify constraints on a curve.
+A point, direction and/or curvature can be specified for a given position using various mouse and keyboard commands.
+SpecificationComponents are visualized as a small black box that can be selected.
+					\item FixedPositionComponent - FixedPositionComponents represent the start and end of a CurveComponent.
+					\item SegmentComponent - A SegmentComponent represents a section of a curve delimited by two PositionedControlComponents.
+The SegmentComponent renders the curve in that segment, but it also allows to modify the length of the curve in that segment.
+					\item DiscreteCurve and DiscreteCurveItem - A DiscreteCurve represents the data necessary to render a curve in the user interface.
+It contains information about the curve's points, direction, and curvature.
+It is represented by a number of DiscreteCurveItems that sample a source curve's point, direction and curvature at a certain positions.
+					The CurveOptimizer yields this type of curve.
+The reason why this kind of discrete curves are used is because the Curve objects and the Optimizer (from the Kurve.Curves subsystem) cannot be used at the same time.
+This, in turn, is caused by the CasADi library, which is not thread safe.
+However, CasADi is used for representing Curves internally, as well as for performing optimization tasks.
+					As Kurve uses a background thread to optimize curves, it is not possible to use the Curve objects yieded by the optimizer to render the curves in the main thread at the same time.
+For that reason, Curve objects are converted to DiscreteCurve objects in that background process.
+					\item CurveOptimizer and OptimizationWorker - As the optimization process required to generate a curve from specifications takes too much time to perform in the main thread (the user interface would be blocked for too long), optimization is performed in a background thread.
+Each CurveComponent uses a CurveOptimizer to submit changed specifications.
+A single shared OptimizationWorker handles the synchronization and makes sure each submitted specification gets processed one by one.
+It is necessary to ensure that no optimization processes run in parallel, as the CasADi library (which is used in this process) is not thread safe.
+				\end{itemize}
+
+			\subsection{Usage}
+
+				The user interface does not focus on user friendliness.
+It's purpose is to test the validity of the approach to curve design researched in this project.
+For that reason, most actions are performed using keyboard shortcuts.
+Many user interface elments can be selected with the mouse; this enables to use keyboard shortcuts that affect that component only.
+It is possible to interact with some components using the mouse.
+However, there is no toolbar or similar user interface elements that would make any of the features more accessible.
+
+				The user interface allows the user to:
+				\begin{itemize}
+					\item add curves
+					\item add specification controls to curves (specifying point, direction and/or curvature at a certain position of the curve) 
+					\item control the curve's length (using specification and segment controls)
+					\item control the degree of the polynomial template and the number of internal segments of the curve
+					\item display a background image (eg. to trace curves in an existing image)
+					\item save and load curves
+				\end{itemize}
+
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/ui_components.pdf}
+					\caption{Overview of the User Interface in Kurve}
+					\label{figure:ui_components}
+				\end{figure}
+
+				\subsubsection{General Commands}
+
+					A curve can be added by pressing `N'.
+A new Curve will appear in window that is initialized with a length of 100, and two specification controls specifying points.
+The curve uses one segment that uses a polynomial curve template with a degree of ten.
+
+					To save all curves in the window, press `S'.
+A save file dialogue will appear that allows you to specify a filename.
+Saved files can later be loaded by pressing `L' which displays an open file dialogue.
+Keep in mind that loading a file dismisses all curves in the current window.
+
+					A background image can be displayed by passing a path to an image file when launching the Kurve application.
+
+				\subsubsection{Curves}
+
+					Click the right mouse button anywhere on a curve to add a specification control at that position.
+The newly added specification control does not yet specify anything.
+To remove a specification control, select it and press `R'.
+See the section \ref{section:specification_control} on how to use specification controls.
+
+					The length of the curve can be modified by selecting specification controls and segment controls and using the mouse wheel while pressing shift.
+If a specification control is selected, length is inserted to the left and right of the specification control; if a segment control is selected, that segment's length is changed.
+
+					The current version of Kurve uses polynoms as the curve template. The template can be modified as follows:
+
+					\begin{itemize}
+						\item To change the number of segments used, press `1' to decrease and `2' to increase the number of segments by one.
+(Keep in mind that these segments are not visible to the user and are used in the optimization process only.
+Adding more segments allows more complicated curves, but has negative effects on the performance.)
+						\item To change the degree of the polynomial used in each segment, press alt + `1' to decrease, and alt + `2' to increase the degree by one.
+					\end{itemize}
+
+				\subsubsection{Specification Control}
+				\label{section:specification_control}
+
+					A specification control on a curve can specify any combination of a point, direction, or curvature at a certain position on the curve.
+Any combination of the three following letters represent which geometric attributes are specified by the specification control.
+
+					\begin{itemize}
+						\item `P' indicates that a point that the curve has to pass through at that position is specified.
+						\item `D' indicates that the direction of the curve at that position is specified.
+						\item `C' indicates that the curvature of the curve at that position is specified.
+					\end{itemize}
+
+					Which attributes are specified can be toggled by selecting a specification component by clicking it and pressing the corresponding key on the keyboard.
+Which attributes are currently being specified is displayed in a label below the specification control.
+For example, if the label is empty, nothing is specified, while `PC' indicates that a point and a curvature are specified.
+
+					When a geometric attribute is specified by a specification control, it can be modified:
+
+					\begin{itemize}
+						\item Point: Thp point can be modified by dragging the specification control with the mouse.
+						\item Direction: The direction can be modified using the mouse wheel.
+						\item Curvature: The curvature can be modified by pressing the windows key while using the mouse wheel.
+					\end{itemize}
+
+					The position of the control can be modified by pressing the control key while using the mouse wheel.
+Keep in mind that when changing the position of a specification control that does not specify a point, the control itself will most likely move (and change it's point).
+
+					To achieve more fine-grained control, it is possible to press the alt key while performing any of the actions described above.
+This changes the affected property of the specification control more slowly.
+For example, if you press alt while dragging a specification control, it will move very slowly.
+
+					As an experimental feature, it is also possible to change the position of a specification control by dragging it with the mouse while pressing shift.
+The position is calculated by the current state of the curve.
+
+				\subsubsection{Segment Control}
+
+					A segment control is a part of a curve delimited by either two specification controls or the start/end of the curve. It can be selected by clicking it.
+
+					If at least one of the delimiting controls specifies a point, segments can be moved by dragging them with the mouse.
+
+					A segment also displays the speed of the curve if it differs from the optimal speed.
+Too high speed is indicated by a red hue; this can usually be fixed by inserting additional length.
+Too low speed is indicated by a blue hue; usually there will be both blue and red sections.
+this usually indicates that the curve is too complex (ie.
+there are too many specifications) to acchieve a good representation with the current curve template.
+This can be fixed by either removing specifications, adding internal segments to the curve, or increasing the degree of the polynomial function template used in the curve.
+
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=\textwidth]{../resources/ui_speeds.pdf}
+					\caption{Example of a Curve that is `Overspecified' (ie. the speed of the curve cannot be kept constant due to the high amount of constraints)}
+					\label{figure:ui_speeds}
+				\end{figure}
+
+				\subsubsection{Command Cheat Sheet}
+					\begin{itemize}
+						\item General
+						\begin{description}
+							\item[`N'] Add a new curve
+							\item[`L'/`S'] Load/Save
+							\item[`1'/`2'] decrease/increase internal segment count
+							\item[alt + `1'/`2'] decrease/increase polynomial template degree
+						\end{description}
+						\item All Controls
+						\begin{description}
+							\item[Click] Selects a control 
+							\item[Shift + Click] Select multiple controls
+							\item[Shift + Scroll (+ alt)] Insert length at component (slowly)
+						\end{description}
+						\item Specification Control
+						\begin{description}
+							\item[Ctrl + Scroll (+ alt)] Change the position of the specification (slowly)
+							\item[Shift + Drag] Change position of specification (Adjusts the point of the specification)
+							\item[`R'] Removes the specification
+							\item[`P'] Triggers point specification
+							\item[Drag (+ alt)] Move point (slowly). Only possible when point specification is enabled.
+							\item[`D'] Triggers direction specification
+							\item[Scroll (+ alt)] Changes the direction (slowly). Only possible when direction specification is enabled.
+							\item[`C'] Triggers curvature specification
+							\item[Win + Scroll (+ alt)] Changes the curvature (slowly). Only possible when curvature specification is enabled.
+							\item[Shift + Click] Select multiple controls
+							\item[Shift + Scroll (+ alt)] Insert length at component (slowly)
+						\end{description}
+						\item Segment Control
+						\begin{description}
+							\item[Right-Click] Add empty specifcation control at that position
+							\item[Drag] Moves any point specifications delimiting the segment. (Only possible if at leats one of the specification controls has the point specification enabled.)
+						\end{description}
+					\end{itemize}
+
 		% choice of platform

### 2014-01-24 idp/documentation/paper.tex
Fixed some errors in documentation.

@@
 
-				The first step of the implementation of the proposed solution was the choice of the platform and development environment that would be used. 
-				As the programming language, C\# was the best option due to the following features:
+				The first step of the implementation of the proposed solution is the choice of the platform and development environment that will be used. 
+				As the programming language, C\# is the best option due to the following features:
 
@@
 				  	\item Platform independence - the project is not bound to any specific platform and can be used on all major operating systems, provided the libraries Kurve depends on are available. 
-				  	\item C\# is a high-level language - his reduced overhead in the development process, for example spending time on manual memory management.
-					\item Ability to interface with native code - native code can be interfaced using reasonable amounts of boilerplate code.
+				  	\item C\# is a high-level language - this reduces overhead in the development process, for example spending time on manual memory management.
+					\item Ability to interface with native code - native code can be interfaced with by using reasonable amounts of boilerplate code.
 As it was clear that native libraries would be used, this is a key requirement.
-					\item Familiarity - the developers were familiar with this language. This choice eliminated overhead that would have been caused by adopting an unfamiliar language.
+					\item Familiarity - the developers are familiar with this language. This choice eliminates overhead that would be caused by adopting an unfamiliar language.
 					\item UI frameworks - part of the project is a proof of concept user interface to work with curves. Several UI frameworks are available that interface directly with C\#.
@@
 				\begin{itemize}
-					\item Any number of point, direction and and curvature specifications. A set of such CurveSpecifications is stored in a BasicSpecification.
+					\item Any number of point, direction and curvature specifications. A set of such CurveSpecifications is stored in a BasicSpecification.
 					\item The curve's total length is also part of a BasicSpecification.
@@
 					The type of function used is represented using a FunctionTermCurveTemplate object; this template allows the creation of FunctionTermCurve objects.  
-					However, this information is only used during optimization; if the number of segments and function template yield a curve with enough degrees of freedom in respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed.
-It is possible to specify the segment count and function template to modify the degrees of freedom that the resulting curves offers (ie.
+					However, this information is only used during optimization; if the number of segments and function template yield a curve with enough degrees of freedom with respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed.
+It is possible to specify the segment count and function template to modify the degrees of freedom that the resulting curves offers (i.e.
 more degrees of freedom mean the curve can satisfy more curve specifications while maintaining constant speed and good fairness, but the optimization process takes longer).
 					\item Curves could be generated from a BasicSpecification alone.
-However, if there are multiple curves that fit a BasicSpecification very well (ie.
+However, if there are multiple curves that fit a BasicSpecification very well (i.e.
 the optimization problem has multiple local optima), different actual curves could be generated from a BasicSpecification.
@@
 That disambiguation information is the result of the optimization, which can be used as a starting position for the optimization process when a curve is loaded.
-This both ensures that the same curve will be generated and speeds the optimization process up as well.
+This both ensures that the same curve will be generated and speeds up the optimization process as well.
 				\end{itemize}
@@
 Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the specifications of the curve are realized using a constraint function.
+This step also computes the first and second derivatives of those functions.
 The concrete values required by the curve specifications are not used; instead, placeholder variables are introduced.
 For the total length of the curve a placeholder is used as well.
-This allows the reuse of the optimization problem if curve specifications or curve length change.
+This allows the reuse of the optimization problem if curve specifications or curve length change; the derivatives do not have to be recomputed.
 Thus, the optimization problem is cached and only rebuilt if the optimiztion segments change or if the number of each type of constraint on the curve changes.
-If the value that a constraint requires changes, but the type of constraint remains the same (eg.
+If the value of a constraint changes, but the type of constraint remains the same (e.g.
 a point constraint requires the curve to pass through a different point than before), the optimization problem does not have to be rebuilt.
@@
 				\subsubsection{Substitutions}
-					In this step, substitutions for each of the placeholder variables (ie.
+					In this step, substitutions for each of the placeholder variables (i.e.
 for each curve specification and the curve length) are created.
@@
 To obtain this, the solver created in the previous step is run; as the starting position, the previous optimization's position is used.
-Using the obainted position, a concrete curve can be obtained from the optimization segments.
+Using the obtained position, a concrete curve can be obtained from the optimization segments.
 This curve is the desired end result returned by the Optimizer.
@@
 
-				Kurve's user interface constructed from Components. Components may contain a number of subcomponents, render themselves into a GTK context, and handle mouse and keyboards events. 
+				Kurve's user interface constructed from Components. Components may contain a number of subcomponents, render themselves into a GTK\# context, and handle mouse and keyboards events. 
 				\begin{itemize} 
-					\item MainWindow - The MainWindow is a GTK window that contains the RootComponent.
-It triggers redraws of the RootComponent when appropriate (eg.
+					\item MainWindow - The MainWindow is a GTK\# window that contains the RootComponent.
+It triggers redraws of the RootComponent when appropriate (e.g.
 when the window is resized) and forwards mouse and keyboard events to the RootComponent.
 					\item Component - Component is the abstract base class of all components that make up Kurve's user interface.
-It provides default implementations for rendering components, handling mouse and keyboard events, and provides a mechanism to trigger redraws of components.
+It provides default implementations for rendering components, handling mouse and keyboard events, and provides a mechanism to trigger redraws of subcomponents.
 					\item RootComponent - The RootComponent manages a list of CurveComponents and implements functionality to add and remove curve components.
-Additionaly, the RootComponent implements save and load functionality.
+Additionally, the RootComponent implements save and load functionality.
 					\item BackgroundComponent - This component can display a background image. This can be used to trace existing images with the Kurve tools.
 					\item CurveComponent - This component represents a curve in the user interface.
-It consists of any number of PositionedControlComponents (any number of SpecificationComponents and a FixedPositionComponent at the start and the end of the curve each).
+It consists of any number of PositionedControlComponents (any number of SpecificationComponents and a FixedPositionComponent at both the start and the end of the curve).
 This yields segments of the curve - one segment between each pair of PositionedControlComponents.
@@
 A CurveComponent uses a CurveOptimizer to create a curve from the specifications given by the SpecificationComponents.
-It implmenets functionality to add and remove SpecificationComponents and to change the curve template used (eg.
+It implements functionality to add and remove SpecificationComponents and to change the curve template used (e.g.
 change the degree of the polynomial template).
@@
 It contains information about the curve's points, direction, and curvature.
-It is represented by a number of DiscreteCurveItems that sample a source curve's point, direction and curvature at a certain positions.
+It is represented by a number of DiscreteCurveItems that sample a source curve's point, direction and curvature at certain positions.
 					The CurveOptimizer yields this type of curve.
-The reason why this kind of discrete curves are used is because the Curve objects and the Optimizer (from the Kurve.Curves subsystem) cannot be used at the same time.
+The reason why this kind of discrete curve is used is because the Curve objects and the Optimizer (from the Kurve.Curves subsystem) cannot be used at the same time.
 This, in turn, is caused by the CasADi library, which is not thread safe.
@@
 				The user interface does not focus on user friendliness.
-It's purpose is to test the validity of the approach to curve design researched in this project.
+Its purpose is to test the validity of the approach to curve design researched in this project.
 For that reason, most actions are performed using keyboard shortcuts.
@@
 					\item add curves
-					\item add specification controls to curves (specifying point, direction and/or curvature at a certain position of the curve) 
+					\item add specification controls to curves (specifying point, direction and/or curvature at a certain position on the curve) 
 					\item control the curve's length (using specification and segment controls)
 					\item control the degree of the polynomial template and the number of internal segments of the curve
-					\item display a background image (eg. to trace curves in an existing image)
+					\item display a background image (e.g. to trace curves in an existing image)
 					\item save and load curves
@@
 A new Curve will appear in window that is initialized with a length of 100, and two specification controls specifying points.
-The curve uses one segment that uses a polynomial curve template with a degree of ten.
+The curve uses one segment that uses a polynomial curve template with a degree of 10.
 
 					To save all curves in the window, press `S'.
-A save file dialogue will appear that allows you to specify a filename.
-Saved files can later be loaded by pressing `L' which displays an open file dialogue.
+A save file dialog will appear that allows the user to specify a filename.
+Saved files can later be loaded by pressing `L' which displays an open file dialog.
 Keep in mind that loading a file dismisses all curves in the current window.
@@
 
-					The current version of Kurve uses polynoms as the curve template. The template can be modified as follows:
+					The current version of Kurve uses polynomials functions in each segment. The template that generates the functions and the number of segments can be modified as follows:
 
 					\begin{itemize}
-						\item To change the number of segments used, press `1' to decrease and `2' to increase the number of segments by one.
-(Keep in mind that these segments are not visible to the user and are used in the optimization process only.
-Adding more segments allows more complicated curves, but has negative effects on the performance.)
+						\item To change the number of segments used, press `1' to decrease and `2' to increase the number of segments by one. ()
 						\item To change the degree of the polynomial used in each segment, press alt + `1' to decrease, and alt + `2' to increase the degree by one.
 					\end{itemize}
-
+					Keep in mind that the segment count and the degree of the polynomial are not directly visible to the user and are used in the optimization process only.
+Adding more segments or increasing the degree of the polynomial allows more complicated curves, but has negative effects on the performance.
 				\subsubsection{Specification Control}
@@
 Which attributes are currently being specified is displayed in a label below the specification control.
-For example, if the label is empty, nothing is specified, while `PC' indicates that a point and a curvature are specified.
+For example, if the label shows `n/a', nothing is specified, while `PC' indicates that a point and a curvature are specified.
 
@@
 					\begin{itemize}
-						\item Point: Thp point can be modified by dragging the specification control with the mouse.
+						\item Point: The point can be modified by dragging the specification control with the mouse.
 						\item Direction: The direction can be modified using the mouse wheel.
@@
 					The position of the control can be modified by pressing the control key while using the mouse wheel.
-Keep in mind that when changing the position of a specification control that does not specify a point, the control itself will most likely move (and change it's point).
+Keep in mind that when changing the position of a specification control that does not specify a point, the control itself will most likely move (and change its point).
 
@@
 This changes the affected property of the specification control more slowly.
-For example, if you press alt while dragging a specification control, it will move very slowly.
+For exa^mple, if you press alt while dragging a specification control, it will move very slowly.
 
@@
 Too low speed is indicated by a blue hue; usually there will be both blue and red sections.
-this usually indicates that the curve is too complex (ie.
-there are too many specifications) to acchieve a good representation with the current curve template.
+This usually indicates that the curve is too complex (i.e.
+there are too many specifications) to achieve a good representation with the current curve template.
 This can be fixed by either removing specifications, adding internal segments to the curve, or increasing the degree of the polynomial function template used in the curve.
@@
 					\includegraphics[width=\textwidth]{../resources/ui_speeds.pdf}
-					\caption{Example of a Curve that is `Overspecified' (ie. the speed of the curve cannot be kept constant due to the high amount of constraints)}
+					\caption{Example of a Curve that is `Overspecified' (i.e. the speed of the curve cannot be kept constant due to the high amount of constraints)}
 					\label{figure:ui_speeds}
@@
 							\item[`1'/`2'] decrease/increase internal segment count
-							\item[alt + `1'/`2'] decrease/increase polynomial template degree
+							\item[Alt + `1'/`2'] decrease/increase polynomial template degree
 						\end{description}
@@
 							\item[Shift + Click] Select multiple controls
-							\item[Shift + Scroll (+ alt)] Insert length at component (slowly)
+							\item[Shift + Scroll (+ Alt)] Insert length at component (slowly)
 						\end{description}
@@
 						\begin{description}
-							\item[Ctrl + Scroll (+ alt)] Change the position of the specification (slowly)
-							\item[Shift + Drag] Change position of specification (Adjusts the point of the specification)
+							\item[Ctrl + Scroll (+ Alt)] Change the position of the specification (slowly)
+							\item[Shift + Drag] Change position of specification (adjusts the point of the specification)
 							\item[`R'] Removes the specification
 							\item[`P'] Triggers point specification
-							\item[Drag (+ alt)] Move point (slowly). Only possible when point specification is enabled.
+							\item[Drag (+ Alt)] Move point (slowly). Only possible when point specification is enabled.
 							\item[`D'] Triggers direction specification
-							\item[Scroll (+ alt)] Changes the direction (slowly). Only possible when direction specification is enabled.
+							\item[Scroll (+ Alt)] Changes the direction (slowly). Only possible when direction specification is enabled.
 							\item[`C'] Triggers curvature specification
-							\item[Win + Scroll (+ alt)] Changes the curvature (slowly). Only possible when curvature specification is enabled.
+							\item[Win + Scroll (+ Alt)] Changes the curvature (slowly). Only possible when curvature specification is enabled.
 							\item[Shift + Click] Select multiple controls
-							\item[Shift + Scroll (+ alt)] Insert length at component (slowly)
+							\item[Shift + Scroll (+ Alt)] Insert length at component (slowly)
 						\end{description}

### 2014-01-24 idp/documentation/paper.tex
Fixed typo.
started working on evaluation section

@@
 		While there has been some promising research into improving curve design tools \cite{thesis-mvc} \cite{thesis-spiro}, it focuses mostly on the aspect of improving the fairness of curves, rather than usability as a whole.
-Unfortunately, it has not had much of an impact on actual vector graphics software used in practice, which continues to focus almost exclusively on Bézier splines to this day.
+Unfortunately, it also has not had much of an impact on actual vector graphics software used in practice, which continues to focus almost exclusively on Bézier splines to this day.
 
@@
 
+				% TODO: explain this in more detail (o--o----------------o, pull last specification, makes wobbly segment at beginning) 
 				With this setup, it seems advisable to also include the curve's total arc length in the specification.
@@
 Generally speaking, the MEC fairness measure is not very sensitive to bumps and inflection points on the curve, as long as they're small.
-The human visual system on the other hand, very much is and would prefer one large circular arc over five small bumps on the curve anytime.
+The human visual system on the other hand, very much is and would always prefer one large circular arc over five small bumps on the curve.
 
@@
 
+			% TODO: make it clear that splines imply segmentation
 			Polynomial curves were chosen since they are mathematically simple, very flexible through the choice of their degree, and well-suited to approximate a variety of smooth curves.
@@
 				\end{equation*}
-				This allows expression of a variety of optimization goals.
+				This allows for the expression of a variety of optimization goals.
 The ones that are important for us are zero-error constraints (using the constraint function and bounds to specify that an error term should be zero), upper bounds on errors (using the constraint function and bounds to specify an upper bound for an error term) and minimization goals (using the objective function to express that an error term should be minimized).
@@
 We also need to state the specification of the total arc length of the curve as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
-Furthermore, the difficulty of expressing the fairness measure, which is naturally an integral over the arc length of the curve, is dependent on the complexity of the covered arc length function.
+Furthermore, the difficulty of expressing the fairness measure, which is often an integral over the arc length of the curve, is dependent on the complexity of the covered arc length function.
 In order to address all these problems, the following error term is used:
@@
 
-		% choice of platform
-		%   C#: platform independent, high-level, possibility to interface with native code, familarity, good UI frameworks (GTK#)
-		%   ipopt
-		%   casadi
-		%   link to googlecode page
-		% numeric optimization
-		%   trapezoid sum to approximate integrals
-		%   formalization of model using symbolic functions in casadi
-		%   optimization using ipopt
-		%   caching
-		%   keeping track of disambiguation information
-		%   input is a set of specifications, disambiguation, etc.
-		%   result is a curve whose segmentation is no longer visible
-		% user interface (include screenshots)
-		%   goal: proof of concept
-		%   interacts with numeric optimizer
-		%   specify any combination of point, direction and curvature at positions on the curve
-		%   change position of a specification
-		%     in case of pointspecifications optionally moving the point along the curve accordingly
-		%   insert/remove curve length at or between specifications
-		%   displays deviation from set velocity, such that the user knows when a segment is over-stretched, or where the curve degree/segment count is too low
-		%   loading/saving of documents
-		%     specification can be serialized from/to XML
 				For Kurve's user interface, GTK\#\footnote{http://www.mono-project.com/GtkSharp} was used.
@@
 This changes the affected property of the specification control more slowly.
-For exa^mple, if you press alt while dragging a specification control, it will move very slowly.
+For example, if you press alt while dragging a specification control, it will move very slowly.
 
@@
 
-		% choice of platform
-		%   C#: platform independent, high-level, possibility to interface with native code, familarity, good UI frameworks (GTK#)
-		%   ipopt
-		%   casadi
-		%   link to googlecode page
-		% numeric optimization
-		%   trapezoid sum to approximate integrals
-		%   formalization of model using symbolic functions in casadi
-		%   optimization using ipopt
-		%   caching
-		%   keeping track of disambiguation information
-		%   input is a set of specifications, disambiguation, etc.
-		%   result is a curve whose segmentation is no longer visible
-		% user interface (include screenshots)
-		%   goal: proof of concept
-		%   interacts with numeric optimizer
-		%   specify any combination of point, direction and curvature at positions on the curve
-		%   change position of a specification
-		%     in case of pointspecifications optionally moving the point along the curve accordingly
-		%   insert/remove curve length at or between specifications
-		%   displays deviation from set velocity, such that the user knows when a segment is over-stretched, or where the curve degree/segment count is too low
-		%   loading/saving of documents
-		%     specification can be serialized from/to XML
+	\section{Evaluation}
+	\label{section:evaluation}
 
-	\section{Evaluation}
+		The goal of this project was to explore possible improvements concerning the usability of curve design tools. The following sections discuss the results in this area.
 
-		% were the goals met? (TODO: systematic approach for describing expressiveness and new features)
+		\subsection{Usability Analysis}
+		\label{section:usability_analysis}
+
+			Having applied the usability criteria from section \ref{section:usability_criteria_curve_design_tools} to Bézier splines and Spiro splines, we will now use them to evaluate usability aspects of our approach.
+
+			\subsubsection{Expressiveness of the Specification Language}
+
+				One of the primary goals of this project was to explore more expressive specification languages than those which are currently used in vector graphics software.
+To that end, our specification language supports specifying points, directions and curvatures at any number of positions on the curve, as well as the total length of the curve.
+This grants the user much flexibility in describing the source curve, allowing them to specify those properties that describe the curve in the most concise fashion.
+
+				The ability to specify curvature constitues a significant improvement over common curve design tools, where this is usually not possible.
+Using this functionality, the user can specify curves which are best expressed in terms of curvature in a natural and concise way.
+Not only does this make the curve design process less tedious, but it may also lead to better curves, since only the essential information about the curve is given to the software.
+This allows the fairness measure to choose the best possible curve, something that is not possible if lots of auxiliary information is supplied to the software in order to indirectly influence the curvature.
+
+				Section \ref{section:specification_language} describes why it makes sense to specify the curve's total arc length.
+While fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, there are actually quite a few upsides.
+For instance, by adjusting the arc length of a curve with a point specifications on each end, one can conveniently produce arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a smooth arc is produced.
+Using common vector graphics software, one usually has to either risk an asymmetric and/or bumpy arc by specifying it using Bézier splines, or deal with some cumbersome tool for constructing elliptical arcs.
+
+				% expressiveness of the specification language
+				%   much more expressive than Bézier and Spiro splines, allows specification of curvature, length
 				%     controlling length is actually a very nice thing to have
 				%       a good way to control circular arcs in terms of radius (fluttershy's ear, mouth)
+				%         compare to using elliptical arcs in inkscape
 				%       a good way to pull curves in place when they don't have direction specifications
+				%       benign behavior
+				%   doesn't allow *not* expressing length, may pose a problem
+				%   demonstrate euler spirals and bézier curves being embedded
+				%     demonstrates expressiveness
+				%     minimization of normal jerk squares prefers linear change in curvature
+				%   a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification
+
+			% human ability to `read' the specification language from curves
+			%   familiar differential geometric properties
+			% human ability to `speak' the specification language to the software
+			%   familiar differential geometric properties
+			% smoothness of curves selected by the fairness measure
+			%   very good, thanks to MVC
+			% minimality of curves selected by the fairness measure
+			%   not guaranteed by optimization, but may usually be obtained by pushing the curve into a different optimum
+			% software's ability to derive curves from descriptions
+			%   optimization is slow and sometimes unstable, still needs some work before it becomes practical
+			%   local optima
+			%   bad stability, infeasible for numeric optimization
+			%   performance
+			%   adjustment of segment count and degree by the user is not desirable
+
+		\subsection{Examples}
+		\label{section:examples}
+
 			% fluttershy's mane
@@
 			%     curvature incontinuities, small bumps
+			%     a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification
 			%   somewhat hampered by poor performance and optimization stability issues, concept seems very powerful though
+			% lugia profile view / lugia's back line (curled up)
+			%   profits from global fairness optimization and curvature continuity guarantees
 			%   some curves are more easily described in terms of curvature than points and/or direction
-		% problems
-		%   local optima
-		%   bad stability, infeasible for numeric optimization
-		%   performance
-		%     possible improvements
-		%       make sure substitutions only happen where and when necessary
-		%       cache partially instantiated IpoptProblem? such that fewer substitutions have to happen when changing a single specification
-		%       use better linear solver with ipopt
-		%       properly configure ipopt options
-		%       adaptive segment density
-		%       adaptive velocity/fairness trapezoid sum
+			% example of stability issues
 
-		% demonstrate euler spirals and bézier curves being embedded
-		%   minimization of normal jerk squares prefers linear change in curvature
-		% take examples from demo
-		% revisit examples made when analyzing Bézier splines and Spiro splines
-		% a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification
+		\subsection{Verdict}
+		\label{section:verdict}
 
 	\section{Conclusion}
+	\label{section:conclusion}
 
+		% most issues stem from the optimization approach
 		% distinguish between further research work and turning the project into practical software)	
@@
 		% investigate into the calculus of variations and whether there may be a possibility to construct MVC curves based on our specification format, once the format itsself has proven to be useful
+		% performance
+
+		%   possible improvements
+		%     make sure substitutions only happen where and when necessary
+		%     cache partially instantiated IpoptProblem? such that fewer substitutions have to happen when changing a single specification
+		%     use better linear solver with ipopt
+		%     properly configure ipopt options
+		%     adaptive segment density
+		%     adaptive velocity/fairness trapezoid sum
 

### 2014-01-25 idp/documentation/paper.tex
continued working on evaluation section

@@
 			\subsubsection{Expressiveness of the Specification Language}
+			\label{section:expressiveness_specification_language}
 
@@
 To that end, our specification language supports specifying points, directions and curvatures at any number of positions on the curve, as well as the total length of the curve.
-This grants the user much flexibility in describing the source curve, allowing them to specify those properties that describe the curve in the most concise fashion.
+This grants the user much flexibility in describing the source curve, allowing them to specify those properties that can describe the curve in the most concise fashion.
 
@@
 For instance, by adjusting the arc length of a curve with a point specifications on each end, one can conveniently produce arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a smooth arc is produced.
-Using common vector graphics software, one usually has to either risk an asymmetric and/or bumpy arc by specifying it using Bézier splines, or deal with some cumbersome tool for constructing elliptical arcs.
+				% Using common vector graphics software, one usually has to either risk an asymmetric and/or bumpy arc by specifying it using Bézier splines, or deal with some cumbersome tool for constructing elliptical arcs.
+				Generally speaking, having control over the arc length of the curve results in benign and predictable curve behavior and is often useful for pulling it in place, not unlike the way one would handle a mechanical spline.
 
-				% expressiveness of the specification language
-				%   much more expressive than Bézier and Spiro splines, allows specification of curvature, length
-				%     controlling length is actually a very nice thing to have
-				%       a good way to control circular arcs in terms of radius (fluttershy's ear, mouth)
-				%         compare to using elliptical arcs in inkscape
-				%       a good way to pull curves in place when they don't have direction specifications
-				%       benign behavior
-				%   doesn't allow *not* expressing length, may pose a problem
-				%   demonstrate euler spirals and bézier curves being embedded
-				%     demonstrates expressiveness
-				%     minimization of normal jerk squares prefers linear change in curvature
-				%   a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification
+				On the other hand though, it is not possible not to specify the curve's arc length, and there may be cases where the user wants to leave it to the fairness measure to choose the best arc length for the curve.
+While it is sometimes possible to work around this issue by placing direction and/or curvature specifications instead of point specifications, that may not always be the case and further investigation into how arc length specification can be made optional while avoiding the problems mentioned in section \ref{section:specification_language} seems advisable.
+
+				As a final way of demonstrating the high expressiveness, it may be noted that it is actually possible to specify segments of the Euler spiral using the specification language.
+As described in section \ref{section:fairness_measure}, the MVC fairness measure prefers constant or linearly changing curvature.
+Since every curve with linearly changing curvature is some segment of the Euler spiral, it is easy to specify those using the specification language by leaving the curve otherwise unconstrained.
 

### 2014-01-25 idp/documentation/paper.tex
continued working on evaluation section

@@
 % TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
-% TODO: expand n't contractions
+% TODO: expand n't and *'s contractions
 % TODO: check if all quotations are done `like this'
@@
 
-			% human ability to `read' the specification language from curves
-			%   familiar differential geometric properties
-			% human ability to `speak' the specification language to the software
-			%   familiar differential geometric properties
-			% smoothness of curves selected by the fairness measure
-			%   very good, thanks to MVC
-			% minimality of curves selected by the fairness measure
-			%   not guaranteed by optimization, but may usually be obtained by pushing the curve into a different optimum
-			% software's ability to derive curves from descriptions
-			%   optimization is slow and sometimes unstable, still needs some work before it becomes practical
-			%   local optima
-			%   bad stability, infeasible for numeric optimization
-			%   performance
-			%   adjustment of segment count and degree by the user is not desirable
+			\subsubsection{Human Interaction with the Specification Language}
+			\label{section:human_interaction_specification_language}
+
+				The specification language is based on the familiar differential geometric properties point, direction, curvature and arc length, all of which have a simple geometric interpretation and are thus easy to understand.
+
+				Points, directions and curvatures are easy to extract from some source curve.
+While extracting the arc length of a long curve may be somewhat difficult, this is usually not required since the arc length is mostly adjusted locally, where it is fairly easy to estimate.
+
+				Since curves can usually be described using only a few specificaitons, communicating this information to the curve design tool is not too difficult either.
+Points are well-suited to be specified using the mouse pointer.
+The current prototype curve design tool only supports very basic input facilities for direction and curvature, making it somewhat difficult to create and modify these specifications.
+This is easily fixed though by utilizing their geometric interpretation, letting the user specify tangent lines and osculating circles instead of raw numeric values.
+Arc lengths are usually specified locally and by repeatedly making small adjustments, making them relatively easy to handle.
+
+			\subsubsection{Smoothness and Minimality of Curves Selected by the Fairness Measure}
+			\label{section:smoothness_minimality_curves_selected_fairness_measure}
+
+				As described in section \ref{section:fairness_measure}, the MVC fairness measure the intuitive notion of smoothness very well, resulting in aesthetically pleasing curves.
+
+				Unfortunately, the nonlinear optimization approach to finding the best curve can only find local optima, such that in many cases, the curve is not minimal, often containing loops and unnecessary inflection points.
+However, this is usually easy to fix by adjusting the specifications in such a way that the global optimum is found.
+For instance, a loop in the curve can often be removed by placing an additional point specification on the loop, pulling it away from the curve until the curve is forced to straighten out and finally removing the point specification again.
+Once the curve has been pushed into the global optimum like this, it is usually very stable.
+
+			\subsubsection{Ability to Derive Curves from Descriptions}
+			\label{section:ability_derive_curves_descriptions}
+
+				As mentioned in section \ref{section:curve_derivation}, using nonlinear optimization has many advantages for research and proof of concept projects like this one.
+Unfortunately, the flexibility it grants comes with a price, in that it is much slower and also less reliable than unique mathematical constructions like the ones used for Bézier splines.
+
+				Even on very powerful computers, running nonlinear optimization on splines with many segments and/or polynomials of high degree is too slow for interactive curve design, with updates taking more than 5 seconds.
+Most curves are very simple and do not need this kind of complexity, but for those that do, the bad performance can be very disruptive.
+
+				Another issue is reliability.
+For some descriptions, the iterative optimization algorithm may converge extremely slowly or not at all, thus not producing a curve.
+In some cases, it is possible that small changes in the description restore the ability to converge on a solution, but sometimes one simply has to start over.
+There are also cases where the optimization gets stuck in a pathological local optimum, in which case one usually has to start over aswell.
+
+				All these issues cause the prototype curve design tool to be somewhat impractical for actual curve design work, making this the primary issue with the chosen approach.
 
@@
 		% most issues stem from the optimization approach
+		%  may make it hard to actually rate usability of the idea behind this
 		% distinguish between further research work and turning the project into practical software)	

### 2014-01-26 idp/documentation/paper.tex
moved math macros
improved finding curves as optimization problem slide

@@
 \usepackage{times}             % replaces Computer Modern font with something resembling Times
-\usepackage{textcomp}          % adds symbols like \textdegree
 \usepackage{graphicx}          % enables including graphics
@@
 
-\input{math.tex}               % some math helper macros
+\input{../common/math.tex}     % some math helper macros
 
@@
 
-				In the following, we will describe how nonlinear optimization can be used for building polynomial splines based on descriptions in the language established in section \ref{section:description_language}.
-
+				In the following, we will describe how nonlinear optimization can be used for finding polynomial splines based on descriptions in the language established in section \ref{section:description_language}.
+
+			% TODO: consider structuring and naming this section similarly to the slides
 			\subsubsection{Segmentation and Notation}
 			\label{section:segmentation_notation}
+
+				%\begin{equation*}
+				%	\apply{\phi_i}{t} = \sum_j a_{i,j} \cdot t^j
+				%\end{equation*}
 
@@
 
+			% idea seems good, optimization issues get in the way
+			% maybe move this to conclusion
+
 	\section{Conclusion}

### 2014-01-26 source/Kurve/Kurve/Components/Controls/SegmentComponent.cs
added curvature visualization, improved discrete curve resolution

@@ -41,7 +41,7 @@ namespace Kurve.Component
 		{
 			if (Curve == null) return;
 
-			foreach (Tuple<double, double> positions in Scalars.GetIntermediateValuesSymmetric(leftComponent.Position, rightComponent.Position, 100).GetRanges()) 
+			foreach (Tuple<double, double> positions in Scalars.GetIntermediateValuesSymmetric(leftComponent.Position, rightComponent.Position, 250).GetRanges()) 
 			{
 				double stretchFactor = Curve.GetSpeed((positions.Item1 + positions.Item2) / 2) / BasicSpecification.CurveLength;
 
@@ -50,12 +50,13 @@ namespace Kurve.Component
 				Krach.Graphics.Color color = StretchedColor(stretchFactor);
 				Drawing.DrawLine(context, Curve.GetPoint(positions.Item1), Curve.GetPoint(positions.Item2), 2, color);
 
-//				Vector2Double point = 0.5 * (Curve.GetPoint(positions.Item1) + Curve.GetPoint(positions.Item2));
-//				Vector2Double direction = Curve.GetDirection((positions.Item1 + positions.Item2) / 2);
-//				Vector2Double angularDirection = new Vector2Double(direction.Y, -direction.X);
-//				double curvature = Curve.GetCurvature((positions.Item1 + positions.Item2) / 2);
-//				Vector2Double curvatureVector = 10000 * curvature * angularDirection;
-//				Drawing.DrawLine(context, point, point + curvatureVector, 2, Colors.Green);
+				Vector2Double point = 0.5 * (Curve.GetPoint(positions.Item1) + Curve.GetPoint(positions.Item2));
+				double direction = Curve.GetDirection((positions.Item1 + positions.Item2) / 2);
+				Vector2Double directionVector = new Vector2Double(Scalars.Cosine(direction), Scalars.Sine(direction));
+				Vector2Double angularDirection = new Vector2Double(directionVector.Y, -directionVector.X);
+				double curvature = Curve.GetCurvature((positions.Item1 + positions.Item2) / 2);
+				Vector2Double curvatureVector = 10000 * curvature * angularDirection;
+				Drawing.DrawLine(context, point, point + curvatureVector, 0.5, Colors.Green);
 			}
 
 			base.Draw(context);

### 2014-01-26 source/Kurve/Kurve/Components/RootComponent.cs
Added first version of SVG export.

@@ -95,6 +95,7 @@ namespace Kurve.Component
 				case Kurve.Interface.Key.N: AddCurve(); break;
 				case Kurve.Interface.Key.L: Load(); break;
 				case Kurve.Interface.Key.S: Save(); break;
+				case Kurve.Interface.Key.E: Export(); break;
 			}
 
 			base.KeyUp(key);
@@ -133,6 +134,36 @@ namespace Kurve.Component
 						select curveComponent.Specification.XElement
 					)
 					.Save(fileChooser.Filename);
+				}
+				
+				fileChooser.Destroy();
+			}
+		}
+		
+		void DoExport(IEnumerable<CurveOptimizer> optimizers, String filename) {			
+			XAttribute style = new XAttribute("style", "fill:none; stroke:black; stroke-width:2");
+			XNamespace @namespace = "http://www.w3.org/2000/svg";
+			XNamespace noNamespace = "";
+			new XElement(
+				@namespace + "svg",
+		        new XAttribute("viewBox", "0 0 500 400"),
+				new XAttribute("version", "1.1"),
+				from optimizer in optimizers
+				select new XElement(@namespace+"path", style, new XAttribute("d", optimizer.GetSvgAttributeString(1000)))
+			)
+			.Save(filename);
+		}
+		void Export() {
+			using (FileChooserDialog fileChooser = new FileChooserDialog("Export", parentWindow, FileChooserAction.Save, "Cancel", ResponseType.Cancel, "Save", ResponseType.Accept))
+			{
+				if ((ResponseType)fileChooser.Run() == ResponseType.Accept) 
+				{
+						
+					IEnumerable<CurveOptimizer> optimizers = 
+						from curveComponent in curveComponents
+						select curveComponent.CurveOptimizer;
+					
+					this.optimizationWorker.SubmitTask(this, () => DoExport(optimizers, fileChooser.Filename));
 				}
 				
 				fileChooser.Destroy();

### 2014-01-26 source/Kurve/Kurve/Components/RootComponent.cs
Added support for exporting curvature indicators.

@@ -140,25 +140,36 @@ namespace Kurve.Component
 			}
 		}
 		
-		void DoExport(IEnumerable<CurveOptimizer> optimizers, String filename) {			
+		void DoExport(IEnumerable<CurveOptimizer> optimizers, String filename) 
+		{
+			Console.WriteLine("exporting to "+filename);
+			
 			XAttribute style = new XAttribute("style", "fill:none; stroke:black; stroke-width:2");
+			XAttribute curvatureStyle = new XAttribute("style", "fill:none; stroke:green; stroke-width:0.5");
 			XNamespace @namespace = "http://www.w3.org/2000/svg";
-			XNamespace noNamespace = "";
+			int width;
+			int height;
+			this.parentWindow.GetSize(out width, out height);
 			new XElement(
 				@namespace + "svg",
-		        new XAttribute("viewBox", "0 0 500 400"),
+		        new XAttribute("viewBox", "0 0 "+width+" "+height),
 				new XAttribute("version", "1.1"),
 				from optimizer in optimizers
-				select new XElement(@namespace+"path", style, new XAttribute("d", optimizer.GetSvgAttributeString(1000)))
+				from item in Enumerables.Create
+				(
+					new XElement(@namespace+"path", style, new XAttribute("d", optimizer.GetSvgAttributeString(1000))),
+					new XElement(@namespace+"path", curvatureStyle, new XAttribute("d", optimizer.GetCurvatureIndicators(250)))
+				)
+				select item
 			)
 			.Save(filename);
 		}
-		void Export() {
+		void Export() 
+		{
 			using (FileChooserDialog fileChooser = new FileChooserDialog("Export", parentWindow, FileChooserAction.Save, "Cancel", ResponseType.Cancel, "Save", ResponseType.Accept))
 			{
 				if ((ResponseType)fileChooser.Run() == ResponseType.Accept) 
 				{
-						
 					IEnumerable<CurveOptimizer> optimizers = 
 						from curveComponent in curveComponents
 						select curveComponent.CurveOptimizer;

### 2014-01-26 source/Kurve/Kurve/CurveOptimizer.cs

@@ -59,6 +59,22 @@ namespace Kurve.Component
 			return svgString+" "+svgComponents;
 		}
 		
+		public String GetCurvatureIndicators(int count) {
+			Kurve.Curves.Curve curve = optimizer.GetCurve(specification);
+			
+			return (
+				from positions in Scalars.GetIntermediateValuesSymmetric(0, 1, 250).GetRanges()
+				let point = 0.5 * (curve.GetPoint(positions.Item1) + curve.GetPoint(positions.Item2))
+				let direction = curve.GetDirection((positions.Item1 + positions.Item2) / 2)
+				let directionVector = new Vector2Double(Scalars.Cosine(direction), Scalars.Sine(direction))
+				let angularDirection = new Vector2Double(directionVector.Y, -directionVector.X)
+				let curvature = curve.GetCurvature((positions.Item1 + positions.Item2) / 2)
+				let curvatureVector = 10000 * curvature * angularDirection
+				select "M "+PointToSvgString(point)+" "+PointToSvgString(point + curvatureVector)+" "
+			)
+			.AggregateString();
+		}
+		
 		String PointToSvgString(Vector2Double point) 
 		{
 			return point.X+","+point.Y;

### 2014-01-27 source/Kurve/Kurve.Curves/Optimization/OptimizationProblem.cs
added experimental soft specifications

@@ -34,47 +34,7 @@ namespace Kurve.Curves.Optimization
 			this.curveSpecifications = curveSpecifications;
 
 			this.curveLength = Terms.Variable("curveLength");
-
-			IEnumerable<ValueTerm> variables = optimizationSegments.Parameters; 
 			
-			ValueTerm segmentLength = Terms.Quotient(curveLength, Terms.Constant(optimizationSegments.Segments.Count()));
-
-			ValueTerm speedError = Terms.Scaling
-			(
-				Terms.Exponentiation(segmentLength, Terms.Constant(-2)),
-				Terms.Sum
-				(
-					from segment in optimizationSegments.Segments
-					let position = Terms.Variable("t")
-					let curve = segment.LocalCurve
-					let speed = curve.Speed.Apply(position)
-					let error = Terms.Square(Terms.Difference(speed, segmentLength))
-					select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
-				)
-			);
-
-			ValueTerm fairnessError = Terms.Scaling
-			(
-				Terms.Exponentiation(segmentLength, Terms.Constant(-4)),
-				Terms.Sum
-				(
-					from segment in optimizationSegments.Segments
-					let position = Terms.Variable("t")
-					let curve = segment.LocalCurve
-					let jerk = curve.Jerk.Apply(position)
-					let normal = Terms.Normal(curve.Velocity.Apply(position))
-					let error = Terms.Square(Terms.DotProduct(jerk, normal))
-					select Terms.IntegrateTrapezoid(error.Abstract(position), new OrderedRange<double>(0, 1), 100)
-				)
-			);
-
-			ValueTerm objectiveValue = Terms.Sum
-			(
-				Terms.Product(Terms.Constant(100000.0), speedError),
-				Terms.Product(Terms.Constant(1), fairnessError)
-			)
-			.Simplify();
-
 			this.pointSpecificationTemplates =
 			(
 				from pointSpecificationIndex in Enumerable.Range(0, curveSpecifications.Count(curveSpecification => curveSpecification is PointCurveSpecification))
@@ -93,6 +53,90 @@ namespace Kurve.Curves.Optimization
 				select SpecificationTemplate.CreateCurvatureSpecificationTemplate(optimizationSegments.Segments, curveLength, curvatureSpecificationIndex)
 			)
 			.ToArray();
+
[…]

### 2014-01-27 source/Kurve/Kurve/Components/RootComponent.cs

@@ -157,7 +157,7 @@ namespace Kurve.Component
 				from optimizer in optimizers
 				from item in Enumerables.Create
 				(
-					new XElement(@namespace+"path", style, new XAttribute("d", optimizer.GetSvgAttributeString(1000))),
+					new XElement(@namespace+"path", style, new XAttribute("d", optimizer.GetSvgAttributeString(100))),
 					new XElement(@namespace+"path", curvatureStyle, new XAttribute("d", optimizer.GetCurvatureIndicators(250)))
 				)
 				select item

### 2014-01-27 idp/documentation/paper.tex
fixed typo in documentation
continued working on slides
minor adjustments of documentation

@@
 Since there is no way to directly specify curvature using Bézier splines, a considerable number of nodes is needed to get close to the desired curve.
+This in turn, may lead to decreased smoothness as discussed in the previous paragraphs.
 
@@
 These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
-This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
+This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to the expression of circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
 Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties.
@@
 				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need the function mapping the covered arc length to the corresponding parameter values in order to state the optimization goals for the specification items.
-Unfortunately, this function is not very well-suited to be used as a term in a smooth opimization problem even for simple curves like polynomials.
+Unfortunately, this function is not very well-suited to be used as a term in a smooth optimization problem even for simple curves like polynomials.
 Also, since the function isn't fixed and varies with the coefficients, there is no way of telling on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
@@
 		% investigate into the calculus of variations and whether there may be a possibility to construct MVC curves based on our specification format, once the format itsself has proven to be useful
-		% performance
-
-		%   possible improvements
+
+		%   possible performance improvements
 		%     make sure substitutions only happen where and when necessary

### 2014-01-27 source/Kurve/Kurve/CurveOptimizer.cs
improved svg export accuracy

@@ -49,8 +49,8 @@ namespace Kurve.Component
 			string svgComponents = 
 			(
 				from range in ranges
-				let controlPoint2 = curve.GetPoint(range.Item1) + 0.33*(1.0/numberOfSegments)*Velocity(curve.GetDirection(range.Item1), curve.GetSpeed(range.Item1))
-				let controlPoint3 = curve.GetPoint(range.Item2) - 0.33*(1.0/numberOfSegments)*Velocity(curve.GetDirection(range.Item2), curve.GetSpeed(range.Item2))
+				let controlPoint2 = curve.GetPoint(range.Item1) + (1.0 / (3 * numberOfSegments)) * Velocity(curve.GetDirection(range.Item1), curve.GetSpeed(range.Item1))
+				let controlPoint3 = curve.GetPoint(range.Item2) - (1.0 / (3 * numberOfSegments)) * Velocity(curve.GetDirection(range.Item2), curve.GetSpeed(range.Item2))
 				let controlPoint4 = curve.GetPoint(range.Item2)
 				select " C"+PointToSvgString(controlPoint2)+" "+PointToSvgString(controlPoint3)+" "+PointToSvgString(controlPoint4)
 			)

### 2014-01-27 source/Kurve/Kurve/DiscreteCurve.cs
removed discrete curve precision

@@ -21,7 +21,7 @@ namespace Kurve
 
 			this.items =
 			(
-				from position in Scalars.GetIntermediateValuesSymmetric(0, 1, 1000)
+				from position in Scalars.GetIntermediateValuesSymmetric(0, 1, 250)
 				select new DiscreteCurveItem
 				(
 					curve.GetPoint(position),

### 2014-01-30 source/Kurve/Kurve.Curves/Curve.cs
moved curve drawing to CurveComponent simplified, centralized and improved segmentation fixed various off-by-one errors fixed an off-by-one error causing wrong velocities in SVG export fixed capturing-related crash

@@ -1,9 +1,5 @@
-using System;
-using System.Linq;
-using System.Collections.Generic;
 using Krach.Basics;
 using Krach.Extensions;
-using Wrappers.Casadi;
 
 namespace Kurve.Curves
 {
@@ -13,6 +9,23 @@ namespace Kurve.Curves
 		public abstract double GetSpeed(double position);
 		public abstract double GetDirection(double position);
 		public abstract double GetCurvature(double position);
+
+		public Vector2Double GetDirectionVector(double position)
+		{
+			double direction = GetDirection(position);
+
+			return new Vector2Double(Scalars.Cosine(direction), Scalars.Sine(direction));
+		}
+		public Vector2Double GetNormalVector(double position)
+		{
+			Vector2Double directionVector = GetDirectionVector(position);
+
+			return new Vector2Double(directionVector.Y, -directionVector.X);
+		}
+		public Vector2Double GetVelocity(double position)
+		{
+			return GetSpeed(position) * GetDirectionVector(position);
+		}
 	}
 }
 

### 2014-01-30 source/Kurve/Kurve/Components/Controls/Abstract/PositionedControlComponent.cs

@@ -26,6 +26,7 @@ namespace Kurve.Component
 
 		public BasicSpecification BasicSpecification { get { return curveComponent.BasicSpecification; } }
 		public Curve Curve { get { return curveComponent.Curve; } }
+		public int SegmentCount { get { return curveComponent.SegmentCount; } }
 		public abstract double Position { get; }
 		public bool IsSelected { get { return isSelected; } set { isSelected = value; } }
 		public bool IsDragging { get { return isDragging; } }

### 2014-01-30 source/Kurve/Kurve/Components/Controls/SegmentComponent.cs

@@ -24,6 +24,17 @@ namespace Kurve.Component
 		readonly PositionedControlComponent rightComponent;
 
 		Orthotope2Double Bounds { get { return new Orthotope2Double(Point - 0.5 * size, Point + 0.5 * size); } }
+		int SegmentSegmentCount
+		{
+			get
+			{
+				double length = rightComponent.Position - leftComponent.Position;
+
+				if (length == 0) return 1;
+
+				return (int)(length * SegmentCount).Ceiling();
+			}
+		}
 
 		public override double Position { get { return (leftComponent.Position + rightComponent.Position) / 2; } }
 		public Vector2Double Point { get { return Curve == null ? Vector2Double.Origin : Curve.GetPoint(Position); } }
@@ -39,41 +50,11 @@ namespace Kurve.Component
 
 		public override void Draw(Context context)
 		{
-			if (Curve == null) return;
-
-			foreach (Tuple<double, double> positions in Scalars.GetIntermediateValuesSymmetric(leftComponent.Position, rightComponent.Position, 250).GetRanges()) 
-			{
-				double stretchFactor = Curve.GetSpeed((positions.Item1 + positions.Item2) / 2) / BasicSpecification.CurveLength;
-
-				if (IsSelected) Drawing.DrawLine(context, Curve.GetPoint(positions.Item1), Curve.GetPoint(positions.Item2), 8, Colors.Green.ReplaceAlpha(0.3));
-
-				Krach.Graphics.Color color = StretchedColor(stretchFactor);
-				Drawing.DrawLine(context, Curve.GetPoint(positions.Item1), Curve.GetPoint(positions.Item2), 2, color);
-
-				Vector2Double point = 0.5 * (Curve.GetPoint(positions.Item1) + Curve.GetPoint(positions.Item2));
-				double direction = Curve.GetDirection((positions.Item1 + positions.Item2) / 2);
-				Vector2Double directionVector = new Vector2Double(Scalars.Cosine(direction), Scalars.Sine(direction));
-				Vector2Double angularDirection = new Vector2Double(directionVector.Y, -directionVector.X);
-				double curvature = Curve.GetCurvature((positions.Item1 + positions.Item2) / 2);
-				Vector2Double curvatureVector = 10000 * curvature * angularDirection;
-				Drawing.DrawLine(context, point, point + curvatureVector, 0.5, Colors.Green);
-			}
+			if (Curve != null && IsSelected)
[…]

### 2014-01-30 source/Kurve/Kurve/Components/Controls/SpecificationComponent.cs

@@ -41,7 +41,8 @@ namespace Kurve.Component
 			{ 
 				if (specifiesPoint) return point;
 				else if (Curve != null) return Curve.GetPoint(Position);
-				else return Vector2Double.Origin;			}
+				else return Vector2Double.Origin;
+			}
 			set 
 			{
 				if (specifiesPoint) point = value;
@@ -136,7 +137,7 @@ namespace Kurve.Component
 				{
 					double closestPosition = 
 					(
-						from position in Scalars.GetIntermediateValuesSymmetric(0, 1, 100)
+						from position in Scalars.GetIntermediateValuesSymmetric(0, 1, SegmentCount + 1)
 						let distance = (Curve.GetPoint(position) - mousePosition).Length
 						orderby distance ascending
 						select position

### 2014-01-30 source/Kurve/Kurve/Components/CurveComponent.cs

@@ -18,9 +18,6 @@ namespace Kurve.Component
 	class CurveComponent : Component
 	{
 		readonly CurveOptimizer curveOptimizer;
-		
-		public CurveOptimizer CurveOptimizer { get { return curveOptimizer; } }
-		
 		readonly List<SpecificationComponent> specificationComponents;
 		readonly List<SegmentComponent> segmentComponents;
 		readonly FixedPositionComponent curveStartComponent;
@@ -100,9 +97,11 @@ namespace Kurve.Component
 		}
 
 		public event Action RemoveCurve;
-
+		
+		public CurveOptimizer CurveOptimizer { get { return curveOptimizer; } }
 		public BasicSpecification BasicSpecification { get { return basicSpecification; } }
 		public Curve Curve { get { return curve; } }
+		public int SegmentCount { get { return curveOptimizer.SegmentCount; } }
 		public Specification Specification { get { return curveOptimizer.Specification; } }
 
 		public CurveComponent(Component parent, OptimizationWorker optimizationWorker, Specification specification) : base(parent)
@@ -124,11 +123,13 @@ namespace Kurve.Component
 
 			curveOptimizer.Submit(nextSpecification);
 
-			IEnumerable<SpecificationComponent> specificationComponents = 
+			IEnumerable<SpecificationComponent> specificationComponents =
+			(
 				from spec in nextSpecification.CurveSpecifications
 				orderby spec.Position ascending
 				group spec by spec.Position into specificationGroup
-				select new SpecificationComponent(this, this, specificationGroup.Key, specificationGroup);
+				select new SpecificationComponent(this, this, specificationGroup.Key, specificationGroup)
+			);
 
 			foreach (SpecificationComponent component in specificationComponents) AddSpecificationComponent(component);
 		}
@@ -306,7 +307,6 @@ namespace Kurve.Component
 			}
 		}
 
-
 		void RebuildSegmentComponents()
 		{
 			IEnumerable<PositionedControlComponent> orderedSpecificationComponents =
@@ -350,6 +350,23 @@ namespace Kurve.Component
 			}
 		}
 
+		public override void Draw(Context context)
+		{
+			if (curve != null)
+				foreach (Tuple<double, double> positions in Scalars.GetIntermediateValuesSymmetric(0, 1, SegmentCount + 1).GetRanges())
[…]

### 2014-01-30 source/Kurve/Kurve/Components/RootComponent.cs

@@ -10,6 +10,7 @@ using Kurve.Curves.Optimization;
 using Kurve.Interface;
 using Gtk;
 using System.Xml.Linq;
+using Krach.Formats.Svg;
 
 namespace Kurve.Component
 {
@@ -142,25 +143,17 @@ namespace Kurve.Component
 		
 		void DoExport(IEnumerable<CurveOptimizer> optimizers, String filename) 
 		{
-			Console.WriteLine("exporting to "+filename);
-			
-			XAttribute style = new XAttribute("style", "fill:none; stroke:black; stroke-width:2");
-			XAttribute curvatureStyle = new XAttribute("style", "fill:none; stroke:green; stroke-width:0.5");
-			XNamespace @namespace = "http://www.w3.org/2000/svg";
-			int width;
-			int height;
-			this.parentWindow.GetSize(out width, out height);
-			new XElement(
-				@namespace + "svg",
-		        new XAttribute("viewBox", "0 0 "+width+" "+height),
+			int width, height;
+			parentWindow.GetSize(out width, out height);
+
+			new XElement
+			(
+				Svg.Namespace + "svg",
+				new XAttribute("viewBox", string.Format("{0} {1} {2} {3}", 0, 0, width, height)),
 				new XAttribute("version", "1.1"),
 				from optimizer in optimizers
-				from item in Enumerables.Create
-				(
-					new XElement(@namespace+"path", style, new XAttribute("d", optimizer.GetSvgAttributeString(100))),
-					new XElement(@namespace+"path", curvatureStyle, new XAttribute("d", optimizer.GetCurvatureIndicators(250)))
-				)
-				select item
+				from path in optimizer.GetSvgPaths()
+				select path
 			)
 			.Save(filename);
 		}
@@ -170,11 +163,14 @@ namespace Kurve.Component
 			{
 				if ((ResponseType)fileChooser.Run() == ResponseType.Accept) 
 				{
-					IEnumerable<CurveOptimizer> optimizers = 
+					IEnumerable<CurveOptimizer> optimizers =
+					(
 						from curveComponent in curveComponents
-						select curveComponent.CurveOptimizer;
+						select curveComponent.CurveOptimizer
+					);
+					string path = fileChooser.Filename;
 					
-					this.optimizationWorker.SubmitTask(this, () => DoExport(optimizers, fileChooser.Filename));
[…]

### 2014-01-30 source/Kurve/Kurve/CurveOptimizer.cs

@@ -12,11 +12,16 @@ using Krach.Maps.Scalar;
 using Krach.Maps;
 using Kurve.Interface;
 using Gtk;
+using System.Xml.Linq;
+using Krach.Formats.Svg;
 
 namespace Kurve.Component
 {
 	class CurveOptimizer
 	{
+		public const double SegmentDensity = 0.2;
+		public const int CurvatureMarkersFactor = 5000;
+
 		readonly Optimizer optimizer;
 		readonly OptimizationWorker optimizationWorker;
 		
@@ -25,6 +30,7 @@ namespace Kurve.Component
 		public event Action<BasicSpecification, Kurve.Curves.Curve> CurveChanged;
 
 		public Specification Specification { get { return specification; } }
+		public int SegmentCount { get { return (int)(SegmentDensity * specification.BasicSpecification.CurveLength).Ceiling(); } }
 
 		public CurveOptimizer(OptimizationWorker optimizationWorker, Specification specification)
 		{
@@ -38,50 +44,33 @@ namespace Kurve.Component
 
 		public void Submit(BasicSpecification basicSpecification)
 		{
-			optimizationWorker.SubmitTask(this, () => this.Optimize(basicSpecification));
+			optimizationWorker.SubmitTask(this, curveOptimizer => curveOptimizer.Optimize(basicSpecification));
 		}
-		public String GetSvgAttributeString(int numberOfSegments) 
+		public IEnumerable<XElement> GetSvgPaths()
 		{
 			Kurve.Curves.Curve curve = optimizer.GetCurve(specification);
-			IEnumerable<Tuple<double, double>> ranges = Scalars.GetIntermediateValuesSymmetric(0, 1, numberOfSegments).GetRanges();
 			
-			string svgString = "M"+PointToSvgString(curve.GetPoint(ranges.First().Item1));
-			string svgComponents = 
+			XAttribute curveStyle = new XAttribute("style", "fill:none; stroke:black; stroke-width:2");
+			IEnumerable<string> curveCommands = Enumerables.Concatenate
 			(
-				from range in ranges
-				let controlPoint2 = curve.GetPoint(range.Item1) + (1.0 / (3 * numberOfSegments)) * Velocity(curve.GetDirection(range.Item1), curve.GetSpeed(range.Item1))
[…]

### 2014-01-30 source/Kurve/Kurve/DiscreteCurve.cs

@@ -15,13 +15,14 @@ namespace Kurve
 
 		public IEnumerable<DiscreteCurveItem> Items { get { return items; } }
 
-		public DiscreteCurve(Curve curve)
+		public DiscreteCurve(Curve curve, int segmentCount)
 		{
 			if (curve == null) throw new ArgumentNullException("curve");
+			if (segmentCount < 1) throw new ArgumentOutOfRangeException("segmentCount");
 
 			this.items =
 			(
-				from position in Scalars.GetIntermediateValuesSymmetric(0, 1, 250)
+				from position in Scalars.GetIntermediateValuesSymmetric(0, 1, segmentCount + 1)
 				select new DiscreteCurveItem
 				(
 					curve.GetPoint(position),

### 2014-01-30 source/Kurve/Kurve/OptimizationWorker.cs

@@ -16,7 +16,7 @@ namespace Kurve
 	{
 		readonly ManualResetEvent workAvailable;
 		readonly Thread workerThread;
-		readonly Dictionary<object, Action> optimizationTasks;
+		readonly Dictionary<CurveOptimizer, Action<CurveOptimizer>> optimizationTasks;
 
 		bool disposed = false;
 		bool running = true;
@@ -26,7 +26,7 @@ namespace Kurve
 			this.workAvailable = new ManualResetEvent(false);
 			this.workerThread = new Thread(Work);
 			this.workerThread.Start();
-			this.optimizationTasks = new Dictionary<object, System.Action>();
+			this.optimizationTasks = new Dictionary<CurveOptimizer, Action<CurveOptimizer>>();
 		}
 
 		public void Dispose()
@@ -44,11 +44,11 @@ namespace Kurve
 				GC.SuppressFinalize(this);
 			}
 		}
-		public void SubmitTask(object key, Action action)
+		public void SubmitTask(CurveOptimizer curveOptimizer, Action<CurveOptimizer> action)
 		{
 			lock (optimizationTasks)
 			{
-				optimizationTasks[key] = action;
+				optimizationTasks[curveOptimizer] = action;
 
 				workAvailable.Set();
 			}
@@ -62,7 +62,7 @@ namespace Kurve
 
 				if (!running) break;
 
-				KeyValuePair<object, Action> task;
+				KeyValuePair<CurveOptimizer, Action<CurveOptimizer>> task;
 
 				lock (optimizationTasks)
 				{
@@ -73,7 +73,7 @@ namespace Kurve
 					if (!optimizationTasks.Any()) workAvailable.Reset();
 				}
 
-				task.Value();
+				task.Value(task.Key);
 			}
 		}
 	}

### 2014-01-30 source/Kurve/Kurve/Components/RootComponent.cs
added specification item visualizations to SVG export

@@ -146,11 +146,9 @@ namespace Kurve.Component
 			int width, height;
 			parentWindow.GetSize(out width, out height);
 
-			new XElement
+			Svg.Scene
 			(
-				Svg.Namespace + "svg",
-				new XAttribute("viewBox", string.Format("{0} {1} {2} {3}", 0, 0, width, height)),
-				new XAttribute("version", "1.1"),
+				new Orthotope2Double(0, 0, width, height),
 				from optimizer in optimizers
 				from path in optimizer.GetSvgPaths()
 				select path

### 2014-01-30 source/Kurve/Kurve/CurveOptimizer.cs

@@ -49,8 +49,7 @@ namespace Kurve.Component
 		public IEnumerable<XElement> GetSvgPaths()
 		{
 			Kurve.Curves.Curve curve = optimizer.GetCurve(specification);
-			
-			XAttribute curveStyle = new XAttribute("style", "fill:none; stroke:black; stroke-width:2");
+
 			IEnumerable<string> curveCommands = Enumerables.Concatenate
 			(
 				Enumerables.Create(Svg.MoveTo(curve.GetPoint(0))),
@@ -60,9 +59,8 @@ namespace Kurve.Component
 				let point2 = curve.GetPoint(positions.Item2)
 				select Svg.CurveTo(controlPoint1, controlPoint2, point2)
 			);
-			yield return new XElement(Svg.Namespace + "path", curveStyle, new XAttribute("d", curveCommands.Separate(" ").AggregateString()));
-			
-			XAttribute curvatureMarkersStyle = new XAttribute("style", "fill:none; stroke:blue; stroke-width:0.5");
+			yield return Svg.PathShape(curveCommands, null, Colors.Black, 2);
+
 			IEnumerable<string> curvatureMarkersCommands =
 			(
 				from positions in Scalars.GetIntermediateValuesSymmetric(0, 1, SegmentCount + 1).GetRanges()
@@ -70,7 +68,40 @@ namespace Kurve.Component
 				let curvatureVector = CurvatureMarkersFactor * curve.GetCurvature(position) * curve.GetNormalVector(position)
 				select Svg.Line(curve.GetPoint(position), curve.GetPoint(position) + curvatureVector)
 			);
-			yield return new XElement(Svg.Namespace + "path", curvatureMarkersStyle, new XAttribute("d", curvatureMarkersCommands.Separate(" ").AggregateString()));
+			yield return Svg.PathShape(curvatureMarkersCommands, null, Colors.Blue, 0.5);
+
+			foreach (PointCurveSpecification pointCurveSpecification in specification.BasicSpecification.CurveSpecifications.OfType<PointCurveSpecification>())
+			{
+				Vector2Double point = curve.GetPoint(pointCurveSpecification.Position);
+
+				yield return Svg.CircleShape(point, 10, Colors.Red, null, 0);
+			}
+			foreach (DirectionCurveSpecification directionCurveSpecification in specification.BasicSpecification.CurveSpecifications.OfType<DirectionCurveSpecification>())
+			{
[…]

### 2014-02-01 idp/documentation/paper.tex
minor refinement on paper and slides
started fixing usage of specification language vs. description language fixed float placing

@@
 			In a similar fashion, one may consider some shortcomings (for instance, insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request curvature continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+These decisions have to be made on a case by case basis, selecting the approach that is better suited to modeling the reality of curve design.
 
@@
 			We conclude that the two ways in which curve design tools can affect the design process, are through their choice of specification language and fairness measure, as well as their practical ability to derive curves from them.
-We call a combination of specification language and fairness measure a description language.
-One may note that while the specification language has both syntactical as well as semantical influence on the language, the fairness measure determines semantics only.
-
-			Ideally, the specification language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
+Note that while syntactically complete, the semantics of the specification language also depend on the fairness measure.
+Thus, it makes sense to consider the combination of a specification language and a fairness measure, which we will call a description language.
+
+			Ideally, the description language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
 The fairness measure should capture the intuitive notion of curve smoothness sufficiently well and select curves that are minimal with respect to the specification.
@@
 			\begin{enumerate}
-				\item expressiveness of the specification language
+				\item expressiveness of the description language
 				\item human ability to `read' the specification language from curves
@@
 Bézier splines are furthermore very simple to derive from their description, fulfilling criterion 6 perfectly.
-Design tools for Bézier splines have some issues with both the fairness measure's smoothness (criterion 4) and the expressiveness of the specification language (criterion 1), both of which shall be analyzed in the following paragraphs.
+Design tools for Bézier splines have some issues with both the fairness measure's smoothness (criterion 4) and the expressiveness of the description language (criterion 1), both of which shall be analyzed in the following paragraphs.
 
@@
 In case the software allows the user to request curvature continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
-Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with specification language expressiveness.
-
-				The specification language used to specify Bézier splines in common design tools is lacking in expressiveness.
+Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with description language expressiveness.
+
+				The description language used to specify Bézier splines in common design tools is lacking in expressiveness.
 Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature).
@@
 
-				\begin{figure}[htb]
+				\begin{figure}[htbp]
 					\centering
-					\includegraphics[width=\textwidth]{../resources/usability_bezier.pdf}
-					\caption{Usability Analysis of Bézier Splines (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
-					\label{figure:usability_bézier}
+					\includegraphics[width=\textwidth]{../resources/examples_bezier.pdf}
+					\caption{Bézier Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
+					\label{figure:examples_bézier}
 				\end{figure}
 
-				Figure \ref{figure:usability_bézier} illustrates some of these issues by showing some examples of Bézier splines as they appear in common design tools.
+				Figure \ref{figure:examples_bézier} illustrates some of these issues by showing some examples of Bézier splines as they appear in common design tools.
 Bézier nodes are displayed as red dots, while Bézier handles are displayed as green lines.
@@
 The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 6 well enough.
-The only real issue lies in the insufficient expressiveness of the specification language (criterion 1), discussed further in the following paragraph.
+The only real issue lies in the insufficient expressiveness of the description language (criterion 1), discussed further in the following paragraph.
 
@@
 
-				\begin{figure}[htb]
+				\begin{figure}[htbp]
 					\centering
-					\includegraphics[width=\textwidth]{../resources/usability_spiro.pdf}
-					\caption{Usability Analysis of Spiro Splines (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
-					\label{figure:usability_spiro}
+					\includegraphics[width=\textwidth]{../resources/examples_spiro.pdf}
+					\caption{Spiro Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
+					\label{figure:examples_spiro}
 				\end{figure}
 
-				Figure \ref{figure:usability_spiro} illustrates some of these issues by showing some examples of Spiro splines as they appear in common design tools. Control points are displayed as red dots.
+				Figure \ref{figure:examples_spiro} illustrates some of these issues by showing some examples of Spiro splines as they appear in common design tools. Control points are displayed as red dots.
 
@@
 
-			\begin{figure}[htb]
+			\begin{figure}[htbp]
 				\centering
@@
 			\label{section:fairness_measure}
+
+				% TODO: fairness measure integral is over the arc length of the curve
+				% TODO: mention that mvc facilitates describing circular arcs and spirals (constant and linearly changing curvature)
+				% TODO: add some illustrations, describe their general behaviour, smoothness
 
@@
 
+				% TODO: sketch overall curve derivation process similar to the summary slide
+
 	\section{Implementation}
@@
 
-				\begin{figure}[htb]
+				\begin{figure}[htbp]
 					\centering
@@
 			\subsection{Specification}
-				\begin{figure}[htb]
+				\begin{figure}[htbp]
 					\centering
@@
 			\subsection{Optimization}
-				\begin{figure}[htb]
+				\begin{figure}[htbp]
 					\centering
@@
 			\subsection{Implementation}
-				\begin{figure}[htb]
+				\begin{figure}[htbp]
 					\centering
@@
 
-				\begin{figure}[htb]
+				\begin{figure}[htbpp]
 					\centering
@@
 
-			Having applied the usability criteria from section \ref{section:usability_criteria_curve_design_tools} to Bézier splines and Spiro splines, we will now use them to evaluate usability aspects of our approach.
-
-			\subsubsection{Expressiveness of the Specification Language}
-			\label{section:expressiveness_specification_language}
-
-				One of the primary goals of this project was to explore more expressive specification languages than those which are currently used in vector graphics software.
+			Having applied the usability criteria from section \ref{section:usability_criteria_curve_design_tools} to Bézier splines and Spiro splines, we will now use them to evaluate usability aspects of the prototype curve design tool.
+
+			\subsubsection{Expressiveness of the Description Language}
+			\label{section:expressiveness_description_language}
+
+				% TODO: mention that using the description (not specification) language, one can now specify linearly increasing curvature
+
+				One of the primary goals of this project was to explore more expressive description languages than those which are currently used in vector graphics software.
 To that end, our specification language supports specifying points, directions and curvatures at any number of positions on the curve, as well as the total length of the curve.
@@
 While fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, there are actually quite a few upsides.
-For instance, by adjusting the arc length of a curve with a point specifications on each end, one can conveniently produce arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a smooth arc is produced.
+For instance, by adjusting the arc length of a curve with a point specifications on each end, one can conveniently produce curcular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced.
 				% Using common vector graphics software, one usually has to either risk an asymmetric and/or bumpy arc by specifying it using Bézier splines, or deal with some cumbersome tool for constructing elliptical arcs.
@@
 
-				Since curves can usually be described using only a few specificaitons, communicating this information to the curve design tool is not too difficult either.
+				Since curves can usually be described using only a few specifications, communicating this information to the curve design tool is not too difficult either.
 Points are well-suited to be specified using the mouse pointer.
@@
 
+			We will now revisit the examples shown in sections \ref{section:bézier_splines} and \ref{section:spiro_splines} using the prototype curve design tool in order to see how well the description language we propose can handle them.
+
+			\begin{figure}[htbp]
+				\centering
+				\includegraphics[width=\textwidth]{../resources/examples_prototype.pdf}
+				\caption{Prototype Curve Design Tool Examples (left: Fluttershy \cite{fluttershy})}
+				\label{figure:examples_prototype}
+			\end{figure}
+
+			% TODO: add illustration to section introducing specification language, add reference to it here
+			Figure \ref{figure:examples_prototype} shows some examples of curves designed using the prototype curve design tool.
+Like in section 1.2.3, point specification items are represented using red dots, direction specification items are represented using green tangent lines and curvature specification items are represented using blue circular arcs taken from the osculating circle.
+
+			Example 1 shows how adjusting the arc length of an otherwise unconstrained curve allows the convenient creation of curcular arcs between two points.
+
+			Examples 2, 3 and 4 show how to make use of the good expressiveness of the description language.
+
 			% fluttershy's mane
 			%   spirals done this way are guaranteed to be smooth and not have any unintentional bumps
+			%     since point specifications occur only at the start and end of the curve, there is no danger of introducing unintentional bumps in the middle of the curve
 			%   comparison with bézier
@@
 			% idea seems good, optimization issues get in the way
+			% usability of the prototype curve design tool is still bad, but that may not mean the description language is useless, maybe there are better ways than optimization to derive curves from descriptions and to fulfill criterion 6 better
 			% maybe move this to conclusion

### 2014-02-01 idp/documentation/paper.tex
first draft of evaluation and conclusion section is complete

@@
 Unfortunately, designing curves using the tools available in current vector graphics software can be very frustrating.
-The artist is often unable to make the curve look the way they want it to.
-To make matters worse it is often unclear why this is the case, as the software seems to put all the necessary tools into the hands of the designer, yet at times, they somehow don't allow the artist to communicate their intention to the software.
+The user is often unable to make the curve look the way they want it to.
+To make matters worse it is often unclear why this is the case, as the software seems to put all the necessary tools into the hands of the designer, yet at times, they somehow don't allow the user to communicate their intention to the software.
 
@@
 
-				% TODO: mention that using the description (not specification) language, one can now specify linearly increasing curvature
-
 				One of the primary goals of this project was to explore more expressive description languages than those which are currently used in vector graphics software.
 To that end, our specification language supports specifying points, directions and curvatures at any number of positions on the curve, as well as the total length of the curve.
+Furthermore, the fairness measure ensures constant and linearly changing curvature where possible.
 This grants the user much flexibility in describing the source curve, allowing them to specify those properties that can describe the curve in the most concise fashion.
@@
 Not only does this make the curve design process less tedious, but it may also lead to better curves, since only the essential information about the curve is given to the software.
-This allows the fairness measure to choose the best possible curve, something that is not possible if lots of auxiliary information is supplied to the software in order to indirectly influence the curvature.
+This allows the fairness measure to choose the best curve, something that is not possible if lots of auxiliary information is supplied to the software in order to indirectly influence the curvature.
+
+				Curvature specifications are especially useful in conjunction with the MVC fairness measure, which selects constant or linearly changing curvature where possible.
+This way, placing curvature specifications on an otherwise unconstrained curve allows the user to precisely create spirals.
+Of course this behavior of the fairness measure also makes it possible to construct circular arcs in various ways, since they are the preferred shape of curve.
 
@@
 While fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, there are actually quite a few upsides.
-For instance, by adjusting the arc length of a curve with a point specifications on each end, one can conveniently produce curcular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced.
-				% Using common vector graphics software, one usually has to either risk an asymmetric and/or bumpy arc by specifying it using Bézier splines, or deal with some cumbersome tool for constructing elliptical arcs.
+For instance, by adjusting the arc length of a curve with a point specification item on each end, one can conveniently produce curcular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced.
 Generally speaking, having control over the arc length of the curve results in benign and predictable curve behavior and is often useful for pulling it in place, not unlike the way one would handle a mechanical spline.
@@
 
-				As a final way of demonstrating the high expressiveness, it may be noted that it is actually possible to specify segments of the Euler spiral using the specification language.
-As described in section \ref{section:fairness_measure}, the MVC fairness measure prefers constant or linearly changing curvature.
-Since every curve with linearly changing curvature is some segment of the Euler spiral, it is easy to specify those using the specification language by leaving the curve otherwise unconstrained.
+				As a final way of demonstrating the high expressiveness of the description language, it may be noted that it actually allows for the description of arbitrary segments of the Euler spiral.
+Since every segment of the Euler spiral exhibits linearly changing curvature between its start and end points, specififying these curvature values on an otherwise unconstrained curve will cause the fairness measure to select the desired segment of the Euler spiral.
 
@@
 
-				As described in section \ref{section:fairness_measure}, the MVC fairness measure the intuitive notion of smoothness very well, resulting in aesthetically pleasing curves.
+				% TODO: expand, figure out what to write here and in the section about the fairness measure
+
+				As described in section \ref{section:fairness_measure}, the MVC fairness measure captures the intuitive notion of smoothness very well, resulting in aesthetically pleasing curves.
 
@@
 
-				Even on very powerful computers, running nonlinear optimization on splines with many segments and/or polynomials of high degree is too slow for interactive curve design, with updates taking more than 5 seconds.
+				Even on very powerful computers, running nonlinear optimization on splines with many segments and/or polynomials of high degree is too slow for interactive curve design, with updates taking seconds.
 Most curves are very simple and do not need this kind of complexity, but for those that do, the bad performance can be very disruptive.
@@
 In some cases, it is possible that small changes in the description restore the ability to converge on a solution, but sometimes one simply has to start over.
-There are also cases where the optimization gets stuck in a pathological local optimum, in which case one usually has to start over aswell.
+There are also cases where the optimization gets stuck in a pathological local optimum, in which case one often has to start over aswell.
 
@@
 
-			We will now revisit the examples shown in sections \ref{section:bézier_splines} and \ref{section:spiro_splines} using the prototype curve design tool in order to see how well the description language we propose can handle them.
+			We will now revisit some of the examples shown in sections \ref{section:bézier_splines} and \ref{section:spiro_splines} using the prototype curve design tool in order to see how well the description language we propose can handle them.
 
@@
 
-			Examples 2, 3 and 4 show how to make use of the good expressiveness of the description language.
-
-			% fluttershy's mane
-			%   spirals done this way are guaranteed to be smooth and not have any unintentional bumps
-			%     since point specifications occur only at the start and end of the curve, there is no danger of introducing unintentional bumps in the middle of the curve
-			%   comparison with bézier
-			%     source image is in flash, already made of bézier splines, easy to trace
-			%     still many more nodes needed
-			%     hard to get actual, perfect spirals right
-			%     curvature incontinuities, small bumps
-			%     a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification
-			%   somewhat hampered by poor performance and optimization stability issues, concept seems very powerful though
-			% lugia profile view / lugia's back line (curled up)
-			%   profits from global fairness optimization and curvature continuity guarantees
-			%   some curves are more easily described in terms of curvature than points and/or direction
-			% example of stability issues
-
-		\subsection{Verdict}
-		\label{section:verdict}
-
-			% idea seems good, optimization issues get in the way
-			% usability of the prototype curve design tool is still bad, but that may not mean the description language is useless, maybe there are better ways than optimization to derive curves from descriptions and to fulfill criterion 6 better
-			% maybe move this to conclusion
+			Examples 2, 3 and 4 show how to make use of the high expressiveness of the description language.
+In example 2, the curve is described mainly using three curvature specification items, the point specification item and the direction specification item acting merely as a fixture for the curve, since the curvature specification items leave it unconstrained with regard to translation and rotation.
+This way, the curve is basically defined by its length and the three curvature specification items, giving the fairness measure maximal freedom controlling the curve, in this case allowing it to establish linearly changing curvature between each pair of curvature specification items.
+
+			Example 3 also starts out with a position that has all 3 types of specification items attached, but instead of specifying the curvature along the way, opts to use a point specification item to enforce the desired shape, resulting in nonlinear but extremely smooth change of curvature.
+This technique is useful in cases where either nonlinear changing curvature is desired, or where controlling the curve's shape using only curvature specifications is too tedious (since little errors in curvature specifications propagate and accumulate along the entire curve, making it hard to get the shape right).
+
+			Example 4 uses a mixture of the techniques used in examples 2 and 3, placing curvature specification items along the curve in order to not unnecessarily restrict the fairness measure's freedom and using a final point specification at the end of the curve to force it into place.
+This approach ensures that no unintentional bumps are introduced (which can happen easily when placing point specification items in the middle of a curve), while still avoiding the tedious placing of perfectly adjusted curvature specification items (which would be necessary if the final point specification item was not there).
+
+			It may also be worth estimating the amount of information given in the description.
+Recall that the curve in example 4 was already designed using Bézier splines in section \ref{section:bézier_splines}, using 7 nodes and 12 handles, and Spiro splines in section \ref{section:spiro_splines}, using 7 control points.
+Here, it is designed using 2 point specifications, 1 direction specification, and 3 curvature specifications and specified arc length.
+Unfortunately, it is somewhat difficult to directly compare these amounts since specification items use fractions of the arc length for positioning instead of node/control point ordering, as is the case with Bézier splines and Spiro splines.
+Fraction of arc length positions do of course contain more information than simple orderings, although this is arguably not the case if said positions lie exactly on the start or end of the curve.
+Furthermore, direction and curvature specification items only specify a scalar value instead of a plane vector, so could be considered to only supply half as much information as a point specification item.
+So we count 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving us a total of 11 scalar values for the entire specification, just shy of the 14 scalar values for the Spiro spline and a lot less than the 33 scalar values for the Bézier spline (where the loss of one degree of freedom for smooth nodes with colinear velocities has been taken into account).
+Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgement on the amount of information that needs to be supplied to define a specific curve.
+
+			Finally, example 5 demonstrates the ability of the MVC fairness to construct aesthetically pleasing, smooth curves.
+The indicators for the specification items were left out in this example to make it easier to judge the curve's smoothness.
 
@@
 
-		% most issues stem from the optimization approach
-		%  may make it hard to actually rate usability of the idea behind this
-		% distinguish between further research work and turning the project into practical software)	
-		% further possibilities to instantiate overall structure (different and/or combined spline primitives, constructive instead of optimization approach)
-		% adaptive line segments, adaptive bézier approximation
-		% steps that are necessary to turn this project into a useful application
-		%   make absolutely sure the concept works by testing it with actual users
-		%   performance improvements
-		%   inkscape plugin
-		% possibility to have curves without fixed length? could be useful for curves which only have specifications at the start and/or end of the curve
-		% investigate into the calculus of variations and whether there may be a possibility to construct MVC curves based on our specification format, once the format itsself has proven to be useful
-
-		%   possible performance improvements
-		%     make sure substitutions only happen where and when necessary
-		%     cache partially instantiated IpoptProblem? such that fewer substitutions have to happen when changing a single specification
-		%     use better linear solver with ipopt
-		%     properly configure ipopt options
-		%     adaptive segment density
-		%     adaptive velocity/fairness trapezoid sum
+		Having analyzed the usability both through theoretical analysis and experimentally using the prototype curve design tool, we think that our approach looks promising.
+The added expressiveness of the proposed description language has certainly proven useful, facilitating the succinct description of curves which, when using conventional curve design tools, are harder to design and/or lack smoothness due to unfitting description languages.
+
+		While we are very satisfied with the MVC fairness measure, we feel that the specification language can still be improved, for instance by making the specification of the curve's total arc length as well as the specification items' positions optional.
+Another possible extension would be to allow specification items to be slightly violated in order to improve the curve's fairness (soft specifications).
+Further experiments using improved versions of the current description language and the prototype curve design tool are needed to make an informed decision here.
+It would also be advisable to do empirical studies with users of vector graphics software to assess the usability of description languages.
+
+		We observe that the prototype curve design tool's usability is still somewhat impaired by performance and stability issues which stem from the use of optimization for solving the problem of curve derivation.
+Once the final decisions on the description language have been made, some research into how curve derivation can be made more stable and efficient is in order.
+
+		Ideally, the final step of this research would then be the incorporation of the developed curve design tool into a piece of vector graphics software.
+The vector graphics editor Inkscape would be a good candidate, mainly because it's free software and thus easy to extend, both technically and from a political/licensing perspective.
 

### 2014-02-01 idp/documents/documentation/paper.tex
restructuring

--- 2014-02-01 idp/documentation/paper.tex
@@
 					\centering
-					\includegraphics[width=\textwidth]{../resources/examples_bezier.pdf}
+					\includegraphics[width=\textwidth]{../content/common/examples_bezier.pdf}
 					\caption{Bézier Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../resources/examples_spiro.pdf}
+					\includegraphics[width=\textwidth]{../content/common/examples_spiro.pdf}
 					\caption{Spiro Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
@@
 				\centering
-				\includegraphics[width=\textwidth]{../resources/description-based_curves.pdf}
+				\includegraphics[width=\textwidth]{../content/common/description-based_curves.pdf}
 				\caption{Description-Based Curves}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../resources/subsystems.pdf}
+					\includegraphics[width=\textwidth]{../content/documentation/subsystems.pdf}
 					\caption{Subsystem Decomposition of the Kurve Project}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../resources/specification_class_diagram.pdf}
+					\includegraphics[width=\textwidth]{../content/documentation/specification_class_diagram.pdf}
 					\caption{Class Diagram of Specification-Related Classes}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../resources/optimization_class_diagram.pdf}
+					\includegraphics[width=\textwidth]{../content/documentation/optimization_class_diagram.pdf}
 					\caption{Class Diagram of Optimization-Related Classes}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../resources/ui_components_class_diagram.pdf}
+					\includegraphics[width=\textwidth]{../content/documentation/ui_components_class_diagram.pdf}
 					\caption{Class Diagram of the User Interface Components}
@@
 
-				\begin{figure}[htbpp]
+				\begin{figure}[htbp]
 					\centering
-					\includegraphics[width=\textwidth]{../resources/ui_components.pdf}
+					\includegraphics[width=\textwidth]{../content/documentation/ui_components.pdf}
 					\caption{Overview of the User Interface in Kurve}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../resources/ui_speeds.pdf}
+					\includegraphics[width=\textwidth]{../content/documentation/ui_speeds.pdf}
 					\caption{Example of a Curve that is `Overspecified' (i.e. the speed of the curve cannot be kept constant due to the high amount of constraints)}
@@
 				\centering
-				\includegraphics[width=\textwidth]{../resources/examples_prototype.pdf}
+				\includegraphics[width=\textwidth]{../content/common/examples_prototype.pdf}
 				\caption{Prototype Curve Design Tool Examples (left: Fluttershy \cite{fluttershy})}

### 2014-02-01 idp/documents/documentation/paper.tex
included some more illustrations from slides in documentation

@@
 				% TODO: mention that mvc facilitates describing circular arcs and spirals (constant and linearly changing curvature)
+				% TODO: MEC cannot handle curvature specifications
 				% TODO: add some illustrations, describe their general behaviour, smoothness
@@
 
+				\begin{figure}[htbp]
+					\centering
+					\includegraphics[width=\textwidth]{../content/documentation/fairness_mec.pdf}
+					\caption{Minimum Energy Curves}
+					\label{figure:minimum_energy_curves}
+				\end{figure}
+
 				In order to adress these issues, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
@@
 
+				\begin{figure}[htbp]
+					\centering
+					\includegraphics[width=\textwidth]{../content/documentation/fairness_mvc.pdf}
+					\caption{Minimum Variation Curves}
+					\label{figure:minimum_variation_curves}
+				\end{figure}
+
 		\subsection{Curve Derivation}
@@
 
-			% TODO: consider structuring and naming this section similarly to the slides
 			\subsubsection{Segmentation and Notation}
@@
 
-				%\begin{equation*}
-				%	\apply{\phi_i}{t} = \sum_j a_{i,j} \cdot t^j
-				%\end{equation*}
-
-				Since the polynomial spline is uniquely determined by the coefficients of each polynomial segment, the optimization domain is a vector consisting of all these coefficients.
-Let \(m\) be the number of segments that make up the polynomial spline.
+				We start by considering the segment polynomials that make up the polynomial spline we want to find using optimization.
+Let \(m\) be the number of segments that make up the polynomial spline and let \(d\) be the degree of each segment polynomial.
+Then the \(i\)th segment polynomial is given by:
+				\begin{equation*}
+					\apply{\phi_i}{t} = \sum_{j = 0}^{d} a_{i,j} \cdot t^j
+				\end{equation*}
 				We will use the functions defined in section \ref{section:parametric_plane_curves} with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline (\(\apply{\phi}{t}\)).
@@
 				That is, the segments are simply laid out consecutively in parameter space, scaled to fit the unit interval, making the parameter mapping function linear and static.
+Figure \ref{figure:linear_segmentation} shows what the segmentation and parameter mapping look like when using 3 segments.
+
+				\begin{figure}[htbp]
+					\centering
+					\includegraphics[width=100mm]{../content/common/linear_segmentation.pdf}
+					\caption{Linear Segmentation}
+					\label{figure:linear_segmentation}
+				\end{figure}
+
+				Since the polynomial spline is uniquely determined by the coefficients of each segment polynomial, the optimization domain is the vector consisting of all the coefficients \(a_{i,j}\).
 

### 2014-02-01 idp/documents/documentation/paper.tex
more restructuring

@@
 
-\input{../common/math.tex}     % some math helper macros
+\input{math.tex}               % some math helper macros
 
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/common/examples_bezier.pdf}
+					\includegraphics[width=\textwidth]{content/output/examples_bezier.pdf}
 					\caption{Bézier Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/common/examples_spiro.pdf}
+					\includegraphics[width=\textwidth]{content/output/examples_spiro.pdf}
 					\caption{Spiro Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
@@
 				\centering
-				\includegraphics[width=\textwidth]{../content/common/description-based_curves.pdf}
+				\includegraphics[width=\textwidth]{content/output/description-based_curves.pdf}
 				\caption{Description-Based Curves}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/fairness_mec.pdf}
+					\includegraphics[width=\textwidth]{content/output/fairness_mec.pdf}
 					\caption{Minimum Energy Curves}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/fairness_mvc.pdf}
+					\includegraphics[width=\textwidth]{content/output/fairness_mvc.pdf}
 					\caption{Minimum Variation Curves}
@@
 					\centering
-					\includegraphics[width=100mm]{../content/common/linear_segmentation.pdf}
+					\includegraphics[width=100mm]{content/output/linear_segmentation.pdf}
 					\caption{Linear Segmentation}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/subsystems.pdf}
+					\includegraphics[width=\textwidth]{content/output/subsystems.pdf}
 					\caption{Subsystem Decomposition of the Kurve Project}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/specification_class_diagram.pdf}
+					\includegraphics[width=\textwidth]{content/output/specification_class_diagram.pdf}
 					\caption{Class Diagram of Specification-Related Classes}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/optimization_class_diagram.pdf}
+					\includegraphics[width=\textwidth]{content/output/optimization_class_diagram.pdf}
 					\caption{Class Diagram of Optimization-Related Classes}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/ui_components_class_diagram.pdf}
+					\includegraphics[width=\textwidth]{content/output/ui_components_class_diagram.pdf}
 					\caption{Class Diagram of the User Interface Components}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/ui_components.pdf}
+					\includegraphics[width=\textwidth]{content/output/ui_components.pdf}
 					\caption{Overview of the User Interface in Kurve}
@@
 					\centering
-					\includegraphics[width=\textwidth]{../content/documentation/ui_speeds.pdf}
+					\includegraphics[width=\textwidth]{content/output/ui_speeds.pdf}
 					\caption{Example of a Curve that is `Overspecified' (i.e. the speed of the curve cannot be kept constant due to the high amount of constraints)}
@@
 				\centering
-				\includegraphics[width=\textwidth]{../content/common/examples_prototype.pdf}
+				\includegraphics[width=\textwidth]{content/output/examples_prototype.pdf}
 				\caption{Prototype Curve Design Tool Examples (left: Fluttershy \cite{fluttershy})}

### 2014-02-02 idp/documents/documentation/paper.tex
updated section on fairness measures

@@
 \usepackage{graphicx}          % enables including graphics
+\usepackage{caption}           % provides float caption formatting
 \usepackage{amsmath}           % ams math module
@@
 \widowpenalty = 10000
+
+% setup caption formatting
+\captionsetup{margin = 10pt, font = small, labelfont = bf}
 
@@
 
-				% TODO: fairness measure integral is over the arc length of the curve
-				% TODO: mention that mvc facilitates describing circular arcs and spirals (constant and linearly changing curvature)
-				% TODO: MEC cannot handle curvature specifications
-				% TODO: add some illustrations, describe their general behaviour, smoothness
-
-				Different fairness measures for plane curves have been proposed throughout history.
-Many of them set out to emulate the behavior of mechanical splines made from wood or steel, which were used for industrial design before the advent of computer aided design.
-The most direct approach of this type uses mechanics to derive a term for the energy in an idealized bent steel band to minimize, arriving at the following functional describing the minimum energy curve (MEC) \cite{paper-mec}:
+				In this section, we introduce two important fairness measures, the minimum energy curve and the minimum variation curve, both of which are characterized by an integral over the arc length of the curve.
+For the sake of simplicity, we will assume that the curve has arc length parametrization, such that the integrals are over the curve's parameter on the unit interval.
+
+				Before the advent of computer aided design, mechanical splines made from wood or steel were used in industrial design.
+Thus, people tried to emulate this behavior using mathematical curves, deriving a term for the bending energy in an idealized steel band, which is subsequently minimized.
+This leads to the following functional describing the minimum energy curve (MEC) \cite{paper-mec}, which tries to minimize the curvature, and thus the bending energy:
 				\begin{equation*}
@@
 				\end{equation*}
-				While this approach is a good idealization of the mechanical spline, it has some disadvantages when used for curve design.
+
+				While the minimum energy curve is a good idealization of the mechanical spline, it has some disadvantages when used for curve design.
 For instance, it will not yield a circular arc when the curve is constrained by co-circular points.
 Generally speaking, the MEC fairness measure is not very sensitive to bumps and inflection points on the curve, as long as they're small.
-The human visual system on the other hand, very much is and would always prefer one large circular arc over five small bumps on the curve.
+The human visual system on the other hand, very much is and would always prefer one large arc over multiple small ones.
+Since the MEC tries to minimize curvature, it is also unsuited for the use with curvature specifications, as it will try to bring the curvature back to zero before and after a nonzero curvature specification.
 
@@
 					\includegraphics[width=\textwidth]{content/output/fairness_mec.pdf}
-					\caption{Minimum Energy Curves}
-					\label{figure:minimum_energy_curves}
-				\end{figure}
-
-				In order to adress these issues, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
-One such approach is that of the minimum variation curve (MVC) \cite{thesis-mvc}, whose functional is given below:
+					\caption{Minimum Energy Curve Examples. Point specification items are represented as red dots, direction specification items are represented as green tangent lines.}
+					\label{figure:minimum_energy_curve_examples}
+				\end{figure}
+
+				Figure \ref{figure:minimum_energy_curve_examples} shows a few examples of minimum energy curves under various constraints.
+Note the characteristic shape of the MEC (rectangular elastica), as well as the fact that, in absence of direction or curvature specification items, its curvature is always zero at the endpoints.
+
+				In order to address the issues of the MEC, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
+One such approach is that of the minimum variation curve (MVC) \cite{thesis-mvc}, which tries to minimize the variation of the curvature.
+The corresponding functional is given below:
 				\begin{equation*}
@@
 				\end{equation*}
+
 				Minimum variation curves capture the intuitive notion of smoothness very well.
 The functional will put straight lines through points that are co-linear and circular arcs through points that are co-circular when otherwise unconstrained.
-When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvature, since that will make the derivative of the curvature a constant and thereby minimize the integral over its square.
+When constraints cause different points on the curve to have different curvatures, the fairness measure will prefer linearly changing curvature, since that will make the derivative of the curvature a constant and thereby minimize the integral over its square.
 In general, the MVC fairness measure is very good at avoiding bumps and inflection points when possible.
-Because of these appealing properties, the MVC fairness measure was chosen as part of our description language.
+It also works nicely with curvature constraints, since constant curvature is perfectly acceptable for the MVC fairness measure.
 
@@
 					\includegraphics[width=\textwidth]{content/output/fairness_mvc.pdf}
-					\caption{Minimum Variation Curves}
-					\label{figure:minimum_variation_curves}
-				\end{figure}
+					\caption{Minimum Variation Curve Examples. Point specification items are represented as red dots, direction specification items are represented as green tangent lines.}
+					\label{figure:minimum_variation_curve_examples}
+				\end{figure}
+
+				Figure \ref{figure:minimum_variation_curve_examples} shows a few examples of minimum variation curves under the same constraints as the curves in figure \ref{figure:minimum_energy_curve_examples}.
+Note that instead of segments from the rectangular elastica, the MVC's characteristic shapes are circular arcs and spirals.
+One may also observe that, instead of trying to bring the curvature to zero towards the endpoints like the MEC does, the MVC tries to establish constant curvature.
+In contrast to the MEC, the MVC constructs a circular arc in the leftmost example and avoids the inflection point in the rightmost example.
+
+				From these considerations, it seems clear that for use as a fairness measure in our approach to curve design, the minimum variation curve is much better suited than the minimum energy curve, so it was chosen as part of our description language.
 
@@
 
-			% TODO: add illustration to section introducing specification language, add reference to it here
 			Figure \ref{figure:examples_prototype} shows some examples of curves designed using the prototype curve design tool.
-Like in section 1.2.3, point specification items are represented using red dots, direction specification items are represented using green tangent lines and curvature specification items are represented using blue circular arcs taken from the osculating circle.
+Point specification items are represented using red dots, direction specification items are represented using green tangent lines and curvature specification items are represented using blue circular arcs taken from the osculating circle.
 

### 2014-02-02 idp/documents/documentation/paper.tex
worked on a few TODOs, fixed some issues with floats

@@
 
-% TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
 % TODO: expand n't and *'s contractions
@@
 
-				\begin{figure}[htbp]
+				\begin{figure}[htb]
 					\centering
@@
 
-			\begin{figure}[htbp]
+			\begin{figure}[htb]
 				\centering
@@
 
-				% TODO: explain this in more detail (o--o----------------o, pull last specification, makes wobbly segment at beginning) 
 				With this setup, it seems advisable to also include the curve's total arc length in the specification.
@@
 Also, with unspecified curve length, changing any specification item may inadvertently change the curve's total arc length, thus possibly adversely affecting other parts of the curve whose specification item positions are given as a fraction of the curve's total arc length.
+For instance, with three point specification items on the curve, pulling the third one further away from the second will increase the total arc length of the curve, thus also increasing the arc length of the curve segment between the first and the second point specification item, thereby unintentionally creating a bulge between those two specification items.
 
@@
 
-			% TODO: make it clear that splines imply segmentation
 			Polynomial curves were chosen since they are mathematically simple, very flexible through the choice of their degree, and well-suited to approximate a variety of smooth curves.
-We opted to use polynomial splines since they allow approximating more complex curves without involving polynomials of excessively high degree, which are known to have poor stability.
+Since polynomials of high degree are known to have poor stability, we opted to use polynomial splines, which allows us to increase the segment count, rather than the degree, in order to approximate more complex curves.
 
@@
 
-				% TODO: sketch overall curve derivation process similar to the summary slide
+				Once the optimization problem is put together, it is passed to the opimization solver, which will return the optimized coefficients of the segment polynomials.
+These segment polynomials are then assembled according to the layout described in section \ref{section:segmentation_notation} in order to obtain the final curve.
 
@@
 			\subsection{Specification}
-				\begin{figure}[htbp]
+				\begin{figure}[htb]
 					\centering
@@
 			\subsection{Optimization}
-				\begin{figure}[htbp]
+				\begin{figure}[htb]
 					\centering
@@
 			\subsection{Implementation}
-				\begin{figure}[htbp]
+				\begin{figure}[htb]
 					\centering
@@
 
-				\begin{figure}[htbp]
+				% TODO: make this smaller, maybe use landscape format
+				\begin{figure}[p]
 					\centering

### 2014-02-02 idp/documents/documentation/paper.tex
started working with subfloats

@@
 \usepackage{caption}           % provides float caption formatting
+\usepackage{subcaption}        % provides subfloat functionality and caption formatting
 \usepackage{amsmath}           % ams math module
@@
 					\centering
-					\includegraphics[width=\textwidth]{content/output/examples_bezier.pdf}
-					\caption{Bézier Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
+					\begin{subfigure}{\textwidth}
+						\includegraphics[width=\textwidth]{content/output/examples_bezier_arc.pdf}
+						\caption{Circular Arc}
+						\label{figure:examples_bézier_arc}
+					\end{subfigure}
+					\begin{subfigure}[b]{0.5 \textwidth}
+						\includegraphics[width=\textwidth]{content/output/examples_bezier_fluttershy.pdf}
+						\caption{Fluttershy \cite{fluttershy}}
+						\label{figure:examples_bézier_fluttershy}
+					\end{subfigure}%
+					\begin{subfigure}[b]{0.5 \textwidth}
+						\includegraphics[width=\textwidth]{content/output/examples_bezier_lugia.pdf}
+						\caption{Lugia \cite{lugia}}
+						\label{figure:examples_bézier_lugia}
+					\end{subfigure}
+					\caption{Bézier splines, as they appear in common curve design tools. Bézier nodes are displayed as red dots, Bézier handles are displayed as green tangent lines.}
 					\label{figure:examples_bézier}
@@
 
-				Figure \ref{figure:examples_bézier} illustrates some of these issues by showing some examples of Bézier splines as they appear in common design tools.
-Bézier nodes are displayed as red dots, while Bézier handles are displayed as green lines.
-
-				Example 1 shows two Bézier splines.
+				Figure \ref{figure:examples_bézier_arc} shows two Bézier splines.
 The left curve is a Bézier spline approximation of a circular arc, exhibiting nearly constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
@@
 
-				Examples 2 and 3 show curves that could be described more efficiently in terms of curvature.
+				Figures \ref{figure:examples_bézier_fluttershy} and \ref{figure:examples_bézier_lugia} show curves that could be described more efficiently in terms of curvature.
 Since there is no way to directly specify curvature using Bézier splines, a considerable number of nodes is needed to get close to the desired curve.
@@
 
-				Example 4 shows a spline with a curvature discontinuity which went unnoticed by the artist at the time of creation.
+				The red curve segment in figure \ref{figure:examples_bézier_lugia} is a Bézier spline with a curvature discontinuity which went unnoticed by the artist at the time of creation.
 In fact, this discontinuity would be even less visible if the curve was not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.
@@
 					\includegraphics[width=\textwidth]{content/output/fairness_mec.pdf}
-					\caption{Minimum Energy Curve Examples. Point specification items are represented as red dots, direction specification items are represented as green tangent lines.}
+					\caption{Minimum Energy Curve Examples. Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.}
 					\label{figure:minimum_energy_curve_examples}
@@
 					\includegraphics[width=\textwidth]{content/output/fairness_mvc.pdf}
-					\caption{Minimum Variation Curve Examples. Point specification items are represented as red dots, direction specification items are represented as green tangent lines.}
+					\caption{Minimum Variation Curve Examples. Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.}
 					\label{figure:minimum_variation_curve_examples}
@@
 			Figure \ref{figure:examples_prototype} shows some examples of curves designed using the prototype curve design tool.
-Point specification items are represented using red dots, direction specification items are represented using green tangent lines and curvature specification items are represented using blue circular arcs taken from the osculating circle.
+Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle.
 

### 2014-02-02 idp/documents/documentation/paper.tex
added more subfloats

@@
 					\centering
-					\begin{subfigure}{\textwidth}
+					\begin{subfigure}[b]{\textwidth / 1}
 						\includegraphics[width=\textwidth]{content/output/examples_bezier_arc.pdf}
@@
 					\end{subfigure}
-					\begin{subfigure}[b]{0.5 \textwidth}
+					\begin{subfigure}[b]{\textwidth / 2}
 						\includegraphics[width=\textwidth]{content/output/examples_bezier_fluttershy.pdf}
@@
 					\end{subfigure}%
-					\begin{subfigure}[b]{0.5 \textwidth}
+					\begin{subfigure}[b]{\textwidth / 2}
 						\includegraphics[width=\textwidth]{content/output/examples_bezier_lugia.pdf}
@@
 					\centering
-					\includegraphics[width=\textwidth]{content/output/examples_spiro.pdf}
-					\caption{Spiro Spline Examples (left: Fluttershy \cite{fluttershy}, right: Lugia \cite{lugia})}
+					\begin{subfigure}[b]{\textwidth / 2}
+						\includegraphics[width=\textwidth]{content/output/examples_spiro_fluttershy.pdf}
+						\caption{Fluttershy \cite{fluttershy}}
+						\label{figure:examples_spiro_fluttershy}
+					\end{subfigure}%
+					\begin{minipage}[b]{\textwidth / 2}
+						\begin{subfigure}[b]{\textwidth}
+							\includegraphics[width=\textwidth]{content/output/examples_spiro_lugia_1.pdf}
+							\caption{Lugia \cite{lugia}}
+							\label{figure:examples_spiro_lugia_1}
+						\end{subfigure}
+						\begin{subfigure}[b]{\textwidth}
+							\includegraphics[width=\textwidth]{content/output/examples_spiro_lugia_2.pdf}
+							\caption{Lugia \cite{lugia}}
+							\label{figure:examples_spiro_lugia_2}
+						\end{subfigure}
+					\end{minipage}
+					\caption{Spiro splines, as they appear in common curve design tools. Control points are displayed as red dots.}
 					\label{figure:examples_spiro}
@@
 
-				Figure \ref{figure:examples_spiro} illustrates some of these issues by showing some examples of Spiro splines as they appear in common design tools. Control points are displayed as red dots.
-
-				Example 1 shows a shape that is in principle well-suited for being represented by parts of the Euler spiral.
+				Figure \ref{figure:examples_spiro_fluttershy} shows a shape that is in principle well-suited for being represented by parts of the Euler spiral.
 Unfortunately, describing the curve using Spiro splines still requires many control points.
@@
 
-				Example 2 shows how some shapes can be described using Spiro splines by placing only very few control points.
-In this case, the shape consists of a circular arc with low curvature (segment a), a segment of linearly increasing curvature (segment b) and finally a circular arc with high curvature (segment c).
+				Figure \ref{figure:examples_spiro_lugia_1} shows how some shapes can be described using Spiro splines by placing only very few control points.
+In this case, the shape consists of a circular arc with low curvature (red), a segment of linearly increasing curvature (green) and finally a circular arc with high curvature (blue).
 This shape can be described conveniently using a Spiro spline with just 4 carefully placed control points.
 
-				Unfortunately, as can be seen in example 3, this scenario does not handle modifications very well.
-For instance, increasing the curvature of segment c has to be done indirectly by moving a control point.
-Unfortunately, this adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
+				Unfortunately, as can be seen in figure \ref{figure:examples_spiro_lugia_2}, this scenario does not handle modifications very well.
+For instance, increasing the curvature of the blue segment has to be done indirectly by moving a control point.
+Unfortunately, this adversely affects the red segment, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
 

### 2014-02-02 idp/documents/documentation/paper.tex
more subfloats

@@
 				\centering
-				\includegraphics[width=\textwidth]{content/output/examples_prototype.pdf}
-				\caption{Prototype Curve Design Tool Examples (left: Fluttershy \cite{fluttershy})}
+				\begin{subfigure}[b]{8 \textwidth / 13}
+					\includegraphics[width=\textwidth]{content/output/examples_prototype_fluttershy.pdf}
+					\caption{Fluttershy \cite{fluttershy}}
+					\label{figure:examples_prototype_fluttershy}
+				\end{subfigure}%
+				\begin{subfigure}[b]{5 \textwidth / 13}
+					\includegraphics[width=\textwidth]{content/output/examples_prototype_fairness.pdf}
+					\caption{Fairness Measure Example}
+					\label{figure:examples_prototype_fairness}
+				\end{subfigure}
+				\caption{Examples of curves designed using the prototype curve design tool.
+Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle.}
 				\label{figure:examples_prototype}
@@
 
-			Figure \ref{figure:examples_prototype} shows some examples of curves designed using the prototype curve design tool.
-Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle.
-
-			Example 1 shows how adjusting the arc length of an otherwise unconstrained curve allows the convenient creation of curcular arcs between two points.
+			Figure \ref{figure:examples_prototype_fluttershy} shows a few curves designed using the prototype curve design tool.
+Example 1 demonstrates how adjusting the arc length of an otherwise unconstrained curve allows the convenient creation of curcular arcs between two points.
 
@@
 
-			Finally, example 5 demonstrates the ability of the MVC fairness to construct aesthetically pleasing, smooth curves.
+			Finally, figure \ref{figure:examples_prototype_fairness} demonstrates the ability of the MVC fairness to construct aesthetically pleasing, smooth curves.
 The indicators for the specification items were left out in this example to make it easier to judge the curve's smoothness.

### 2014-02-02 idp/documents/documentation/paper.tex
split more illustrations into subfloats

@@
 					\centering
-					\includegraphics[width=\textwidth]{content/output/fairness_mec.pdf}
+					\begin{subfigure}[b]{\textwidth / 3}
+						\includegraphics[width=\textwidth]{content/output/fairness_mec_1.pdf}
+					\end{subfigure}%
+					\begin{subfigure}[b]{\textwidth / 3}
+						\includegraphics[width=\textwidth]{content/output/fairness_mec_2.pdf}
+					\end{subfigure}%
+					\begin{subfigure}[b]{\textwidth / 3}
+						\includegraphics[width=\textwidth]{content/output/fairness_mec_3.pdf}
+					\end{subfigure}
 					\caption{Minimum Energy Curve Examples. Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.}
@@
 					\centering
-					\includegraphics[width=\textwidth]{content/output/fairness_mvc.pdf}
+					\begin{subfigure}[b]{\textwidth / 3}
+						\includegraphics[width=\textwidth]{content/output/fairness_mvc_1.pdf}
+					\end{subfigure}%
+					\begin{subfigure}[b]{\textwidth / 3}
+						\includegraphics[width=\textwidth]{content/output/fairness_mvc_2.pdf}
+					\end{subfigure}%
+					\begin{subfigure}[b]{\textwidth / 3}
+						\includegraphics[width=\textwidth]{content/output/fairness_mvc_3.pdf}
+					\end{subfigure}
 					\caption{Minimum Variation Curve Examples. Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.}

### 2014-02-02 idp/documents/documentation/paper.tex
fixed minor mistakes, added abstract

@@
 % TODO: check if all quotations are done `like this'
-% TODO: consistent use of polynomial, planar, plane, parametric, curve
 
@@
 
-		\noindent TODO
+		\noindent We analyze the usability of curve design tools in current vector graphics software and suggest an abstract approach for designing improved, usability-oriented tools.
+We then instantiate this approach using a powerful description language and the minimum variation curve fairness measure on top of nonlinear optimization, developing a prototype curve design tool in the process.
+Finally, we evaluate this prototype's usability and suggest ideas for further research and development.
 
@@
 				That is, the segments are simply laid out consecutively in parameter space, scaled to fit the unit interval, making the parameter mapping function linear and static.
+Note that these segments are independent of the specification items and are, in fact, completely invisible to the user.
 Figure \ref{figure:linear_segmentation} shows what the segmentation and parameter mapping look like when using 3 segments.
@@
 				\begin{itemize}
+					% TODO: this section could be shortened by simply stating that Ipopt can solve optimization problems as described in section \ref{section:optimization_problem}
 				  	\item Ipopt\footnote{https://projects.coin-or.org/Ipopt} is a library that does numeric non-linear optimization.
@@
 
-				% TODO: expand, figure out what to write here and in the section about the fairness measure
-
 				As described in section \ref{section:fairness_measure}, the MVC fairness measure captures the intuitive notion of smoothness very well, resulting in aesthetically pleasing curves.
@@
 
-	\clearpage
-
 	\appendix
@@
 
-	\clearpage
-
 	\begin{thebibliography}{}

### 2014-02-02 idp/documents/documentation/paper.tex
fixed a few minor mistakes

@@
 We then instantiate this approach using a powerful description language and the minimum variation curve fairness measure on top of nonlinear optimization, developing a prototype curve design tool in the process.
-Finally, we evaluate this prototype's usability and suggest ideas for further research and development.
+Finally, we evaluate the usability of our approach based on the prototype and suggest ideas for further research and development.
 
@@
 					\end{subfigure}
-					\caption{Minimum Energy Curve Examples. Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.}
+					\caption{Examples of minimum energy curves under various constraints.
+Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.
+The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
 					\label{figure:minimum_energy_curve_examples}
@@
 
-				Figure \ref{figure:minimum_energy_curve_examples} shows a few examples of minimum energy curves under various constraints.
-Note the characteristic shape of the MEC (rectangular elastica), as well as the fact that, in absence of direction or curvature specification items, its curvature is always zero at the endpoints.
+				Note the characteristic rectangular elastica shapes of the minimum energy curves in figure \ref{figure:minimum_energy_curve_examples}, as well as the fact that, in absence of both direction and curvature specification items, the curvature of a minimum energy curve is always zero at the endpoints.
 
@@
 					\end{subfigure}
-					\caption{Minimum Variation Curve Examples. Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.}
+					\caption{Examples of minimum variation curves under the same constraints as the curves in figure \ref{figure:minimum_energy_curve_examples}.
+Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.
+The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
 					\label{figure:minimum_variation_curve_examples}
@@
 
-				Figure \ref{figure:minimum_variation_curve_examples} shows a few examples of minimum variation curves under the same constraints as the curves in figure \ref{figure:minimum_energy_curve_examples}.
-Note that instead of segments from the rectangular elastica, the MVC's characteristic shapes are circular arcs and spirals.
+				Note that instead of segments from the rectangular elastica, the characteristic shapes of the minimum variation curves in figure \ref{figure:minimum_variation_curve_examples} are circular arcs and spirals.
 One may also observe that, instead of trying to bring the curvature to zero towards the endpoints like the MEC does, the MVC tries to establish constant curvature.
@@
 				\caption{Examples of curves designed using the prototype curve design tool.
-Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle.}
+Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle, or tangent lines if the curve is locally a straght line.
+The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
 				\label{figure:examples_prototype}

### 2014-02-02 idp/documents/documentation/paper.tex
fixed some spelling errors

@@
 
-% TODO: expand n't and *'s contractions
-% TODO: check if all quotations are done `like this'
-
 \begin{document}
@@
 		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings so as to find a way to improve it without making the same or similar mistakes.
-In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves.
+In the context of curve design tools, this can be achieved by analyzing and modeling the process that people use when designing curves.
 Once such a model has been obtained, it can be used to propose criteria for good curve design tools as well as to analyze usability aspects of existing tools.
@@
 Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve, in step 4, the design tool will have to choose the curve that the user was most likely describing from a possibly infinite set of curves that exhibit the supplied properties.
-This notion of `most likely' is usually modelled in terms of a fairness measure, fairer curves being regarded as more likely.
-Fairness ususally encodes intuitive notions such as smoothness and minimality using mathematical concepts such as continuity and variation of curvature.
+This notion of `most likely' is usually modeled in terms of a fairness measure, fairer curves being regarded as more likely.
+Fairness usually encodes intuitive notions such as smoothness and minimality using mathematical concepts such as continuity and variation of curvature.
 Here, the software has another opportunity to influence the design process both positively and negatively depending on what fairness measure is chosen and how good the software is at reliably finding a curve that has a high fairness.
@@
 			Note that there may be multiple ways to describe a given curve design tool in terms of this process.
-Instead of modelling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify.
+Instead of modeling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify.
 For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve.
 Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \xp{p_1 - p_0} t\), thus uniquely identifying the desired curve.
-Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
+Modeling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.
 
@@
 
-				When applying the model introduced in section \ref{section:curve_design_process} to Bézier spline design tools, there are two possibilities for modelling the fairness measure.
+				When applying the model introduced in section \ref{section:curve_design_process} to Bézier spline design tools, there are two possibilities for modeling the fairness measure.
 One way to view things is to assume that the user, in supplying the four coefficient points of each curve segment, uniquely specifies each segment's Bézier curve and thus the entire spline, leaving no room for choosing a curve according to some fairness measure.
@@
 Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created or moved in order to change the overall curve.
-It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes.
+It also sacrifices smoothness, since it's easy to accidentally introduce subtle bumps in the curve when using a large number of nodes.
 Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous.
@@
 				Again, the model from section \ref{section:curve_design_process} is applied.
-With Spiro splines, the modelling is fairly straightforward.
+With Spiro splines, the modeling is fairly straightforward.
 The user supplies a set of points for which the software computes an interpolating spline.
@@
 The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them are needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
-The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criteria 4 and 5).
+The fairness measure leads to beautifully smooth curves that are usually sufficiently close to what a human would expect given the control points (criteria 4 and 5).
 The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 6 well enough.
@@
 
-				Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability.
+				Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversely affects usability.
 Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature.
 These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
-This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to the expression of circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
+This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itself excellently to the expression of circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
 Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties.
@@
 These issues are somewhat alleviated by the guaranteed curvature continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
-That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
+That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidentally create small bumps as with Bézier splines.
 
@@
 Let \(\hat{t}_i\) be the position of the \(i\)th specification item.
-If the \(i\)th item is a point specification, the point is denoted \(\hat{\phi}_i\), if it is a direction specification, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification, the curvature is doneted \(\hat{\chi}_i\).
+If the \(i\)th item is a point specification, the point is denoted \(\hat{\phi}_i\), if it is a direction specification, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification, the curvature is denoted \(\hat{\chi}_i\).
 The specified total arc length for the curve is denoted \(\hat{\lambda}\).
@@
 
-				Once the optimization problem is put together, it is passed to the opimization solver, which will return the optimized coefficients of the segment polynomials.
+				Once the optimization problem is put together, it is passed to the optimization solver, which will return the optimized coefficients of the segment polynomials.
 These segment polynomials are then assembled according to the layout described in section \ref{section:segmentation_notation} in order to obtain the final curve.
@@
 
-				The optimization process is perfomed in a number of steps. Depending on the change of the specification, only certain steps have to be run again as previous results can be reused.
+				The optimization process is performed in a number of steps. Depending on the change of the specification, only certain steps have to be run again as previous results can be reused.
 
@@
 This allows the reuse of the optimization problem if curve specifications or curve length change; the derivatives do not have to be recomputed.
-Thus, the optimization problem is cached and only rebuilt if the optimiztion segments change or if the number of each type of constraint on the curve changes.
+Thus, the optimization problem is cached and only rebuilt if the optimization segments change or if the number of each type of constraint on the curve changes.
 If the value of a constraint changes, but the type of constraint remains the same (e.g.
@@
 However, CasADi is used for representing Curves internally, as well as for performing optimization tasks.
-					As Kurve uses a background thread to optimize curves, it is not possible to use the Curve objects yieded by the optimizer to render the curves in the main thread at the same time.
+					As Kurve uses a background thread to optimize curves, it is not possible to use the Curve objects yielded by the optimizer to render the curves in the main thread at the same time.
 For that reason, Curve objects are converted to DiscreteCurve objects in that background process.
@@
 For that reason, most actions are performed using keyboard shortcuts.
-Many user interface elments can be selected with the mouse; this enables to use keyboard shortcuts that affect that component only.
+Many user interface elements can be selected with the mouse; this enables to use keyboard shortcuts that affect that component only.
 It is possible to interact with some components using the mouse.
@@
 					To save all curves in the window, press `S'.
-A save file dialog will appear that allows the user to specify a filename.
+A save file dialog will appear that allows the user to specify a file name.
 Saved files can later be loaded by pressing `L' which displays an open file dialog.
@@
 					\includegraphics[width=\textwidth]{content/output/ui_speeds.pdf}
-					\caption{Example of a Curve that is `Overspecified' (i.e. the speed of the curve cannot be kept constant due to the high amount of constraints)}
+					\caption{Example of a Curve that is `overspecified' (i.e. the speed of the curve cannot be kept constant due to the high amount of constraints)}
 					\label{figure:ui_speeds}
@@
 						\begin{description}
-							\item[Right-Click] Add empty specifcation control at that position
-							\item[Drag] Moves any point specifications delimiting the segment. (Only possible if at leats one of the specification controls has the point specification enabled.)
+							\item[Right-Click] Add empty specification control at that position
+							\item[Drag] Moves any point specifications delimiting the segment. (Only possible if at least one of the specification controls has the point specification enabled.)
 						\end{description}
@@
 
-				The ability to specify curvature constitues a significant improvement over common curve design tools, where this is usually not possible.
+				The ability to specify curvature constitutes a significant improvement over common curve design tools, where this is usually not possible.
 Using this functionality, the user can specify curves which are best expressed in terms of curvature in a natural and concise way.
@@
 While fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, there are actually quite a few upsides.
-For instance, by adjusting the arc length of a curve with a point specification item on each end, one can conveniently produce curcular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced.
+For instance, by adjusting the arc length of a curve with a point specification item on each end, one can conveniently produce circular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced.
 Generally speaking, having control over the arc length of the curve results in benign and predictable curve behavior and is often useful for pulling it in place, not unlike the way one would handle a mechanical spline.
@@
 				As a final way of demonstrating the high expressiveness of the description language, it may be noted that it actually allows for the description of arbitrary segments of the Euler spiral.
-Since every segment of the Euler spiral exhibits linearly changing curvature between its start and end points, specififying these curvature values on an otherwise unconstrained curve will cause the fairness measure to select the desired segment of the Euler spiral.
+Since every segment of the Euler spiral exhibits linearly changing curvature between its start and end points, specifying these curvature values on an otherwise unconstrained curve will cause the fairness measure to select the desired segment of the Euler spiral.
 
@@
 In some cases, it is possible that small changes in the description restore the ability to converge on a solution, but sometimes one simply has to start over.
-There are also cases where the optimization gets stuck in a pathological local optimum, in which case one often has to start over aswell.
+There are also cases where the optimization gets stuck in a pathological local optimum, in which case one often has to start over as well.
 
@@
 				\caption{Examples of curves designed using the prototype curve design tool.
-Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle, or tangent lines if the curve is locally a straght line.
+Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle, or tangent lines if the curve is locally a straight line.
 The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
@@
 			Figure \ref{figure:examples_prototype_fluttershy} shows a few curves designed using the prototype curve design tool.
-Example 1 demonstrates how adjusting the arc length of an otherwise unconstrained curve allows the convenient creation of curcular arcs between two points.
+Example 1 demonstrates how adjusting the arc length of an otherwise unconstrained curve allows the convenient creation of circular arcs between two points.
 
@@
 So we count 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving us a total of 11 scalar values for the entire specification, just shy of the 14 scalar values for the Spiro spline and a lot less than the 33 scalar values for the Bézier spline (where the loss of one degree of freedom for smooth nodes with colinear velocities has been taken into account).
-Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgement on the amount of information that needs to be supplied to define a specific curve.
+Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgment on the amount of information that needs to be supplied to define a specific curve.
 
@@
 			It follows immediately that every parametric curve that is \(\contp{n}\) continuous is also \(\contg{n}\) continuous.
-One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itsself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve.
+One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve.
 In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, we can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.

### 2014-02-02 idp/documents/documentation/paper.tex
fixed minor mistakes

@@
 We then instantiate this approach using a powerful description language and the minimum variation curve fairness measure on top of nonlinear optimization, developing a prototype curve design tool in the process.
-Finally, we evaluate the usability of our approach based on the prototype and suggest ideas for further research and development.
+Finally, we evaluate our approach and its instantiation based on the prototype and suggest ideas for further research and development.
 
@@
 The user is often unable to make the curve look the way they want it to.
-To make matters worse it is often unclear why this is the case, as the software seems to put all the necessary tools into the hands of the designer, yet at times, they somehow don't allow the user to communicate their intention to the software.
+To make matters worse it is often unclear why this is the case, as the software seems to put all the necessary tools into the hands of the designer, yet at times, they somehow do not allow the user to communicate their intention to the software.
 
@@
 Note that while syntactically complete, the semantics of the specification language also depend on the fairness measure.
-Thus, it makes sense to consider the combination of a specification language and a fairness measure, which we will call a description language.
+Thus, it makes sense to consider the combination of a specification language and a fairness measure, which will be called a description language.
 
@@
 Bézier splines are furthermore very simple to derive from their description, fulfilling criterion 6 perfectly.
-Design tools for Bézier splines have some issues with both the fairness measure's smoothness (criterion 4) and the expressiveness of the description language (criterion 1), both of which shall be analyzed in the following paragraphs.
+We think that design tools for Bézier splines have some issues with both the fairness measure's smoothness (criterion 4) and the expressiveness of the description language (criterion 1), both of which shall be analyzed in the following paragraphs.
 
@@
 		Having analyzed the usability aspect of curve design tools in section \ref{section:problem_analysis}, it's time to think about how to properly address their current issues.
-In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as our specific ideas for developing a piece of software using this approach.
+In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as specific ideas for developing a piece of software using this approach.
 
@@
 				An obvious choice for specifying a position on a curve would be by giving a parameter value for the underlying parametric curve.
-However, this would violate our design idea of using differential geometric properties of curves.
+However, this would violate the design idea of using differential geometric properties of curves.
 A way of specifying a position on a curve that is independent from its parametrization is the arc length of the curve.
@@
 
-				In this section, we introduce two important fairness measures, the minimum energy curve and the minimum variation curve, both of which are characterized by an integral over the arc length of the curve.
-For the sake of simplicity, we will assume that the curve has arc length parametrization, such that the integrals are over the curve's parameter on the unit interval.
+				In this section, two important fairness measures are introduced, the minimum energy curve and the minimum variation curve, both of which are characterized by an integral over the arc length of the curve.
+For the sake of simplicity, the curve will be assumed to have arc length parametrization, such that the integrals are over the curve's parameter on the unit interval.
 
@@
 
-				From these considerations, it seems clear that for use as a fairness measure in our approach to curve design, the minimum variation curve is much better suited than the minimum energy curve, so it was chosen as part of our description language.
+				We think that the minimum variation curve fairness measure is better at producing curves that appear smooth to the human visual system than the minimum energy curve fairness measure.
+Thus, it produces curves that both look good and are close to what the user expects, given the specification, causing less surprise and frustration when designing curves.
+Furthermore, it complements the chosen specification language nicely, since, in contrast to the minimum energy curve fairness measure, it works well with curvature specifications.
+Given these considerations, we have chosen to use the minimum variation curve fairness measure.
 
@@
 
-			Nonlinear optimization allowed us to stay relatively flexible with respect to the description language and enabled us to experiment with different specification languages and fairness measures, assessing their usability before settling on the choices presented in section \ref{section:description_language}.
+			Nonlinear optimization grants some flexibility with respect to the description language, enabling experimentation with different specification languages and fairness measures.
+This made it possible to assess their usability before settling on the choices presented in section \ref{section:description_language}.
 
 			Polynomial curves were chosen since they are mathematically simple, very flexible through the choice of their degree, and well-suited to approximate a variety of smooth curves.
-Since polynomials of high degree are known to have poor stability, we opted to use polynomial splines, which allows us to increase the segment count, rather than the degree, in order to approximate more complex curves.
+Since polynomials of high degree are known to have poor stability, polynomial splines were used, which allows increasing the segment count, rather than the degree, in order to approximate more complex curves.
 
@@
 				\end{equation*}
-				This allows for the expression of a variety of optimization goals.
-The ones that are important for us are zero-error constraints (using the constraint function and bounds to specify that an error term should be zero), upper bounds on errors (using the constraint function and bounds to specify an upper bound for an error term) and minimization goals (using the objective function to express that an error term should be minimized).
-
-				Our implementation of choice uses smooth optimization, or, more specifically, an interior point method.
+				This allows for the expression of a variety of optimization goals, like zero-error constraints (using the constraint function and bounds to specify that an error term should be zero), upper bounds on errors (using the constraint function and bounds to specify an upper bound for an error term) and minimization goals (using the objective function to express that an error term should be minimized).
+
+				The implementation we chose uses smooth optimization, or, more specifically, an interior point method.
 This type of algorithm further requires the functions \(f\) and \(g\) to be at least once, ideally twice continuously differentiable.
 
-				In the following, we will describe how nonlinear optimization can be used for finding polynomial splines based on descriptions in the language established in section \ref{section:description_language}.
+				The following sections describe how nonlinear optimization can be used for finding polynomial splines based on descriptions in the language established in section \ref{section:description_language}.
 
@@
 
-				We start by considering the segment polynomials that make up the polynomial spline we want to find using optimization.
+				The first step is to consider the segment polynomials that make up the polynomial spline used in the optimization.
 Let \(m\) be the number of segments that make up the polynomial spline and let \(d\) be the degree of each segment polynomial.
@@
 				\end{equation*}
-				We will use the functions defined in section \ref{section:parametric_plane_curves} with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline (\(\apply{\phi}{t}\)).
+				The functions defined in section \ref{section:parametric_plane_curves} will be used with subscript \(i\) when referring to the \(i\)th segment (\(\apply{\phi_i}{t}\)) and without subscript when referring to the whole spline (\(\apply{\phi}{t}\)).
 Mapping the parameter \(t\) between the segments and the whole spline works as follows:
@@
 
-				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need the function mapping the covered arc length to the corresponding parameter values in order to state the optimization goals for the specification items.
+				Since the position of a specification item is given as a fraction of the curve's total arc length, the function mapping the covered arc length to the corresponding parameter values is needed in order to state the optimization goals for the specification items.
 Unfortunately, this function is not very well-suited to be used as a term in a smooth optimization problem even for simple curves like polynomials.
 Also, since the function isn't fixed and varies with the coefficients, there is no way of telling on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
-We also need to state the specification of the total arc length of the curve as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
+The specification of the total arc length of the curve also needs to be stated as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
 Furthermore, the difficulty of expressing the fairness measure, which is often an integral over the arc length of the curve, is dependent on the complexity of the covered arc length function.
@@
 				The error term given above is not well-suited for smooth optimization because it incorporates finding a maximum.
-Since the optimization problem will allow the original error term to be brought close to zero, we can approximate it using the following error term:
+Since the optimization problem will allow the original error term to be brought close to zero, it can be approximated using the following error term:
 				\begin{equation*}
@@
 				As mentioned in section \ref{section:description_language}, we are not concerned with properties beyond the second derivative of the curve.
-Thus, we will only require \(\contg{2}\) continuity for the whole spline.
+Thus, only \(\contg{2}\) continuity will be required.
 While the polynomial curves used as segments are smooth (\(\contp{\infty}\) continuity), the resulting spline in general is not.
@@
 				Having taken care of the somewhat technical details arising from using polynomial splines, it's time to state the error terms for the specification items.
-Since the optimization goal in section \ref{section:constant_speed} makes sure that \(\apply{\lambda}{t} \approx \hat{\lambda}t\), we know that the whole spline is parametrized in such a way that the parameter \(t\) is a fraction of the curve's total arc length.
+Since the optimization goal in section \ref{section:constant_speed} makes sure that \(\apply{\lambda}{t} \approx \hat{\lambda}t\), the whole spline is parametrized in such a way that the parameter \(t\) is a fraction of the curve's total arc length.
 This fits perfectly with the fact that the position of each specification item is also specified as a fraction of the curve's total arc length.
-Thus, for each specification item (\(0 \leq i \leq n\)), we add one of the following error terms, depending on the type of the specification item:
+Thus, for each specification item (\(0 \leq i \leq n\)), one of the following error terms is added, depending on the type of the specification item:
 				\begin{equation*}
@@
 Thus, it is not suitable to be constrained to zero or bounded error and has to be stated as a minimization goal.
-Note that similar to the error term for constant speed, we added the factor \(\hat{\lambda}^2\), which causes the fairness measure to be invariant under scaling of the whole curve, enabling static weighting of error terms in the objective function.
+Note that similar to the error term for constant speed, the factor \(\hat{\lambda}^2\) is added, which causes the fairness measure to be invariant under scaling of the whole curve, enabling static weighting of error terms in the objective function.
 
@@
 
-			Having applied the usability criteria from section \ref{section:usability_criteria_curve_design_tools} to Bézier splines and Spiro splines, we will now use them to evaluate usability aspects of the prototype curve design tool.
+			Having applied the usability criteria from section \ref{section:usability_criteria_curve_design_tools} to Bézier splines and Spiro splines, it is now time to use them to evaluate usability aspects of the prototype curve design tool.
 
@@
 				One of the primary goals of this project was to explore more expressive description languages than those which are currently used in vector graphics software.
-To that end, our specification language supports specifying points, directions and curvatures at any number of positions on the curve, as well as the total length of the curve.
+To that end, the chosen specification language supports specifying points, directions and curvatures at any number of positions on the curve, as well as the total length of the curve.
 Furthermore, the fairness measure ensures constant and linearly changing curvature where possible.
@@
 
-				All these issues cause the prototype curve design tool to be somewhat impractical for actual curve design work, making this the primary issue with the chosen approach.
+				All these issues cause the prototype curve design tool to be somewhat impractical for actual curve design work, making this the primary issue with the current implementation.
 
@@
 
-			We will now revisit some of the examples shown in sections \ref{section:bézier_splines} and \ref{section:spiro_splines} using the prototype curve design tool in order to see how well the description language we propose can handle them.
+			It's time to revisit some of the examples shown in sections \ref{section:bézier_splines} and \ref{section:spiro_splines} using the prototype curve design tool in order to see how well the description language we propose can handle them.
 
@@
 Recall that the curve in example 4 was already designed using Bézier splines in section \ref{section:bézier_splines}, using 7 nodes and 12 handles, and Spiro splines in section \ref{section:spiro_splines}, using 7 control points.
-Here, it is designed using 2 point specifications, 1 direction specification, and 3 curvature specifications and specified arc length.
+Here, it is designed using 2 point specifications, 1 direction specification, 3 curvature specifications and specified arc length.
 Unfortunately, it is somewhat difficult to directly compare these amounts since specification items use fractions of the arc length for positioning instead of node/control point ordering, as is the case with Bézier splines and Spiro splines.
@@
 Furthermore, direction and curvature specification items only specify a scalar value instead of a plane vector, so could be considered to only supply half as much information as a point specification item.
-So we count 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving us a total of 11 scalar values for the entire specification, just shy of the 14 scalar values for the Spiro spline and a lot less than the 33 scalar values for the Bézier spline (where the loss of one degree of freedom for smooth nodes with colinear velocities has been taken into account).
+Thus, there is 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving a total of 11 scalar values for the entire specification, just short of the 14 scalar values for the Spiro spline and a lot less than the 33 scalar values for the Bézier spline (where the loss of one degree of freedom for smooth nodes with colinear velocities has been taken into account).
 Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgment on the amount of information that needs to be supplied to define a specific curve.
@@
 
-		Having analyzed the usability both through theoretical analysis and experimentally using the prototype curve design tool, we think that our approach looks promising.
-The added expressiveness of the proposed description language has certainly proven useful, facilitating the succinct description of curves which, when using conventional curve design tools, are harder to design and/or lack smoothness due to unfitting description languages.
+		Having analyzed the usability both through theoretical analysis and experimentally using the prototype curve design tool, we think that the chosen description language looks promising.
+The added expressiveness has certainly proven useful, facilitating the succinct description of curves which, when using conventional curve design tools, are hard to design and/or lack smoothness due to unfitting description languages.
 
@@
 The represented curve is then the image of this function.
-While there are many ways to parametrize a given curve, that is, represent it using a parametrization function, we are usually interested in those which yield a sufficiently smooth parametrization function, allowing for the concise description of curve properties using differential geometry.
+While there are many ways to parametrize a given curve, that is, represent it using a parametrization function, one is usually interested in those which yield a sufficiently smooth parametrization function, allowing for the concise description of curve properties using differential geometry.
 
@@
 
-			In the special case of parametric plane curves, that is, parametric curves in the euclidean plane, we can define the following functions:
+			In the special case of parametric plane curves, that is, parametric curves in the euclidean plane, one can define the following functions:
 			\begin{align*}
@@
 One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve.
-In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, we can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.
+In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, one can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.
 

### 2014-02-04 idp/documents/documentation/paper.tex
minor corrections

@@
 		\noindent We analyze the usability of curve design tools in current vector graphics software and suggest an abstract approach for designing improved, usability-oriented tools.
-We then instantiate this approach using a powerful description language and the minimum variation curve fairness measure on top of nonlinear optimization, developing a prototype curve design tool in the process.
-Finally, we evaluate our approach and its instantiation based on the prototype and suggest ideas for further research and development.
+We instantiate this approach using a powerful description language and the minimum variation curve fairness measure on top of nonlinear optimization on polynomials, developing a prototype curve design tool in the process.
+We evaluate our approach and its instantiation based on the prototype and suggest ideas for further research and development.
 
@@
 
-		Raster graphics images (for instance, digital photos and most digitally painted artworks) consist of a finite number of pixels (picture elements) which discretely specify the colors at each point of the image.
+		Raster graphics images (for instance, digital photos and most digitally painted artworks) consist of a finite number of pixels (picture elements) which discretely specify the colors of points on the image.
 In contrast, vector graphics designs consist of primitives like curves and surfaces, which describe the shapes present in the design using mathematical functions.
-This has various advantages over the raster approach, since designs do not suffer from discretization artifacts and can, for instance, be arbitrarily scaled or repeatedly modified without any loss of quality.
+This has various advantages over the raster approach, since designs do not suffer from discretization artifacts and can, for instance, be arbitrarily scaled and repeatedly modified without any loss of quality.
 
@@
 			Step 1 is of course independent of any software that may be used by the designer.
-Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as specifications for curves (the specification language the software uses).
+Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it uses for the specification of curves (the specification language the software uses).
 Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve, in step 4, the design tool will have to choose the curve that the user was most likely describing from a possibly infinite set of curves that exhibit the supplied properties.
@@
 
-			We conclude that the two ways in which curve design tools can affect the design process, are through their choice of specification language and fairness measure, as well as their practical ability to derive curves from them.
+			We conclude that the two ways in which curve design tools can affect the design process, are through their choice of specification language and fairness measure, as well as their practical ability to derive curves from specifications and the fairness measure.
 Note that while syntactically complete, the semantics of the specification language also depend on the fairness measure.
@@
 			Ideally, the description language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently.
-The fairness measure should capture the intuitive notion of curve smoothness sufficiently well and select curves that are minimal with respect to the specification.
+The fairness measure should capture the intuitive notion of curve smoothness sufficiently well and select curves that are minimal with respect to the given specification.
 These goals are competing with the software's ability to derive curves from descriptions reliably and efficiently, a process which becomes increasingly difficult with more expressive specification languages and more complex fairness measures.
@@
 				\item minimality of curves selected by the fairness measure
-				\item software's ability to derive curves from descriptions
+				\item ability of the software to derive curves from descriptions
 			\end{enumerate}
@@
 
-			Using the model of the curve design process as well as the criteria for usability of curve design software that were established in sections \ref{section:curve_design_process} and \ref{section:usability_criteria_curve_design_tools}, various tools available in current vector graphics software can be analyzed.
+			Using the model of the curve design process as well as the criteria for usability of curve design software that were established in sections \ref{section:curve_design_process} and \ref{section:usability_criteria_curve_design_tools}, some curve design tools available in common vector graphics software are analyzed.
 
@@
 One way to view things is to assume that the user, in supplying the four coefficient points of each curve segment, uniquely specifies each segment's Bézier curve and thus the entire spline, leaving no room for choosing a curve according to some fairness measure.
-Since users usually don't think about Bézier splines that way, it is more practical to assume that the user only supplies both point and velocity at the start and end of each curve segment, with a fairness measure designed to always choose the cubic Bézier curve that is uniquely specified by the given properties as the fairest curve.
+Since users usually do not think about Bézier splines that way, it is more practical to assume that the user only supplies both point and velocity at the start and end of each curve segment, with a fairness measure designed to always choose the cubic Bézier curve that is uniquely specified by the given properties as the fairest curve.
 
@@
 Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created or moved in order to change the overall curve.
-It also sacrifices smoothness, since it's easy to accidentally introduce subtle bumps in the curve when using a large number of nodes.
+It also sacrifices smoothness, since it is easy to accidentally introduce subtle bumps in the curve when using a large number of nodes.
 Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous.
@@
 The left curve is a Bézier spline approximation of a circular arc, exhibiting nearly constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
-Since the difference can be hard to see, it's easy to make mistakes when trying to approximate circular arcs using Bézier splines.
+Since the difference can be hard to see, it is easy to make mistakes when trying to approximate circular arcs using Bézier splines.
 
@@
 Since there is no way to directly specify curvature using Bézier splines, a considerable number of nodes is needed to get close to the desired curve.
-This in turn, may lead to decreased smoothness as discussed in the previous paragraphs.
+This, in turn, makes the curve hard to create and/or modify and may lead to decreased smoothness as discussed in the previous paragraphs.
 
@@
 Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties.
-Unfortunately, this is rarely practical, since it both requires a lot of time to set up and doesn't behave very well when modified.
+Unfortunately, this is rarely practical, since it both requires a lot of time to set up and does not behave very well when modified.
 These issues are somewhat alleviated by the guaranteed curvature continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
-That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidentally create small bumps as with Bézier splines.
+That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it is not as easy to accidentally create small bumps as with Bézier splines.
 
@@
 Unfortunately, describing the curve using Spiro splines still requires many control points.
-Note that even though this example has the same number of specified on-curve points (7) as the corresponding example in section \ref{section:bézier_splines}, the latter has an additional 12 velocity specifications, proving that even though curvature cannot be specified explicitly, Spiro splines are a much better choice for curves that are best expressed in terms of curvature than Bézier splines.
+Note that even though this example has the same number of specified on-curve points (7) as the corresponding example in section \ref{section:bézier_splines}, the latter has an additional 12 velocity specifications, proving that even though curvature cannot be specified explicitly, Spiro splines are a much better choice for shapes with approximately linearly changing curvature than Bézier splines.
 
@@
 				Unfortunately, as can be seen in figure \ref{figure:examples_spiro_lugia_2}, this scenario does not handle modifications very well.
-For instance, increasing the curvature of the blue segment has to be done indirectly by moving a control point.
-Unfortunately, this adversely affects the red segment, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
+For instance, increasing the curvature of the blue segment has to be done indirectly by moving a control point, which adversely affects the red segment, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.
 
@@
 
-		Having analyzed the usability aspect of curve design tools in section \ref{section:problem_analysis}, it's time to think about how to properly address their current issues.
+		Having analyzed the usability aspect of curve design tools in section \ref{section:problem_analysis}, it is time to think about how to properly address their current issues.
 In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as specific ideas for developing a piece of software using this approach.
@@
 			Figure \ref{figure:description-based_curves} shows such a process.
-Note that this approach doesn't give any specifics yet, since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like.
+Note that this approach does not give any specifics yet, since the previous usability analysis did not give a complete answer as to what the optimal curve design tool looks like.
 It has yet to be determined what the description language should be, as well as how to derive curves from these descriptions.
@@
 
-			After establishing the general approach to curve design tools in section \ref{section:description-based_curves}, it's time to decide on the specifics in order to create an instance of this model.
+			After establishing the general approach to curve design tools in section \ref{section:description-based_curves}, it is time to decide on the specifics in order to create an instance of this model.
 The first step is to choose a description language, consisting of both a specification language and a fairness measure.
@@
 
-				When only concerned with point specifications, it's sometimes enough to require the curve to go through all of them in order, although sometimes, this will lead to undesirable results.
-Even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points.
-When specifying direction and/or curvature without an associated point, it becomes a necessity to associate this specification with some kind of positional information in order to anchor it to some point on the curve.
-If this was not done, the specification could move freely along the curve and would thus behave very unpredictably.
+				When only concerned with point specifications, it may be enough to require the curve to go through all of them in order, although sometimes, even this will lead to undesirable results.
+That is because even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points.
+When specifying direction and/or curvature without an associated point, it becomes a necessity to associate these specifications with some kind of positional information in order to anchor them to points on the curve.
+If this was not done, the specifications could move freely along the curve and would thus behave very unpredictably.
 We thus chose to associate each combination of point, direction and curvature with a position on the curve.
 
-				With that in mind, it is now possible to abandon this notion of a combination of point, direction and curvature and instead talk about specification items, a specification item being either a positioned point, a positioned direction or a positioned curvature.
-For example, a combination of point and curvature at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
+				With that in mind, it is now possible to abandon this notion of a combination of point, direction and curvature specifications and instead talk about specification items, a specification item being a positioned point, direction or curvature specification.
+For example, a combination of point and curvature specification at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
 
@@
 				With this setup, it seems advisable to also include the curve's total arc length in the specification.
-This way, one can specify curves that don't have any specification items associated to their start and end points.
+This way, one can sensibly specify curves that have only one point specification item.
 Also, with unspecified curve length, changing any specification item may inadvertently change the curve's total arc length, thus possibly adversely affecting other parts of the curve whose specification item positions are given as a fraction of the curve's total arc length.
-For instance, with three point specification items on the curve, pulling the third one further away from the second will increase the total arc length of the curve, thus also increasing the arc length of the curve segment between the first and the second point specification item, thereby unintentionally creating a bulge between those two specification items.
+For instance, with three point specification items on the curve, pulling the third one further away from the second one would increase the total arc length of the curve, thus also increasing the arc length of the curve segment between the first and the second point specification item, thereby unintentionally creating a bulge between those two specification items.
 
@@
 						& \mathrm{Positions}          && = \unit\\
-						& \mathrm{Points}             && = \R{2}\\
-						& \mathrm{Directions}         && = \RR\\
-						& \mathrm{Curvatures}         && = \RR\\
+						& \mathrm{Points}             && = \xc{1} \times \R{2}\\
+						& \mathrm{Directions}         && = \xc{2} \times \RR\\
+						& \mathrm{Curvatures}         && = \xc{3} \times \RR\\
 						& \mathrm{SpecificationItems} && = \mathrm{Positions} \times \left(\mathrm{Points} \cup \mathrm{Directions} \cup \mathrm{Curvatures}\right)\\
@@
 				Before the advent of computer aided design, mechanical splines made from wood or steel were used in industrial design.
-Thus, people tried to emulate this behavior using mathematical curves, deriving a term for the bending energy in an idealized steel band, which is subsequently minimized.
+People tried to emulate this behavior using mathematical curves, deriving a term for the bending energy in an idealized steel band, which is subsequently minimized.
 This leads to the following functional describing the minimum energy curve (MEC) \cite{paper-mec}, which tries to minimize the curvature, and thus the bending energy:
@@
 Generally speaking, the MEC fairness measure is not very sensitive to bumps and inflection points on the curve, as long as they're small.
-The human visual system on the other hand, very much is and would always prefer one large arc over multiple small ones.
-Since the MEC tries to minimize curvature, it is also unsuited for the use with curvature specifications, as it will try to bring the curvature back to zero before and after a nonzero curvature specification.
+The human visual system on the other hand, very much is and would always prefer one large arc to multiple small ones.
+Since the MEC tries to minimize curvature, it is also not suited for the use with curvature specifications, as it will try to bring the curvature back to zero before and after a nonzero curvature specification.
 
@@
 					\caption{Examples of minimum energy curves under various constraints.
-Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.
+Point specifications are displayed as red dots, direction specifications are displayed as green tangent lines.
 The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
@@
 
-				Note the characteristic rectangular elastica shapes of the minimum energy curves in figure \ref{figure:minimum_energy_curve_examples}, as well as the fact that, in absence of both direction and curvature specification items, the curvature of a minimum energy curve is always zero at the endpoints.
+				Note the characteristic rectangular elastica shapes of the minimum energy curves in figure \ref{figure:minimum_energy_curve_examples}, as well as the fact that, in absence of both direction and curvature specifications, the curvature of a minimum energy curve is always zero at the endpoints.
 
@@
 In general, the MVC fairness measure is very good at avoiding bumps and inflection points when possible.
-It also works nicely with curvature constraints, since constant curvature is perfectly acceptable for the MVC fairness measure.
+It also works nicely with curvature specifications, since constant curvature is perfectly acceptable for the MVC fairness measure.
 
@@
 					\caption{Examples of minimum variation curves under the same constraints as the curves in figure \ref{figure:minimum_energy_curve_examples}.
-Point specification items are displayed as red dots, direction specification items are displayed as green tangent lines.
+Point specifications are displayed as red dots, direction specifications are displayed as green tangent lines.
 The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
@@
 				We think that the minimum variation curve fairness measure is better at producing curves that appear smooth to the human visual system than the minimum energy curve fairness measure.
-Thus, it produces curves that both look good and are close to what the user expects, given the specification, causing less surprise and frustration when designing curves.
+It produces curves that both look good and are close to what the user would expect, given the specification, causing less surprise and frustration when designing curves.
 Furthermore, it complements the chosen specification language nicely, since, in contrast to the minimum energy curve fairness measure, it works well with curvature specifications.
@@
 			Polynomial curves were chosen since they are mathematically simple, very flexible through the choice of their degree, and well-suited to approximate a variety of smooth curves.
-Since polynomials of high degree are known to have poor stability, polynomial splines were used, which allows increasing the segment count, rather than the degree, in order to approximate more complex curves.
+Since polynomials of high degree are known to behave poorly, polynomial splines were used, which allows increasing the segment count, rather than the degree, in order to approximate more complex curves.
 
@@
 Let \(\hat{t}_i\) be the position of the \(i\)th specification item.
-If the \(i\)th item is a point specification, the point is denoted \(\hat{\phi}_i\), if it is a direction specification, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification, the curvature is denoted \(\hat{\chi}_i\).
+If the \(i\)th item is a point specification item, the point is denoted \(\hat{\phi}_i\), if it is a direction specification item, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification item, the curvature is denoted \(\hat{\chi}_i\).
 The specified total arc length for the curve is denoted \(\hat{\lambda}\).
@@
 Unfortunately, this function is not very well-suited to be used as a term in a smooth optimization problem even for simple curves like polynomials.
-Also, since the function isn't fixed and varies with the coefficients, there is no way of telling on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
+Also, since the function is not fixed and varies with the coefficients, there is no way of telling on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
 The specification of the total arc length of the curve also needs to be stated as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
@@
 
-				Having taken care of the somewhat technical details arising from using polynomial splines, it's time to state the error terms for the specification items.
+				Having taken care of the somewhat technical details arising from using polynomial splines, it is time to state the error terms for the specification items.
 Since the optimization goal in section \ref{section:constant_speed} makes sure that \(\apply{\lambda}{t} \approx \hat{\lambda}t\), the whole spline is parametrized in such a way that the parameter \(t\) is a fraction of the curve's total arc length.
@@
 				\end{equation*}
-				It's clear that in many cases, this error term will not be zero, and in fact, may be arbitrarily large.
+				It is clear that in many cases, this error term will not be zero, and in fact, may be arbitrarily large.
 Thus, it is not suitable to be constrained to zero or bounded error and has to be stated as a minimization goal.
@@
 
-				Having gathered all the necessary error terms, it's time to put them together.
+				Having gathered all the necessary error terms, it is time to put them together.
 Terms with zero-error constraints, as well as terms with upper bounds on the error are added as component functions of the constraint function \(g\) together with appropriate bound components in \(g_l\) and \(g_u\).
@@
 				\subsubsection{Optimization Problem}
+					% TODO: \ref{section:fairness_measure} should be \ref{section:fairness_error}
 					Next, the optimization problem is created.
@@
 Furthermore, the fairness measure ensures constant and linearly changing curvature where possible.
-This grants the user much flexibility in describing the source curve, allowing them to specify those properties that can describe the curve in the most concise fashion.
+This grants the user much flexibility in describing the source curve, allowing them to specify those properties which can describe the curve in the most concise fashion.
 
@@
 Not only does this make the curve design process less tedious, but it may also lead to better curves, since only the essential information about the curve is given to the software.
-This allows the fairness measure to choose the best curve, something that is not possible if lots of auxiliary information is supplied to the software in order to indirectly influence the curvature.
+This allows the fairness measure to choose the best curve, something that is not possible if lots of information is supplied to the software in order to indirectly influence the curvature.
 
@@
 				On the other hand though, it is not possible not to specify the curve's arc length, and there may be cases where the user wants to leave it to the fairness measure to choose the best arc length for the curve.
-While it is sometimes possible to work around this issue by placing direction and/or curvature specifications instead of point specifications, that may not always be the case and further investigation into how arc length specification can be made optional while avoiding the problems mentioned in section \ref{section:specification_language} seems advisable.
+While it is sometimes possible to work around this issue by placing direction and/or curvature specification items instead of point specification items, that may not always be the case and further investigation into how arc length specification can be made optional while avoiding the problems mentioned in section \ref{section:specification_language} is necessary.
 
@@
 
-				Since curves can usually be described using only a few specifications, communicating this information to the curve design tool is not too difficult either.
+				Since curves can usually be described using only a few specification items, communicating this information to the curve design tool is not too difficult either.
 Points are well-suited to be specified using the mouse pointer.
-The current prototype curve design tool only supports very basic input facilities for direction and curvature, making it somewhat difficult to create and modify these specifications.
+The current prototype curve design tool only supports very basic input facilities for direction and curvature, making it somewhat difficult to create and modify these specification items.
 This is easily fixed though by utilizing their geometric interpretation, letting the user specify tangent lines and osculating circles instead of raw numeric values.
@@
 				Unfortunately, the nonlinear optimization approach to finding the best curve can only find local optima, such that in many cases, the curve is not minimal, often containing loops and unnecessary inflection points.
-However, this is usually easy to fix by adjusting the specifications in such a way that the global optimum is found.
-For instance, a loop in the curve can often be removed by placing an additional point specification on the loop, pulling it away from the curve until the curve is forced to straighten out and finally removing the point specification again.
+However, this is usually easy to fix by adjusting the specification in such a way that the global optimum is found.
+For instance, a loop in the curve can often be removed by placing an additional point specification item on the loop, pulling it away from the curve until it is forced to straighten out and finally removing the point specification item again.
 Once the curve has been pushed into the global optimum like this, it is usually very stable.
@@
 
-			It's time to revisit some of the examples shown in sections \ref{section:bézier_splines} and \ref{section:spiro_splines} using the prototype curve design tool in order to see how well the description language we propose can handle them.
+			It is time to revisit some of the examples shown in sections \ref{section:bézier_splines} and \ref{section:spiro_splines} using the prototype curve design tool in order to see how well the description language we propose can handle them.
 
@@
 				\caption{Examples of curves designed using the prototype curve design tool.
-Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle, or tangent lines if the curve is locally a straight line.
+Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle, or blue tangent lines if the curve is locally a straight line.
 The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
@@
 			Example 3 also starts out with a position that has all 3 types of specification items attached, but instead of specifying the curvature along the way, opts to use a point specification item to enforce the desired shape, resulting in nonlinear but extremely smooth change of curvature.
-This technique is useful in cases where either nonlinear changing curvature is desired, or where controlling the curve's shape using only curvature specifications is too tedious (since little errors in curvature specifications propagate and accumulate along the entire curve, making it hard to get the shape right).
-
-			Example 4 uses a mixture of the techniques used in examples 2 and 3, placing curvature specification items along the curve in order to not unnecessarily restrict the fairness measure's freedom and using a final point specification at the end of the curve to force it into place.
+This technique is useful in cases where either nonlinear changing curvature is desired, or where controlling the curve's shape using only curvature specification items is too tedious (since little errors in curvature specification items propagate and accumulate along the entire curve, making it hard to get the shape right).
+
+			Example 4 uses a mixture of the techniques used in examples 2 and 3, placing curvature specification items along the curve in order to not unnecessarily restrict the fairness measure's freedom and using a final point specification item at the end of the curve to force it into place.
 This approach ensures that no unintentional bumps are introduced (which can happen easily when placing point specification items in the middle of a curve), while still avoiding the tedious placing of perfectly adjusted curvature specification items (which would be necessary if the final point specification item was not there).
@@
 Recall that the curve in example 4 was already designed using Bézier splines in section \ref{section:bézier_splines}, using 7 nodes and 12 handles, and Spiro splines in section \ref{section:spiro_splines}, using 7 control points.
-Here, it is designed using 2 point specifications, 1 direction specification, 3 curvature specifications and specified arc length.
-Unfortunately, it is somewhat difficult to directly compare these amounts since specification items use fractions of the arc length for positioning instead of node/control point ordering, as is the case with Bézier splines and Spiro splines.
+Here, it is designed using 2 point specification items, 1 direction specification item, 3 curvature specification items and specified arc length.
+Unfortunately, it is somewhat difficult to directly compare these numbers since specification items use fractions of the arc length for positioning instead of node/control point ordering, as is the case with Bézier splines and Spiro splines.
 Fraction of arc length positions do of course contain more information than simple orderings, although this is arguably not the case if said positions lie exactly on the start or end of the curve.
 Furthermore, direction and curvature specification items only specify a scalar value instead of a plane vector, so could be considered to only supply half as much information as a point specification item.
-Thus, there is 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving a total of 11 scalar values for the entire specification, just short of the 14 scalar values for the Spiro spline and a lot less than the 33 scalar values for the Bézier spline (where the loss of one degree of freedom for smooth nodes with colinear velocities has been taken into account).
-Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgment on the amount of information that needs to be supplied to define a specific curve.
+Thus, there is 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving a total of 11 scalar values for the entire specification, just short of the 14 scalar values used in the Spiro spline example and a lot less than the 33 scalar values used in the Bézier spline example (where the loss of one degree of freedom for smooth nodes with colinear velocities has already been taken into account).
+Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgment of the amount of information that needs to be supplied to define a specific curve.
 
@@
 		Ideally, the final step of this research would then be the incorporation of the developed curve design tool into a piece of vector graphics software.
-The vector graphics editor Inkscape would be a good candidate, mainly because it's free software and thus easy to extend, both technically and from a political/licensing perspective.
+The vector graphics editor Inkscape would be a good candidate, mainly because it is free software and thus easy to extend, both technically and from a political/licensing perspective.
 
@@
 The represented curve is then the image of this function.
-While there are many ways to parametrize a given curve, that is, represent it using a parametrization function, one is usually interested in those which yield a sufficiently smooth parametrization function, allowing for the concise description of curve properties using differential geometry.
+While there are many ways to parametrize a given curve, that is, represent it using a parametrization function, one is usually interested in those which yield a sufficiently smooth parametrization function, allowing for the convenient description of curve properties using differential geometry.
 
@@
 			One may notice that this notion of continuity is a little too strict.
-A curve may look arbitrarily smooth, while its parametrization does not even have \(\contp{1}\) continuity, since the derivatives of the parametrization function may have discontinuities that do not manifest themselves in the curve that they represent.
+A curve may look arbitrarily smooth, while its parametrization does not even have \(\contp{1}\) continuity, since the derivatives of the parametrization function may have discontinuities that do not manifest themselves in the curve that it represents.
 If that is the case, one can always find a different parametrization of the same curve that does not have these discontinuities.
@@
 			\emph{Minimum Curvature Variation Curves, Networks, and Surfaces for Fair Free-Form Shape Design}\\
-			Hentry Packard Moreton\\
+			Henry Packard Moreton\\
 			University of California, Berkeley, 1992

### 2014-02-06 idp/documents/documentation/paper.tex
fixed reference

@@
 				\subsubsection{Optimization Problem}
-					% TODO: \ref{section:fairness_measure} should be \ref{section:fairness_error}
 					Next, the optimization problem is created.
-The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_measure}.
+The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_error}.
 Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the specifications of the curve are realized using a constraint function.

### 2014-02-12 idp/documents/documentation/paper.tex
Updated diagrams, added verbatim text, other work on documentation.

@@
 
+		In this section, the implementation of our prototype, referred to as Kurve in short, will be described.
+
 		\subsection{Architecture}
@@
 				Kurve offers its own APIs in C\# for working with terms and optimization problems.
-However, the implementations of the functionalities of these APIs are provided by third-party libraries.
+However, the implementation of the functionality of these APIs is provided by third-party libraries.
 Hence, Kurve relies on several libraries, including:
@@
 				\begin{itemize}
-					% TODO: this section could be shortened by simply stating that Ipopt can solve optimization problems as described in section \ref{section:optimization_problem}
 				  	\item Ipopt\footnote{https://projects.coin-or.org/Ipopt} is a library that does numeric non-linear optimization.
-This library finds optima of optimization problems based on an objective function and its first two derivatives.
-In addition, it allows the specification of hard constraint functions (and their first two derivatives) that the solutions have to match.
-This allows Kurve to use hard specifications on curves.
+It allows to solve optimization problems as described in \ref{section:optimization_problem}.
 					\item CasADi\footnote{http://casadi.org} provides functionality to perform automatic differentiation of terms.
@@
 					\centering
-					\includegraphics[width=\textwidth]{content/output/subsystems.pdf}
+					\includegraphics[width=110mm]{content/output/subsystems.pdf}
 					\caption{Subsystem Decomposition of the Kurve Project}
@@
 
-				Kurve's implementation is split into two primary subsystems: Kurve contains the user interface related code and has a dependency on GTK\# as well as Kurve.Curves; Kurve.Curves is responsible for generating curves from specifications using CasADi through Wrappers.Casadi.
-
-				Wrappers.Casadi and Wrappers.Casadi.Native contain a wrapper for CasADi using C\# and C++ code respectively.
-Wrappers.Casadi uses the functions defined in Wrappers.Casadi.Native.
-Wrappers.Casadi and Wrappers.Casadi.Native are now part of the general-purpose library Krach to allow easy reuse in other projects.
-
-		\section{Numerical Optimization}
-			The Curve.Curves project contains the code required to transform input specifications into curves.
-It relies on CasADi (through Krach.Wrappers.Casadi) to represent curves symbolically and on Ipopt for optimization.
-
-			\subsection{Specification}
-				\begin{figure}[htb]
-					\centering
-					\includegraphics[width=\textwidth]{content/output/specification_class_diagram.pdf}
+				Kurve's implementation is split into two primary subsystems: \verb|Kurve| contains the user interface related code and has a dependency on GTK\# as well as \verb|Kurve.Curves|; \verb|Kurve.Curves| is responsible for generating curves from specifications using CasADi through \verb|Wrappers.Casadi|.
+
+				 \verb|Wrappers.Casadi| and \verb|Wrappers.Casadi.Native| contain a wrapper for CasADi using C\# and C++ code respectively. \verb|Wrappers.Casadi| relies on \verb|Wrappers.Casadi.Native|.
+
+				 \verb|Wrappers.Casadi| and \verb|Wrappers.Casadi.Native| are now part of the general-purpose library Krach to allow easy reuse in other projects.
+
+			\subsection{Numerical Optimization}
+				The \verb|Curve.Curves| project contains the code required to transform input specifications into curves.
+It relies on CasADi (through \verb|Krach.Wrappers.Casadi|) to represent curves symbolically and on Ipopt for optimization.
+
+				\subsubsection{Specification}
+					\begin{figure}[htb]
+						\centering
+						\includegraphics[width=135mm]{content/output/specification_class_diagram.pdf}
 						\caption{Class Diagram of Specification-Related Classes}
@@
 					\begin{itemize}
-					\item Any number of point, direction and curvature specifications. A set of such CurveSpecifications is stored in a BasicSpecification.
-					\item The curve's total length is also part of a BasicSpecification.
-					\item The curves are represented using a number of segments, each of which is represented by some function, realized using the FunctionTermCurve class.
-Because of that, the number of segments as well as the function type that is used to represent each segment are part of the BasicSpecification as well.
-					The type of function used is represented using a FunctionTermCurveTemplate object; this template allows the creation of FunctionTermCurve objects.  
+						\item Any number of point, direction and curvature specifications. These \verb|CurveSpecification|s are stored in a \verb|BasicSpecification|.
+						\item The curve's total length is also part of a \verb|BasicSpecification|.
+						\item The curves are represented using a number of segments, each of which is represented by some function, realized using the \verb|FunctionTermCurve| class.
+Because of that, the number of segments as well as the function type that is used to represent each segment are part of \verb|BasicSpecification| as well.
+						The type of function used is represented using a \verb|FunctionTermCurveTemplate| object; this template allows the creation of \verb|FunctionTermCurve| objects.  
 						However, this information is only used during optimization; if the number of segments and function template yield a curve with enough degrees of freedom with respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed.
@@
 more degrees of freedom mean the curve can satisfy more curve specifications while maintaining constant speed and good fairness, but the optimization process takes longer).
-					\item Curves could be generated from a BasicSpecification alone.
-However, if there are multiple curves that fit a BasicSpecification very well (i.e.
-the optimization problem has multiple local optima), different actual curves could be generated from a BasicSpecification.
-This results in a number of problems: for example, when saving a curve, if only the BasicSpecification were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process.
+						\item Curves could be generated from a \verb|BasicSpecification| alone.
+However, if there are multiple curves that fit a \verb|BasicSpecification| very well (i.e.
+the optimization problem has multiple local optima), different actual curves could be generated from a \verb|BasicSpecification|.
+This results in a number of problems: eg., when saving a curve, if only the \verb|BasicSpecification| were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process.
 This is clearly undesirable and a way to disambiguate between the different optima is required.
-					For that reason, a full Specification consists of a BasicSpecification as well as some disambiguation information.
+						For that reason, a full \verb|Specification| consists of a \verb|BasicSpecification| as well as some disambiguation information.
 That disambiguation information is the result of the optimization, which can be used as a starting position for the optimization process when a curve is loaded.
@@
 
-			\subsection{Optimization}
-				\begin{figure}[htb]
-					\centering
-					\includegraphics[width=\textwidth]{content/output/optimization_class_diagram.pdf}
+				\subsubsection{Optimization}
+					\begin{figure}[htb]
+						\centering
+						\includegraphics[width=110mm]{content/output/optimization_class_diagram.pdf}
 						\caption{Class Diagram of Optimization-Related Classes}
@@
 
-				The Optimizer class handles the optimization process of a curve.
-By supplying a Specification, a concrete curve as well as a new Specification with updated disambiguation information can be obtained.
-The resulting curve is represented by a Curve object which allows access to the curve's points, speed, direction and curvature, but carries no information about the internal segmentation and function type used.
+					The \verb|Optimizer| class handles the optimization process of a curve.
+By supplying a \verb|Specification|, a concrete curve as well as a new \verb|Specification| with updated disambiguation information can be obtained.
+The resulting curve is represented by a \verb|Curve| object which allows access to the curve's points, speed, direction and curvature, but carries no information about the internal segmentation and function type used.
 
@@
 
-				\subsubsection{Segments}
-					The first step, OptimizationSegments, creates curve segments based on the number of segments and the segment template specified in the BasicSpecification.
+					\begin{description}
+						\item[Segments]
+						In the first step curve segments are created based on the number of segments and the segment template specified in the \verb|BasicSpecification|.
 A list of parameters of all segments is created; by applying concrete values to each of the parameters, a concrete curve can be obtained.
@@
 
-				\subsubsection{Optimization Problem}
+						\item[Optimization Problem]
 						Next, the optimization problem is created.
@@
 
-				\subsubsection{Substitutions}
+						\item[Substitutions]
 						In this step, substitutions for each of the placeholder variables (i.e.
@@
 
-				\subsubsection{Solver}
+						\item[Solver]
 						In this step, a solver for the optimization problem is created.
@@
 A new solver needs to be created whenever either the optimization problem or the substitutions change.
-
-				\subsubsection{Position}
+						\item[Position]
 							The position represents a result of an optimization process.
@@
 This curve is the desired end result returned by the Optimizer.
-
-		\section{User Interface}
-			\subsection{Implementation}
+					\end{description}
+
+			\subsection{User Interface Implementation}
 				\begin{figure}[htb]
@@
 				\begin{itemize} 
-					\item MainWindow - The MainWindow is a GTK\# window that contains the RootComponent.
-It triggers redraws of the RootComponent when appropriate (e.g.
-when the window is resized) and forwards mouse and keyboard events to the RootComponent.
-					\item Component - Component is the abstract base class of all components that make up Kurve's user interface.
+					\item \verb|MainWindow| - The \verb|MainWindow| is a GTK\# window that contains the \verb|RootComponent|.
+It triggers redraws of the \verb|RootComponent| when appropriate (e.g.
+when the window is resized) and forwards mouse and keyboard events to the \verb|RootComponent|.
+					\item \verb|Component| - \verb|Component| is the abstract base class of all components that make up Kurve's user interface.
 It provides default implementations for rendering components, handling mouse and keyboard events, and provides a mechanism to trigger redraws of subcomponents.
-					\item RootComponent - The RootComponent manages a list of CurveComponents and implements functionality to add and remove curve components.
+					\item \verb|RootComponent| - The \verb|RootComponent| manages a list of \verb|CurveComponent|s and implements functionality to add and remove curve components.
 Additionally, the RootComponent implements save and load functionality.
-					\item BackgroundComponent - This component can display a background image. This can be used to trace existing images with the Kurve tools.
-					\item CurveComponent - This component represents a curve in the user interface.
-It consists of any number of PositionedControlComponents (any number of SpecificationComponents and a FixedPositionComponent at both the start and the end of the curve).
-This yields segments of the curve - one segment between each pair of PositionedControlComponents.
-These segments are represented by SegmentComponents.
-A CurveComponent uses a CurveOptimizer to create a curve from the specifications given by the SpecificationComponents.
-It implements functionality to add and remove SpecificationComponents and to change the curve template used (e.g.
+					\item \verb|BackgroundComponent| - This component can display a background image. This can be used to trace existing images with the Kurve tools.
+					\item \verb|CurveComponent| - This component represents a curve in the user interface.
+It consists of several \verb|PositionedControlComponent|s (any number of \verb|SpecificationComponent|s and a \verb|FixedPositionComponent| at both the start and the end of the curve).
+This yields segments of the curve - one segment between each pair of \verb|PositionedControlComponent|s.
+These segments are represented by \verb|SegmentComponent|s.
+A \verb|CurveOptimizer| is used to create a curve from the specifications given by the \verb|SpecificationComponent|s.
+\verb|CurveComponent| implements functionality to add and remove \verb|SpecificationComponent|s and to change the curve template used (e.g.
 change the degree of the polynomial template).
-					\item LengthControlComponent - A LengthControlComponent is a Component that can modify the length of the Curve in a certain area.
-					\item PositionedControlComponent - A PositionedControlComponent is a component that has a specific position on the curve. 
-					\item SpecificationComponent - A SpecificationComponent allows to specify constraints on a curve.
+					\item \verb|LengthControlComponent| - A \verb|LengthControlComponent| is a \verb|Component| that can modify the length of the \verb|Curve| in a certain area.
+					\item \verb|PositionedControlComponent| - A \verb|PositionedControlComponent| is a component that has a specific position on the curve. 
+					\item \verb|SpecificationComponent| - A \verb|SpecificationComponent| allows to specify constraints on a curve.
 A point, direction and/or curvature can be specified for a given position using various mouse and keyboard commands.
-SpecificationComponents are visualized as a small black box that can be selected.
-					\item FixedPositionComponent - FixedPositionComponents represent the start and end of a CurveComponent.
-					\item SegmentComponent - A SegmentComponent represents a section of a curve delimited by two PositionedControlComponents.
+\verb|SpecificationComponent|s are visualized as a small black box that can be selected.
+					\item \verb|FixedPositionComponent| - \verb|FixedPositionComponent|s represent the start and end of a CurveComponent.
+					\item \verb|SegmentComponent| - A \verb|SegmentComponent| represents a section of a curve delimited by two PositionedControlComponents.
 The SegmentComponent renders the curve in that segment, but it also allows to modify the length of the curve in that segment.
-					\item DiscreteCurve and DiscreteCurveItem - A DiscreteCurve represents the data necessary to render a curve in the user interface.
+					\item \verb|DiscreteCurve| and \verb|DiscreteCurveItem| - A \verb|DiscreteCurve| represents the data necessary to render a curve in the user interface.
 It contains information about the curve's points, direction, and curvature.
-It is represented by a number of DiscreteCurveItems that sample a source curve's point, direction and curvature at certain positions.
-					The CurveOptimizer yields this type of curve.
-The reason why this kind of discrete curve is used is because the Curve objects and the Optimizer (from the Kurve.Curves subsystem) cannot be used at the same time.
+It is represented by a number of \verb|DiscreteCurveItem|s that sample a source curve's point, direction and curvature at certain positions.
+					The \verb|CurveOptimizer| yields this type of curve.
+The reason why this kind of discrete curve is used is because the \verb|Curve| objects and the \verb|Optimizer| (from the \verb|Kurve.Curves| subsystem) cannot be used at the same time.
 This, in turn, is caused by the CasADi library, which is not thread safe.
-However, CasADi is used for representing Curves internally, as well as for performing optimization tasks.
-					As Kurve uses a background thread to optimize curves, it is not possible to use the Curve objects yielded by the optimizer to render the curves in the main thread at the same time.
-For that reason, Curve objects are converted to DiscreteCurve objects in that background process.
-					\item CurveOptimizer and OptimizationWorker - As the optimization process required to generate a curve from specifications takes too much time to perform in the main thread (the user interface would be blocked for too long), optimization is performed in a background thread.
-Each CurveComponent uses a CurveOptimizer to submit changed specifications.
-A single shared OptimizationWorker handles the synchronization and makes sure each submitted specification gets processed one by one.
+However, CasADi is used for representing \verb|Curve|s internally, as well as for performing optimization tasks.
+					As Kurve uses a background thread to optimize curves, it is not possible to use the \verb|Curve| objects yielded by the optimizer to render the curves in the main thread at the same time.
+For that reason, \verb|Curve| objects are converted to \verb|DiscreteCurve| objects in that background process.
+					\item \verb|CurveOptimizer| and \verb|OptimizationWorker| - As the optimization process required to generate a curve from specifications takes too much time to perform in the main thread (the user interface would be blocked for too long), optimization is performed in a background thread.
+Each \verb|CurveComponent| uses a \verb|CurveOptimizer| to submit changed specifications.
+A single shared \verb|OptimizationWorker| handles the synchronization and makes sure each submitted specification gets processed one by one.
 It is necessary to ensure that no optimization processes run in parallel, as the CasADi library (which is used in this process) is not thread safe.
@@
 
-			\subsection{Usage}
+			\subsection{Using Kurve}
 
@@
 
-				% TODO: make this smaller, maybe use landscape format
-				\begin{figure}[p]
-					\centering
-					\includegraphics[width=\textwidth]{content/output/ui_components.pdf}
+				\begin{figure}[htb]
+					\centering
+					\includegraphics[width=130mm]{content/output/ui_components.pdf}
 					\caption{Overview of the User Interface in Kurve}
@@
 					A curve can be added by pressing `N'.
-A new Curve will appear in window that is initialized with a length of 100, and two specification controls specifying points.
+A new curve will appear in window that is initialized with a length of 100, and two specification controls specifying points.
 The curve uses one segment that uses a polynomial curve template with a degree of 10.
@@
 
-				% TODO: make this smaller, maybe use landscape format
-				\begin{figure}[p]
+				\begin{figure}[htb]
 					\centering

### 2014-02-12 idp/documents/documentation/paper.tex
Updated description based curves diagram, other fixes.

@@
 				\centering
-				\includegraphics[width=\textwidth]{content/output/description-based_curves.pdf}
+				\includegraphics[width=115mm]{content/output/description-based_curves.pdf}
 				\caption{Description-Based Curves}
@@
 
-				Kurve's user interface constructed from Components. Components may contain a number of subcomponents, render themselves into a GTK\# context, and handle mouse and keyboards events. 
+				Kurve's user interface is constructed from \verb|Component|s.
+\verb|Component|s may contain a number of subcomponents, render themselves into a GTK\# context, and handle mouse and keyboards events.
 				\begin{itemize} 

### 2014-02-12 idp/documents/documentation/paper.tex
fixed some small mistakes

@@
 
-					The current version of Kurve uses polynomials functions in each segment. The template that generates the functions and the number of segments can be modified as follows:
+					The current version of Kurve uses polynomial functions in each segment. The template that generates the functions and the number of segments can be modified as follows:
 
@@
 					\end{itemize}
-					Keep in mind that the segment count and the degree of the polynomial are not directly visible to the user and are used in the optimization process only.
-Adding more segments or increasing the degree of the polynomial allows more complicated curves, but has negative effects on the performance.
+					Keep in mind that the segment count and the degree of the polynomials are not directly visible to the user and are used in the optimization process only.
+Adding more segments or increasing the degree of the polynomials allows more complicated curves, but has negative effects on the performance.
 				\subsubsection{Specification Control}

### 2014-02-13 idp/documents/documentation/paper.tex
spell checking, fixed some verbatim environments

@@
 the optimization problem has multiple local optima), different actual curves could be generated from a \verb|BasicSpecification|.
-This results in a number of problems: eg., when saving a curve, if only the \verb|BasicSpecification| were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process.
+This results in a number of problems: e.g., when saving a curve, if only the \verb|BasicSpecification| were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process.
 This is clearly undesirable and a way to disambiguate between the different optima is required.
@@
 					\item \verb|RootComponent| - The \verb|RootComponent| manages a list of \verb|CurveComponent|s and implements functionality to add and remove curve components.
-Additionally, the RootComponent implements save and load functionality.
+Additionally, the \verb|RootComponent| implements save and load functionality.
 					\item \verb|BackgroundComponent| - This component can display a background image. This can be used to trace existing images with the Kurve tools.
@@
 \verb|SpecificationComponent|s are visualized as a small black box that can be selected.
-					\item \verb|FixedPositionComponent| - \verb|FixedPositionComponent|s represent the start and end of a CurveComponent.
-					\item \verb|SegmentComponent| - A \verb|SegmentComponent| represents a section of a curve delimited by two PositionedControlComponents.
-The SegmentComponent renders the curve in that segment, but it also allows to modify the length of the curve in that segment.
+					\item \verb|FixedPositionComponent| - \verb|FixedPositionComponent|s represent the start and end of a \verb|CurveComponent|.
+					\item \verb|SegmentComponent| - A \verb|SegmentComponent| represents a section of a curve delimited by two \verb|PositionedControlComponents|.
+The \verb|SegmentComponent| renders the curve in that segment, but it also allows to modify the length of the curve in that segment.
 					\item \verb|DiscreteCurve| and \verb|DiscreteCurveItem| - A \verb|DiscreteCurve| represents the data necessary to render a curve in the user interface.

### 2014-02-18 source/Kurve/Kurve/Components/CurveComponent.cs
relaxed polynomial template degree restriction

@@ -284,9 +284,9 @@ namespace Kurve.Component
 		{
 			if (Selected)
 			{
-				if (newDegree < 4)
+				if (newDegree < 1)
 				{
-					Console.WriteLine("polynomial template degree cannot be less than 4");
+					Console.WriteLine("polynomial template degree cannot be less than 1");
 
 					return;
 				}

### 2014-02-18 source/Kurve/Kurve/Components/RootComponent.cs

@@ -146,9 +146,9 @@ namespace Kurve.Component
 			int width, height;
 			parentWindow.GetSize(out width, out height);
 
-			Svg.Scene
+			Svg.Document
 			(
-				new Orthotope2Double(0, 0, width, height),
+				new Orthotope2Double(0, width, 0, height),
 				from optimizer in optimizers
 				from path in optimizer.GetSvgPaths()
 				select path

### 2014-02-20 idp/documents/documentation/paper.tex
froze date
added page breaks
fixed minor grammar error

@@
 \author{Julian Asamer, Julian Brunner}
-\date{\today}
+\date{February 13, 2014}
 
@@
 				Applying the criteria established in section \ref{section:usability_criteria_curve_design_tools}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice.
-The curve chosen by the fairness measure is also reasonably minimal (criterion 5), since the spline's complexity is limited by the number of segments that are used, which is in direct correspondence to the specification.
+The curve chosen by the fairness measure is also reasonably minimal (criterion 5), since the spline's complexity is limited by the number of segments that are used, which is in direct correspondence with the specification.
 Bézier splines are furthermore very simple to derive from their description, fulfilling criterion 6 perfectly.
@@
 
+	\clearpage
+
 	\begin{thebibliography}{}

### 2014-04-02 idp/documents/documentation/paper.tex
Included Yannick's feedback.

@@
 					\item Ability to interface with native code - native code can be interfaced with by using reasonable amounts of boilerplate code.
-As it was clear that native libraries would be used, this is a key requirement.
+Native libraries are used in Kurve, therefore this is a key requirement.
 					\item Familiarity - the developers are familiar with this language. This choice eliminates overhead that would be caused by adopting an unfamiliar language.
@@
 					\item CasADi\footnote{http://casadi.org} provides functionality to perform automatic differentiation of terms.
-It interfaces with Ipopt directly, which makes it a perfect match for Kurve's symbolic differentiation needs.
+It interfaces with Ipopt directly, which makes it the preferred choice for the symbolic differentiation required in Kurve.
 CasADi is written in C++, but unfortunately only C functions can be called easily from C\#.
@@
 
-				 Kurve has strong requirements for both libraries: Terms used in the functions that need to be optimized in Kurve are quite large and plentiful, but differentiation and optimization still need to be performed in a short period of time to allow real-time editing of curves.
+				 Kurve has strong requirements for both libraries: Terms used in the functions that need to be optimized in Kurve are large and plentiful, but differentiation and optimization still need to be performed in a short period of time to allow real-time editing of curves.
 Fortunately, CasADi and Ipopt are very efficient and allow Kurve to offer close to real time editing of curves.
@@
 				For Kurve's user interface, GTK\#\footnote{http://www.mono-project.com/GtkSharp} was used.
-GTK\# is used to draw the curves and controls into a window, as well as handling keyboard and mouse events.
+GTK\# provides functionality to draw the curves and controls into a window, as well as handling keyboard and mouse events.
 
@@
 			\subsection{Numerical Optimization}
-				The \verb|Curve.Curves| project contains the code required to transform input specifications into curves.
+				The \verb|Kurve.Curves| project contains the code required to transform input specifications into curves.
 It relies on CasADi (through \verb|Krach.Wrappers.Casadi|) to represent curves symbolically and on Ipopt for optimization.
@@
 Because of that, the number of segments as well as the function type that is used to represent each segment are part of \verb|BasicSpecification| as well.
-						The type of function used is represented using a \verb|FunctionTermCurveTemplate| object; this template allows the creation of \verb|FunctionTermCurve| objects.  
-						However, this information is only used during optimization; if the number of segments and function template yield a curve with enough degrees of freedom with respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed.
+						The type of function used is represented using a \verb|FunctionTermCurveTemplate| object. This template allows the creation of \verb|FunctionTermCurve| objects.  
+						However, this information is only used during optimization.
+If the number of segments and function template yield a curve with enough degrees of freedom with respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed.
 It is possible to specify the segment count and function template to modify the degrees of freedom that the resulting curves offers (i.e.
@@
 the optimization problem has multiple local optima), different actual curves could be generated from a \verb|BasicSpecification|.
-This results in a number of problems: e.g., when saving a curve, if only the \verb|BasicSpecification| were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process.
-This is clearly undesirable and a way to disambiguate between the different optima is required.
+One problem that can arise when saving a curve, if only the \verb|BasicSpecification| were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process.
+This is undesirable and a way to disambiguate between the different optima is required.
 						For that reason, a full \verb|Specification| consists of a \verb|BasicSpecification| as well as some disambiguation information.
 That disambiguation information is the result of the optimization, which can be used as a starting position for the optimization process when a curve is loaded.
-This both ensures that the same curve will be generated and speeds up the optimization process as well.
+This ensures that the same curve will be generated and speeds up the optimization process as well.
 					\end{itemize}
@@
 						In the first step curve segments are created based on the number of segments and the segment template specified in the \verb|BasicSpecification|.
-A list of parameters of all segments is created; by applying concrete values to each of the parameters, a concrete curve can be obtained.
-The generated segments are cached and do not have to be regenerated if the number of segments and segment template specified in the basic specification do not change.
+A list of parameters of all segments is created.
+By applying concrete values to each of the parameters, a curve can be obtained.
+The generated segments are cached and do not have to be regenerated if the number of segments and segment template specified in the \verb|BasicSpecification| do not change.
 
@@
 The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_error}.
-Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the specifications of the curve are realized using a constraint function.
+Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the curve's specificationsare realized using a constraint function.
 This step also computes the first and second derivatives of those functions.
-The concrete values required by the curve specifications are not used; instead, placeholder variables are introduced.
+The concrete values required by the curve specifications are not used.
+Instead, placeholder variables are introduced.
 For the total length of the curve a placeholder is used as well.
-This allows the reuse of the optimization problem if curve specifications or curve length change; the derivatives do not have to be recomputed.
+This allows the reuse of the optimization problem if curve specifications or curve length change.
+The derivatives do not have to be recomputed.
 Thus, the optimization problem is cached and only rebuilt if the optimization segments change or if the number of each type of constraint on the curve changes.
@@
 							The position represents a result of an optimization process.
-To obtain this, the solver created in the previous step is run; as the starting position, the previous optimization's position is used.
+To obtain this, the solver created in the previous step is run.
+As the starting position, the previous optimization's position is used.
 Using the obtained position, a concrete curve can be obtained from the optimization segments.
@@
 					The \verb|CurveOptimizer| yields this type of curve.
-The reason why this kind of discrete curve is used is because the \verb|Curve| objects and the \verb|Optimizer| (from the \verb|Kurve.Curves| subsystem) cannot be used at the same time.
+The reason why this kind of discrete curve is used is due to the lack of the ability to use \verb|Curve| and \verb|Optimizer| (from the \verb|Kurve.Curves| subsystem) at the same time.
 This, in turn, is caused by the CasADi library, which is not thread safe.
@@
 For that reason, \verb|Curve| objects are converted to \verb|DiscreteCurve| objects in that background process.
-					\item \verb|CurveOptimizer| and \verb|OptimizationWorker| - As the optimization process required to generate a curve from specifications takes too much time to perform in the main thread (the user interface would be blocked for too long), optimization is performed in a background thread.
+					\item \verb|CurveOptimizer| and \verb|OptimizationWorker| - The optimization process required to generate a curve from specifications takes too much time to perform in the main thread (the user interface would be blocked for too long), therefore optimization is performed in a background thread.
 Each \verb|CurveComponent| uses a \verb|CurveOptimizer| to submit changed specifications.
@@
 For that reason, most actions are performed using keyboard shortcuts.
-Many user interface elements can be selected with the mouse; this enables to use keyboard shortcuts that affect that component only.
+Many user interface elements can be selected with the mouse.
+This enables to use keyboard shortcuts that affect that component only.
 It is possible to interact with some components using the mouse.
@@
 					\begin{itemize}
-						\item To change the number of segments used, press `1' to decrease and `2' to increase the number of segments by one. ()
+						\item To change the number of segments used, press `1' to decrease and `2' to increase the number of segments by one.
 						\item To change the degree of the polynomial used in each segment, press alt + `1' to decrease, and alt + `2' to increase the degree by one.
@@
 					A segment also displays the speed of the curve if it differs from the optimal speed.
-Too high speed is indicated by a red hue; this can usually be fixed by inserting additional length.
-Too low speed is indicated by a blue hue; usually there will be both blue and red sections.
-This usually indicates that the curve is too complex (i.e.
+Too high speed is indicated by a red hue.
+This can usually be fixed by inserting additional length.
+Too low speed is indicated by a blue hue.
+Usually there will be both blue and red sections.
+This may indicate that the curve is too complex (i.e.
 there are too many specifications) to achieve a good representation with the current curve template.

### 2014-04-02 idp/documents/documentation/paper.tex
added some corrections to idp paper

@@
 All elements in two-dimensional vector graphics are described in terms of curves, so the design process of curves is the single most important part of two-dimensional vector graphics.
-Unfortunately, designing curves using the tools available in current vector graphics software can be very frustrating.
-The user is often unable to make the curve look the way they want it to.
+Unfortunately, designing curves using the tools available in current vector graphics software can be very frustrating, with the user being unable to achieve the desired curve shape.
 To make matters worse it is often unclear why this is the case, as the software seems to put all the necessary tools into the hands of the designer, yet at times, they somehow do not allow the user to communicate their intention to the software.
@@
 		While there has been some promising research into improving curve design tools \cite{thesis-mvc} \cite{thesis-spiro}, it focuses mostly on the aspect of improving the fairness of curves, rather than usability as a whole.
-Unfortunately, it also has not had much of an impact on actual vector graphics software used in practice, which continues to focus almost exclusively on Bézier splines to this day.
-
-		Thus, the objective of this project is to identify the exact nature and cause of the shortcomings of current curve design tools in order to use this knowledge to develop a curve design tool that does not exhibit them.
+Additionally, it has not had much of an impact on actual vector graphics software used in practice, which continues to focus almost exclusively on Bézier splines to this day.
+
+		Thus, the objective of this project is to identify the exact nature and cause of the shortcomings of current curve design tools in order to use this knowledge to develop better ones.
 
@@
 
-		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings so as to find a way to improve it without making the same or similar mistakes.
+		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings in order to find a way to improve it without making the same or similar mistakes.
 In the context of curve design tools, this can be achieved by analyzing and modeling the process that people use when designing curves.
@@
 			Step 1 is of course independent of any software that may be used by the designer.
-Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it uses for the specification of curves (the specification language the software uses).
+Step 2 and 3 are strongly influenced by the choice of design tool, as the software dictates which properties it uses for the specification of curves (the specification language the software uses).
 Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve, in step 4, the design tool will have to choose the curve that the user was most likely describing from a possibly infinite set of curves that exhibit the supplied properties.
@@
 Fairness usually encodes intuitive notions such as smoothness and minimality using mathematical concepts such as continuity and variation of curvature.
-Here, the software has another opportunity to influence the design process both positively and negatively depending on what fairness measure is chosen and how good the software is at reliably finding a curve that has a high fairness.
+At this point, the software has another opportunity to influence the design process both positively and negatively depending on what fairness measure is chosen as well as the ability of the software to find a curve that has a high fairness.
 
@@
 Instead of modeling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify.
-For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve.
+For example, in a curve design tool which takes two points from the user and connects them using a straight line, one possible view is that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve.
 Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \xp{p_1 - p_0} t\), thus uniquely identifying the desired curve.
@@
 
-			In a similar fashion, one may consider some shortcomings (for instance, insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request curvature continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).
+			In a similar fashion, one may attribute certain shortcomings (for instance, insufficient continuity) to either the specification language (the language does not allow the user to request curvature continuity) or the fairness measure (the fairness measure does not select sufficiently smooth curves).
 These decisions have to be made on a case by case basis, selecting the approach that is better suited to modeling the reality of curve design.
@@
 				Applying the criteria established in section \ref{section:usability_criteria_curve_design_tools}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice.
-The curve chosen by the fairness measure is also reasonably minimal (criterion 5), since the spline's complexity is limited by the number of segments that are used, which is in direct correspondence with the specification.
+The curve chosen by the fairness measure is reasonably minimal (criterion 5), as the spline's complexity is limited by the number of segments that are used, which is in direct correspondence with the specification.
 Bézier splines are furthermore very simple to derive from their description, fulfilling criterion 6 perfectly.
@@
 The problem is aggravated by the fact that curvature discontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes.
-In case the software allows the user to request curvature continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
-Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with description language expressiveness.
-
-				The description language used to specify Bézier splines in common design tools is lacking in expressiveness.
+In case the software does support the user by allowing them to request curvature continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result.
+Placing more nodes leads to some disadvantages that will be discussed further in the next paragraph.
+
+				The description language used to specify Bézier splines in common design tools lacks expressiveness.
 Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature).
 Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment.
-There is also no way to directly specify constant or linearly changing curvature.
-This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result.
-Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created or moved in order to change the overall curve.
+Additionally, there is no way to directly specify constant or linearly changing curvature.
+This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to achieve the desired result.
+Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, as many nodes and handles have to be created or moved in order to change the overall curve.
 It also sacrifices smoothness, since it is easy to accidentally introduce subtle bumps in the curve when using a large number of nodes.
@@
 				Figure \ref{figure:examples_bézier_arc} shows two Bézier splines.
-The left curve is a Bézier spline approximation of a circular arc, exhibiting nearly constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature.
-Since the difference can be hard to see, it is easy to make mistakes when trying to approximate circular arcs using Bézier splines.
+The left curve is a Bézier spline approximation of a circular arc, exhibiting nearly constant curvature, while the right one is not, changing from low curvature to high curvature and back to low curvature.
+As the difference can be hard to see, it is easy to make mistakes when trying to approximate circular arcs using Bézier splines.
 
@@
 				The red curve segment in figure \ref{figure:examples_bézier_lugia} is a Bézier spline with a curvature discontinuity which went unnoticed by the artist at the time of creation.
-In fact, this discontinuity would be even less visible if the curve was not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.
+In fact, this discontinuity would be even less visible if the curve was not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of the problem's origin.
 
@@
 
-				Spiro splines were added to Inkscape in 2009.
+				Spiro splines \cite{thesis-spiro} were added to Inkscape in 2009.
 Even though they were integrated using path effects and thus have a somewhat cumbersome user interface, they constitute a commendable effort of incorporating results from curve research into vector graphics software, something that has hardly happened this far.
-More information on Spiro splines can be found in the dissertation of Raphael Levien \cite{thesis-spiro}.
 
@@
 
-				Again, the model from section \ref{section:curve_design_process} is applied.
-With Spiro splines, the modeling is fairly straightforward.
+				In order to evaluate the usability of Spiro splines, the model from section \ref{section:curve_design_process} is applied once more.
+In the case of Spiro splines, the modeling is fairly straightforward.
 The user supplies a set of points for which the software computes an interpolating spline.
-The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is curvature continuity and no excessive winding.
+The fairness measure chooses the segments of this spline to be parts of the Euler spiral, ensuring curvature continuity and trying to avoid excessive winding.
 
 				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_criteria_curve_design_tools}.
-The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them are needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
+With the specification language of Spiro splines consisting of points on the curve, plus the fact that on sufficiently smooth curves, not very many of them are needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3).
 The fairness measure leads to beautifully smooth curves that are usually sufficiently close to what a human would expect given the control points (criteria 4 and 5).
-The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 6 well enough.
+The software is also reliable and quick with the construction of these curves.
+There are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves.
+It thus fulfills criterion 6 satisfactorily.
 The only real issue lies in the insufficient expressiveness of the description language (criterion 1), discussed further in the following paragraph.
@@
 Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature.
-These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
-This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itself excellently to the expression of circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
+These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties.
+This results in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature.
+This seems especially unfortunate, considering that the spline primitive of Spiro splines, the Euler spiral, would lend itself excellently to the expression of circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature.
 Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties.
 Unfortunately, this is rarely practical, since it both requires a lot of time to set up and does not behave very well when modified.
-These issues are somewhat alleviated by the guaranteed curvature continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
+These issues are somewhat alleviated by the guaranteed curvature continuity and the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved.
 That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it is not as easy to accidentally create small bumps as with Bézier splines.
@@
 Unfortunately, describing the curve using Spiro splines still requires many control points.
-Note that even though this example has the same number of specified on-curve points (7) as the corresponding example in section \ref{section:bézier_splines}, the latter has an additional 12 velocity specifications, proving that even though curvature cannot be specified explicitly, Spiro splines are a much better choice for shapes with approximately linearly changing curvature than Bézier splines.
+Note that even though this example has the same number of specified on-curve points (7) as the corresponding example for Bézier splines shown in figure \ref{figure:examples_bézier_fluttershy}, the latter has an additional 12 velocity specifications.
+This illustrates that even though curvature cannot be specified explicitly, Spiro splines are a much better choice for shapes with approximately linearly changing curvature than Bézier splines.
 
@@
 		Having analyzed the usability aspect of curve design tools in section \ref{section:problem_analysis}, it is time to think about how to properly address their current issues.
-In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as specific ideas for developing a piece of software using this approach.
+In the following sections, we propose both an abstract approach for designing usability-oriented curve design tools, as well as specific ideas for developing software using this approach.
 
@@
 
-			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority.
-Together with a good fairness measure, it forms a description language that, if it allows for reliable and efficient derivation of curves, will then lead to usable curve design tools.
+			We think that in order to get good usability, using a powerful specification language which does not rely on implementation details has to be the top priority.
+In combination with a good fairness measure, it forms a description language that, if it allows for reliable and efficient derivation of curves, will then lead to usable curve design tools.
 Since it is not immediately clear which description language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on descriptions.
@@
 				The specification language we propose uses differential geometric properties of parametric curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use.
-The specification language will focus on the properties point, direction and curvature as defined in \ref{section:parametric_plane_curves}.
+The specification language will focus on the properties point, direction and curvature as defined in section \ref{section:parametric_plane_curves}.
 In order to provide maximal flexibility, the specification language should allow to specify any combination of these three properties at any number of positions on the curve.
 
-				When only concerned with point specifications, it may be enough to require the curve to go through all of them in order, although sometimes, even this will lead to undesirable results.
+				When only concerned with point specifications, requiring the curve to go through all of them in order may be enough to allow the user to express their intent, although sometimes, even this will lead to undesirable results.
 That is because even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points.
 When specifying direction and/or curvature without an associated point, it becomes a necessity to associate these specifications with some kind of positional information in order to anchor them to points on the curve.
-If this was not done, the specifications could move freely along the curve and would thus behave very unpredictably.
+If this were not done, the specifications could move freely along the curve and would therefore behave very unpredictably.
 We thus chose to associate each combination of point, direction and curvature with a position on the curve.
 
-				With that in mind, it is now possible to abandon this notion of a combination of point, direction and curvature specifications and instead talk about specification items, a specification item being a positioned point, direction or curvature specification.
+				With that in mind, it is now possible to abandon this notion of a combination of point, direction and curvature specifications and instead talk about specification items, a specification item being a positioned point-, direction- or curvature specification.
 For example, a combination of point and curvature specification at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.
 
-				An obvious choice for specifying a position on a curve would be by giving a parameter value for the underlying parametric curve.
+				The most simple way of specifying a position on a curve would be by giving a parameter value for the underlying parametric curve.
 However, this would violate the design idea of using differential geometric properties of curves.
 A way of specifying a position on a curve that is independent from its parametrization is the arc length of the curve.
-Instead of directly using the arc length to give the positions of specification items, we chose to specify positions as fractions of the total arc length of the curve.
+Instead of directly using the arc length to describe the positions of specification items, we chose to specify positions as fractions of the total arc length of the curve.
 This way, the position of a specification item relative to the start and end points of the curve is independent of the total arc length of the curve.
@@
 Also, with unspecified curve length, changing any specification item may inadvertently change the curve's total arc length, thus possibly adversely affecting other parts of the curve whose specification item positions are given as a fraction of the curve's total arc length.
-For instance, with three point specification items on the curve, pulling the third one further away from the second one would increase the total arc length of the curve, thus also increasing the arc length of the curve segment between the first and the second point specification item, thereby unintentionally creating a bulge between those two specification items.
-
-				Thus, the final specification language is as follows:
+For instance, with three point specification items on the curve, pulling the third one further away from the second one would increase the total arc length of the curve.
+This would also increase the arc length of the curve segment between the first and the second point specification item, thereby unintentionally creating a bulge between those two specification items.
+
+				One way to formally represent the final specification language is as follows:
 				\begin{equation*}
@@
 
-				In order to address the issues of the MEC, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves.
+				In order to address the issues of the MEC, there has been some research into fairness measures whose design is motivated more directly by the human perception of smoothness in plane curves.
 One such approach is that of the minimum variation curve (MVC) \cite{thesis-mvc}, which tries to minimize the variation of the curvature.
@@
 The functional will put straight lines through points that are co-linear and circular arcs through points that are co-circular when otherwise unconstrained.
-When constraints cause different points on the curve to have different curvatures, the fairness measure will prefer linearly changing curvature, since that will make the derivative of the curvature a constant and thereby minimize the integral over its square.
+When constraints cause different points on the curve to have different curvatures, the fairness measure will prefer linearly changing curvature, as this will result in the derivative of the curvature being constant, thereby minimizing the integral over its square.
 In general, the MVC fairness measure is very good at avoiding bumps and inflection points when possible.
@@
 
-				The implementation we chose uses smooth optimization, or, more specifically, an interior point method.
+				The implementation we chose uses smooth optimization, specifically an interior point method.
 This type of algorithm further requires the functions \(f\) and \(g\) to be at least once, ideally twice continuously differentiable.
@@
 
-				Let \(n\) be the number of specification items.
+				In the following sections, let \(n\) be the number of specification items.
 Let \(\hat{t}_i\) be the position of the \(i\)th specification item.
@@
 
-				Since the position of a specification item is given as a fraction of the curve's total arc length, the function mapping the covered arc length to the corresponding parameter values is needed in order to state the optimization goals for the specification items.
+				As established in section \ref{section:specification_language}, the position of a specification item is given as a fraction of the curve's total arc length, the function mapping the covered arc length to the corresponding parameter values is needed in order to state the optimization goals for the specification items.
 Unfortunately, this function is not very well-suited to be used as a term in a smooth optimization problem even for simple curves like polynomials.
-Also, since the function is not fixed and varies with the coefficients, there is no way of telling on which segment each specification item will end up, something that is also difficult to state in a smooth optimization problem.
-The specification of the total arc length of the curve also needs to be stated as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
+Also, the function is not fixed and varies with the coefficients, so there is no way of telling on which segment each specification item will end up, something else that is difficult to state in a smooth optimization problem.
+Another problem is that the specification of the total arc length of the curve needs to be stated as an optimization goal, which may be difficult if the the covered arc length function is hard to express.
 Furthermore, the difficulty of expressing the fairness measure, which is often an integral over the arc length of the curve, is dependent on the complexity of the covered arc length function.
@@
 				For polynomial splines with sufficient degree and segment count, this term will be brought close to zero in the optimization, making the speed approximately constant with value \(\hat{\lambda}\).
-This, in turn, makes the covered arc length function approximately linear (\(\apply{\lambda}{t} \approx \hat{\lambda}t\)), facilitating the simple statement of the optimization goals for the specification items and the fairness measure as well as fixing the segment for each specification item.
-It also causes the total arc length of the curve to be approximately equal to \(\hat{\lambda}\), as requested by the specification.
+This, in turn, makes the covered arc length function approximately linear (\(\apply{\lambda}{t} \approx \hat{\lambda}t\)), facilitating the simple statement of the optimization goals for the specification items and the fairness measure.
+It also makes the segment for each specification item fixed.
+Additionally, it causes the total arc length of the curve to be approximately equal to \(\hat{\lambda}\), as requested by the specification.
 
@@
 				\end{equation*}
-				Since this term is intended to be used as a minimization goal, the factor \(\hat{\lambda}^{-2}\) was added, which makes it scale invariant.
-This enables the error term to be statically weighted against other terms in the objective function without this weighting being shifted because of curve scaling.
+				As this term is intended to be used as a minimization goal, the factor \(\hat{\lambda}^{-2}\) was added, which makes it scale invariant.
+This enables the error term to be statically weighted against other terms in the objective function without this weighting being shifted due to curve scaling.
 
@@
 Thus, to ensure \(\contg{2}\) continuity for the whole curve, it is sufficient to ensure \(\contg{2}\) continuity at the segment connection points.
-To do this, the following error terms are added for each connection point (\(0 \leq i \leq m - 1\)):
+To achieve this, the following error terms are added for each connection point (\(0 \leq i \leq m - 1\)):
 				\begin{equation*}
@@
 
-				Since it is not acceptable for the curve segments to not touch, have sharp corners or curvature discontinuities, the error terms above are added as zero-error constraints.
+				It is not acceptable for the curve segments to not touch, have sharp corners or curvature discontinuities, so the error terms above are added as zero-error constraints.
 
@@
 
-				Having taken care of the somewhat technical details arising from using polynomial splines, it is time to state the error terms for the specification items.
-Since the optimization goal in section \ref{section:constant_speed} makes sure that \(\apply{\lambda}{t} \approx \hat{\lambda}t\), the whole spline is parametrized in such a way that the parameter \(t\) is a fraction of the curve's total arc length.
+				With the technical details arising from the use of polynomial splines taken care of, the error terms for the specification items can be stated.
+Since the optimization goal in section \ref{section:constant_speed} ensures that \(\apply{\lambda}{t} \approx \hat{\lambda}t\), the whole spline is parametrized such that the parameter \(t\) is a fraction of the curve's total arc length.
 This fits perfectly with the fact that the position of each specification item is also specified as a fraction of the curve's total arc length.
@@
 
-				The user usually wants to specify exact curve behavior rather than indirectly influencing properties, so these error terms are also added as zero-error constraints.
+				The user usually wants to specify exact curve behavior rather than indirectly influencing properties, so these error terms are added as zero-error constraints.
 
@@
 
-				Since constant speed was established in section \ref{section:constant_speed}, stating the error term for the fairness chosen in section \ref{section:fairness_measure} is fairly straightforward:
+				After establishing constant speed in section \ref{section:constant_speed}, the error term for the fairness chosen in section \ref{section:fairness_measure} is well-suited for smooth optimization:
 				\begin{equation*}
@@
 				\end{equation*}
-				It is clear that in many cases, this error term will not be zero, and in fact, may be arbitrarily large.
-Thus, it is not suitable to be constrained to zero or bounded error and has to be stated as a minimization goal.
-Note that similar to the error term for constant speed, the factor \(\hat{\lambda}^2\) is added, which causes the fairness measure to be invariant under scaling of the whole curve, enabling static weighting of error terms in the objective function.
+				Whenever there is non-constant curvature, this error term will not be zero, and in fact, may be arbitrarily large.
+Thus, it is not suitable to be constrained to zero or be used as a bounded error and has to be stated as a minimization goal.
+Note that similar to the error term for constant speed, the factor \(\hat{\lambda}^2\) is added, which causes the fairness measure to be invariant under scaling of the whole curve.
+This enables static weighting of error terms in the objective function.
 
@@
 
-				Having gathered all the necessary error terms, it is time to put them together.
+				Having gathered all the necessary error terms, they are assembled in a nonlinear optimization problem.
 Terms with zero-error constraints, as well as terms with upper bounds on the error are added as component functions of the constraint function \(g\) together with appropriate bound components in \(g_l\) and \(g_u\).
 
-				Minimization goals, on the other hand, are assembled in a weighted sum which defines the the objective function \(f\).
+				Minimization goals, on the other hand, are assembled in a weighted sum which defines the objective function \(f\).
 Since the error terms were designed to be scale invariant, this weighting can be done statically.
@@
 
-			Having applied the usability criteria from section \ref{section:usability_criteria_curve_design_tools} to Bézier splines and Spiro splines, it is now time to use them to evaluate usability aspects of the prototype curve design tool.
+			Having applied the usability criteria from section \ref{section:usability_criteria_curve_design_tools} to Bézier splines and Spiro splines, they are finally used to evaluate the usability aspects of the prototype curve design tool.
 
@@
 Using this functionality, the user can specify curves which are best expressed in terms of curvature in a natural and concise way.
-Not only does this make the curve design process less tedious, but it may also lead to better curves, since only the essential information about the curve is given to the software.
+Not only does this make the curve design process less tedious, but it may also lead to better curves, since only the essential information about the curve is provided to the software.
 This allows the fairness measure to choose the best curve, something that is not possible if lots of information is supplied to the software in order to indirectly influence the curvature.
@@
 This way, placing curvature specifications on an otherwise unconstrained curve allows the user to precisely create spirals.
-Of course this behavior of the fairness measure also makes it possible to construct circular arcs in various ways, since they are the preferred shape of curve.
-
-				Section \ref{section:specification_language} describes why it makes sense to specify the curve's total arc length.
-While fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, there are actually quite a few upsides.
-For instance, by adjusting the arc length of a curve with a point specification item on each end, one can conveniently produce circular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced.
+This behavior of the fairness measure also makes it possible to construct circular arcs in various ways, since they are the preferred shape of curve.
+
+				Section \ref{section:specification_language} describes why it makes sense to force the specification of the curve's total arc length.
+Fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, however, there are actually quite a few upsides.
+For instance, by adjusting the arc length of a curve with a point specification item on each end, one can conveniently produce circular arcs between two points.
+This is because increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced.
 Generally speaking, having control over the arc length of the curve results in benign and predictable curve behavior and is often useful for pulling it in place, not unlike the way one would handle a mechanical spline.
@@
 				On the other hand though, it is not possible not to specify the curve's arc length, and there may be cases where the user wants to leave it to the fairness measure to choose the best arc length for the curve.
-While it is sometimes possible to work around this issue by placing direction and/or curvature specification items instead of point specification items, that may not always be the case and further investigation into how arc length specification can be made optional while avoiding the problems mentioned in section \ref{section:specification_language} is necessary.
+While it is sometimes possible to work around this issue by placing direction and/or curvature specification items instead of point specification items, that may not always be the case.
+Further investigation into how arc length specification can be made optional while avoiding the problems mentioned in section \ref{section:specification_language} is necessary.
 
@@
 
-				Points, directions and curvatures are easy to extract from some source curve.
-While extracting the arc length of a long curve may be somewhat difficult, this is usually not required since the arc length is mostly adjusted locally, where it is fairly easy to estimate.
+				Points, directions and curvatures are easy to extract from a given source curve.
+While extracting the arc length of a long curve may be difficult in some cases, this is usually not required since the arc length is mostly adjusted locally, where it is fairly easy to estimate.
 
@@
 Points are well-suited to be specified using the mouse pointer.
-The current prototype curve design tool only supports very basic input facilities for direction and curvature, making it somewhat difficult to create and modify these specification items.
-This is easily fixed though by utilizing their geometric interpretation, letting the user specify tangent lines and osculating circles instead of raw numeric values.
+The current prototype curve design tool only supports very basic input facilities for direction and curvature, resulting in some difficulties when creating and modifying these specification items.
+This could easily be fixed by utilizing their geometric interpretation, letting the user specify tangent lines and osculating circles instead of raw numeric values.
 Arc lengths are usually specified locally and by repeatedly making small adjustments, making them relatively easy to handle.
@@
 In example 2, the curve is described mainly using three curvature specification items, the point specification item and the direction specification item acting merely as a fixture for the curve, since the curvature specification items leave it unconstrained with regard to translation and rotation.
-This way, the curve is basically defined by its length and the three curvature specification items, giving the fairness measure maximal freedom controlling the curve, in this case allowing it to establish linearly changing curvature between each pair of curvature specification items.
+This way, the curve is basically defined by its length and the three curvature specification items, giving the fairness measure maximal freedom controlling the curve.
+In this case, this enables the fairness measure to establish linearly changing curvature between each pair of curvature specification items.
 
 			Example 3 also starts out with a position that has all 3 types of specification items attached, but instead of specifying the curvature along the way, opts to use a point specification item to enforce the desired shape, resulting in nonlinear but extremely smooth change of curvature.
-This technique is useful in cases where either nonlinear changing curvature is desired, or where controlling the curve's shape using only curvature specification items is too tedious (since little errors in curvature specification items propagate and accumulate along the entire curve, making it hard to get the shape right).
+This technique is useful in cases where either nonlinear change of curvature is desired, or where controlling the curve's shape using only curvature specification items is too tedious.
+The latter may be the case since little errors in curvature specification items propagate and accumulate along the entire curve, making it hard to get the shape right.
 
@@
 Fraction of arc length positions do of course contain more information than simple orderings, although this is arguably not the case if said positions lie exactly on the start or end of the curve.
-Furthermore, direction and curvature specification items only specify a scalar value instead of a plane vector, so could be considered to only supply half as much information as a point specification item.
-Thus, there is 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving a total of 11 scalar values for the entire specification, just short of the 14 scalar values used in the Spiro spline example and a lot less than the 33 scalar values used in the Bézier spline example (where the loss of one degree of freedom for smooth nodes with colinear velocities has already been taken into account).
+Furthermore, direction and curvature specification items only specify a scalar value instead of a plane vector, so they could be considered to only supply half as much information as a point specification item.
+Thus, there is 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving a total of 11 scalar values for the entire specification.
+This is just short of the 14 scalar values used in the Spiro spline example and a lot less than the 33 scalar values used in the Bézier spline example (where the loss of one degree of freedom for smooth nodes with colinear velocities has already been taken into account).
 Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgment of the amount of information that needs to be supplied to define a specific curve.
@@
 Another possible extension would be to allow specification items to be slightly violated in order to improve the curve's fairness (soft specifications).
-Further experiments using improved versions of the current description language and the prototype curve design tool are needed to make an informed decision here.
+Further experiments using improved versions of the current description language and the prototype curve design tool are needed to make an informed decision.
 It would also be advisable to do empirical studies with users of vector graphics software to assess the usability of description languages.
@@
 
-		Ideally, the final step of this research would then be the incorporation of the developed curve design tool into a piece of vector graphics software.
+		Ideally, the final step of this research would be the incorporation of the developed curve design tool into vector graphics software.
 The vector graphics editor Inkscape would be a good candidate, mainly because it is free software and thus easy to extend, both technically and from a political/licensing perspective.
@@
 			Parametric curves are described using functions \(\phi : \function{\unit}{\R{n}}\), called the parametrization function.
-The represented curve is then the image of this function.
+The represented curve is the image of this function.
 While there are many ways to parametrize a given curve, that is, represent it using a parametrization function, one is usually interested in those which yield a sufficiently smooth parametrization function, allowing for the convenient description of curve properties using differential geometry.
@@
 A curve may look arbitrarily smooth, while its parametrization does not even have \(\contp{1}\) continuity, since the derivatives of the parametrization function may have discontinuities that do not manifest themselves in the curve that it represents.
-If that is the case, one can always find a different parametrization of the same curve that does not have these discontinuities.
+If that is the case, it is always possible to find a different parametrization of the same curve that does not have these discontinuities.
 This insight gives rise to the definition of geometric continuity, denoted by \(\contg{n}\), which only requires that there exists a reparametrization of the parametric curve which is \(\contp{n}\) continuous.
@@
 			It follows immediately that every parametric curve that is \(\contp{n}\) continuous is also \(\contg{n}\) continuous.
-One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve.
+Observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve.
 In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, one can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.
